### 2023-10-10 niceAi.Rmd
initial set up using historical HMM modelling

---
title: "Untitled"
author: "Ismail Guennouni"
date: "2023-05-16"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(depmixS4)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
origin_dat <- read.csv("data/longFormData.csv")

origin_dat <- origin_dat %>% mutate(investment_bin = cut(investment, breaks = c(-1,2.5,7.5,12.5,17.5,21)),
                                    return_pct_bin = cut(return_pct, breaks = c(-.1,.125,.325,.51,.755,1.1)),
                                    next_investment = lead(investment, default=0),
                                            invPnL = return - investment,
                                            trPnL = 3*investment - return)
```

# Support functions and discrete Gaussian class construction 
```{r supportHMM, include=FALSE}
# order the states of the HMM to allow it to order them as low ret/mid/ret/high ret. 
label_switch <- function(mod,labels) {
  # labels is vector, first element is new label for original state 1, second is new label for original state 2, etc.
  if(!is(mod,"depmix") || !is(mod,"depmix.fitted")) stop("this function is for depmix models")
  n_states <- mod@nstates
  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
    stop("labels needs to be a vector of unique integers between 1 and", n_states)
  }
  inv_labels <- sapply(1:n_states,function(x) which(labels == x))
  tmp <- mod
  # relabel prior
  ppars <- getpars(mod@prior)
  fpars <- getpars(mod@prior,which="fixed")
  out_pars <- as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels]))
  out_fixed <- as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels]))
  if(!tmp@prior@family$link=="identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
  # relabel transition
  for(i in 1:n_states) {
    ppars <- getpars(mod@transition[[inv_labels[i]]])
    fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
    out_pars <- c(out_pars,as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels])))
    out_fixed <- c(out_fixed,as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels])))
    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
    if(!tmp@transition[[i]]@family$link=="identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
    #out_pars <- c(out_pars,getpars(mod@transition[[inv_labels[i]]]))
  }
  # relabel response
  for(i in 1:n_states) {
    out_pars <- c(out_pars,unlist(lapply(mod@response[[inv_labels[i]]],getpars)))
    out_fixed <- c(out_fixed,unlist(lapply(mod@response[[inv_labels[i]]],getpars,which="fixed")))
  }
  tmp <- setpars(tmp,out_fixed,which="fixed")
  tmp <- setpars(tmp,out_pars)
  if(is(tmp,"depmix.fitted")) tmp@posterior <- viterbi(tmp)
  return(tmp)
}

# ORDERING STATES WHEN RESPONSE IS BINS
order_mod <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation (5 bins, for binned returns)
    sum[i] = sum(1:5*mod@response[[i]][[1]]@parameters$coefficients)
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum))
  return(mod)
}

# ORDERING STATES WHEN RESPONSE IS GAUSSIAN
order_mod_gauss <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation 
    sum[i] = mod@response[[i]][[1]]@parameters$mu
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum))
  return(mod)
}

```

```{r, include=F}
# define a response class which only contains the standard slots, no additional slots
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

# define a generic for the method defining the response class

setGeneric("discgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("discgaus"))

# define the method that creates the response class

setMethod("discgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("discgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks)
            mod
          }
)

setMethod("show","discgaus",
          function(object) {
            cat("Gaussian with discrete support\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

setMethod("dens","discgaus",
          function(object,log=FALSE) {
            p <- pnorm(object@breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@breaks[-length(object@breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

setMethod("setpars","discgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values)!=npar) stop("length of 'values' must be",npar)
            # determine whether parameters or fixed constraints are being set
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

setMethod("getpars","discgaus",
          function(object,which="pars",...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

setMethod("fit","discgaus",
          function(object,w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object,c(pars[1], exp(pars[2])))
            object
          }
)

setClass("truncdiscgaus", contains="discgaus", slots=c(min="numeric", max="numeric"))

setMethod("dens","truncdiscgaus",
          function(object,log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            p <- p/sum(p)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

setGeneric("truncdiscgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("truncdiscgaus"))

setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20, ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)
```

```{r}
# Specify simple models to copy transition and prior
inv_simple <- list()
for(i in 2:6) {
  inv_simple[[i]] <- depmix(investment_bin ~ 1, data = origin_dat, nstates = i, transition = ~ invPnL, family=multinomial("mlogit"), ntimes = rep(10,381))
}

# Specify truncated gaussian models for 3 state model 
rModels3 <- list(
  list(
    truncdiscgaus(origin_dat$investment,pstart=c(5,2), min=-.5, max=20.5) # first guassian initialised with mean of 5 and sd of 2
  ),
  list(
    truncdiscgaus(origin_dat$investment,pstart=c(10,2), min=-.5, max=20.5)
  ),
  list(
    truncdiscgaus(origin_dat$investment,pstart=c(15,2), min=-.5, max=20.5)
  )
)

inv_dmod_3 <- makeDepmix(response=rModels3,transition=inv_simple[[3]]@transition,prior=inv_simple[[3]]@prior,ntimes = rep(10,381), homogeneous=FALSE)

inv_fdmod_3 <- fit(inv_dmod_3, emcontrol=em.control(random.start = FALSE))  #random.start = FALSE -> use parameters supplied as starting values

mod_3S <- order_mod_gauss(inv_fdmod_3)

```

```{r, include=F}
#ORDER THE STATES OF THE HMM

state1 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[1]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))

state2 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[2]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))

state3 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[3]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))

investment <- seq(0:20) - 1

inv_resp_3S <- as.data.frame(cbind(investment,state1,state2,state3)) %>% 
  pivot_longer(cols=c("state1","state2","state3"),
                    names_to='investor_state',
                    values_to='probability')

```

```{r}
inv_resp_3S %>% mutate( investor_state = fct_recode(investor_state, "low-trust" = "state1", "medium-trust" = "state2", "high-trust" = "state3"))%>%
        ggplot( aes(x = investment,
                    y = probability,
                    fill = investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state') +
  theme_bw() + 
  theme(legend.position = "bottom")

```

Let's write a function to plot transition functions given parameters

```{r}
# Here pars takes the form of a list ns matrices, one for each state from 1 to ns.
plot_transitions_PnL <- function(fmod) {

  ns <- nstates(fmod)

  trans_prob <- data.frame(
    from = rep(1:ns, each=80*ns),
    to = rep(1:ns, each=80),
    PnL = seq(-20,60,length=80),
    prob = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=80)

  for(from in 1:ns) {
    pars <- matrix(getpars(fmod)[seq(ns + 1 + (from - 1)*2*ns, ns + 1 + (from - 1)*2*ns + 2*ns - 1)],ncol=2)
    for(to in 1:ns) {
      x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"PnL"]
      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)
    for(to in 1:ns) {
      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  trans_prob <- trans_prob %>% mutate(from=recode(from, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust"),
                        to=recode(to, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust")) %>% 
    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))

#     iris %>%
#    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>%
# ggplot() + 
#    geom_histogram(aes(Petal.Width))+ 
#    facet_grid(Species~.)
#   
  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1)) +
      theme(legend.position = "bottom")
}
```

```{r}
plot_transitions_PnL(mod_3S)
```
```{r}
ns <- 3
pars <- NULL

for (from in 1:ns){
  Tr_mat <- matrix(getpars(mod_3S)[seq(ns + 1 + (from - 1)*2*ns, ns + 1 + (from - 1)*2*ns + 2*ns - 1)],ncol=2)

colnames(Tr_mat) <- c("intercept","invPnL")
rownames(Tr_mat) <- c("to_1", "to_2", "to_3")
pars[[from]] <- Tr_mat

}

pars[[2]]["to_2","invPnL"]

```

## Write function to plot transition probs based on input of trans matrix

```{r}
plot_transitions_input <- function(pars_list) {

  ns <- nrow(pars_list[[1]])  # number of states

  trans_prob <- data.frame(
    from = rep(1:ns, each=80*ns),
    to = rep(1:ns, each=80),
    PnL = seq(-20,60,length=80),
    prob = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=80)

  for(from in 1:ns) {
    pars <- pars_list[[from]]
    for(to in 1:ns) {
      x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"PnL"]
      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)
    for(to in 1:ns) {
      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  trans_prob <- trans_prob %>% mutate(from=recode(from, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust"),
                        to=recode(to, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust")) %>% 
    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))

  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))+
      theme(legend.position = "bottom")
}

```

```{r}
# pars has the format [[from]]["to","x"]
pars[[1]]["to_1","invPnL"] <- 0
pars[[1]]["to_1","intercept"] <-  10

pars[[2]]["to_1","intercept"] <- 3
#pars
plot_transitions_input(pars)
```

### 2024-02-13 screenCode.Rmd
more screening

---
title: "screeningAnalysis"
author: "Ismail Guennouni"
date: "2023-11-20"
output: pdf_document
---

```{r setup2, include=FALSE}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(depmixS4)
library(lubridate)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
#rsq_dat <- read.csv("data/RSQ_Feb06_24.csv") 
rsq_dat <- read.csv("data/RSQ_Feb13_24.csv")

# Step 1: Remove rows 1 and 2
df_screening <- rsq_dat[-c(1, 2), ]

df_screening <- df_screening %>%
  mutate(
    MH = if_else(as.Date(ymd_hms(StartDate)) > as.Date("2023-11-28"), 1, 0),
    MH.f = factor(MH, levels = c(0, 1), labels = c("Gen pop", "MH Condition"))
  )

# Rename columns Q4 to Q12 to Q1 to Q9
old_names <- paste0("Q", 4:12)
new_names <- paste0("Q", 1:9)
names(df_screening)[names(df_screening) %in% old_names] <- new_names

# Rename the "Duration..in.seconds." column to "duration"
names(df_screening)[names(df_screening) == "Duration..in.seconds."] <- "duration"

# Step 2: Filter rows
df_screening <- df_screening[df_screening$Q1 == "I accept" & !grepl('test', df_screening$Q2) & df_screening$DistributionChannel ==  'anonymous', ]
cat("valid submissions: ",  nrow(df_screening), "\n")

# Calculate the number of valid responses (non-NA in all columns)
#valid_responses_count <- sum(complete.cases(df_screening))

###########################
# Step 3: Identifying responses where both Q_att_1 and Q_att_2 are less than or equal to 6
unsuccessful_responses <- df_screening$Q_att_1 < 6 | df_screening$Q_att_2 < 6

# Calculating the number of unsuccessful responses
unsuccessful_count <- sum(unsuccessful_responses)

# Print the number of unsuccessful responses
cat("Number of responses failing attention checks:", unsuccessful_count, "\n")

###########################

# # Convert duration to numeric (if it's not already)
# df_screening$duration <- as.numeric(as.character(df_screening$duration))
# 
# # Calculate the median duration for the whole dataset
# median_duration_total <- median(df_screening$duration, na.rm = TRUE)
# 
# # Calculate the median duration for those who failed the attention checks
# median_duration_unsuccessful <- median(df_screening$duration[unsuccessful_responses], na.rm = TRUE)
# 
# # Print the median durations
# cat("Median duration for the whole dataset:", median_duration_total, "seconds\n")
# cat("Median duration for unsuccessful responses:", median_duration_unsuccessful, "seconds\n")

#########################

# Filtering out these unsuccessful responses
df_screening<- df_screening[!unsuccessful_responses, ]

# Print the number of valid responses
cat("Number of valid responses:", nrow(df_screening), "\n")

```
```{r}
# Step 4: Calculate rejection sensitivity score
# Identifying columns ending with _1 and _2, ignoring Q_att_1 and Q_att_2
one_columns <- grep('_1$', names(df_screening), value = TRUE)
two_columns <- grep('_2$', names(df_screening), value = TRUE)
one_columns <- setdiff(one_columns, c('Q_att_1', 'Q_att_2'))
two_columns <- setdiff(two_columns, c('Q_att_1', 'Q_att_2'))

# The scale is from 1 to 6 for b columns, we reverse it by subtracting from 7
df_screening[two_columns] <- lapply(df_screening[two_columns], function(x) 7 - as.numeric(x))

# Calculate the rejection sensitivity score for each situation
for (i in 1:length(one_columns)) {
    score_col <- paste0(substr(one_columns[i], 1, nchar(one_columns[i])-2), '_score')
    df_screening[[score_col]] <- as.numeric(df_screening[[one_columns[i]]]) * as.numeric(df_screening[[two_columns[i]]])
}

# Step 5: Create a new column for the mean of the rejection sensitivity scores
score_columns <- grep('_score$', names(df_screening), value = TRUE)
df_screening$RS_score <- rowMeans(df_screening[, score_columns], na.rm = TRUE)

# Display the first few rows of the updated dataframe
head(df_screening)

```

```{r}

# Plot histogram of mean rejection sensitivity with overlaid density plot
ggplot(df_screening, aes(x = RS_score)) +
    geom_histogram(aes(y = after_stat(density)), binwidth = 1, fill = 'skyblue', color = 'black', alpha = 0.6) +
    geom_density(color = 'red', size = 1.5, alpha = 0.7) +
    theme_minimal() +
    labs(title = 'Histogram and Density of Mean Rejection Sensitivity Scores',
         x = 'Mean Rejection Sensitivity Score',
         y = 'Density') +
    theme(plot.title = element_text(hjust = 0.5),
          axis.title.x = element_text(size = 12, face = "bold"),
          axis.title.y = element_text(size = 12, face = "bold")) + 
   facet_wrap(~ MH.f)

```

```{r}
# Calculate the mean of the RS_score 
mean_score <- round(mean(df_screening$RS_score, na.rm = TRUE),2)

# Calculate the standard deviation
std_dev <- round(sd(df_screening$RS_score, na.rm = TRUE),2)

# Count the number of respondents with scores above 15 (inclusive)
count_above_15 <- sum(df_screening$RS_score >= 15, na.rm = TRUE)

# Count the number of respondents more than 1 standard deviation above the mean
count_above_1sd <- sum(df_screening$RS_score > (mean_score + std_dev), na.rm = TRUE)

# Count the number of respondents more than 2 standard deviations above the mean
count_above_2sd <- sum(df_screening$RS_score > (mean_score + 2 * std_dev), na.rm = TRUE)

# Print the results
cat("Mean of the mean rejection sensitivity scores:", mean_score, "\n")
cat("SD:", std_dev, "\n")
cat("Number of respondents with scores >= 15:", count_above_15, "\n")
cat("Number of respondents > 1 standard deviation above the mean:", count_above_1sd, "\n")
cat("Number of respondents > 2 standard deviations above the mean:", count_above_2sd, "\n")

```
```{r}
# Count the number of respondents with scores above 15 (inclusive) in both samples
count_above_15_normal <- round(
  (df_screening %>% dplyr::filter(MH == 0, RS_score >= 15) %>% nrow()) / 
  (df_screening %>% dplyr::filter(MH == 0) %>% nrow()) * 100, 1)

count_above_15_MH <- round(
  (df_screening %>% dplyr::filter(MH == 1, RS_score >= 15) %>% nrow()) / 
  (df_screening %>% dplyr::filter(MH == 1) %>% nrow()) * 100, 1)

cat("Proportion of HEALTHY respondents with scores >= 15 on RSQ :", count_above_15_normal, "%\n")
cat("Proportion of MH respondents with scores >= 15 on RSQ :", count_above_15_MH, "%\n")

```

```{r}
# Return Prolific IDs for those above a certain score:

high_RS <- df_screening %>% dplyr::filter(RS_score >= 15) %>% dplyr::select(PROLIFIC_PID)

high_RS

write.csv(high_RS, "data/high_RS.csv", sep = ",", row.names = FALSE,quote = FALSE)

```
```{r}
# Old high RS participants identified in screening in November/December 2023
high_RS_old <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) != as.Date("2024-02-06") ) %>% dplyr::select(PROLIFIC_PID)

# Newly identified 06 feb 2024
high_RS_new <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) == as.Date("2024-02-06") ) %>% dplyr::select(PROLIFIC_PID)

# String them together and add to prolific
new_ids_string <- paste(high_RS_new, collapse = ",")
ids_string
new_ids_string

```

```{r}
# Load csv of IDs of already collected data from high RSQ in early December
highRS_collected <- read.csv("data/highRSQ_collected.csv") 
```

```{r}
# Find IDs in high_RS that are not in highRS_collected
missing_ids <- !high_RS$PROLIFIC_PID %in% highRS_collected$Participant.id

# Subset high_RS to get the dataframe of people you haven't collected data for
not_collected <- high_RS[missing_ids, ]

nrow(not_collected)

ids_string <- paste(not_collected, collapse = ",")
ids_string
```

# Newly identified high RS on the 13th feb 2024.

```{r}

high_RS_13feb <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) == as.Date("2024-02-13") ) %>% dplyr::select(PROLIFIC_PID)

nrow(high_RS_13feb)

# String them together and add to prolific
feb13_ids_string <- paste(high_RS_13feb$PROLIFIC_PID, collapse = ",")

feb13_ids_string
```

# Select those with low rejection sensitivity defined as below mean of 10 on the RSQ (...by the 6th of feb)
```{r}

# Return Prolific IDs for those above a certain score:

low_RS <- df_screening %>% 
          dplyr::filter(RS_score < 10) %>% 
          dplyr::select(PROLIFIC_PID)

# Extract the PROLIFIC_PID values as a character vector
low_RS_vector <- low_RS$PROLIFIC_PID

# Create the comma-separated string
low_RS_string <- paste(low_RS_vector, collapse = ",")

# Count the number of elements in the extracted PROLIFIC_PID vector
num_ids <- length(low_RS_vector)

cat("Number of extracted IDs: ", num_ids)

```

### 2024-02-13 article.Rmd
more screening

---
title: "(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games"
author:
  "Ismail Guennouni^1^, Christoph Korn, Georgia Koppe"

keywords:
- Social Trust
- Repair of Cooperation
- Cognitive Intervention 
- Economic Games 
- Hidden Markov Models

abstract: |

bibliography: "bib/cogIntervention.bib"
biblio-style: spphys

output: pdf_document
header-includes:
  - \usepackage{lscape}
---

```{r setupCoax, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
knitr::opts_chunk$set(out.width = "\\textwidth")
```

```{r load-packages05, include = FALSE}
library(papaja)
library(kableExtra)
require(knitr)
#require(citr)
require(bookdown)

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(grid)
library(gridExtra)
library(forcats)
library(ggsignif)
#library(magick)

```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(42)
options(tinytex.verbose = TRUE)
```

```{r}

d_finished <- readRDS("data/d_finished.csv")
exposure_intervention <- paste0(as.character(d_finished %>% dplyr::select(treatment.intervention) %>% unique()) , as.character(d_finished %>% dplyr::filter(gameNumber == "11", isLastRound) %>% dplyr::select(game_opponent) %>% unique()))

d_finished <- d_finished  %>% mutate(gameNum.f = factor(gameNumber,labels = c("pre", "expo1", "expo2", "expo3","post"),levels=c("1", "11", "12", "13", "2")), 
                              condition  = exposure_intervention) %>% 
                        rename(# Before Exposure
                            rating_cooperative_1 = `ratePLayer_trustGame_1.How cooperative do you think the other player is?`,
                            rating_forgiving_1 = `ratePLayer_trustGame_1.How forgiving do you think the other player is?`,
                            rating_playAgain_1 = `ratePLayer_trustGame_1.If given the choice, would you like to play again with the other player?`,

                            # During Exposure
                            rating_cooperative_11 = `ratePLayer_trustGame_11.How cooperative do you think the other player is?`,
                            rating_forgiving_11 = `ratePLayer_trustGame_11.How forgiving do you think the other player is?`,
                            rating_playAgain_11 = `ratePLayer_trustGame_11.If given the choice, would you like to play again with the other player?`,
                            rating_cooperative_12 = `ratePLayer_trustGame_12.How cooperative do you think the other player is?`,
                            rating_forgiving_12 = `ratePLayer_trustGame_12.How forgiving do you think the other player is?`,
                            rating_playAgain_12 = `ratePLayer_trustGame_12.If given the choice, would you like to play again with the other player?`,
                            rating_cooperative_13 = `ratePLayer_trustGame_13.How cooperative do you think the other player is?`,
                            rating_forgiving_13 = `ratePLayer_trustGame_13.How forgiving do you think the other player is?`,
                            rating_playAgain_13 = `ratePLayer_trustGame_13.If given the choice, would you like to play again with the other player?`,

                            # After Exposure
                            rating_cooperative_2 = `ratePLayer_trustGame_2.How cooperative do you think the other player is?`,
                            rating_forgiving_2 = `ratePLayer_trustGame_2.How forgiving do you think the other player is?`,
                            rating_playAgain_2 = `ratePLayer_trustGame_2.If given the choice, would you like to play again with the other player?`,

                            )

#######################    Pre-processing player ratings

# Pivot longer ratings on each attribute 
df_coop <- d_finished %>% dplyr::select(playerId,condition,contains("cooperative")) %>%  
  pivot_longer(cols=contains("cooperative"), names_to = c("gameNumber"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()

df_forgiv <- d_finished %>% dplyr::select(playerId,condition,contains("forgiving")) %>%
  pivot_longer(cols=contains("forgiving"), names_to = c("gameNumber"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()

df_playAgain <- d_finished %>% dplyr::select(playerId,condition,contains("again")) %>%  
  pivot_longer(cols=contains("again"), names_to = c("gameNumber"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()

#merge all data frames together
datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% reduce(full_join, by=c('playerId','gameNumber','condition') )

# Group means for each rating 
mu <- datRatings %>% group_by(gameNumber, condition) %>% summarise(mean_coop = mean(rating_coop),
                                                  mean_forgiv = mean(rating_forgiving),
                                                  mean_playAgain = mean(rating_playAgain)
                                                  )
mu

# ggplot(mu,aes(x=gameNumber)) + 
#   geom_bar(aes(y=mean_forgiv, stat = "identity", position = "dodge", color="blue")) + 
#   geom_bar(aes(y=mean_coop, stat = "identity", position = "dodge", color="darkgreen")) +
#   geom_bar(aes(y=mean_playAgain, stat = "identity", position = "dodge", color="red")) +
#   ggtitle("Player ratings after each game") +
#   xlab("Game Number") +
#   theme_minimal() + 
#   theme(legend.position = "bottom")
#

```

```{r}
library(ggplot2)
library(reshape2)

# First, you may need to convert the data from wide to long format
df_long <- melt(datRatings, id.vars = c("playerId", "condition", "gameNumber"),
                measure.vars = c("rating_coop", "rating_forgiving", "rating_playAgain"),
                variable.name = "rating_type", value.name = "rating_value")

# Order the gameNumber factor as you specified
df_long$gameNumber <- factor(df_long$gameNumber, levels = c("1", "11", "12", "13", "2"))

# Create a boxplot for each rating

  ggplot(df_long, aes(x = gameNumber, y = rating_value, fill = rating_type)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "darkred", show.legend = TRUE) +
  facet_wrap(~rating_type, scales = "free") +
  labs(x = "Game Number", y = "Rating") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r}
# Filter only trust rounds and create % returns and investments

avg_ret_df <- d_finished %>%
  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition) %>% 
  filter(roundType=="trust",!is.na(roundNum)) %>% 
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_scaled = scale(investment)) %>% 
  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)))

# Number of people left for analysis. (there are 51 rounds in each game)
#num_participants <- nrow(avg_ret_df)/51
num_participants <- length(unique(avg_ret_df$playerId))

num_participants

```

```{r}
# average investment for first 10 rounds pre intervention
avg_pct <- avg_ret_df %>% filter(roundNum < 11 , gameNum.f == "pre") %>% summarise(avg_inv = mean(inv_pct), avg_ret = mean(ret_pct_0, na.rm=TRUE))
avg_pct
```
```{r}
# Plot Invs and Rets in absolute values for pre and post
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=roundNum)) + 
  stat_summary(aes(y=returns, color="Returns"), fun = "mean", geom = "line", group = 1) +
  stat_summary(aes(y=investment, color="Investment"), fun = "mean", geom = "line", group = 1) +
  facet_wrap(~gameNum.f) +
  scale_color_manual(values = c("Returns" = "red", "Investment" = "blue")) +
  labs(color = "") +
  xlab("Round Number")+
  theme_bw() +
  theme(legend.position = "bottom",  # Move legend to bottom
        legend.title.align = 0.5,  # Center-align the legend title
        legend.key = element_blank(),
        plot.title = element_text(hjust = 0.5),  # Center-align the plot title
        axis.text = element_text(color = "black"),  # Customize axis text
        axis.title = element_text(color = "black"))  # Customize axis title

# # Plot Invs and Rets in percentages for pre and post
# ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")),aes(x=roundNum)) + 
#   stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=ret_pct_0)) + 
#   stat_summary(fun = "mean", geom = "line",group=1, color = "blue", alpha =0.6,position = position_dodge2(width=0.8),aes(y=inv_pct)) + 
#   geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
#   facet_wrap(~gameNum.f)

ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=roundNum)) + 
  stat_summary(aes(y=ret_pct_0, color="Returns"), fun = "mean", geom = "line", group = 1) +
  stat_summary(aes(y=inv_pct, color="Investment"), fun = "mean", geom = "line", group = 1) +
  geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
  facet_wrap(~gameNum.f) +
  scale_color_manual(values = c("Returns" = "red", "Investment" = "blue")) +
  labs(color = "") +
  xlab("Round Number")+
  ylab("Investment / Returns (%)")+
  theme_bw() +
  theme(legend.position = "bottom",  # Move legend to bottom
        legend.title.align = 0.5,  # Center-align the legend title
        plot.title = element_text(hjust = 0.5),  # Center-align the plot title
        axis.text = element_text(color = "black"),  # Customize axis text
        axis.title = element_text(color = "black"))  # Customize axis title

```
```{r}
# Plot Invs and Rets in absolute values for exposure
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")),aes(x=roundNum)) + 
  stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=returns)) + 
  #stat_summary(fun.data = "mean_ci", geom = "errorbar",alpha =0.4, width = 0.2, aes(y=ret_pct) ) + 
  stat_summary(fun = "mean", geom = "line",group=1, color = "blue",aes(y=investment)) + 
  facet_wrap(~gameNum.f)

# Plot Invs and Rets in percentages for exposure
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")),aes(x=roundNum)) + 
  stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=ret_pct_0)) + 
  #stat_summary(fun.data = "mean_ci", geom = "errorbar",alpha =0.4, width = 0.2, aes(y=ret_pct) ) + 
  stat_summary(fun = "mean", geom = "line",group=1, color = "blue", alpha =0.6,position = position_dodge2(width=0.8),aes(y=inv_pct)) + 
  geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
  facet_wrap(~gameNum.f)

```

```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. The blue line shows the returns pre-manipulation and the green line post-manipulation. We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}

ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_0, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.key = element_blank(),
        #legend.key = element_rect(fill = NA), # Make legend key background transparent
        legend.margin = margin(t = 0, b = 0, unit = "pt"), # Adjust top and bottom margin of each legend
        legend.box.margin = margin(t = 0, b = 0, unit = "pt")) + # Adjust top and bottom margin of the legend box
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") + # Hide the fill legend
  guides(color = guide_legend(override.aes = list(fill = NA))) # Remove fill for color legend keys

```

### 2024-02-13 niceAi.Rmd
more screening

@@
 ```{r supportHMM, include=FALSE}
-# order the states of the HMM to allow it to order them as low ret/mid/ret/high ret. 
-label_switch <- function(mod,labels) {
-  # labels is vector, first element is new label for original state 1, second is new label for original state 2, etc.
-  if(!is(mod,"depmix") || !is(mod,"depmix.fitted")) stop("this function is for depmix models")
-  n_states <- mod@nstates
-  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
-    stop("labels needs to be a vector of unique integers between 1 and", n_states)
-  }
-  inv_labels <- sapply(1:n_states,function(x) which(labels == x))
-  tmp <- mod
-  # relabel prior
-  ppars <- getpars(mod@prior)
-  fpars <- getpars(mod@prior,which="fixed")
-  out_pars <- as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels]))
-  out_fixed <- as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels]))
-  if(!tmp@prior@family$link=="identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
-  # relabel transition
-  for(i in 1:n_states) {
-    ppars <- getpars(mod@transition[[inv_labels[i]]])
-    fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
-    out_pars <- c(out_pars,as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels])))
-    out_fixed <- c(out_fixed,as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels])))
-    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
-    if(!tmp@transition[[i]]@family$link=="identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
-    #out_pars <- c(out_pars,getpars(mod@transition[[inv_labels[i]]]))
-  }
-  # relabel response
-  for(i in 1:n_states) {
-    out_pars <- c(out_pars,unlist(lapply(mod@response[[inv_labels[i]]],getpars)))
-    out_fixed <- c(out_fixed,unlist(lapply(mod@response[[inv_labels[i]]],getpars,which="fixed")))
-  }
-  tmp <- setpars(tmp,out_fixed,which="fixed")
-  tmp <- setpars(tmp,out_pars)
-  if(is(tmp,"depmix.fitted")) tmp@posterior <- viterbi(tmp)
-  return(tmp)
-}
-
-# ORDERING STATES WHEN RESPONSE IS BINS
-order_mod <- function(mod) {
-  ns <- nstates(mod)
-  sum <- rep(0.0,ns)
-  for(i in 1:ns) {
-    # Expected return in each state calculation (5 bins, for binned returns)
-    sum[i] = sum(1:5*mod@response[[i]][[1]]@parameters$coefficients)
-  }
-  # reordering the states
-  mod <- label_switch(mod, rank(sum))
-  return(mod)
-}
-
-# ORDERING STATES WHEN RESPONSE IS GAUSSIAN
-order_mod_gauss <- function(mod) {
-  ns <- nstates(mod)
-  sum <- rep(0.0,ns)
-  for(i in 1:ns) {
-    # Expected return in each state calculation 
-    sum[i] = mod@response[[i]][[1]]@parameters$mu
-  }
-  # reordering the states
-  mod <- label_switch(mod, rank(sum))
-  return(mod)
-}
+# # order the states of the HMM to allow it to order them as low ret/mid/ret/high ret. 
+# label_switch <- function(mod,labels) {
+#   # labels is vector, first element is new label for original state 1, second is new label for original state 2, etc.
+#   if(!is(mod,"depmix") || !is(mod,"depmix.fitted")) stop("this function is for depmix models")
+#   n_states <- mod@nstates
+#   if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
+#     stop("labels needs to be a vector of unique integers between 1 and", n_states)
+#   }
+#   inv_labels <- sapply(1:n_states,function(x) which(labels == x))
+#   tmp <- mod
+#   # relabel prior
+#   ppars <- getpars(mod@prior)
+#   fpars <- getpars(mod@prior,which="fixed")
+#   out_pars <- as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels]))
+#   out_fixed <- as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels]))
+#   if(!tmp@prior@family$link=="identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
+#   # relabel transition
+#   for(i in 1:n_states) {
+#     ppars <- getpars(mod@transition[[inv_labels[i]]])
+#     fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
+#     out_pars <- c(out_pars,as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels])))
+#     out_fixed <- c(out_fixed,as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels])))
+#     tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
+#     if(!tmp@transition[[i]]@family$link=="identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
+#     #out_pars <- c(out_pars,getpars(mod@transition[[inv_labels[i]]]))
+#   }
+#   # relabel response
+#   for(i in 1:n_states) {
+#     out_pars <- c(out_pars,unlist(lapply(mod@response[[inv_labels[i]]],getpars)))
+#     out_fixed <- c(out_fixed,unlist(lapply(mod@response[[inv_labels[i]]],getpars,which="fixed")))
+#   }
+#   tmp <- setpars(tmp,out_fixed,which="fixed")
+#   tmp <- setpars(tmp,out_pars)
+#   if(is(tmp,"depmix.fitted")) tmp@posterior <- viterbi(tmp)
+#   return(tmp)
+# }
+# 
+# # ORDERING STATES WHEN RESPONSE IS BINS
+# order_mod <- function(mod) {
+#   ns <- nstates(mod)
+#   sum <- rep(0.0,ns)
+#   for(i in 1:ns) {
+#     # Expected return in each state calculation (5 bins, for binned returns)
+#     sum[i] = sum(1:5*mod@response[[i]][[1]]@parameters$coefficients)
+#   }
+#   # reordering the states
+#   mod <- label_switch(mod, rank(sum))
+#   return(mod)
+# }
+# 
+# # ORDERING STATES WHEN RESPONSE IS GAUSSIAN
+# order_mod_gauss <- function(mod) {
+#   ns <- nstates(mod)
+#   sum <- rep(0.0,ns)
+#   for(i in 1:ns) {
+#     # Expected return in each state calculation 
+#     sum[i] = mod@response[[i]][[1]]@parameters$mu
+#   }
+#   # reordering the states
+#   mod <- label_switch(mod, rank(sum))
+#   return(mod)
+# }
 
@@
           signature(y="ANY"), 
-          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), ...) {
+          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), ...) {
             y <- matrix(y,length(y))
@@
               parameters$sigma <- pstart[2]
+            } else {
+              parameters <- list(mu = 10, sigma = 3)
             }
@@
             breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
+            prec <- pnorm(object@max, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@min, mean = object@parameters$mu, sd = object@parameters$sigma)
+            if(prec < 1e-12) {
+              # return uniform probabilities as precision is likely off
+              p <- rep(1/(length(breaks)-1), length(object@y))
+              if(log) return(log(p)) else return(p)
+            } else {
               p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
@@
               if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
+            }
           }
@@
           signature(y="ANY"), 
-          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20, ...) {
+          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), min=-0.5, max=20.5, ...) {
             y <- matrix(y,length(y))
@@
 )
+
+# define a response class which only contains the standard slots, no additional slots
+setClass("vtdgaus", contains="response", slots=c(yield="numeric"))
+
+# define a generic for the method defining the response class
+
+setGeneric("vtdgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("vtdgaus"))
+
+# define the method that creates the response class
+
+setMethod("vtdgaus", 
+          signature(y="ANY"), 
+          function(y, yield, pstart=NULL,fixed=NULL, ...) {
+            y <- matrix(y,length(y))
+            x <- matrix(1)
+            parameters <- list()
+            npar <- 2
+            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
+            if(!is.null(pstart)) {
+              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
+              parameters$mu <- pstart[1]
+              parameters$sigma <- pstart[2]
+            } else {
+              parameters <- list(mu=.5,sigma=1)
+            }
+            mod <- new("vtdgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, yield=yield)
+            mod
+          }
+)
+
+setMethod("show","vtdgaus",
+  function(object) {
+    cat("Gaussian with variable discrete support for percentage responses\n")
+    cat("Parameters: \n")
+    cat("mu: ", object@parameters$mu, "\n")
+    cat("sigma: ", object@parameters$sigma, "\n")
+  }
+)
+
+setMethod("dens","vtdgaus",
+  function(object,log=FALSE) {
+      # determine cuts for pnorm based on 
+      #p <- pnorm(sapply(object@y + .5*(1/object@yield), function(x) min(x,1.001)), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(sapply(object@y - .5*(1/object@yield),function(x) max(x,-0.001)), mean = object@parameters$mu, sd = object@parameters$sigma)
+      #norm <- (pnorm(1 + .001, mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(-0.001, mean = object@parameters$mu, sd = object@parameters$sigma))
+      #p <- p/norm
+
+      # alternatively, normalize factor based on size of cuts
+
+      prec <- pnorm(1 + .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma)
+      # check for issues with precision, using highest yield
+      if(prec < 1e-12) {
+        p <- 1/(object@yield + 1)
+      } else {
+        p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
+        norm <- pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
+        p <- p/norm
+      }
+      # probability when yield == 0 is always 1 
+      p[object@yield == 0] <- 1
+      if(log) return(log(p)) else return(p)
+    }
+)
+
+setMethod("setpars","vtdgaus",
+          function(object, values, which="pars", ...) {
+            npar <- npar(object)
+            if(length(values)!=npar) stop("length of 'values' must be",npar)
+            # determine whether parameters or fixed constraints are being set
+            nms <- names(object@parameters)
+            switch(which,
+                   "pars"= {
+                     object@parameters$mu <- values[1]
+                     object@parameters$sigma <- values[2]
+                   },
+                   "fixed" = {
+                     object@fixed <- as.logical(values)
+                   }
+            )
+            names(object@parameters) <- nms
+            return(object)
+          }
+)
+
+setMethod("getpars","vtdgaus",
+          function(object,which="pars",...) {
+            switch(which,
+                   "pars" = {
+                     parameters <- numeric()
+                     parameters <- unlist(object@parameters)
+                     pars <- parameters
+                   },
+                   "fixed" = {
+                     pars <- object@fixed
+                   }
+            )
+            return(pars)
+          }
+)
+
+setMethod("fit","vtdgaus",
+          function(object,w) {
+            if(missing(w)) w <- NULL
+            if(!is.null(w)) {
+              negLL <- function(pars) {
+                object <- setpars(object, c(pars[1], exp(pars[2])))
+                -sum(w*log(dens(object)))
+              }
+            } else {
+              negLL <- function(pars) {
+                object <- setpars(object, c(pars[1], exp(pars[2])))
+                -sum(log(dens(object)))
+              }
+            }
+            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
+            object <- setpars(object,c(pars[1], exp(pars[2])))
+            object
+          }
+)
 ```
@@
 ```{r}
-# Specify simple models to copy transition and prior
-inv_simple <- list()
-for(i in 2:6) {
-  inv_simple[[i]] <- depmix(investment_bin ~ 1, data = origin_dat, nstates = i, transition = ~ invPnL, family=multinomial("mlogit"), ntimes = rep(10,381))
+#| label: HMM-support-functions
+#| include: false
+# order the states of the HMM to allow it to order them as low ret/mid/ret/high ret. 
+label_switch <- function(mod,labels) {
+  # labels is vector, first element is new label for original state 1, second is new label for original state 2, etc.
+  if(!is(mod,"depmix") || !is(mod,"depmix.fitted")) stop("this function is for depmix models")
+  n_states <- mod@nstates
+  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
+    stop("labels needs to be a vector of unique integers between 1 and", n_states)
+  }
+  inv_labels <- sapply(1:n_states, function(x) which(labels == x))
+  tmp <- mod
+  # relabel prior
+  ppars <- getpars(mod@prior)
+  fpars <- getpars(mod@prior, which="fixed")
+  out_pars <- as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels]))
+  out_fixed <- as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels]))
+  if(!tmp@prior@family$link=="identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
+  # relabel transition
+  for(i in 1:n_states) {
+    ppars <- getpars(mod@transition[[inv_labels[i]]])
+    fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
+    out_pars <- c(out_pars,as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels])))
+    out_fixed <- c(out_fixed,as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels])))
+    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
+    if(!tmp@transition[[i]]@family$link=="identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
+    #out_pars <- c(out_pars,getpars(mod@transition[[inv_labels[i]]]))
+  }
+  # relabel response
+  for(i in 1:n_states) {
+    out_pars <- c(out_pars,unlist(lapply(mod@response[[inv_labels[i]]],getpars)))
+    out_fixed <- c(out_fixed,unlist(lapply(mod@response[[inv_labels[i]]],getpars,which="fixed")))
+  }
+  tmp <- setpars(tmp,out_fixed,which="fixed")
+  tmp <- setpars(tmp,out_pars)
+  if(is(tmp,"depmix.fitted")) tmp@posterior <- viterbi(tmp)
+  return(tmp)
 }
 
-# Specify truncated gaussian models for 3 state model 
-rModels3 <- list(
-  list(
-    truncdiscgaus(origin_dat$investment,pstart=c(5,2), min=-.5, max=20.5) # first guassian initialised with mean of 5 and sd of 2
-  ),
-  list(
-    truncdiscgaus(origin_dat$investment,pstart=c(10,2), min=-.5, max=20.5)
-  ),
-  list(
-    truncdiscgaus(origin_dat$investment,pstart=c(15,2), min=-.5, max=20.5)
-  )
-)
+# ordering states for a truncated discretised Gaussian
+order_mod_truncdiscgaus <- function(mod) {
+  ns <- nstates(mod)
+  sum <- rep(0.0,ns)
+  for(i in 1:ns) {
+    # Expected return in each state calculation
+    tpars <- getpars(mod@response[[i]][[1]])
+    dmod <- truncdiscgaus(seq(0, 20), pstart = tpars, min=-0.5, max=20.5)
+    sum[i] <- sum(seq(0,20)*dens(dmod))
+  }
+  cat("Expected values for states: ",sum, "\n")
+  # reordering the states
+  mod <- label_switch(mod, rank(sum, ties.method = "first"))
+  return(mod)
+}
 
-inv_dmod_3 <- makeDepmix(response=rModels3,transition=inv_simple[[3]]@transition,prior=inv_simple[[3]]@prior,ntimes = rep(10,381), homogeneous=FALSE)
+# ordering states for a variable response truncated discretised Gaussian
+order_mod_vtdgaus <- function(mod) {
+  ns <- nstates(mod)
+  sum <- rep(0.0,ns)
+  for(i in 1:ns) {
+    # Expected return in each state calculation
+    tpars <- getpars(mod@response[[i]][[1]])
+    dmod <- vtdgaus(seq(0, 1, length=61), pstart = tpars, yield=rep(60, 61))
+    sum[i] <- sum(seq(0, 1, length=61)*dens(dmod))
+  }
+  # reordering the states
+  mod <- label_switch(mod, rank(sum, ties.method = "first"))
+  return(mod)
+}
+```
 
-inv_fdmod_3 <- fit(inv_dmod_3, emcontrol=em.control(random.start = FALSE))  #random.start = FALSE -> use parameters supplied as starting values
+```{r}
+#| label: fit-HMM-investor-models
+#| include: false
+#| cache: true
 
-mod_3S <- order_mod_gauss(inv_fdmod_3)
+inv_fdmod <- list()
+
+set.seed(20230824)
+
+for(i in 2:4) {
+  inv_simple <- depmix(investment_bin ~ 1, data = origin_dat, nstates = i, transition = ~ invPnL, family=multinomial("mlogit"), ntimes = rep(10,381))
+
+  mus <- (1:i)*(20/(i+1))
+  sigmas <- rep((20/(i+1)), i)
+  rModels_inv <- list()
+  for(s in 1:i) {
+    rModels_inv[[s]] <- list(truncdiscgaus(origin_dat$investment, pstart=c(mus[s],sigmas[s]), min=-0.5, max=20.5))
+  }
+
+  #rModels_inv <- rep(list(list(truncdiscgaus(origin_dat$investment, pstart=c(10,5), min=-0.5, max=20.5))), each=i)
+
+  inv_dmod <- makeDepmix(response=rModels_inv, transition=inv_simple@transition, prior=inv_simple@prior, ntimes = rep(10,381), homogeneous=FALSE)
+
+  inv_fdmod[[i]] <- fit(inv_dmod, emcontrol = em.control(random.start = FALSE))
+
+  #inv_fdmod[[i]] <- multistart(inv_dmod, nstart=20)
+}
+
+#ORDER THE STATES OF THE HMM
+mod_3S  <- order_mod_truncdiscgaus(inv_fdmod[[3]])
+
+```
+
+```{r}
+# # Specify simple models to copy transition and prior
+# inv_simple <- list()
+# for(i in 2:6) {
+#   inv_simple[[i]] <- depmix(investment_bin ~ 1, data = origin_dat, nstates = i, transition = ~ invPnL, family=multinomial("mlogit"), ntimes = rep(10,381))
+# }
+# 
+# 
+# # Specify truncated gaussian models for 3 state model 
+# rModels3 <- list(
+#   list(
+#     truncdiscgaus(origin_dat$investment,pstart=c(5,2), min=-.5, max=20.5) # first guassian initialised with mean of 5 and sd of 2
+#   ),
+#   list(
+#     truncdiscgaus(origin_dat$investment,pstart=c(10,2), min=-.5, max=20.5)
+#   ),
+#   list(
+#     truncdiscgaus(origin_dat$investment,pstart=c(15,2), min=-.5, max=20.5)
+#   )
+# )
+# 
+# inv_dmod_3 <- makeDepmix(response=rModels3,transition=inv_simple[[3]]@transition,prior=inv_simple[[3]]@prior,ntimes = rep(10,381), homogeneous=FALSE)
+# 
+# inv_fdmod_3 <- fit(inv_dmod_3, emcontrol=em.control(random.start = FALSE))  #random.start = FALSE -> use parameters supplied as starting values
+# 
+# mod_3S <- order_mod_gauss(inv_fdmod_3)
 
@@
 ```{r, include=F}
-#ORDER THE STATES OF THE HMM
 
-state1 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[1]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))
+# 
+# state1 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[1]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))
+# 
+# state2 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[2]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))
+# 
+# state3 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[3]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))
+# 
+# investment <- seq(0:20) - 1
+# 
+# inv_resp_3S <- as.data.frame(cbind(investment,state1,state2,state3)) %>% 
+#   pivot_longer(cols=c("state1","state2","state3"),
+#                     names_to='investor_state',
+#                     values_to='probability')
 
-state2 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[2]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))
+#####################
+inv_policy <- list()
+for(i in 1:3) {
+  inv_policy[[i]] <- dens(truncdiscgaus(seq(0,20),  pstart=c(unlist(mod_3S@response[[i]][[1]]@parameters)), breaks = c(-Inf, seq(0, 19) + .5, Inf), min=-0.5, max=20.5))
+}
 
-state3 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[3]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))
+investment <- seq(0,20)
 
-investment <- seq(0:20) - 1
+df_inv_policy <- data.frame(investor_state = rep(1:3, each=length(investment)),
+                            investment,
+                            probability = unlist(inv_policy))
 
-inv_resp_3S <- as.data.frame(cbind(investment,state1,state2,state3)) %>% 
-  pivot_longer(cols=c("state1","state2","state3"),
-                    names_to='investor_state',
-                    values_to='probability')
+#####################
 
@@
 ```{r}
-inv_resp_3S %>% mutate( investor_state = fct_recode(investor_state, "low-trust" = "state1", "medium-trust" = "state2", "high-trust" = "state3"))%>%
-        ggplot( aes(x = investment,
+#| label: fig-policy-plot-3-state-HMM
+#| echo: false
+#| fig.cap: Probability mass function of the investor's policy conditional on its latent state as an output of the best fitting 3 state HMM using a truncated discretised Gaussian as a response function
+#| fig.align: center
+#| fig.width: 8
+#| fig.height: 4
+
+plotinvHMM <- ggplot(df_inv_policy ,                            
+       aes(x = investment,
            y = probability,
-                    fill = investor_state)) +
+           fill = factor(paste0("s",investor_state)))) +
   geom_bar(stat = "identity",
@@
   labs(fill='Latent investor state') +
-  theme_bw() + 
+  theme_minimal() + 
   theme(legend.position = "bottom")
+
+plotinvHMM
+
+```
+
+```{r}
+# inv_resp_3S %>% mutate( investor_state = fct_recode(investor_state, "low-trust" = "state1", "medium-trust" = "state2", "high-trust" = "state3"))%>%
+#         ggplot( aes(x = investment,
+#                     y = probability,
+#                     fill = investor_state)) +
+#   geom_bar(stat = "identity",
+#            position = "dodge") + 
+#   labs(fill='Latent investor state') +
+#   theme_bw() + 
+#   theme(legend.position = "bottom")
 
@@
 ```{r}
-# Here pars takes the form of a list ns matrices, one for each state from 1 to ns.
+#| label: transition-plot-function
+#| include: false
+#| warning: false
 plot_transitions_PnL <- function(fmod) {
@@
 
-  trans_prob <- trans_prob %>% mutate(from=recode(from, 
-                         `1`="low-trust",
-                         `2`="medium-trust",
-                         `3`="high-trust"),
-                        to=recode(to, 
-                         `1`="low-trust",
-                         `2`="medium-trust",
-                         `3`="high-trust")) %>% 
-    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
-    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))
+  trans_prob <- trans_prob %>%
+    mutate(from = paste0("s", from),
+           to = paste0("s", to))
 
-#     iris %>%
+  # trans_prob <- trans_prob %>% mutate(from=recode(from, 
+  #                        `1`="low-trust",
+  #                        `2`="medium-trust",
+  #                        `3`="high-trust"),
+  #                       to=recode(to, 
+  #                        `1`="low-trust",
+  #                        `2`="medium-trust",
+  #                        `3`="high-trust")) %>% 
   #   mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
-# ggplot() + 
-#    geom_histogram(aes(Petal.Width))+ 
-#    facet_grid(Species~.)
-#   
-  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1)) +
-      theme(legend.position = "bottom")
+  #   mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))
+  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))
 }
@@
 ```{r}
-plot_transitions_PnL(mod_3S)
+#| label: transition-plot-investor-HMM-3-state
+#| echo: false
+#| fig.cap: Transition probabilities of the investor state as a function of investor gain in the current round.
+Each panel is the state transitioned from, and each line shows the probability of transitioning to the state identified by the line color.
+#| fig.align: center
+#| fig.width: 8
+#| fig.height: 4
+plot_transitions_PnL(mod_3S) + labs(x = "Investor gain", y = "Probability", color='Transition to') + theme_bw() + theme(legend.position = "bottom") 
 ```
+
+## Write function to plot transition probs based on input of trans matrix
+
+```{r}
+# plot_transitions_input <- function(pars_list) {
+# 
+#   ns <- nrow(pars_list[[1]])  # number of states
+#   
+#   trans_prob <- data.frame(
+#     from = rep(1:ns, each=80*ns),
+#     to = rep(1:ns, each=80),
+#     PnL = seq(-20,60,length=80),
+#     prob = 0
+#   )
+# 
+#   y <- matrix(0.0,ncol=ns, nrow=80)
+#   
+#   for(from in 1:ns) {
+#     pars <- pars_list[[from]]
+#     for(to in 1:ns) {
+#       x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"PnL"]
+#       y[,to] <- exp(pars[to,1] + pars[to,2]*x)
+#     }
+#     y <- y/rowSums(y)
+#     for(to in 1:ns) {
+#       trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
+#     }
+#   }
+#   
+#   trans_prob <- trans_prob %>% mutate(from=recode(from, 
+#                          `1`="low-trust",
+#                          `2`="medium-trust",
+#                          `3`="high-trust"),
+#                         to=recode(to, 
+#                          `1`="low-trust",
+#                          `2`="medium-trust",
+#                          `3`="high-trust")) %>% 
+#     mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
+#     mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))
+#    
+#   ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))+
+#       theme(legend.position = "bottom")
+# }
+#
+
+```
+
+```{r}
+plot_HMM_transitions <- function(ns, pars_mat) {
+
+  trans_prob <- data.frame(
+    from = rep(1:ns, each=100*ns),
+    to = rep(1:ns, each=100),
+    ret = seq(-20,60,length=100),
+    probs = 0
+  )
+
+  y <- matrix(0.0,ncol=ns, nrow=100)
+
+  for(from in 1:ns) {
+  pars <- matrix(pars_mat[[from]], ncol=2)
+  # print(pars)
+
+    for(to in 1:ns) {
+        x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"ret"]
+        y[,to] <- exp(pars[to,1] + pars[to,2]*x)
+    }
+    y <- y/rowSums(y)
+
+    for(to in 1:ns) {
+      trans_prob$probs[trans_prob$from == from & trans_prob$to == to] <- y[,to]
+    }
+  }
+
+  df <- as.data.frame(trans_prob) %>% 
+    mutate(from = recode(from, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust"),
+           to = recode(to, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust") ) %>% 
+    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
+    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))
+
+    # Create a separate data frame with the background colors
+  bg_colors <- data.frame(
+    from = factor(c("low-trust", "medium-trust", "high-trust"), levels=c("low-trust","medium-trust","high-trust"))
+  )
+
+  # plotting code...
+  ggplot() +
+    geom_rect(data = bg_colors, aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = from), alpha = 0.1) +
+    geom_line(data = df, aes(x = ret, y = probs, colour = as.factor(to))) +
+    facet_wrap(~from, labeller = labeller(from = function(x) paste("From", x, "state on trial t"))) +
+    ylim(c(0,1)) +
+    scale_fill_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue")) +
+    labs(x = "Investor's net return on trial t", y = "Transition probability to \nState on trial t+1", color = 'State transitioned to') +
+    theme_bw() +
+    theme(legend.position = "bottom", 
+          axis.title.y = element_text(size = 8, angle = 90, vjust = 1.5))
+
+}
+
+```
+
 ```{r}
@@
 
-pars[[2]]["to_2","invPnL"]
+# pars has the format [[from]]["to","x"]
+pars[[2]]
 
-```
-
-## Write function to plot transition probs based on input of trans matrix
-
-```{r}
-plot_transitions_input <- function(pars_list) {
-
-  ns <- nrow(pars_list[[1]])  # number of states
-
-  trans_prob <- data.frame(
-    from = rep(1:ns, each=80*ns),
-    to = rep(1:ns, each=80),
-    PnL = seq(-20,60,length=80),
-    prob = 0
-  )
-
-  y <- matrix(0.0,ncol=ns, nrow=80)
-
-  for(from in 1:ns) {
-    pars <- pars_list[[from]]
-    for(to in 1:ns) {
-      x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"PnL"]
-      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
-    }
-    y <- y/rowSums(y)
-    for(to in 1:ns) {
-      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
-    }
-  }
-
-  trans_prob <- trans_prob %>% mutate(from=recode(from, 
-                         `1`="low-trust",
-                         `2`="medium-trust",
-                         `3`="high-trust"),
-                        to=recode(to, 
-                         `1`="low-trust",
-                         `2`="medium-trust",
-                         `3`="high-trust")) %>% 
-    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
-    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))
-
-  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))+
-      theme(legend.position = "bottom")
-}
+plot_HMM_transitions (3,pars)
 
@@
 # pars has the format [[from]]["to","x"]
-pars[[1]]["to_1","invPnL"] <- 0
-pars[[1]]["to_1","intercept"] <-  10
+new_pars <- pars
+new_pars[[1]]["to_1","invPnL"] <- 0
+new_pars[[1]]["to_1","intercept"] <-  -10
+new_pars[[1]]["to_3","intercept"] <-  -10
 
-pars[[2]]["to_1","intercept"] <- 3
-#pars
-plot_transitions_input(pars)
+#new_pars[[2]]["to_1","intercept"] <- 5
+plot_HMM_transitions (3,new_pars)
 ```

### 2024-02-15 article.Rmd
loaded data only to check equal participants between conditions

@@
 
-d_finished <- readRDS("data/d_finished.csv")
-exposure_intervention <- paste0(as.character(d_finished %>% dplyr::select(treatment.intervention) %>% unique()) , as.character(d_finished %>% dplyr::filter(gameNumber == "11", isLastRound) %>% dplyr::select(game_opponent) %>% unique()))
+library(dplyr)
+d_finished <- readRDS("../scriptMongo/data/d_finished.csv")
 
-d_finished <- d_finished  %>% mutate(gameNum.f = factor(gameNumber,labels = c("pre", "expo1", "expo2", "expo3","post"),levels=c("1", "11", "12", "13", "2")), 
-                              condition  = exposure_intervention) %>% 
+# Step 1: Identify the expo_opponent for each playerId where gameNumber == "11" (exposure game) & isLastRound == TRUE (to select a particular)
+expo_opponent_df <- d_finished %>%
+  filter(gameNumber == "11", isLastRound == TRUE) %>%
+  dplyr::select(id, expo_opponent = game_opponent) %>%
+  distinct() # Ensure uniqueness in case of duplicates
+
+# Step 2: Join this information back to the original dataset
+d_finished <- d_finished %>%
+  left_join(expo_opponent_df, by = "id") %>%
+  mutate(condition.f = as.factor(expo_opponent)) %>%
+  dplyr::select(-expo_opponent) # Optionally, remove the expo_opponent column if it's no longer needed
+
+###### get whether high or low RS and score from df_screening
+
+# Load the "df_screening.RData" file
+load("data/df_screening.RData")
+
+RS_data <- df_screening %>% mutate(high_RS = as.factor(ifelse(RS_score > 15, 1,0))) %>% 
+                            dplyr::select(id, MH.f, RS_score, high_RS)
+
+# Find IDs for those who participated in experiment and add RS_score and high_RS as new columsn to d_finished.
+# Merging d_finished with relevant columns from RS_data based on matching ids
+d_finished <- d_finished %>%
+  inner_join(RS_data %>% dplyr::select(id, RS_score, high_RS), by = "id")
+
+# Summarize to count the number of participants for each Condition within each high_RS level
+temp <- d_finished %>% dplyr::select(id, condition.f, high_RS) %>% 
+                      unique() %>%
+                      group_by(high_RS, condition.f) %>%
+                      summarize(participant_count = n(), .groups = 'drop')
+
+# View the summarized counts
+print(temp)
+```
+```{r}
+# Filter and Select Columns
+result <- d_finished %>%
+          filter(high_RS == 1) %>%
+          dplyr::select(id, RS_score)%>% 
+          unique()
+
+# Print the Result
+print(result)
+```
+
+```{r}
+
+# d_finished now has the 'Condition' column with the value equal to expo_opponent for all rounds and rows per playerId
+
+d_finished <- d_finished  %>% mutate( gameNum.f = factor(gameNumber,labels = c("pre", "expo1", "expo2", "expo3","post"),
+                                      levels=c("1", "11", "12", "13", "2"))) %>% 
                               rename(# Before Exposure
@@
                             )
+```
+
+```{r}
+
+d_finished %>% dplyr::filter(id=="601d2aa3cf8c480009fc0719")  %>% dplyr::select(Turing.justification)
+```
+
+```
+
+```{r}
 
@@
 #   
-
 ```
@@
 
-# First, you may need to convert the data from wide to long format
+# First, we need to convert the data from wide to long format
 df_long <- melt(datRatings, id.vars = c("playerId", "condition", "gameNumber"),

### 2024-02-15 screenCode.Rmd
loaded data only to check equal participants between conditions

@@
 names(df_screening)[names(df_screening) == "Duration..in.seconds."] <- "duration"
+names(df_screening)[names(df_screening) == "Q2"] <- "id"
 
 # Step 2: Filter rows
-df_screening <- df_screening[df_screening$Q1 == "I accept" & !grepl('test', df_screening$Q2) & df_screening$DistributionChannel ==  'anonymous', ]
+df_screening <- df_screening[df_screening$Q1 == "I accept" & !grepl('test', df_screening$id) & df_screening$DistributionChannel ==  'anonymous', ]
 cat("valid submissions: ",  nrow(df_screening), "\n")
@@
 head(df_screening)
+
+# Save the df_screening data frame to a file named "df_screening.RData"
+save(df_screening, file = "data/df_screening.RData")
+
+```
+
+#DUPLICATES?
+```{r}
+library(dplyr)
+
+# Find duplicates based on 'id'
+duplicates <- df_screening %>%
+  group_by(id) %>%
+  summarize(count = n(), .groups = 'drop') %>%
+  filter(count > 1)
+
+# View the duplicates and their counts
+print(duplicates)
 

### 2024-03-18 article.Rmd
first draft paper

@@
 ---
-title: "(Re)building cooperation: Effects of a cognitive intervention on cooperative behavior in games"
+title: "Can Exposure To Forgiving AI Foster Cooperative Play?"
 author:
-  "Ismail Guennouni^1^, Christoph Korn, Georgia Koppe"
+  "Ismail Guennouni, Georgia Koppe, Christoph Korn"
 
@@
 - Repair of Cooperation
-- Cognitive Intervention 
+- Cognitive Manipulation 
 - Economic Games 
@@
 
-bibliography: "bib/cogIntervention.bib"
+bibliography: "bib/exposure.bib"
 biblio-style: spphys
 
-output: pdf_document
+output:
+  bookdown::pdf_document2:
+    toc: false
 header-includes:
@@
 ```{r setupCoax, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
+knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, include=FALSE) 
 knitr::opts_chunk$set(out.width = "\\textwidth")
@@
 
-```{r load-packages05, include = FALSE}
+```{r load-packages05}
 library(papaja)
@@
 library(ggsignif)
-#library(magick)
+library(magick)
 
@@
 
-```{r}
+```{r emmOptions}
+emm_options(pbkrtest.limit = 6210)
+emm_options(lmerTest.limit = 6210)
+
+```
+
+# Introduction
+
+<!-- Trust-based cooperation is an important element of human social interactions, facilitating seamless interpersonal and intergroup relations -->
+
+<!-- <!-- ### Lack of trust and social dysfunction  -->
+
+<!-- Research into the determinants of psychopathology has also linked trust-based constructs to the emergence of mental health disorders [@fonagy_mentalizing_2017].
+Many studies have drawn attention to difficulties in social cognition as a crucial concern within personality disorders (See review in @herpertz_social-cognitive_2014).
+Individuals with these disorders often find it challenging to initiate and sustain social interactions, leading to behaviors that undermine their relationships.
+Specifically, individuals facing such social cognitive challenges tend to act uncooperatively, which has been identified as a key marker for the severity and presence of  a cluster of Personality Disorder (PD) symptoms [@mulder_relationship_1999].
+-->
+
+<!-- <!-- ### Could lack of trust be due to sampling bias for uncaring early caregivers ? -->
+
+<!-- One perspective of how these difficulties in the social domain emerge is through experience with early caregivers.
+Attachment theory [@bowlby_attachment_1978] posits that the quality of early caregiver relationships shapes an individual's future ability to form secure attachments and trust others.
+Research strongly suggests a connection between insecure attachment styles (characterized by anxiety or avoidance) and lower levels of trust in relationships.
+Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998].
+Similarly, @fonagy_role_2014 posit that when a young learner grows up surrounded by unreliable communicators, they may adapt by becoming mistrustful of social knowledge.
+This mistrust serves as a protective mechanism, particularly within environments where caregivers lack benign intention.
+-->
+
+<!-- <!-- In their model, drawn on Bowlby’s (1973) attachment theory, Downey and colleagues (1996, 1998) proposed that RS may evolve through early experiences of repeated rejection from primary caregivers, which lead individuals to form the expectation that they are likely to be rejected by others.
+This expectation is associated with hypervigilance for rejection-relevant cues, which in turn leads to negative emotional responses and maladaptive behavioral reactions.
+Even if RS is more strongly activated in threatening situations, it is assumed to be a stable disposition, a sort of implicit theory about the self (i.e., rejectable) and others (i.e., rejecters).
+As such, it can lead to long-term negative effects on individual functioning (e.g., self-regulation, mental health, per- sonality, social cognition) and undermine relationship quality (Gao, Assink, Cipriani, & Lin, 2017).
+-->
+
+<!-- <!-- ### Exposure as a way to correct the sampling bias.  -->
+
+<!-- If this adaptive mistrust is the source of social dysfunction, we can ask whether exposing those who exhibit it to cooperative and forgiving interaction partners might correct this bias.
+Research in the fields of behavioral economics and psychology has explored how positive social interactions influence trust and cooperation.
+The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995].
+In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors and trust levels, even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005].
+@fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior.
+Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior.
+These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust.
+-->
+
+<!-- <!-- ### Describe experiment -->
+
+<!-- In this study, using a randomised control trial design, we assess the effectiveness of a manipulation aimed at repairing the potential breakdown of cooperation in the RTG from a low investment sent by a computerized investor.
+The manipulation focuses on repeatedly exposing participants to agents that have been designed to have a limited propensity to retaliate to a non-cooperative action by the participants.
+In order to select a sample that might exhibit difficulties in building relationships based on Trust dues to adverse early experiences, we focused on stratifying participants based on Rejection Sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection.
+It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997].
+Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017).
+However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas.
+By startifying participants based on RS, we can directly examine whether the manipulation is particularly effective  for those most vulnerable to trust-related issues.
+Additionally, comparing outcomes for those with high and low RS will help us determine if the exposure-based manipulation has differential impacts depending on an individual's underlying sensitivity to rejection.
+The computerized investor in the RTG was programmed to play according to a hidden Markov model estimated from real players' data in prior research.
+A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
+Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
+-->
+
+Trust is fundamental to human social interactions, enabling smooth relationships at both interpersonal and intergroup levels.
+The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders [@fonagy_mentalizing_2017].
+Individuals with personality disorders often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors – a marker for the severity of PD symptoms [@herpertz_social-cognitive_2014; @mulder_relationship_1999].
+
+One explanation for such social challenges lies in early caregiver experiences.
+Attachment theory [@bowlby_attachment_1978] suggests that the quality of these relationships shapes our capacity for secure attachments and trust.
+Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998].
+Similarly, learners exposed to unreliable communicators could develop mistrust of social knowledge as a protective strategy [@fonagy_role_2014].
+
+If this adaptive mistrust is the source of social dysfunction, we can ask whether exposing those who exhibit it to cooperative and forgiving interaction partners might correct this bias.
+Research in the fields of behavioral economics and psychology has explored how positive social interactions influence trust and cooperation.
+The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995].
+In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors and trust levels, even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005].
+@fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior.
+Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior.
+These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust.
+
+In this study, we use a randomized control trial to test a manipulation aimed at repairing potential breakdowns in RTG cooperation.
+Participants are exposed to agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
+We stratify our sample based on rejection sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection.
+It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997].
+Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017).
+However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas.
+Finally, the computerized investor in the RTG was programmed to play according to a hidden Markov model estimated from real players' data in prior research.
+A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
+Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
+
+# Methods
+
+```{r loadCSV}
 
@@
   left_join(expo_opponent_df, by = "id") %>%
-  mutate(condition.f = as.factor(expo_opponent)) %>%
-  dplyr::select(-expo_opponent) # Optionally, remove the expo_opponent column if it's no longer needed
+  mutate(condition.f = factor(expo_opponent, levels= c("AI_HMM","AI_HMM_nice"),
+                              labels=c("Control","Manipulation"))) %>%
+  dplyr::select(-expo_opponent) # remove the expo_opponent column if it's no longer needed
 
@@
 
-RS_data <- df_screening %>% mutate(high_RS = as.factor(ifelse(RS_score > 15, 1,0))) %>% 
+RS_data <- df_screening %>%  mutate(high_RS = factor(ifelse(RS_score > 15, 1, 0), 
+                                    levels = c(0, 1), 
+                                    labels = c("low RS", "high RS")))  %>% 
                              dplyr::select(id, MH.f, RS_score, high_RS)
@@
 ```
-```{r}
-# Filter and Select Columns
-result <- d_finished %>%
-          filter(high_RS == 1) %>%
-          dplyr::select(id, RS_score)%>% 
-          unique()
 
-# Print the Result
-print(result)
+## Participants
+
+A total of 206 participants were recruited on the Prolific Academic platform (prolific.co).
+<!-- The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years, with a `r round(sd(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years standard deviation.
+-->
+Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £XXX.
+Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire to form two similarly sized groups: One with high (RSQ score > 15) and the other with low rejection sensitivity (RSQ score < 10).
+
+## Design and Procedure
+
+The experiment had a 2 (Condition: Manipulation or Control) by 2 (Rejection Sensitivity : high or low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor.
+Participants within each pre-screened group were randomly assigned to one of the two levels of the first factor.
+The games were designed and implemented online using Empirica v1 [@almaatouq_empirica_2021].
+
+## Tasks and Measures
+
+```{r, include=FALSE}
+Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
+0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
+0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
+0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
+0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
+0.0000037665630, 0.0000009965755)
+
+Neutral <- c(0.003393529, 0.006930874, 0.013053553, 0.022671228,
+0.036310144, 0.053627606, 0.073039389, 0.091734929, 0.106248217,
+0.113479701, 0.111769796, 0.101517390, 0.085028780, 0.065675083,
+0.046778251, 0.030725245, 0.018610323, 0.010394861, 0.005354126,
+0.002543094, 0.001113881)
+
+Pro_social <- c(0.003162697, 0.001057162, 0.001410197, 0.001881016,
+0.002508877, 0.003346113, 0.004462480, 0.005950950, 0.007935434,
+0.010581067, 0.014107909, 0.018809194, 0.025075644, 0.033427845,
+0.044559370, 0.059394206, 0.079163228, 0.105506029, 0.140606514,
+0.187373417, 0.249680651)
+
+investment <- seq(0:20) -1
+
+response_probs <- as.data.frame(cbind(investment,Anti_social,Neutral,Pro_social)) %>% 
+  rename("low-trust" = "Anti_social", "medium-trust"="Neutral", "high-trust" = "Pro_social") %>%
+  pivot_longer(cols=c("low-trust","medium-trust","high-trust"),
+                    names_to='Investor_state',
+                    values_to='probability') %>% 
+   mutate(across(Investor_state, factor, levels=c("low-trust","medium-trust","high-trust")))
 ```
 
-```{r}
+```{r, include=F}
+
+plotinvHMM <- ggplot(response_probs,                            
+       aes(x = investment,
+           y = probability,
+           fill = Investor_state)) +
+  geom_bar(stat = "identity",
+           position = "dodge") + 
+  labs(fill='Latent investor state', x = "Investment", y= "Probability") + 
+  theme_bw() + 
+  theme(legend.position = "bottom",  legend.text = element_text(size = 12),  legend.title = element_text(size = 14))
+
+plotinvHMM
+```
+
+```{r, include=FALSE}
+
+# Parameters for HMM human-like Transition Function
+unhappy_pars <- rbind(c(0,0), c(-3.366027, 0.40910797 ), c(-3.572619,0.08137274)) 
+neutral_pars <- rbind(c(0,0), c(3.3142637, 0.3763408), c(0.9169736, 0.4502838))
+happy_pars   <- rbind(c(0,0), c(0.7134085, 0.02101626), c(2.2215478 ,0.16162964))
+
+pars_inv <- list(unhappy_pars, neutral_pars, happy_pars)
+
+# Parameters for HMM nice Transition Function
+unhappy_pars_nice <- rbind(c(-10,0), c(-3.366027, 0.40910797 ), c(-3.572619,0.08137274)) 
+neutral_pars_nice <- rbind(c(0,0), c(3.3142637, 0.3763408), c(0.9169736, 0.4502838))
+happy_pars_nice   <- rbind(c(0,0), c(0.7134085, 0.02101626), c(2.2215478 ,0.16162964))
+
+pars_inv_nice <- list(unhappy_pars_nice, neutral_pars_nice, happy_pars_nice)
+
+plot_HMM_transitions <- function(ns, pars_mat) {
+
+  trans_prob <- data.frame(
+    from = rep(1:ns, each=100*ns),
+    to = rep(1:ns, each=100),
+    ret = seq(-20,60,length=100),
+    probs = 0
+  )
+
+  y <- matrix(0.0,ncol=ns, nrow=100)
+
+  for(from in 1:ns) {
+  pars <- matrix(pars_mat[[from]], ncol=2)
+  # print(pars)
+
+    for(to in 1:ns) {
+        x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"ret"]
+        y[,to] <- exp(pars[to,1] + pars[to,2]*x)
+    }
+    y <- y/rowSums(y)
+
+    for(to in 1:ns) {
+      trans_prob$probs[trans_prob$from == from & trans_prob$to == to] <- y[,to]
+    }
+  }
+
+  df <- as.data.frame(trans_prob) %>% 
+    mutate(from = recode(from, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust"),
+           to = recode(to, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust") ) %>% 
+    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
+    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))
+
+    # Create a separate data frame with the background colors
+  bg_colors <- data.frame(
+    from = factor(c("low-trust", "medium-trust", "high-trust"), levels=c("low-trust","medium-trust","high-trust"))
+  )
+
+  # plotting code...
+  ggplot() +
+    geom_rect(data = bg_colors, aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = from), alpha = 0.1) +
+    geom_line(data = df, aes(x = ret, y = probs, colour = as.factor(to))) +
+    facet_wrap(~from, labeller = labeller(from = function(x) paste("From", x, "state on trial t"))) +
+    ylim(c(0,1)) +
+    scale_fill_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
+                    name = "From state") +  # Changed legend title for 'fill' here
+    scale_color_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
+                       labels = c("low-trust", "medium-trust", "high-trust"),
+                       name = "State transitioned to") +
+    labs(x = "Investor's net return on trial t", y = "Transition probability to \nState on trial t+1", color = 'State transitioned to') +
+    theme_bw() +
+    theme(legend.position = "bottom",
+          legend.text = element_text(size = 12),
+          legend.key.size = unit(1, 'lines'),
+          legend.spacing.x = unit(0.1, 'in'),
+          legend.title = element_text(size = 14),
+          legend.margin = margin(t = 0.2, b = 0, unit = 'cm'),
+          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
+          strip.text = element_text(size = 10),
+          legend.box = "vertical" # Arrange legends vertically
+    ) +
+    guides(fill = guide_legend(order = 1, title.position = "left", title.hjust = 0.3),
+           color = guide_legend(order = 2, title.position = "left", title.hjust = 0.3))
+}
+
+```
+
+```{r, include=F}
+
+plotInvTran <- plot_HMM_transitions(3, pars_inv) 
+print(plotInvTran)
+
+plotInvTranNice <- plot_HMM_transitions(3, pars_inv_nice) 
+print(plotInvTranNice)
+
+```
+
+### Repeated Trust Game
+
+Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor.
+On each round the investor is endowed with 20 units and decides how much of that endowment to invest.
+This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor.
+If the trustee returns more than one third of the amount, the investor makes a gain.
+Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right (Figure \@ref(fig:HMMPanels).A).
+The participants were able to choose the icon that represents them at the start of the experiment.
+The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner.
+Participants were not told they were facing computerised co-players.
+
+The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) co-player.
+Full detail on the datasets used in the Supplementary Information.
+Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states.
+Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \@ref(fig:HMMPanels).B).
+These distributions reflect "low-trust", "medium-trust", or "high-trust".
+Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C).
+In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) or round 13 (post-manipulation).
+On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
+The initial state for the HMM investor in each instance of the game was the "mid-trust" state.
+
+\begin{landscape}
+```{r HMMPanels, fig.cap= "\\small{A: Screenshot of the repeated Trust Game.
+The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee.
+The investor is endowed with 20 units at the start of each round.
+They need to decide how much of that endowment they want to invest with the trustee.
+The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor.
+Shown here is the stage at which the trustee decides how much to send back to the investor.
+Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees.
+From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel B.
+The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C.
+Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
+Panel D shows the transition probability of the forgiving HMM agent, where we can see on the left plot that the agent always transitions out of the low-trust state.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
+
+library(magick)
+
+# Read in the images
+image1 <- image_read("figures/trust_game.png")
+image2 <- image_read("figures/HMMinvPolicy2.png")
+image3 <- image_read("figures/HMMinvTrans2.png")
+image4 <- image_read("figures/HMMniceInvTrans.png")
+
+# Determine the target width for all images, for example, the width of the widest image
+target_width <- max(image_info(image1)$width, image_info(image2)$width, 
+                    image_info(image3)$width, image_info(image4)$width)
+
+target_height <- max(image_info(image1)$height, image_info(image2)$height, 
+                    image_info(image3)$height, image_info(image4)$height)
+
+# Scale images to have the same width
+image1 <- image_scale(image1, paste0(target_width, "x", target_height))
+image2 <- image_scale(image2, paste0(target_width, "x", target_height))
+image3 <- image_scale(image3, paste0(target_width, "x", target_height))
+image4 <- image_scale(image4, paste0(target_width, "x", target_height))
+
+# Annotate each image with a label
+image1 <- image_annotate(image1, text = "A", location = "+200+10", size = 90, color = "black", gravity = "northwest")
+image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
+image3 <- image_annotate(image3, text = "C", location = "-0+10", size = 90, color = "black", gravity = "northwest")
+image4 <- image_annotate(image4, text = "D", location = "+0+10", size = 90, color = "black", gravity = "northwest")
+
+# Assuming a 300 DPI resolution for good print quality
+dpi <- 300
+
+# Spacing of 0.1 inches converted to pixels
+spacing <- 0.05 * dpi
+
+# Calculating the width and height for each image
+width_per_image <- ((12 - 0.1) / 2) * dpi
+height_per_image <- ((7.4 - 0.1) / 2) * dpi
+# The geometry string for each image, including spacing (assuming no borders)
+geometry_string <- paste0(width_per_image, 'x', height_per_image, '+', spacing, '+', spacing)
+
+# Now create the montage with the specified geometry
+combined_image <- image_montage(
+  c(image1, image2, image3, image4), 
+  tile = "2x2",
+  geometry = geometry_string
+)
+
+# Display the combined image
+plot(combined_image)
+```
+\end{landscape}
+
+## Manipulation
+
+To design a forgiving and ultimately more cooperative agent, we change the transition function of the investor HMM to make it impossible to remain in a low trust state once the agent transitions there.
+This is achieved by choosing the parameters of the transition function to make the probability of remaining in the "low-trust" state, when the agent is in that state, effectively nil.
+The policies conditional on the state and the transition function in the other states remain unchanged.
+The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
+
+## Procedure
+
+At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player.
+Participants were told their goal was to maximise the number of points in all phases.
+They were not told the number of rounds of each phase.
+Phase one was a 15 round Repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds.
+On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round.
+After completing 15 rounds of the RTG, participants rated how cooperative, forgiving they perceived the investor to be, and whether they would like to play with them again (all on a scale from 1 to 10).
+
+After phase one, participants in the manipulation condition played three games of 7 rounds each against the forgiving HMM agent, whilst those in the control condition faced the human-like HMM agent.
+To keep the experience similar to the pre-manipulation game, the agent in the control condition was also designed to send a very low investment in round 5 of each of the three games.
+Subsequent phase two was similar to phase one, with participants being told they would face a new player.
+Participants then completed the Levels of Personality Functioning Scale (LPFS) questionnaire (see the supplement for details).
+Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.
+
+## Statistical analysis
+
+To explore whether participants behaved differently in the RTG after the manipulation compared to the control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model as described below:
+
+\[
+\begin{split}
+\text{R}_{ij} = & \, \beta_0 + \beta_1 \text{ Phase}_i + \beta_2 \text{ Condition}_i + \beta_3 \text{ Investment}_i + \beta_4 \text{ RS}_i + \\
+& \beta_5 (\text{Phase} \times \text{Condition})_i + \beta_6 (\text{Phase} \times \text{Investment})_i + \beta_7 (\text{Phase} \times \text{RS})_i + \\
+& \beta_8 (\text{Condition} \times \text{Investment})_i + \beta_9 (\text{Condition} \times \text{RS})_i + \beta_{10} (\text{Investment} \times \text{RS})_i + \\
+& \beta_{11} (\text{Phase} \times \text{Condition} \times \text{Investment})_i + \beta_{12} (\text{Phase} \times \text{Condition} \times \text{RS})_i + \\
+& \beta_{13} (\text{Phase} \times \text{Investment} \times \text{RS})_i + \beta_{14} (\text{Condition} \times \text{Investment} \times \text{RS})_i + \\
+& \beta_{15} (\text{Phase} \times \text{Condition} \times \text{Investment} \times \text{RS})_i + \\
+& b_{0j} + b_{1j} \text{ (Phase)}_i+ \epsilon_{ij}
+\end{split}
+\]
+
+where:
+
+- \( \text{R}_{ij} \): percentage of tripled investment returned to investor for participant \( j \) in observation \( i \)
+- \( \beta_0 \): intercept
+- \( \beta_1 \) to \( \beta_4 \): main effects of Phase (RTG game pre vs. post-manipulation), Condition (manipulation vs. control), Investment, and RS (High vs Low RS), respectively
+- \( \beta_5 \) to \( \beta_{10} \): interaction effects between each pair of the four factors, showing how the relationship between one factor and the return percentage not available changes depending on the level of another factor
+- \( \beta_{11} \) to \( \beta_{14} \): three-way interaction effects among the four factors, indicating how the interaction between two factors is further modified by the third factor
+- \( \beta_{15} \): four-way interaction effect between Phase, Condition, Investment, and RS, describing how the interaction among three factors is modified by the fourth factor
+- \( b_{0j} \): player-wise random intercept for player \( j \)
+- \( b_{1j} \): player-wise random slope for Phase for player \( j \)
+- \( \epsilon_{ij} \): error term for player \( j \) in observation \( i \)
+
+The model was estimated using the `afex` package [@singmann_afex_2022] in R.
+More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek_balancing_2017].
+A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses.
+For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex".
+We Z-transform the Investment variable as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
+
+<!-- To model participants' returns in the RTG across games and conditions, we fit various hidden Markov models [@visser2022hidden] to participants' returns using the depmixS4 package [@visser_depmixs4_2021] for R.
+The transition between latent states is assumed to depend on the investment received and a dummy variable to characterise the group that the participant belongs to.
+Details on how the models are constructed can be found in the supplement.
+We fit models with different numbers of hidden states, and use the Bayesian Information Criterion [@schwarz_estimating_1978-1] to select the best model.
+-->
+
+```{r ratingsProcess}
 
@@
 d_finished <- d_finished  %>% mutate( gameNum.f = factor(gameNumber,labels = c("pre", "expo1", "expo2", "expo3","post"),
-                                      levels=c("1", "11", "12", "13", "2"))) %>% 
+                                      levels=c("1", "11", "12", "13", "2")),
+                                      phase.f = factor(ifelse(gameNumber %in% c("1", "2"), 
+                                                       ifelse(gameNumber == "1", "pre", "post"), NA), levels = c("pre", "post"))) %>% 
                               rename(# Before Exposure
@@
 
-```{r}
-
-d_finished %>% dplyr::filter(id=="601d2aa3cf8c480009fc0719")  %>% dplyr::select(Turing.justification)
-```
-
-```
-
-```{r}
+```{r datRatings}
 
@@
 # Pivot longer ratings on each attribute 
-df_coop <- d_finished %>% dplyr::select(playerId,condition,contains("cooperative")) %>%  
-  pivot_longer(cols=contains("cooperative"), names_to = c("gameNumber"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()
+df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative")) %>%  
+  pivot_longer(cols=contains("cooperative"), names_to = c("game"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()
 
-df_forgiv <- d_finished %>% dplyr::select(playerId,condition,contains("forgiving")) %>%
-  pivot_longer(cols=contains("forgiving"), names_to = c("gameNumber"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()
+df_forgiv <- d_finished %>% dplyr::select(playerId,condition.f,contains("forgiving")) %>%
+  pivot_longer(cols=contains("forgiving"), names_to = c("game"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()
 
-df_playAgain <- d_finished %>% dplyr::select(playerId,condition,contains("again")) %>%  
-  pivot_longer(cols=contains("again"), names_to = c("gameNumber"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()
+df_playAgain <- d_finished %>% dplyr::select(playerId,condition.f,contains("again")) %>%  
+  pivot_longer(cols=contains("again"), names_to = c("game"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()
 
 #merge all data frames together
-datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% reduce(full_join, by=c('playerId','gameNumber','condition') )
+datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% 
+              reduce(full_join, by=c('playerId','game','condition.f')) %>%
+              mutate(gameNum.f = factor(game,labels = c("pre", "expo1", "expo2", "expo3","post"),
+                                      levels=c("1", "11", "12", "13", "2")))
 
 # Group means for each rating 
-mu <- datRatings %>% group_by(gameNumber, condition) %>% summarise(mean_coop = mean(rating_coop),
+mu <- datRatings %>% group_by(gameNum.f, condition.f) %>% summarise(mean_coop = mean(rating_coop),
                                                   mean_forgiv = mean(rating_forgiving),
@@
 
-# ggplot(mu,aes(x=gameNumber)) + 
+# ggplot(mu,aes(x=gameNum.f)) + 
 #   geom_bar(aes(y=mean_forgiv, stat = "identity", position = "dodge", color="blue")) + 
@@
 
-```{r}
-library(ggplot2)
-library(reshape2)
+```{r LPFS}
+library(dplyr)
 
-# First, we need to convert the data from wide to long format
-df_long <- melt(datRatings, id.vars = c("playerId", "condition", "gameNumber"),
-                measure.vars = c("rating_coop", "rating_forgiving", "rating_playAgain"),
-                variable.name = "rating_type", value.name = "rating_value")
+# Step 1: Convert LPFS columns to numeric
+d_finished <- d_finished %>%
+  mutate(across(starts_with("LPFS"), as.numeric))
 
-# Order the gameNumber factor as you specified
-df_long$gameNumber <- factor(df_long$gameNumber, levels = c("1", "11", "12", "13", "2"))
+# Step 2: Calculate means
+# Identify LPFS columns
+lpfs_cols = grep("^LPFS", names(d_finished), value = TRUE)
+lpfs_1_to_6_cols = grep("^LPFS\\.[1-6]$", names(d_finished), value = TRUE)
+lpfs_7_to_12_cols = grep("^LPFS\\.(7|8|9|10|11|12)$", names(d_finished), value = TRUE)
 
-# Create a boxplot for each rating
+# Calculate means using vectorized operations
+d_finished <- d_finished %>%
+  mutate(
+    LPFS_score = rowMeans(dplyr::select(., all_of(lpfs_cols)), na.rm = TRUE),
+    LPFS_self_score = rowMeans(dplyr::select(., all_of(lpfs_1_to_6_cols)), na.rm = TRUE),
+    LPFS_other_score = rowMeans(dplyr::select(., all_of(lpfs_7_to_12_cols)), na.rm = TRUE)
+  )
 
-  ggplot(df_long, aes(x = gameNumber, y = rating_value, fill = rating_type)) +
-  geom_boxplot() +
-  stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "darkred", show.legend = TRUE) +
-  facet_wrap(~rating_type, scales = "free") +
-  labs(x = "Game Number", y = "Rating") +
+```
+
+```{r dfCorr}
+
+library(dplyr)
+library(GGally)
+
+df_corr <- d_finished %>% dplyr::select(id, RS_score, LPFS_score, LPFS_other_score,payoffTrust1,payoffTrust2) %>%
+  unique()
+
+# df_corr %>% summarise(
+#   Pearson_correlation = cor(LPFS_score, RS_score, method = "pearson", use = "complete.obs"),
+#   Spearman_correlation = cor(LPFS_score, RS_score, method = "spearman", use = "complete.obs")
+# )
+# 
+# # View the correlations
+# print(correlations)
+# 
+# # Now plot the scatter plot for the entire dataset (not by individual id)
+# ggplot(df_corr, aes(x = LPFS_score, y = RS_score)) +
+#   geom_point() +
+#   geom_smooth(method = "lm", se = FALSE) +
+#   labs(x = "LPFS Score", y = "RS Score", title = "Scatterplot of LPFS Score vs. RS Score") +
+#   theme_minimal()
+
+# Select only the numerical columns for the correlation matrix (excluding the 'id' column)
+numerical_data <- df_corr %>% dplyr::select(-id)
+
+# Calculate the correlation matrix
+correlation_matrix <- cor(numerical_data, use = "complete.obs")
+
+# Print the correlation matrix
+print(correlation_matrix)
+
+# Now create a scatterplot matrix
+# Install the package if you haven't already
+if (!require("GGally")) install.packages("GGally")
+
+ggpairs(numerical_data)
+
+```
+
+```{r corrPayoffs}
+
+temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
+  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f) %>%
+  unique()
+
+ggplot(data = temp2, aes(x = payoffTrust1, y = payoffTrust2)) +
+  geom_point() +  # Color points by id
+  labs(x = "Payoff Trustee 1", y = "Payoff Trustee 2", color = "ID") +
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "right")
+
+cor(temp2$payoffTrust1,temp2$payoffTrust2)
 
@@
 ```{r}
+library(ggplot2)
+library(dplyr)
+
+# First, reshape data into a longer format.
+long_data_temp <- d_finished %>%
+  dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
+  dplyr::select(id, payoffTrust1, payoffTrust2) %>%
+  gather(key = "game", value = "payoff", -id) %>%
+  mutate(game = factor(game, levels = c("payoffTrust1", "payoffTrust2"))) %>% 
+  unique()
+
+# Now, create the dumbbell plot
+ggplot(long_data_temp, aes(x = game, y = payoff, group = id)) +
+  geom_line(aes(group = id), color = "gray") + # Connect payoffs with a line
+  geom_point(aes(color = id), size = 3, alpha = 0.6) + # Payoff points
+  theme_minimal() +
+  theme(legend.position = "none") + # Hide legend if too many playerids
+  labs(x = "Game", y = "Payoff", title = "Payoff Comparison between Game 1 and Game 2")
+
+```
+
+```{r}
+
+df_long_ratings <- datRatings %>%
+  pivot_longer(
+    cols = starts_with("rating_"), 
+    names_to = "rating_type", 
+    values_to = "rating_value"
+  ) %>%
+  dplyr::select(playerId, condition.f, gameNum.f, rating_type, rating_value)
+
+# First, calculate the mean and SEM for each group
+summary_df <- df_long_ratings %>%
+  group_by(gameNum.f, condition.f, rating_type) %>%
+  summarise(
+    mean = mean(rating_value, na.rm = TRUE),
+    sem = sd(rating_value, na.rm = TRUE)/sqrt(n())
+  ) %>%
+  ungroup()
+
+```
+
+<!-- ## Linear mixed effects Model for returns  -->
+
+```{r avgRetDf}
 # Filter only trust rounds and create % returns and investments
@@
 avg_ret_df <- d_finished %>%
-  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition) %>% 
+  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS) %>% 
   filter(roundType=="trust",!is.na(roundNum)) %>% 
   mutate(roundNum = as.numeric(as.character(roundNum))) %>%
-  mutate(inv_scaled = scale(investment)) %>% 
-  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)))
+  mutate(inv_scaled = as.vector(scale(investment))) %>% 
+  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)),ret_pct_na = ifelse(investment==0,NA,returns/(3*investment)))
 
@@
 ```{r}
-# average investment for first 10 rounds pre intervention
-avg_pct <- avg_ret_df %>% filter(roundNum < 11 , gameNum.f == "pre") %>% summarise(avg_inv = mean(inv_pct), avg_ret = mean(ret_pct_0, na.rm=TRUE))
+# average investment for first 10 rounds pre manipulation
+avg_pct <- avg_ret_df %>% filter(roundNum < 11 , gameNum.f == "pre") %>% summarise(avg_inv = mean(inv_pct), avg_ret = mean(ret_pct_na, na.rm=TRUE))
 avg_pct
 ```
+
 ```{r}
@@
         axis.text = element_text(color = "black"),  # Customize axis text
-        axis.title = element_text(color = "black"))  # Customize axis title
+        axis.title = element_text(color = "black")) + # Customize axis title
+  facet_wrap(~condition.f*gameNum.f)
 
@@
         axis.text = element_text(color = "black"),  # Customize axis text
-        axis.title = element_text(color = "black"))  # Customize axis title
+        axis.title = element_text(color = "black"))+ # Customize axis title
+  facet_wrap(~condition.f*gameNum.f)
 
 ```
+
 ```{r}
@@
   stat_summary(fun = "mean", geom = "line",group=1, color = "blue",aes(y=investment)) + 
-  facet_wrap(~gameNum.f)
+  facet_wrap(~condition.f*gameNum.f)
 
@@
   geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
-  facet_wrap(~gameNum.f)
+  facet_wrap(~condition.f*gameNum.f)
 
@@
 
+<!-- Linear mixed effects model for all returns  -->
+
+# Behavioral Results
+
+```{r afexModCoop, cache=TRUE}
+
+mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+#summary(mod_rating_coop)
+
+```
+
+```{r emmCoop, cache=TRUE}
+# Contrast 1: Pre vs. Post
+contrast1 <- c(-1, 0, 0, 0, 1)
+
+# Contrast 2: Pre vs. (Expo1, Expo2, Expo3) 
+contrast2 <- c(-3, 1, 1, 1, 0)
+
+# Contrast 3: Post vs. (Expo1, Expo2, Expo3) 
+contrast3 <- c(0, -1, -1, -1, 3)
+
+# Combine contrasts into a matrix
+contrast_matrix <- rbind(contrast1, contrast2, contrast3)
+colnames(contrast_matrix) <- levels(datRatings$gameNum.f)
+
+# Apply contrasts to the model using 'emmeans'
+emm_coop <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f)
+
+# Custom comparisons: use 'contrast' directly
+comp1_coop <- contrast(emm_coop, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
+comp2_coop <- contrast(emm_coop, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
+comp3_coop <- contrast(emm_coop, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
+
+comp1_coop
+
+comp2_coop
+
+comp3_coop
+
+```
+
+```{r afexModForg}
+
+mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+#summary(mod_rating_forg)
+
+```
+
+```{r emmForg}
+
+# Apply contrasts to the model using 'emmeans'
+emm_forg <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f)
+
+# Custom comparisons: use 'contrast' directly
+comp1_forg <- contrast(emm_forg, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
+comp2_forg <- contrast(emm_forg, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
+comp3_forg <- contrast(emm_forg, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
+
+comp1_forg
+
+comp2_forg
+
+comp3_forg
+
+```
+
+```{r afexModAgain}
+
+mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+#summary(mod_rating_again)
+
+```
+
+```{r emmAgain}
+
+# Apply contrasts to the model using 'emmeans'
+emm_again <- emmeans(mod_rating_again, ~ gameNum.f | condition.f)
+
+# Custom comparisons: use 'contrast' directly
+comp1_again <- contrast(emm_again, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
+comp2_again <- contrast(emm_again, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
+comp3_again <- contrast(emm_again, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
+
+comp1_again
+
+comp2_again
+
+comp3_again
+
+```
+
+## Player ratings
+
+Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition.
+Compared to the first RTG game, participants in the manipulation condition rated the investors they faced in the exposure phase (the forgiving AI) as more Cooperative `r papaja::apa_print(comp2_coop)$full_result$Manipulation_PrevsExpo`.
+No difference in ratings on forgiveness and whether they would like to face them again.
+Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop)$full_result$Control_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg)$full_result$Control_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again)$full_result$Control_PrevsExpo`.
+
+```{r plotRatings,include=T, echo=FALSE, fig.cap="Averages and standard errors of the participants ratings of the opponent for each game and condition.
+Pre and post are the 15 round repeated trust games before and after the exposure phase respectively.
+The games titled expo 1 to 3 are the three 7 round games during the exposure phase.
+We note that absent the exposure to the forgiving AI, the ratings get worse on aggregate even through the participant faces the same human like HMM.
+In the manipulation condition, participants rate the nicer HMM as more cooperative but not more forgiving.
+When they face the human-like HMM again, its rating are considerably worse.
+",fig.align="center", fig.width=6, fig.height = 6}
+# Now create a plot using geom_errorbar for SEM and geom_point for the mean
+ggplot(summary_df, aes(x = gameNum.f, y = mean, group = rating_type, color = rating_type)) +
+  geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem), width = 0.2) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  facet_wrap(~~rating_type*condition.f, nrow = 3, ncol = 2, scales = "free") +
+  labs(x = "Game Number", y = "Mean Rating") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
+Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative`r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
+Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
+
+Finally, comparing ratings of the investors before and after the exposure phase, those in the manipulation condition rated them similarly on cooperation, lower on forgiveness `r papaja::apa_print(comp1_forg)$full_result$Manipulation_PrevsPost` and lower on willingness to face them again `r papaja::apa_print(comp1_again)$full_result$Manipulation_PrevsPost`.
+Those in the control condition rated the investors post the exposure phase lower on all three attributes.
+
+```{r}
+d_finished %>% 
+  dplyr::select(id,Turing.choice) %>% 
+  unique() %>% 
+  group_by(Turing.choice) %>%
+  summarise(Count = n()) %>%
+  mutate(Percentage = (Count / sum(Count)) * 100)
+```
+
+When asked during debrief whether they thought the investors they faced were Human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player.
+When asked to justify their choice, many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.
+
+## Analysis of participant returns
+
+```{r modAllReturns, include=FALSE, cache=TRUE}
+
+prepost_df <- avg_ret_df %>% filter(gameNum.f %in% c("pre","post"))
+
+mod_returns_pct_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+
+summary(mod_returns_pct_RS)
+
+```
+
+```{r prepostByCond, cache=TRUE}
+prepost_bycond <- pairs(emmeans::emmeans(mod_returns_pct_RS, c("phase.f"), by = "condition.f", pbkrtest.limit = 6300))
+```
+
+```{r}
+library(ggplot2)
+
+# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
+ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
+  geom_point(alpha = 0.6) + # Adjust point transparency with alpha
+  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
+  labs(title = "Effect of inv_scaled on Outcome by Condition",
+       x = "inv_scaled", y = "Outcome") +
+  theme_minimal() +
+  facet_wrap(~condition.f)
+
+```
+
+```{r modemmeans, cache=TRUE, include=F}
+library(emmeans)
+
+# Compute EMMs for the interaction of interest
+emms <- emmeans(mod_returns_pct_RS, ~ phase.f * condition.f | high_RS, pbkrtest.limit = 6300)
+
+contrasts <- contrast(emms, "pairwise", by = c("condition.f", "high_RS"))
+summary(contrasts)
+
+results_df <- as.data.frame(summary(contrasts))
+```
+
+On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60% of the endowment for investments and 35-50% of the total yield for returns, as reported in previous studies [@charness_investment_2008; @fiedler_social_2011].
+
+Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs.
+Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the first RTG compared to the second.
+Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.
+post-intervention), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
+Post-hoc tests show a decrease in the percentage returned only in the intervention condition, pre - post,
+`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.
+Looking at this interaction effect for the two levels of RS, we find a three way interaction between Phase, Condition and RS, such that this decrease in returns in the manipulation condition is present in the low RS group `r papaja::apa_print(contrasts)$full_result$Manipulation_LowRS_Pre_post` but not in the high RS group.
+
 ```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round.
 The blue line shows the returns pre-manipulation and the green line post-manipulation.
-We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the intervention condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}
+We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the manipulation condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 6}
 
-ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_0, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
+# Create a data frame for the vertical linesr
+vline_data <- data.frame(
+  xintercept = c(12, 13),
+  Defection_round = factor(c("Pre-manipulation", "Post-manipulation"), levels = c("Pre-manipulation", "Post-manipulation"))
+)
+
+# Plot
+ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
   stat_summary(fun = mean, geom = "line") +
@@
                alpha = 0.3, linetype = 0) +
+  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
+  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
   labs(x = "Round",
@@
   theme(legend.position = "bottom",
+        legend.box = "vertical", # This will stack the legends vertically
         legend.key = element_blank(),
-        #legend.key = element_rect(fill = NA), # Make legend key background transparent
-        legend.margin = margin(t = 0, b = 0, unit = "pt"), # Adjust top and bottom margin of each legend
-        legend.box.margin = margin(t = 0, b = 0, unit = "pt")) + # Adjust top and bottom margin of the legend box
+        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
+        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10)) +
   scale_color_manual(values = c("darkblue", "darkgreen"),
@@
                     guide = "none") + # Hide the fill legend
-  guides(color = guide_legend(override.aes = list(fill = NA))) # Remove fill for color legend keys
+  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
+  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
+         linetype = guide_legend(title = "Defection round")) +
+  facet_wrap(~high_RS*condition.f)
 
 ```
+
+There was also a significant main effect of Investment, 
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`,
+such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, indicates that returns were more affected by investments in the control condition.
+We also find a three way interaction between Phase,  Investment and RS
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post intervention).
+
+<!-- Focusing on post defection only:  -->
+### Post Defection Trials
+
+```{r postDefMod, cache=TRUE}
+post_def_data <- avg_ret_df %>% filter(roundType=="trust",!is.na(gameNum.f),(roundNum >= 12 & gameNum.f =="pre")| (roundNum >= 13 & gameNum.f =="post"))
+
+mod_retPostDef_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled*high_RS + (1 + phase.f| playerId), post_def_data , REML= TRUE, method="KR")
+
+summary(mod_retPostDef_RS)
+```
+
+```{r PairsPostDef}
+phasePairs <- pairs(emmeans::emmeans(mod_retPostDef_RS, c("phase.f"), pbkrtest.limit = 1449))
+phasePairs
+```
+
+Running the same mixed effects model only on the trials following the pre-programmed defection by the HMM agent, we find a significant main effect of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` with returns lower in the second game post defection trials compared to the first.
+We also find a main effect of Investment
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_retPostDef_RS)$full_result$inv_scaled)` 
+where participants continued to return higher proportions when receiving higher investments. 
+Finally, we still find an Investment by Condition interaction 
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the manipulation condition compared to the control condition in post-defection trials.
+
+<!-- Checking whether Investor behavior has changed dramatically between pre and post: -->
+
+```{r}
+library(ggplot2)
+
+# Plotting
+ggplot(prepost_df %>% filter(condition.f=="Manipulation", high_RS == "high RS", roundNum < 12), aes(x = roundNum, y = investment, group = playerId, color = as.factor(gameNum.f))) +
+  geom_line(alpha = 0.5) +  # Set alpha for transparency to handle overlap
+  facet_wrap(~ playerId, scales = 'free_y') + # Faceting by playerID, each gets its own y-axis scale
+  theme_minimal() +
+  labs(color = 'Game Number') +
+  theme(
+    strip.text = element_blank(), # Remove facet labels since there will be too many
+    axis.text.x = element_text(angle = 90, hjust = 1) # Rotate x-axis labels for readability
+  )
+
+```
+
+### HMM investments
+<!-- Linear mixed effects model for investments -->
+
+```{r invMod, cache=TRUE}
+
+mod_invs_RS <- mixed( investment ~ phase.f*condition.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+
+summary(mod_invs_RS)
+
+```
+
+To explore the HMM investors' behavior across games and conditions, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
+This shows no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.
+
+```{r invAfexPlot, cache=TRUE}
+# inv_all <- afex_plot(mod_invs_RS, x = "phase.f", trace = "condition.f", dodge = 0.8, error = "within",
+#             mapping = c("linetype", "shape", "fill"),
+#             data_geom = geom_violin,
+#             data_arg = list(width = 0.5),
+#             factor_levels = list(phase.f = c("Pre", "Post")), 
+#             legend_title = "Condition") + 
+#   theme_bw() + 
+#   ggtitle("All rounds investments") +
+#   labs(y = "HMM Investment", x = "Trust Game Phase") + 
+#   theme(plot.title = element_text(size = 16)) +
+#   theme(legend.position = "bottom") 
+# 
+# 
+# inv_all
+```
+
+<!-- Extracting Investor latent state from data -->
+
+```{r extractInvStates}
+
+library(tidyverse)
+
+# Convert dataset from wide to long format, focusing on columns that start with "State_"
+d_long <- d_finished %>%
+  gather(key = "state_col", value = "investorState", starts_with("State_")) %>%
+  mutate(
+    round_extracted = as.numeric(str_extract(state_col, "(?<=State_r)\\d+")),
+    game_extracted = as.numeric(str_extract(state_col, "(?<=game_)\\d+"))
+  ) %>%
+  # Filter to keep rows where extracted round/game numbers match the actual round/game numbers
+  filter(round_extracted == roundNum, game_extracted == gameNumber) %>%
+  # Select and rename the desired columns
+  dplyr::select(id, condition.f, gameNum.f, gameNumber, roundNum, investorState)
+
+# View the resulting dataset
+print(d_long)
+
+```
+
+```{r InvStatePlot}
+library(ggplot2)
+library(dplyr)
+
+# Filter for only games 1 and 2
+filtered_data <- d_long %>%
+  filter(gameNum.f %in% c("pre", "post"))
+
+# Define the colors for the levels of investorState
+
+state_colors <- c("happy" = "Blue", "neutral" = "#77DD77", "unhappy" = "#FF6961")
+
+# Create the histogram with stacked bars and specified colors
+ggplot(filtered_data, aes(x = as.factor(roundNum), fill = as.factor(investorState))) +
+  geom_bar(aes(group = investorState), position = position_stack(reverse = FALSE), stat = "count") +
+  scale_fill_manual(values = state_colors) +
+  facet_wrap(~condition.f*gameNum.f, scales = "free_x", labeller = label_both) +
+  labs(x = "Round", fill = "Investor State", title = "Histogram of Latent States per Game and Round") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+counts <- filtered_data %>%
+  group_by(condition.f, gameNum.f, investorState,id) %>%
+  summarise(count = n()) %>%
+    summarise(count = sum(count))
+
+# Calculate the average count per state per condition
+total_rounds <- filtered_data  %>%
+  group_by(condition.f, gameNum.f) %>%
+  summarise(total_rounds = n())
+
+state_summary  <- counts %>%
+  left_join(total_rounds, by = c("condition.f", "gameNum.f")) %>% 
+  mutate(state_percentage = count/total_rounds)
+
+ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = as.factor(investorState))) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~condition.f) +
+  labs(x = "Condition", y = "Proportion", fill = "Investor State") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+```
+
+# Discussion
+
+The use of HMM-based artificial agents in economic games led to similar investment and returns to those recorded in human dyadic interactions.
+Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.
+This validates the use of these artificial agents to probe the effectiveness of interventions whilst keeping a high degree of experimental control.
+Following the exposure intervention, participants reduced their returns overall.
+When breaking down the groups into RS subgroups, the reduction in returns was mostly from the group with low Rejection Sensitivity.
+The returns of those in the control group did not change between the pre and post phase of the experiment.
+Why did participants reduce their returns even though they were repeatedly exposed to a more cooperative and more forgiving AI?
+A look at how the participants rated their co-players might shed some light on what is driving this reduction in returns for those exposed to the forgiving AI.
+
+Those exposed to the forgiving AI rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure.
+Indeed, the biggest ratings change in the manipulation group was for the "post" player which were lower than ratings for both the pre-exposure and exposure players.
+This is likely due to a *negative contrast effect*.
+The Contrast Effect occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people.
+If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favorable or even negative by comparison [@kobre_negative_1972].
+As the most recently faced opponents were highly cooperative, this negative contrast effect may have compensated any learning transfer from being repeatedly exposed to cooperative and forgiving AI [@zentall_within-trial_2005].
+If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.
+Participants also perceived the forgiving AI during the exposure phase as more cooperative, but not more forgiving.
+The latter might be the result of the participants not having many opportunities to test the co-player's forgiveness propensity since most decided to continue cooperating.
+
+## Other notable points:
+
+Those in the control group perceived the human-like HMM as less cooperative, forgiving and were less willing to face it after the first interaction.
+This might indicate either *satiation/hedonic adaptation* (the process by which a novel interaction becomes less enjoyable  after too much exposure), or a  *negativity bias* (humans pay more attention to and give more weight to negative rather than positive experiences.
+Over time, as we interact more with someone, we might start noticing more of their flaws or negative traits, which could lead to lower overall ratings or more negative judgements).
+This negative rating did not translate however into lower proportioanl returns in the post-exposure phase.
+
+The study found that RSQ (Rejection Sensitivity Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game.
+This gap highlights a need for further research, despite theoretical connections suggested in the literature review.
+
+\pagebreak
+
+# References

### 2024-03-18 screenCode.Rmd
first draft paper

@@
 ```{r load data}
-#rsq_dat <- read.csv("data/RSQ_Feb06_24.csv") 
-rsq_dat <- read.csv("data/RSQ_Feb13_24.csv")
+
+#rsq_dat <- read.csv("data/RSQ_Feb15_24.csv") 
+rsq_dat <- read.csv("data/RSQ_Feb18_24.csv")
 
@@
 ```
+
 ```{r}
@@
 
-# Newly identified high RS on the 13th feb 2024.
+# Newly identified high RS on the 18th feb 2024.
 
@@
 
-high_RS_13feb <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) == as.Date("2024-02-13") ) %>% dplyr::select(PROLIFIC_PID)
+high_RS_18feb <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) == as.Date("2024-02-18") ) %>% dplyr::select(PROLIFIC_PID)
 
-nrow(high_RS_13feb)
+nrow(high_RS_18feb)
 
 # String them together and add to prolific
-feb13_ids_string <- paste(high_RS_13feb$PROLIFIC_PID, collapse = ",")
+feb18_ids_string <- paste(high_RS_18feb$PROLIFIC_PID, collapse = ",")
 
-feb13_ids_string
+feb18_ids_string
 ```

### 2024-03-27 efa.Rmd
v2 article

---
title: "EFA"
author: "Ismail Guennouni"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include =F)
library(tidyverse)
library(dplyr)
```

<!-- Load dataset 1 (RSQ) and select questionnaires -->
```{r rsq}
load("data/df_screening.RData")

# Identifying columns ending with _1 and _2, ignoring Q_att_1 and Q_att_2
one_columns <- grep('_1$', names(df_screening), value = TRUE)
two_columns <- grep('_2$', names(df_screening), value = TRUE)
one_columns <- setdiff(one_columns, c('Q_att_1', 'Q_att_2'))
two_columns <- setdiff(two_columns, c('Q_att_1', 'Q_att_2'))

df_rsq <- df_screening %>% dplyr::select(id, all_of(one_columns), all_of(two_columns))

```

<!-- Load dataset 2 (LPFS) and select questionnaires -->
```{r LPFS}

load("data/d_finished.RData")
df_lpfs <- d_finished %>% dplyr::select(id,contains("LPFS.")) %>% unique()

# Right join the data frames on the "id" column
df_efa <- merge(df_rsq, df_lpfs, by = "id", all.y = TRUE) %>% unique() %>% dplyr::select(-id)

```

```{r, include=T}
# Convert empty strings to NA and ensure numeric columns are indeed numeric.
# Replace the missing cells with column median
df_clean <- df_efa %>%
  mutate(across(everything(), ~as.numeric(as.character(.)))) %>%
  mutate(across(everything(), ~ifelse(is.na(.), median(., na.rm = TRUE), .)))

```

<!-- Duble checking non numeric and missing value cells -->
```{r}

str(df_clean) # all columns are numeric now 
colSums(is.na(df_clean)) # alls sums are 0 so no more missing values

```

```{r nFactors, include =T}

library(nFactors)
library(psych)

ev <- eigen(cor(df_clean)) # get eigenvalues

ev$values

scree(df_clean, pc=FALSE)  # Scree plot of the eigen values. Use pc=FALSE for factor analysis

```

```{r, include =T}
# another way to plot the eigenvalue
fa.parallel(df_clean, fa="fa")
```
The eigenvalue method (“Kaiser’s rule”) is telling us that 8 factors may be best. The scree plot is putting us somewhere between three and five factors. Parallel analysis suggests similar. THink 3 is adequate for such a small dataset but we can also use other methods such as CNG.

```{r, include =T}
Nfacs <- 3  # Think 3 factors is adequate for this small dataset. You can change this as needed.

# start with assumption of correlation btw factos. so using promax.
fit <- factanal(df_clean, Nfacs, rotation="promax")

print(fit, digits=2, cutoff=0.3, sort=TRUE)
```
```{r, include =T }
load <- fit$loadings[,1:2]
plot(load,type="n") # set up plot
text(load,labels=names(df_clean),cex=.7)
```
```{r, include =T}
loads <- fit$loadings

fa.diagram(loads)
```

```{r, include=T}
dim(fit$loadings) 
round(fit$loadings[ 1:30,], 2)
```

<!-- # Using the Polycor Package  -->

<!-- ```{r} -->
<!-- library(polycor) -->
<!-- library(nFactors) -->
<!-- hetcor_matrix <- hetcor(df_clean, polyserial = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ev_poly <- eigen(hetcor_matrix$correlations) # get eigenvalues -->

<!-- ev_poly$values -->

<!-- # Applying the CNG test -->
<!-- cng_results <- efa.CNG(hetcor_matrix$correlations, n.obs = nrow(df)) -->

<!-- # The CNG test results include the suggested number of factors to retain -->
<!-- print(cng_results) -->

<!-- ``` -->

### 2024-03-27 article.Rmd
v2 article

@@
 
-In this study, we use a randomized control trial to test a manipulation aimed at repairing potential breakdowns in RTG cooperation.
-Participants are exposed to agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
-We stratify our sample based on rejection sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection.
+In this study, we use a randomized control trial to test an exposure manipulation aimed at repairing potential breakdowns in RTG cooperation.
+We stratify the sample of participants based on rejection sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection.
 It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997].
@@
 However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas.
-Finally, the computerized investor in the RTG was programmed to play according to a hidden Markov model estimated from real players' data in prior research.
+
+Participants are exposed to artificial agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
+The computerized investor in the RTG was programmed to play using a generative model of how humans play fit to real players' data from prior studies.
 A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
-Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
+An advantage of having a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation.
+Such generative  model of behavior, informed by empirical data and integrating a responsive trust mechanism, represent a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
 
@@
 
+save(d_finished, file = "data/d_finished.RData")
+
 # Summarize to count the number of participants for each Condition within each high_RS level
@@
 -->
-Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £XXX.
+Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.
 Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire to form two similarly sized groups: One with high (RSQ score > 15) and the other with low rejection sensitivity (RSQ score < 10).
@@
 For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex".
-We Z-transform the Investment variable as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
+We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
 
@@
 # Pivot longer ratings on each attribute 
-df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative")) %>%  
+df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative"),high_RS) %>%  
   pivot_longer(cols=contains("cooperative"), names_to = c("game"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()
 
-df_forgiv <- d_finished %>% dplyr::select(playerId,condition.f,contains("forgiving")) %>%
+df_forgiv <- d_finished %>% dplyr::select(playerId,condition.f,contains("forgiving"),high_RS) %>%
   pivot_longer(cols=contains("forgiving"), names_to = c("game"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()
 
-df_playAgain <- d_finished %>% dplyr::select(playerId,condition.f,contains("again")) %>%  
+df_playAgain <- d_finished %>% dplyr::select(playerId,condition.f,contains("again"),high_RS) %>%  
   pivot_longer(cols=contains("again"), names_to = c("game"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()
@@
 datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% 
-              reduce(full_join, by=c('playerId','game','condition.f')) %>%
+              reduce(full_join, by=c('playerId','game','condition.f','high_RS')) %>%
               mutate(gameNum.f = factor(game,labels = c("pre", "expo1", "expo2", "expo3","post"),
@@
 temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
-  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f) %>%
+  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f, RS_score, LPFS_score) %>%
+  mutate(perf = payoffTrust1 + payoffTrust2) %>% 
   unique()
+
+# #                            Bonus Calculation
+# avg_perf <- mean(temp2$perf)
+# lowRS_bonus <- temp2 %>% filter(high_RS== "low RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
+# write.csv(lowRS_bonus,"data/bonusLow.csv")
+# 
+# highRS_bonus <- temp2 %>% filter(high_RS== "high RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
+# write.csv(highRS_bonus,"data/bonusHigh.csv")
 
@@
 
-cor(temp2$payoffTrust1,temp2$payoffTrust2)
-
+```
+
+```{r}
+# Sanity checks
+
+# prolific_highRS <- read.csv(file="/Users/mbp17/Downloads/prolific_export_highRS.csv")
+# 
+# # Which Ids in highRS file but not in Prolific? 
+# 
+# # Find IDs in highRS_bonus that are not in prolific_highRS
+# ids_in_highRS_not_in_prolific <- setdiff(highRS_bonus$id, prolific_highRS$Participant.id)
+# 
+# # Find IDs in prolific_highRS that are not in highRS_bonus
+# ids_in_prolific_not_in_highRS <- setdiff(prolific_highRS$Participant.id, highRS_bonus$id)
+# 
+# # ids_in_highRS_not_in_prolific now contains IDs that are in highRS_bonus but not in prolific_highRS
+# # ids_in_prolific_not_in_highRS now contains IDs that are in prolific_highRS but not in highRS_bonus
+# 
+# # print these variables to see the IDs
+# print(ids_in_highRS_not_in_prolific)
+# print(ids_in_prolific_not_in_highRS)
 ```
@@
   ) %>%
-  dplyr::select(playerId, condition.f, gameNum.f, rating_type, rating_value)
+  dplyr::select(playerId, condition.f, gameNum.f, rating_type, rating_value, high_RS)
 
@@
 summary_df <- df_long_ratings %>%
-  group_by(gameNum.f, condition.f, rating_type) %>%
+  group_by(gameNum.f, condition.f, rating_type, high_RS) %>%
   summarise(
@@
 avg_ret_df <- d_finished %>%
-  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS) %>% 
+  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score) %>% 
   filter(roundType=="trust",!is.na(roundNum)) %>% 
@@
 
-mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")
-
-#summary(mod_rating_coop)
+mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+summary(mod_rating_coop)
 
@@
 
+```{r}
+# Compute the estimated marginal means (EMMs) for the interaction
+emm_coop_interaction <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f * high_RS)
+
+comp1_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp2_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
+# comp3_coop_interaction <- contrast(emm_coop_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Output the results
+print("Pre vs Post")
+comp1_coop_interaction
+print("Pre vs Expo")
+test(comp2_coop_interaction, adjust="bonferroni")
+test(comp2_coop_interaction)
+
+# test_comp3_coop_interaction
+```
+
 ```{r afexModForg}
 
-mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")
+mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
 
@@
 
+```{r}
+# Compute the estimated marginal means (EMMs) for the interaction
+emm_forg_interaction <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f * high_RS)
+
+comp1_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp2_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
+# comp3_forg_interaction <- contrast(emm_forg_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Output the results
+print("Pre vs Post")
+comp1_forg_interaction
+print("Pre vs Expo")
+comp2_forg_interaction
+
+# test_comp3_forg_interaction
+```
+
 ```{r afexModAgain}
 
-mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")
+mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
 
@@
 
+```{r}
+# Compute the estimated marginal means (EMMs) for the interaction
+emm_again_interaction <- emmeans(mod_rating_again, ~ gameNum.f | condition.f * high_RS)
+
+comp1_again_interaction <- contrast(emm_again_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp2_again_interaction <- contrast(emm_again_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
+# comp3_again_interaction <- contrast(emm_again_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Output the results
+print("Pre vs Post")
+comp1_again_interaction
+print("Pre vs Expo")
+comp2_again_interaction
+
+# test_comp3_again_interaction
+```
+
 ## Player ratings
 
-Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition.
-Compared to the first RTG game, participants in the manipulation condition rated the investors they faced in the exposure phase (the forgiving AI) as more Cooperative `r papaja::apa_print(comp2_coop)$full_result$Manipulation_PrevsExpo`.
-No difference in ratings on forgiveness and whether they would like to face them again.
-Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop)$full_result$Control_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg)$full_result$Control_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again)$full_result$Control_PrevsExpo`.
+Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition and RS group.
+We will focus on two contrasts to analyse the ratings, by Condition and RS group.
+The first is between the rating in the first phase ("pre") when participants phase the human-like HMM, and the average rating during the exposure phase (average of "expo1", "expo2" and "expo3") where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition).
+The second contrast is between the "pre" and "post" phases of the experiment where in both conditions participants face the same human-like HMM.
+
+### Comparing pre and exposure ratings
+
+For those with high Rejection Sensitivity, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`.
+There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
+Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
+
+For those with low Rejection Sensitivity, there was no difference in any of the ratings between the "pre" and "exposure" phases for the Manipulation condition.
+For the Control condition, participants indicated less willingness to face the exposure co-player compared to the "pre" player, `r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo` but also did not differ on cooperation and forgiveness ratings.
+In summary, those with low rejection sensitivity had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving.
+In contrast, we see that exposing participants with high Rejection Sensitivity to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.
 
@@
 ",fig.align="center", fig.width=6, fig.height = 6}
-# Now create a plot using geom_errorbar for SEM and geom_point for the mean
-ggplot(summary_df, aes(x = gameNum.f, y = mean, group = rating_type, color = rating_type)) +
-  geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem), width = 0.2) +
-  geom_point(position = position_dodge(width = 0.5), size = 3) +
-  facet_wrap(~~rating_type*condition.f, nrow = 3, ncol = 2, scales = "free") +
-  labs(x = "Game Number", y = "Mean Rating") +
+
+# ggplot(summary_df, aes(x = gameNum.f, y = mean, group = rating_type, color = rating_type)) +
+#   geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem), width = 0.2) +
+#   geom_point(position = position_dodge(width = 0.5), size = 3) +
+#   facet_wrap(~~rating_type*condition.f, nrow = 3, ncol = 2, scales = "free") +
+#   labs(x = "Game Number", y = "Mean Rating") +
+#   theme_minimal() +
+#   theme(legend.position = "bottom")
+
+ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f), color = rating_type)) +
+  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1, position = position_dodge(width = 0.25)) +
+  geom_line(linetype = "dashed", position = position_dodge(width = 0.25)) +
+  geom_point(position = position_dodge(width = 0.25), size = 3) +
+  facet_wrap(~high_RS*condition.f, scales = "free_x") +
+  labs(x = "Phase", y = "Mean Rating", color = "Rating Type") +
+  scale_color_manual(values = c("rating_coop" = "blue", "rating_forgiving" = "red", "rating_playAgain" = "green"),
+                     labels = c("Cooperative", "Forgiving", "Play again")) +
   theme_minimal() +
@@
 
-Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative`r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
+```{r}
+library(ggplot2)
+
+ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f))) +
+  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem, color = condition.f), 
+                width = .1, position = position_dodge(width = 0.25)) +
+  geom_line(aes(color = condition.f), linetype = "dashed", position = position_dodge(width = 0.25)) +
+  geom_point(aes(color = condition.f), position = position_dodge(width = 0.25), size = 3) +
+  facet_grid(high_RS ~ rating_type, scales = "free_x") +
+  scale_color_manual(values = c("Control" = "blue", "Manipulation" = "red")) +
+  labs(x = "Phase", y = "Mean Rating", color = "Condition") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
+### Comparing pre and post ratings
+<!-- Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
+<!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
 Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
-
-Finally, comparing ratings of the investors before and after the exposure phase, those in the manipulation condition rated them similarly on cooperation, lower on forgiveness `r papaja::apa_print(comp1_forg)$full_result$Manipulation_PrevsPost` and lower on willingness to face them again `r papaja::apa_print(comp1_again)$full_result$Manipulation_PrevsPost`.
-Those in the control condition rated the investors post the exposure phase lower on all three attributes.
+-->
+
+Participants high on Rejection Sensitivity in the Manipulation condition rated the co-players in the "post" phase similarly on cooperation, lower on 
+forgiveness `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
+and lower on willingness to face them again `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`. 
+Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
+(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
+Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
+Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).
+
+For participants low on Rejection Sensitivity, those in the Manipulation condition rated the co-players in the "post" phase lower on all three attributes 
+(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
+Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
+Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`). 
+Those in the Control condition did not differ in their ratings of the "pre" and "post" phase players.
+
+In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the control condition.
+In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the control condition even though the player continues to use the same strategy.
+After exposure to the forgiving HMM, both groups  rates the "post" co-player worse than the "pre" even though they are the same.
 
@@
 
-When asked during debrief whether they thought the investors they faced were Human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player.
-When asked to justify their choice, many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.
-
 ## Analysis of participant returns
@@
 
+```
+
+```{r}
+saveRDS(summary(mod_returns_pct_RS), "data/mod_returns_pct_RS.RDS")
+```
+
+```{r}
+contrasts(prepost_df$phase.f)
+contrasts(prepost_df$condition.f)
+contrasts(prepost_df$high_RS)
+
+```
+```{r}
+RS_model <- lm( RS_score ~ ret_pct_na*inv_scaled + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
 ```
@@
 
-```{r}
-library(ggplot2)
-
-# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
-ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
-  geom_point(alpha = 0.6) + # Adjust point transparency with alpha
-  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
-  labs(title = "Effect of inv_scaled on Outcome by Condition",
-       x = "inv_scaled", y = "Outcome") +
-  theme_minimal() +
-  facet_wrap(~condition.f)
-
-```
-
 ```{r modemmeans, cache=TRUE, include=F}
@@
 results_df <- as.data.frame(summary(contrasts))
+```
+
+<!-- EFFECT SIZES -->
+```{r}
+# Load the necessary packages
+
+library(sjstats)
+effectsize::eta_squared(mod_returns_pct_RS)
 ```
@@
 Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.
-post-intervention), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
-Post-hoc tests show a decrease in the percentage returned only in the intervention condition, pre - post,
+post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
+As shown in Figure \ref{fig:violinPlots}, post-hoc tests confirm a decrease in the percentage returned only in the manipulation condition, pre - post,
 `r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.
-Looking at this interaction effect for the two levels of RS, we find a three way interaction between Phase, Condition and RS, such that this decrease in returns in the manipulation condition is present in the low RS group `r papaja::apa_print(contrasts)$full_result$Manipulation_LowRS_Pre_post` but not in the high RS group.
+We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.
 
@@
 The blue line shows the returns pre-manipulation and the green line post-manipulation.
-We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the manipulation condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 6}
+We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the manipulation condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}
 
@@
          linetype = guide_legend(title = "Defection round")) +
-  facet_wrap(~high_RS*condition.f)
+  facet_wrap(~condition.f)
 
@@
 such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, indicates that returns were more affected by investments in the control condition.
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, reflected  that returns were more affected by investments in the control condition.
 We also find a three way interaction between Phase,  Investment and RS
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post intervention).
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation).
+Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
+
+```{r fourwayInteractionPlot, include=T, fig.cap ="Effect of investment on returns by condition, phase and RS group" }
+
+library(ggplot2)
+
+# 
+# ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
+#   # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
+#   geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
+#   labs(title = "Effect of inv_scaled on Outcome by Condition",
+#        x = "inv_scaled", y = "Outcome") +
+#   theme_minimal() +
+#   facet_wrap(~high_RS)
+
+# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
+ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
+  # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
+  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
+  labs(x = "Investment Z-values", y = "Return proportion", color = "Experiment Phase") +
+  theme_minimal() +
+  facet_wrap(~condition.f*high_RS)+
+  theme(legend.position = "bottom")
+
+```
+
+To interpret this interaction we plot the effect of the Investment on returns for each level of Condition, Phase and RS.
+Figure \ref{fig:fourwayInteractionPlot} shows that while the effect of investment on returns is similar across RS groups and across phases in the Control condition, the Manipulation condition shows a higher effect of investment in the post phase compared to pre in the low RS group, but a lower effect of investment compared to before for the high RS group.
+This suggests that high RS participants based their return less on the investment after being exposed to the forgiving AI compared to the low RS group.
+
+```{r invMod, cache=TRUE}
+
+mod_invs_RS <- mixed( investment ~ phase.f*condition.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+
+summary(mod_invs_RS)
+
+```
+
+```{r violPlotsDef, cache=TRUE}
+ret_all <- afex_plot(mod_returns_pct_RS, x = "phase.f", trace = "condition.f", dodge = 0.8, error = "within",
+            mapping = c("linetype", "shape", "fill"),
+            data_geom = ggplot2::geom_boxplot,
+            data_arg = list(width = 0.5),
+            factor_levels = list(phase.f = c("Pre", "Post")),
+            legend_title = "Condition") +
+  theme_bw() +
+  labs(y = "% Returns", x = "Trust Game Phase") +
+  theme(plot.title = element_text(size = 16)) +
+  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 0.82, tip_length = 0.03,textsize = 3) + 
+  theme(legend.position = "bottom")
+
+inv_all <- afex_plot(mod_invs_RS, x = "phase.f", trace = "condition.f", dodge = 0.8, error = "within",
+            mapping = c("linetype", "shape", "fill"),
+            data_geom = ggplot2::geom_boxplot,
+            data_arg = list(width = 0.5),
+            factor_levels = list(phase.f = c("Pre", "Post")),
+            legend_title = "Condition") +
+  theme_bw() +
+  labs(y = "HMM Investment", x = "Trust Game Phase") +
+  theme(plot.title = element_text(size = 16)) +
+  theme(legend.position = "bottom")
+```
+
+```{r violinPlots, include=T,fig.cap="Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition.
+The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns.
+The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.",fig.align="center", fig.width=6, fig.height = 7}
+
+library(patchwork)
+
+ret_all/inv_all +  plot_layout(guides = 'collect') & theme(legend.position = "bottom")
+
+```
 
@@
 
-mod_retPostDef_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled*high_RS + (1 + phase.f| playerId), post_def_data , REML= TRUE, method="KR")
+mod_retPostDef_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled + (1 + phase.f| playerId), post_def_data , REML= TRUE, method="KR")
 
@@
 
-Running the same mixed effects model only on the trials following the pre-programmed defection by the HMM agent, we find a significant main effect of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` with returns lower in the second game post defection trials compared to the first.
+Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor?.
+To explore this question, we restrict the analysis to the trials following the pre-programmed defection by the HMM agent in both the "pre" (trials 12 to 15) and the "post" phases (trials 13 to 15).
+We fit the same mixed effects model as for all the trials with the exception of the RS variable.
+This is because RS did not show main or interaction effects in the main model, and also due to the necessity of running a simpler model to accommodate the low number of trials.
+We find a significant main effect of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` with returns lower in the second game post defection trials compared to the first.
 We also find a main effect of Investment
@@
 
-### HMM investments
+### HMM investor
 <!-- Linear mixed effects model for investments -->
-
-```{r invMod, cache=TRUE}
-
-mod_invs_RS <- mixed( investment ~ phase.f*condition.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
-
-summary(mod_invs_RS)
-
-```
-
-To explore the HMM investors' behavior across games and conditions, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
-This shows no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.
-
-```{r invAfexPlot, cache=TRUE}
-# inv_all <- afex_plot(mod_invs_RS, x = "phase.f", trace = "condition.f", dodge = 0.8, error = "within",
-#             mapping = c("linetype", "shape", "fill"),
-#             data_geom = geom_violin,
-#             data_arg = list(width = 0.5),
-#             factor_levels = list(phase.f = c("Pre", "Post")), 
-#             legend_title = "Condition") + 
-#   theme_bw() + 
-#   ggtitle("All rounds investments") +
-#   labs(y = "HMM Investment", x = "Trust Game Phase") + 
-#   theme(plot.title = element_text(size = 16)) +
-#   theme(legend.position = "bottom") 
-# 
-# 
-# inv_all
-```
+To test whether the HMM investors' behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
+As seen in Figure \@ref(fig:violinPlots), we find no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.
+
+When asked during debrief whether they thought the investors they faced were Human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player.
+When asked to justify their choice, many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.
+
+### Questionnaire scores and performance
+
+<!-- We found no correlation between either the RS score and task performance, as defined by the total points accumulated in both the "pre" and "post phase of the RTG.
+The LPFS score was also not correlated with task performance.
+These two scores exhibited a positive correlation  (Rho = 0.52, p < 0.001) -->
+
+```{r}
+cor.test(temp2$LPFS_score, temp2$perf, method = "spearman")
+cor.test(temp2$RS_score, temp2$perf, method = "spearman")
+
+cor_result <- cor.test(temp2$RS_score, temp2$LPFS_score, method = "spearman")
+
+```
+
+Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity score `r papaja::apa_print(cor_result)$full_result`, there was no correlation between these and participant's return or overall task performance.
 
@@
   theme(legend.position = "bottom")
+
+state_colors <- c("happy" = "Blue", "neutral" = "#77DD77", "unhappy" = "#FF6961")
+
+# Create the histogram with stacked bars and specified colors
+ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = as.factor(investorState))) +
+  geom_bar(aes(group = investorState), position = position_stack(reverse = FALSE), stat = "identity") +
+  scale_fill_manual(values = state_colors) +
+  facet_wrap(~condition.f*gameNum.f, scales = "free_x", labeller = label_both) +
+  labs(x = "Round", fill = "Investor State", title = "Histogram of Latent States per Game and Round") +
+  theme_bw() +
+  theme(legend.position = "bottom")
 ```
@@
 
+This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of Hidden Markov Models (HMMs) as artificial agents in economic games.
 The use of HMM-based artificial agents in economic games led to similar investment and returns to those recorded in human dyadic interactions.
 Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.
-This validates the use of these artificial agents to probe the effectiveness of interventions whilst keeping a high degree of experimental control.
-Following the exposure intervention, participants reduced their returns overall.
-When breaking down the groups into RS subgroups, the reduction in returns was mostly from the group with low Rejection Sensitivity.
-The returns of those in the control group did not change between the pre and post phase of the experiment.
+This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control.
+
+Following the exposure manipulation, participants reduced their returns overall whilst the returns of those in the control group did not change between the pre and post phase of the experiment.
 Why did participants reduce their returns even though they were repeatedly exposed to a more cooperative and more forgiving AI?
@@
 Those exposed to the forgiving AI rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure.
-Indeed, the biggest ratings change in the manipulation group was for the "post" player which were lower than ratings for both the pre-exposure and exposure players.
-This is likely due to a *negative contrast effect*.
-The Contrast Effect occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people.
-If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favorable or even negative by comparison [@kobre_negative_1972].
-As the most recently faced opponents were highly cooperative, this negative contrast effect may have compensated any learning transfer from being repeatedly exposed to cooperative and forgiving AI [@zentall_within-trial_2005].
+One possible explanation for this drop in rating is that participants exhibited a negative contrast effect.
+This occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people.
+If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favourable or even negative by comparison [@kobre_negative_1972].
+As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving AI [@zentall_within-trial_2005].
 If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.
-Participants also perceived the forgiving AI during the exposure phase as more cooperative, but not more forgiving.
+
+<!-- Participants also perceived the forgiving AI during the exposure phase as more cooperative, but not more forgiving.
 The latter might be the result of the participants not having many opportunities to test the co-player's forgiveness propensity since most decided to continue cooperating.
-
-## Other notable points:
-
-Those in the control group perceived the human-like HMM as less cooperative, forgiving and were less willing to face it after the first interaction.
-This might indicate either *satiation/hedonic adaptation* (the process by which a novel interaction becomes less enjoyable  after too much exposure), or a  *negativity bias* (humans pay more attention to and give more weight to negative rather than positive experiences.
-Over time, as we interact more with someone, we might start noticing more of their flaws or negative traits, which could lead to lower overall ratings or more negative judgements).
-This negative rating did not translate however into lower proportioanl returns in the post-exposure phase.
-
-The study found that RSQ (Rejection Sensitivity Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game.
+-->
+
+It is worth highlighting that Rejection Sensitivity was not correlated to overall performance in the task.
+Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
+However, in examining the player ratings by rejection sensitivity (RS) group more closely, we observe that individuals with high rejection sensitivity demonstrate a heightened attunement to changes in the behavior of their AI co-players.
+Specifically, when exposed to a more forgiving AI agent, these participants accurately increased their ratings of the agent, indicating a sensitive and appropriate response to the behavioral manipulation.
+This adjustment reflects a nuanced perception of social cues and a capacity to modify judgments based on the behavior of interaction partners, potentially indicating a perceived alignment or support from the co-player that mitigates their heightened sensitivity to potential social rebuffs.
+High RS participants' lower ratings of essentially similar co-players as they continued to face them may indicate that the occasional pre-programmed defections of the AI agents failed to mitigate concerns over rejection and possibly exacerbated perceptions of uncooperativeness and unforgiveness, leading to a stronger preference against future interactions.
+Individuals with low RS were less affected by AI behavioral variations, showing a less volatile baseline of social perception.
+They still exhibit behavior consistent with a contrast effect in the manipulation condition.
+This dichotomy highlights the potential of tailored interactions, mediated by advanced AI agents, to address and modulate specific social and psychological predispositions in human participants.
+
+<!-- Interestingly, for participants with low RS, the manipulation did not significantly alter their perceptions of the AI's cooperativeness, forgiveness, or their willingness to interact in the future.
+This group's ratings remained largely unchanged between the pre- and post-exposure phases, regardless of the AI's behavior.
+This stability suggests that individuals with lower sensitivity to rejection may base their ratings more on consistent behavioral patterns or other criteria not directly influenced by the forgiving nature of the AI agents.
+-->
+
+Importantly, our approach utilizing HMMs to simulate interactive partners in social dilemmas offers a promising avenue for future research, especially in understanding and potentially mitigating trust deficits in individuals with high rejection sensitivity.
+The nuanced behaviors elicited through the interaction with HMM agents highlight the complexity of trust dynamics and the potential for computational models to offer novel insights into human social behaviors.
+
+<!-- In conclusion, this study demonstrates the value of integrating HMM-based artificial agents in the study of trust and cooperation.
+By showcasing the potential of these agents to both simulate realistic human interactions and influence real human behavior, our research paves the way for future investigations into the mechanisms of trust repair and the broader implications of artificial agents in social psychology.
+-->
+
+<!-- ## Other notable points: -->
+
+<!-- The study found that RSQ (Rejection Sensitivity Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game.
 This gap highlights a need for further research, despite theoretical connections suggested in the literature review.
+-->
 

### 2024-03-27 supplement.Rmd
v2 article

---
title: "Can Exposure To Forgiving AI Foster Cooperative Play?"
subtitle: "Supplementary Information"

output:
  bookdown::pdf_document2:
   toc: false

bibliography: "bib/exposure.bib"
biblio-style: spphys

header-includes:
  - \renewcommand{\thefigure}{S\arabic{figure}}
---

```{r setupCoaxsupp, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
knitr::opts_chunk$set(out.width = "\\textwidth")
library(papaja)
library(kableExtra)
library(knitr)

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(gridExtra)
library(forcats)
library(ggsignif)
```

# Repeated Trust Game

Participants played a 15-round repeated trust game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. Each player was represented with an icon with the participant always on the left of the screen and the opponent on the right. The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the opponent changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised opponents. We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a random time lapse (randomly chosen between 5 and 10 seconds).

```{r trustGameScreenshot, include=F, fig.cap = "Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee. The trustee in turn needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee makes a decision of how much to send back the investor", out.width="70%", out.height="70%", fig.align='center'}

knitr::include_graphics("figures/trust_game.png")

```

# Hidden markov Model used to simulate the Investor's actions

The HMM assumes that the probability of each investment $I_t=0,…,20$, at each trial t, conditional on the current state of the investor $S_t$, is dependent on an underlying normal distribution with mean $\mu_s$ and standard deviation $\sigma_s$. The probability of each discrete investment was determined from the cumulative normal distribution $\Phi$, computing the probability of a Normal variate falling between the midway points of the response options. As responses were bounded at 0 and 20, we normalized these probabilities further by taking the endpoints into account. For instance, the probability of an investment $I_t=2$ is defined as:

$$P(I_t=2 | S_t=s)= \frac{\Phi(2.5 | \mu_s,\sigma_s ) - \Phi(1.5 | \mu_s,\sigma_s )} {\Phi(20.5 | \mu_s,\sigma_s ) - \Phi(-0.5 | \mu_s,\sigma_s )}$$

Note that the denominator truncates the distribution between 0 and 20. To estimate the transition probability between states for the investor, a multinomial logistic regression model was fitted to the investor's data such as:

$$P(S_{t+1} = s' | S_t=s , X_t = x)= \frac{\exp(\beta_{0,s,s'} + \beta_{1,s,s'} x)}{\sum_{s''} \exp(\beta_{0,s,s''} + \beta_{1,s,s''} x)}$$

where $X_t=R_t-I_t$ is the net return to the investor with  $R_t$ the amount returned by the trustee and $I_t$ is the Investment sent.

The advantages of this approach is that it does not require any a priori assumptions about the model features. The number of states, the policy conditional on the state, and the transition function between states can all determined in a purely data-driven way. These HMMs can in turn be used to simulate a human-like agent playing the trust game. This agent may transition to a new state depending on the other player's actions and adopt a policy reflecting its state, thus simulating changes in emotional dispositions of human players during a repeated game. When the investor gains from the interaction, they become more likely to transition to a state where their policy is more "trusting" with generally higher investments. However, faced with losses, the investor is more likely to transition to a more cautious policy with generally lower investments. The policies and the transitions between states are sufficient to build an agent that reflects this type of adaptive behavior and reacts to the trustee's action choices in a way that mimics a human player.

We estimated a three-state model for investor's behaviour, using maximum likelihood estimation via the Expectation-Maximisation algorithm as implemented in the depmixS4 (Visser & Speekenbrink) package for R. The model was estimated using investments from existing datasets of human dyads playing 10 rounds of the RTG with the same trustee. The dataset consisted of a total of 381 games from two data sources: First, a total of 93 repeated trust games with healthy investors and a mix of healthy trustees and trustees diagnosed with Borderline Personality Disorder (BPD) [@king-casas_rupture_2008]. The second source was from data collected as part of a project investigating social exchanges at the Montague lab at Virginia Tech and consists of 288 games.

## Mixed-effects model results for participant returns

We fit a linear mixed effects model to participant returns as a proportion of the multiplied investment received. The results of the model are presented in attached table.

```{r, include=F}

# Create a summary object
model_summary <- readRDS("data/mod_returns_pct_RS.RDS")

# Extract the fixed effects as a data frame
mod_df <- as.data.frame(model_summary$coefficients)

mod_df <- mod_df %>%
  rownames_to_column(var = "Term") %>% 
  mutate(Term = gsub("condition.f1", "Condition", Term),
         Term = gsub("phase.f1", "Phase", Term),
         Term = gsub("high_RS1", "high_RS", Term),
         Term = gsub("inv_scaled", "Investment", Term))

# Create a summary table using apa_table
summary_table <- papaja::apa_table(
  mod_df,
  caption = "Summary of Mixed-Effects Model of participant returns over all rounds",
  align = c("l", "c", "c", "c", "c", "c")
)

```

```{r myTable, results='asis'}
summary_table
```

\pagebreak

# References

### 2024-06-05 article.Rmd
version 3

@@
 author:
-  "Ismail Guennouni, Georgia Koppe, Christoph Korn"
+  # "Ismail Guennouni, Georgia Koppe, Christoph Korn"
 
@@
 
-Trust is fundamental to human social interactions, enabling smooth relationships at both interpersonal and intergroup levels.
+Trust is fundamental to human social interactions, facilitating seamless relations at both interpersonal and intergroup levels.
 The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders [@fonagy_mentalizing_2017].
@@
 Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998].
+These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection [@downey_early_1997].
+RS has been linked to the development of various mental health conditions, including depression, anxiety, personality disorders, and self-harm [@gao_associations_2017].
+Given the potential influence of RS on trust-related dynamics, it becomes particularly relevant to our understanding of social dysfunction.
+Could an ingrained sensitivity to rejection contribute to a mistrust bias?
 Similarly, learners exposed to unreliable communicators could develop mistrust of social knowledge as a protective strategy [@fonagy_role_2014].
@@
 
-In this study, we use a randomized control trial to test an exposure manipulation aimed at repairing potential breakdowns in RTG cooperation.
-We stratify the sample of participants based on rejection sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection.
-It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997].
-Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017).
-However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas.
-
-Participants are exposed to artificial agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
-The computerized investor in the RTG was programmed to play using a generative model of how humans play fit to real players' data from prior studies.
-A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
+<!-- In this study, we use a randomized control trial to test whether exposing participants to forgiving and more cooperative co-players can produce more trustworthy behavior and repair potential breakdowns in RTG cooperation.
+We stratify the sample of participants based on rejection sensitivity (RS).
+In order to simulate realistic social interaction while maintaining a high degree of experimental control, the computerized investor in the RTG was designed to play using a generative model of how humans play.
+We fit Hidden markov Models (HMMs) to real players' data from prior studies to build a responsive yet consistent co-player.
+A key aspect of this agent is that its actions depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
 An advantage of having a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation.
 Such generative  model of behavior, informed by empirical data and integrating a responsive trust mechanism, represent a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
+In this study, we leverage this flexibility to expose participants in between trust games to artificial agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
+In the control condition, the participants play the human like agents throughout.
+We would expect those in the manipulation condition to behave more cooperatively and have a lower propensity to retaliate to the co-player's defection.
+We would also expect those exhibiting higher rejection sensitivity to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
+-->
+
+In this study, we use a randomized control online experiment to test whether exposing participants with varying levels of RS to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation.
+In order to simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players data.
+A key aspect of these agents is that their actions depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
+An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation.
+In this study, participants were given the role of the trustee.
+After playing a 15 rounds RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition.
+In the Manipulation condition, participants were exposed over three RTGs to HMM investors designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
+In the Control condition, participants, played three RTGs against the same human-like HMM.
+After this exposure phase, all participants played another 15 round RTG with a human-like HMM investor.
+We hypothesise that those in the manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase.
+We also expect those exhibiting higher RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
 
@@
 
-A total of 206 participants were recruited on the Prolific Academic platform (prolific.co).
-<!-- The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years, with a `r round(sd(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years standard deviation.
--->
+A total of 206 participants (56% female) were recruited on the Prolific Academic platform (prolific.co).
+The mean age of participants was 34.6 years, with an 11.9 years standard deviation.
 Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.
-Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire to form two similarly sized groups: One with high (RSQ score > 15) and the other with low rejection sensitivity (RSQ score < 10).
+Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire (RSQ) to form two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10).
 
@@
 
-The experiment had a 2 (Condition: Manipulation or Control) by 2 (Rejection Sensitivity : high or low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor.
+The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : high or low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A).
 Participants within each pre-screened group were randomly assigned to one of the two levels of the first factor.
@@
 
-Participants played a 15-round Repeated Trust Game [@joyce_trust_1995] in the trustee role against a computer-programmed investor.
+Participants played a 15-round RTG [@joyce_trust_1995] in the trustee role against a computer-programmed investor.
 On each round the investor is endowed with 20 units and decides how much of that endowment to invest.
@@
 If the trustee returns more than one third of the amount, the investor makes a gain.
-Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right (Figure \@ref(fig:HMMPanels).A).
+Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right.
 The participants were able to choose the icon that represents them at the start of the experiment.
@@
 
-The strategy of the computerised investor was modelled on behavior of human investors in the Repeated Trust Game (RTG)  over 10-rounds with the same (human) co-player.
-Full detail on the datasets used in the Supplementary Information.
-Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states.
-Each latent state was associated with a state-conditional distribution over the possible investments from 0 to 20 (Figure \@ref(fig:HMMPanels).B).
-These distributions reflect "low-trust", "medium-trust", or "high-trust".
-Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C).
-In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) or round 13 (post-manipulation).
-On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
-The initial state for the HMM investor in each instance of the game was the "mid-trust" state.
-
 \begin{landscape}
-```{r HMMPanels, fig.cap= "\\small{A: Screenshot of the repeated Trust Game.
-The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee.
+```{r HMMPanels, fig.cap= "\\small{A: A timeline of the experiment for both conditions.
+The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor.
 The investor is endowed with 20 units at the start of each round.
@@
 The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor.
-Shown here is the stage at which the trustee decides how much to send back to the investor.
-Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round Repeated Trust Game against human trustees.
-From the fitted HMM, we get the distribution of investments by the artificial investor agent conditional on its latent state as shown in Panel B.
+The difference between conditions is the type of agents participants are exposed to in the exposure phase.
+Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG against human trustees.
+From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B.
 The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C.
 Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
-Panel D shows the transition probability of the forgiving HMM agent, where we can see on the left plot that the agent always transitions out of the low-trust state.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
+Panel D shows the transition probabilities of the forgiving HMM agent which are identical to those of the human-like agent, except in the low trust state.
+Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
 
@@
 # Read in the images
-image1 <- image_read("figures/trust_game.png")
+image1 <- image_read("figures/timeline.png")
 image2 <- image_read("figures/HMMinvPolicy2.png")
 image3 <- image_read("figures/HMMinvTrans2.png")
-image4 <- image_read("figures/HMMniceInvTrans.png")
+image4 <- image_read("figures/HMM_switch.png")
 
@@
 # Annotate each image with a label
-image1 <- image_annotate(image1, text = "A", location = "+200+10", size = 90, color = "black", gravity = "northwest")
+image1 <- image_annotate(image1, text = "A", location = "+0+10", size = 90, color = "black", gravity = "northwest")
 image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
@@
 
+The strategy of the computerised investor was modelled on the behavior of human investors in the RTG using an existing dataset consisting of 388  ten round games with the same co-player.
+Full detail on the datasets used in the Supplementary Information.
+Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states.
+The HMM consists of three outputs.
+First, a distribution over the possible investments from 0 to 20 for each latent state of the investor (Figure \@ref(fig:HMMPanels).B).
+These distributions can be interpreted as reflecting "low-trust", "medium-trust", or "high-trust" dispositions.
+The HMM also specifies how  the investor can move between these latent states.
+The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C).
+Finally, the HMM also provides as an output the initial distribution of latent states for the investor.
+In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) or round 13 (post-manipulation).
+On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
+The initial state for the HMM investor in each instance of the game was the "mid-trust" state.
+
 ## Manipulation
 
-To design a forgiving and ultimately more cooperative agent, we change the transition function of the investor HMM to make it impossible to remain in a low trust state once the agent transitions there.
-This is achieved by choosing the parameters of the transition function to make the probability of remaining in the "low-trust" state, when the agent is in that state, effectively nil.
-The policies conditional on the state and the transition function in the other states remain unchanged.
+The HMM resulting from fitting the model to human investor play in the RTG (Human-like HMM) isused as a basis to design a forgiving and ultimately more cooperative agent.
+To achieve that, we changed the transition function of the investor HMM to make it impossible to remain in a low trust state once the agent transitions there.
+This is done by selecting  parameters of the transition function to make the probability of remaining in the "low-trust" state, when the agent is in that state, effectively nil.
 The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
+The policies conditional on the latent states and the transition function in the other latent states remain unchanged.
 
@@
 They were not told the number of rounds of each phase.
-Phase one was a 15 round Repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds.
+Participants were randomly assigned to either a Control or Manipulation condition.
+The timeline of the experiment is shown in Figure \@ref(fig:HMMPanels).A.
+Phase one ("pre") consisted of a 15 round RTG in which participants took the role of trustee, facing the same investor over all 15 rounds.
 On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round.
-After completing 15 rounds of the RTG, participants rated how cooperative, forgiving they perceived the investor to be, and whether they would like to play with them again (all on a scale from 1 to 10).
-
-After phase one, participants in the manipulation condition played three games of 7 rounds each against the forgiving HMM agent, whilst those in the control condition faced the human-like HMM agent.
-To keep the experience similar to the pre-manipulation game, the agent in the control condition was also designed to send a very low investment in round 5 of each of the three games.
-Subsequent phase two was similar to phase one, with participants being told they would face a new player.
+After completing 15 rounds of the RTG, participants rated how cooperative, forgiving they perceived the investor to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
+
+Phase 2 ("exposure") consisted of three 7-round RTGs.
+Participants in the Manipulation condition faced the forgiving HMM investor and rated the agent on the same attributes as in the pre-manipulation phase.
+Those in the Control condition faced the same human-like HMM agent as in the "pre" phase and rated each co-player on the same attributes.
+To keep the experience similar to the "pre" phase, the agent in the control condition was also designed to send a very low investment in round 5 of each of the three games.
+In the post-manipulation phase ("post"), participants in both conditions were told they would face a new player, and faced the same human-like HMM as in "pre" phase.
 Participants then completed the Levels of Personality Functioning Scale (LPFS) questionnaire (see the supplement for details).
@@
 
-To explore whether participants behaved differently in the RTG after the manipulation compared to the control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model as described below:
-
-\[
-\begin{split}
-\text{R}_{ij} = & \, \beta_0 + \beta_1 \text{ Phase}_i + \beta_2 \text{ Condition}_i + \beta_3 \text{ Investment}_i + \beta_4 \text{ RS}_i + \\
-& \beta_5 (\text{Phase} \times \text{Condition})_i + \beta_6 (\text{Phase} \times \text{Investment})_i + \beta_7 (\text{Phase} \times \text{RS})_i + \\
-& \beta_8 (\text{Condition} \times \text{Investment})_i + \beta_9 (\text{Condition} \times \text{RS})_i + \beta_{10} (\text{Investment} \times \text{RS})_i + \\
-& \beta_{11} (\text{Phase} \times \text{Condition} \times \text{Investment})_i + \beta_{12} (\text{Phase} \times \text{Condition} \times \text{RS})_i + \\
-& \beta_{13} (\text{Phase} \times \text{Investment} \times \text{RS})_i + \beta_{14} (\text{Condition} \times \text{Investment} \times \text{RS})_i + \\
-& \beta_{15} (\text{Phase} \times \text{Condition} \times \text{Investment} \times \text{RS})_i + \\
-& b_{0j} + b_{1j} \text{ (Phase)}_i+ \epsilon_{ij}
-\end{split}
-\]
-
-where:
-
-- \( \text{R}_{ij} \): percentage of tripled investment returned to investor for participant \( j \) in observation \( i \)
-- \( \beta_0 \): intercept
-- \( \beta_1 \) to \( \beta_4 \): main effects of Phase (RTG game pre vs. post-manipulation), Condition (manipulation vs. control), Investment, and RS (High vs Low RS), respectively
-- \( \beta_5 \) to \( \beta_{10} \): interaction effects between each pair of the four factors, showing how the relationship between one factor and the return percentage not available changes depending on the level of another factor
-- \( \beta_{11} \) to \( \beta_{14} \): three-way interaction effects among the four factors, indicating how the interaction between two factors is further modified by the third factor
-- \( \beta_{15} \): four-way interaction effect between Phase, Condition, Investment, and RS, describing how the interaction among three factors is modified by the fourth factor
-- \( b_{0j} \): player-wise random intercept for player \( j \)
-- \( b_{1j} \): player-wise random slope for Phase for player \( j \)
-- \( \epsilon_{ij} \): error term for player \( j \) in observation \( i \)
+To explore whether participants behaved differently in the RTG after the manipulation compared to the control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.
+post-manipulation), Condition (manipulation vs.
+control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase.
 
@@
 More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek_balancing_2017].
-A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses.
+A similar process was used to establish the random effects structures of  linear mixed-effects models used to analyse the HMM agent investments as well as the participants' ratings of the co-players.
 For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex".
 We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
-
-<!-- To model participants' returns in the RTG across games and conditions, we fit various hidden Markov models [@visser2022hidden] to participants' returns using the depmixS4 package [@visser_depmixs4_2021] for R.
-The transition between latent states is assumed to depend on the investment received and a dummy variable to characterise the group that the participant belongs to.
-Details on how the models are constructed can be found in the supplement.
-We fit models with different numbers of hidden states, and use the Bayesian Information Criterion [@schwarz_estimating_1978-1] to select the best model.
--->
 
@@
 
-```{r afexModCoop, cache=TRUE}
-
-mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
-
-summary(mod_rating_coop)
-
-```
-
-```{r emmCoop, cache=TRUE}
-# Contrast 1: Pre vs. Post
-contrast1 <- c(-1, 0, 0, 0, 1)
-
-# Contrast 2: Pre vs. (Expo1, Expo2, Expo3) 
-contrast2 <- c(-3, 1, 1, 1, 0)
-
-# Contrast 3: Post vs. (Expo1, Expo2, Expo3) 
-contrast3 <- c(0, -1, -1, -1, 3)
-
-# Combine contrasts into a matrix
-contrast_matrix <- rbind(contrast1, contrast2, contrast3)
-colnames(contrast_matrix) <- levels(datRatings$gameNum.f)
-
-# Apply contrasts to the model using 'emmeans'
-emm_coop <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f)
-
-# Custom comparisons: use 'contrast' directly
-comp1_coop <- contrast(emm_coop, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
-comp2_coop <- contrast(emm_coop, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
-comp3_coop <- contrast(emm_coop, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
-
-comp1_coop
-
-comp2_coop
-
-comp3_coop
-
-```
-
-```{r}
-# Compute the estimated marginal means (EMMs) for the interaction
-emm_coop_interaction <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f * high_RS)
-
-comp1_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Post" = contrast_matrix[1,]))
-comp2_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
-# comp3_coop_interaction <- contrast(emm_coop_interaction, list("Post vs Expo" = contrast_matrix[3,]))
-
-# Output the results
-print("Pre vs Post")
-comp1_coop_interaction
-print("Pre vs Expo")
-test(comp2_coop_interaction, adjust="bonferroni")
-test(comp2_coop_interaction)
-
-# test_comp3_coop_interaction
-```
-
-```{r afexModForg}
-
-mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
-
-#summary(mod_rating_forg)
-
-```
-
-```{r emmForg}
-
-# Apply contrasts to the model using 'emmeans'
-emm_forg <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f)
-
-# Custom comparisons: use 'contrast' directly
-comp1_forg <- contrast(emm_forg, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
-comp2_forg <- contrast(emm_forg, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
-comp3_forg <- contrast(emm_forg, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
-
-comp1_forg
-
-comp2_forg
-
-comp3_forg
-
-```
-
-```{r}
-# Compute the estimated marginal means (EMMs) for the interaction
-emm_forg_interaction <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f * high_RS)
-
-comp1_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Post" = contrast_matrix[1,]))
-comp2_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
-# comp3_forg_interaction <- contrast(emm_forg_interaction, list("Post vs Expo" = contrast_matrix[3,]))
-
-# Output the results
-print("Pre vs Post")
-comp1_forg_interaction
-print("Pre vs Expo")
-comp2_forg_interaction
-
-# test_comp3_forg_interaction
-```
-
-```{r afexModAgain}
-
-mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
-
-#summary(mod_rating_again)
-
-```
-
-```{r emmAgain}
-
-# Apply contrasts to the model using 'emmeans'
-emm_again <- emmeans(mod_rating_again, ~ gameNum.f | condition.f)
-
-# Custom comparisons: use 'contrast' directly
-comp1_again <- contrast(emm_again, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
-comp2_again <- contrast(emm_again, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
-comp3_again <- contrast(emm_again, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
-
-comp1_again
-
-comp2_again
-
-comp3_again
-
-```
-
-```{r}
-# Compute the estimated marginal means (EMMs) for the interaction
-emm_again_interaction <- emmeans(mod_rating_again, ~ gameNum.f | condition.f * high_RS)
-
-comp1_again_interaction <- contrast(emm_again_interaction, list("Pre vs Post" = contrast_matrix[1,]))
-comp2_again_interaction <- contrast(emm_again_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
-# comp3_again_interaction <- contrast(emm_again_interaction, list("Post vs Expo" = contrast_matrix[3,]))
-
-# Output the results
-print("Pre vs Post")
-comp1_again_interaction
-print("Pre vs Expo")
-comp2_again_interaction
-
-# test_comp3_again_interaction
-```
-
-## Player ratings
-
-Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition and RS group.
-We will focus on two contrasts to analyse the ratings, by Condition and RS group.
-The first is between the rating in the first phase ("pre") when participants phase the human-like HMM, and the average rating during the exposure phase (average of "expo1", "expo2" and "expo3") where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition).
-The second contrast is between the "pre" and "post" phases of the experiment where in both conditions participants face the same human-like HMM.
-
-### Comparing pre and exposure ratings
-
-For those with high Rejection Sensitivity, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`.
-There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
-Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
-
-For those with low Rejection Sensitivity, there was no difference in any of the ratings between the "pre" and "exposure" phases for the Manipulation condition.
-For the Control condition, participants indicated less willingness to face the exposure co-player compared to the "pre" player, `r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo` but also did not differ on cooperation and forgiveness ratings.
-In summary, those with low rejection sensitivity had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving.
-In contrast, we see that exposing participants with high Rejection Sensitivity to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.
-
-```{r plotRatings,include=T, echo=FALSE, fig.cap="Averages and standard errors of the participants ratings of the opponent for each game and condition.
-Pre and post are the 15 round repeated trust games before and after the exposure phase respectively.
-The games titled expo 1 to 3 are the three 7 round games during the exposure phase.
-We note that absent the exposure to the forgiving AI, the ratings get worse on aggregate even through the participant faces the same human like HMM.
-In the manipulation condition, participants rate the nicer HMM as more cooperative but not more forgiving.
-When they face the human-like HMM again, its rating are considerably worse.
-",fig.align="center", fig.width=6, fig.height = 6}
-
-# ggplot(summary_df, aes(x = gameNum.f, y = mean, group = rating_type, color = rating_type)) +
-#   geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem), width = 0.2) +
-#   geom_point(position = position_dodge(width = 0.5), size = 3) +
-#   facet_wrap(~~rating_type*condition.f, nrow = 3, ncol = 2, scales = "free") +
-#   labs(x = "Game Number", y = "Mean Rating") +
-#   theme_minimal() +
-#   theme(legend.position = "bottom")
-
-ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f), color = rating_type)) +
-  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1, position = position_dodge(width = 0.25)) +
-  geom_line(linetype = "dashed", position = position_dodge(width = 0.25)) +
-  geom_point(position = position_dodge(width = 0.25), size = 3) +
-  facet_wrap(~high_RS*condition.f, scales = "free_x") +
-  labs(x = "Phase", y = "Mean Rating", color = "Rating Type") +
-  scale_color_manual(values = c("rating_coop" = "blue", "rating_forgiving" = "red", "rating_playAgain" = "green"),
-                     labels = c("Cooperative", "Forgiving", "Play again")) +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-```
-
-```{r}
-library(ggplot2)
-
-ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f))) +
-  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem, color = condition.f), 
-                width = .1, position = position_dodge(width = 0.25)) +
-  geom_line(aes(color = condition.f), linetype = "dashed", position = position_dodge(width = 0.25)) +
-  geom_point(aes(color = condition.f), position = position_dodge(width = 0.25), size = 3) +
-  facet_grid(high_RS ~ rating_type, scales = "free_x") +
-  scale_color_manual(values = c("Control" = "blue", "Manipulation" = "red")) +
-  labs(x = "Phase", y = "Mean Rating", color = "Condition") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-```
-
-### Comparing pre and post ratings
-<!-- Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
-<!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
-Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
--->
-
-Participants high on Rejection Sensitivity in the Manipulation condition rated the co-players in the "post" phase similarly on cooperation, lower on 
-forgiveness `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
-and lower on willingness to face them again `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`. 
-Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
-(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
-Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
-Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).
-
-For participants low on Rejection Sensitivity, those in the Manipulation condition rated the co-players in the "post" phase lower on all three attributes 
-(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
-Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
-Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`). 
-Those in the Control condition did not differ in their ratings of the "pre" and "post" phase players.
-
-In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the control condition.
-In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the control condition even though the player continues to use the same strategy.
-After exposure to the forgiving HMM, both groups  rates the "post" co-player worse than the "pre" even though they are the same.
-
-```{r}
-d_finished %>% 
-  dplyr::select(id,Turing.choice) %>% 
-  unique() %>% 
-  group_by(Turing.choice) %>%
-  summarise(Count = n()) %>%
-  mutate(Percentage = (Count / sum(Count)) * 100)
-```
-
 ## Analysis of participant returns
@@
 ```{r}
-RS_model <- lm( RS_score ~ ret_pct_na*inv_scaled + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+#RS_model <- lm( RS_score ~ ret_pct_na*inv_scaled + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
 ```
@@
 prepost_bycond <- pairs(emmeans::emmeans(mod_returns_pct_RS, c("phase.f"), by = "condition.f", pbkrtest.limit = 6300))
+prepost_bycond
 ```
@@
 
-Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs.
-Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the first RTG compared to the second.
-Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.
-post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
-As shown in Figure \ref{fig:violinPlots}, post-hoc tests confirm a decrease in the percentage returned only in the manipulation condition, pre - post,
-`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.
-We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.
-
-```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round.
-The blue line shows the returns pre-manipulation and the green line post-manipulation.
-We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the manipulation condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}
+```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received (y-axis) by Condition, Phase, and game round (x-axis).
+The blue line shows the returns in the Pre phase and the green line those in the Post phase.
+The left Panel shows returns in the Control condition and the right one those in the Manipulation condition.
+The dotted lines identify the rounds where the pre-programmed one-off low investment occurs.
+We note lower average returns post vs pre in the manipulation condition, whilst returns in the control condition are similar between the two phases.",fig.align="center", fig.width=6, fig.height = 4}
 
@@
 
+Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs.
+Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the Pre RTG compared to the Post.
+Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.
+post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
+As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the manipulation condition, pre - post,
+`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.
+We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.
+
 There was also a significant main effect of Investment, 
@@
 Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
-
-```{r fourwayInteractionPlot, include=T, fig.cap ="Effect of investment on returns by condition, phase and RS group" }
+For completeness, these effects are discussed in more detail the supplement.
+
+<!-- # USING LMER TO LEVERAGE CAPABILITIES WITH EMTRENDS -->
+
+```{r}
+library(lmerTest)
+library(emmeans)
+
+# Refitting the model directly
+direct_model <- lmer(ret_pct_na ~ phase.f * condition.f * inv_scaled * high_RS + (1 + phase.f | playerId), data = prepost_df, REML = TRUE)
+
+emm_res <- emmeans(direct_model, specs = ~ phase.f * condition.f * high_RS * inv_scaled)
+trends <- emtrends(direct_model, specs = ~ phase.f * condition.f * high_RS, var = "inv_scaled")
+
+# Print results
+summary(emm_res)
+summary(trends)
+
+```
+
+```{r fourwayInteractionPlot, include=F, fig.cap ="Visualising the four way interaction between Phase, RS group, Condition and Investment.
+The plot shows the effect of the investment on participant returns for each condition, phase and RS group.
+The interesting effect is in the Manipulation Condition: " }
 
@@
 
-To interpret this interaction we plot the effect of the Investment on returns for each level of Condition, Phase and RS.
+<!-- To interpret this interaction we plot the effect of the Investment on returns for each level of Condition, Phase and RS.
 Figure \ref{fig:fourwayInteractionPlot} shows that while the effect of investment on returns is similar across RS groups and across phases in the Control condition, the Manipulation condition shows a higher effect of investment in the post phase compared to pre in the low RS group, but a lower effect of investment compared to before for the high RS group.
 This suggests that high RS participants based their return less on the investment after being exposed to the forgiving AI compared to the low RS group.
+-->
 
@@
 ```{r violPlotsDef, cache=TRUE}
-ret_all <- afex_plot(mod_returns_pct_RS, x = "phase.f", trace = "condition.f", dodge = 0.8, error = "within",
+library(afex)
+library(ggplot2)
+library(ggsignif)
+
+# Generate the plot with points and error bars
+ret_all <- afex_plot(mod_returns_pct_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
                      mapping = c("linetype", "shape", "fill"),
-            data_geom = ggplot2::geom_boxplot,
-            data_arg = list(width = 0.5),
-            factor_levels = list(phase.f = c("Pre", "Post")),
-            legend_title = "Condition") +
+                     factor_levels = list(condition.f = c("Control", "Manipulation")),
+                     legend_title = "Phase")
+
+# Extract the data used in afex_plot
+plot_data <- ret_all$data
+
+# Generate the plot
+ret_all <- ggplot(plot_data, aes(x = condition.f, y = y, color = phase.f, group = phase.f)) +
+  geom_point(position = position_dodge(width = 0.8)) +
+  geom_errorbar(aes(ymin = y - SE, ymax = y + SE), width = 0.2, position = position_dodge(width = 0.8)) +
   theme_bw() +
-  labs(y = "% Returns", x = "Trust Game Phase") +
-  theme(plot.title = element_text(size = 16)) +
-  geom_signif(xmin = 1.2, xmax = 2.2, annotations = "***", y_position = 0.82, tip_length = 0.03,textsize = 3) + 
-  theme(legend.position = "bottom")
-
-inv_all <- afex_plot(mod_invs_RS, x = "phase.f", trace = "condition.f", dodge = 0.8, error = "within",
+  labs(y = "% Returns", x = "Condition", color = "Trust Game Phase") +
+  theme(plot.title = element_text(size = 16), legend.position = "bottom") +
+  geom_signif(
+              annotations = "***", 
+              xmin = 1.8, xmax = 2.2,
+              y_position = 0.47, 
+              tip_length = 0.03, 
+              textsize = 3, 
+              color = "black")
+
+######################################################################################################
+
+# Generate the plot with points and error bars
+inv_all <- afex_plot(mod_invs_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
                      mapping = c("linetype", "shape", "fill"),
-            data_geom = ggplot2::geom_boxplot,
-            data_arg = list(width = 0.5),
-            factor_levels = list(phase.f = c("Pre", "Post")),
-            legend_title = "Condition") +
+                     factor_levels = list(condition.f = c("Control", "Manipulation")),
+                     legend_title = "Phase")
+
+# Extract the data used in afex_plot
+plot_data <- inv_all$data
+
+# Generate the plot
+inv_all <- ggplot(plot_data, aes(x = condition.f, y = y, color = phase.f, group = phase.f)) +
+  geom_point(position = position_dodge(width = 0.8)) +
+  geom_errorbar(aes(ymin = y - SE, ymax = y + SE), width = 0.2, position = position_dodge(width = 0.8)) +
   theme_bw() +
-  labs(y = "HMM Investment", x = "Trust Game Phase") +
-  theme(plot.title = element_text(size = 16)) +
-  theme(legend.position = "bottom")
-```
-
-```{r violinPlots, include=T,fig.cap="Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition.
+  labs(y = "HMM Investment", x = "Condition", color = "Trust Game Phase") +
+  theme(plot.title = element_text(size = 16), legend.position = "bottom")+
+  ylim(9, 11)
+
+```
+
+```{r boxPlots, include=T,fig.cap="Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition.
 The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns.
-The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.",fig.align="center", fig.width=6, fig.height = 7}
+The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.",fig.align="center", fig.width=6, fig.height = 8}
 
@@
 `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the manipulation condition compared to the control condition in post-defection trials.
+However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.
 
@@
 <!-- Linear mixed effects model for investments -->
-To test whether the HMM investors' behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
-As seen in Figure \@ref(fig:violinPlots), we find no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.
+To test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
+As seen in Figure \@ref(fig:boxPlots), we find no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.
 
@@
 cor_result <- cor.test(temp2$RS_score, temp2$LPFS_score, method = "spearman")
-
-```
-
-Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity score `r papaja::apa_print(cor_result)$full_result`, there was no correlation between these and participant's return or overall task performance.
+# Format the result using papaja
+apa_result <- papaja::apa_print(cor_result)
+
+print(apa_result)
+# Extract the correlation coefficient and p-value
+spearman_rho <- apa_result$estimate
+
+```
+
+Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity score (RSQ), Spearman's `r spearman_rho`, $p < 0.001$, there was no correlation between these questionnaire scores and participant's return or overall task performance.
+
+<!---------------------- Modelling of co-player ratings -------------------------------------->
+
+```{r afexModCoop, cache=TRUE}
+
+mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+summary(mod_rating_coop)
+
+```
+
+```{r emmCoop, cache=TRUE}
+# Contrast 1: Pre vs. Post
+contrast1 <- c(-1, 0, 0, 0, 1)
+
+# Contrast 2: Pre vs. (Expo1, Expo2, Expo3) 
+contrast2 <- c(-3, 1, 1, 1, 0)
+
+# Contrast 3: Post vs. (Expo1, Expo2, Expo3) 
+contrast3 <- c(0, -1, -1, -1, 3)
+
+# Combine contrasts into a matrix
+contrast_matrix <- rbind(contrast1, contrast2, contrast3)
+colnames(contrast_matrix) <- levels(datRatings$gameNum.f)
+
+# Apply contrasts to the model using 'emmeans'
+emm_coop <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f)
+
+# Custom comparisons: use 'contrast' directly
+comp1_coop <- contrast(emm_coop, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
+comp2_coop <- contrast(emm_coop, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
+comp3_coop <- contrast(emm_coop, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
+
+comp1_coop
+
+comp2_coop
+
+comp3_coop
+
+```
+
+```{r}
+# Compute the estimated marginal means (EMMs) for the interaction
+emm_coop_interaction <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f * high_RS)
+
+comp1_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp2_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
+# comp3_coop_interaction <- contrast(emm_coop_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Output the results
+print("Pre vs Post")
+comp1_coop_interaction
+print("Pre vs Expo")
+test(comp2_coop_interaction, adjust="bonferroni")
+test(comp2_coop_interaction)
+
+# test_comp3_coop_interaction
+```
+
+```{r afexModForg}
+
+mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+#summary(mod_rating_forg)
+
+```
+
+```{r emmForg}
+
+# Apply contrasts to the model using 'emmeans'
+emm_forg <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f)
+
+# Custom comparisons: use 'contrast' directly
+comp1_forg <- contrast(emm_forg, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
+comp2_forg <- contrast(emm_forg, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
+comp3_forg <- contrast(emm_forg, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
+
+comp1_forg
+
+comp2_forg
+
+comp3_forg
+
+```
+
+```{r}
+# Compute the estimated marginal means (EMMs) for the interaction
+emm_forg_interaction <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f * high_RS)
+
+comp1_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp2_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
+# comp3_forg_interaction <- contrast(emm_forg_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Output the results
+print("Pre vs Post")
+comp1_forg_interaction
+print("Pre vs Expo")
+comp2_forg_interaction
+
+# test_comp3_forg_interaction
+```
+
+```{r afexModAgain}
+
+mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")
+
+#summary(mod_rating_again)
+
+```
+
+```{r emmAgain}
+
+# Apply contrasts to the model using 'emmeans'
+emm_again <- emmeans(mod_rating_again, ~ gameNum.f | condition.f)
+
+# Custom comparisons: use 'contrast' directly
+comp1_again <- contrast(emm_again, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
+comp2_again <- contrast(emm_again, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
+comp3_again <- contrast(emm_again, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")
+
+comp1_again
+
+comp2_again
+
+comp3_again
+
+```
+
+```{r}
+# Compute the estimated marginal means (EMMs) for the interaction
+emm_again_interaction <- emmeans(mod_rating_again, ~ gameNum.f | condition.f * high_RS)
+
+comp1_again_interaction <- contrast(emm_again_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp2_again_interaction <- contrast(emm_again_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
+# comp3_again_interaction <- contrast(emm_again_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Output the results
+print("Pre vs Post")
+comp1_again_interaction
+print("Pre vs Expo")
+comp2_again_interaction
+
+# test_comp3_again_interaction
+```
+
+## Player ratings
+
+Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition and RS group.
+We will focus on two contrasts to analyse the ratings by Condition and RS group.
+The first is between the rating in the first phase ("pre") when participants phase the human-like HMM, and the average rating during the exposure phase (average of "expo1", "expo2" and "expo3") where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition).
+The second contrast is between the "pre" and "post" phases of the experiment where in both conditions participants face the same human-like HMM.
+
+### Comparing pre and exposure ratings
+
+For those with high RS, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`.
+There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
+Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
+
+For those with low RS, there was no difference in any of the ratings between the "pre" and "exposure" phases for the Manipulation condition.
+For the Control condition, participants indicated less willingness to face the exposure co-player compared to the "pre" player, `r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo` but also did not differ on cooperation and forgiveness ratings.
+In summary, those with low RS had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving.
+In contrast, we see that exposing participants with high RS to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.
+
+```{r plotRatings,include=T, echo=FALSE, fig.pos='H', fig.cap="Averages and standard errors of the participants ratings of the opponent (y-axis) by each game and condition for each phase (x-axis).
+Pre and post are the 15 round RTGs before and after the exposure phase respectively.
+The games titled expo 1 to 3 are the three 7 round games during the exposure phase.
+The blue line represents participants' perception of co-player cooperation, the red line indicates perceived co-player forgiveness, and the green line shows the participants' willingness to play again with the same co-player.
+Cooperation, forgiveness, and willingness to play again ratings remained relatively stable for the low RS group except post-manipulation, where they were lower.
+In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the manipulation condition, but decreased over time in the control group,.
+Ratings were also markedly lower in the post-manipulation game for this group.",fig.align="center", fig.width=6, fig.height = 6}
+
+ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f), color = rating_type)) +
+  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1, position = position_dodge(width = 0.25)) +
+  geom_line(linetype = "dashed", position = position_dodge(width = 0.25)) +
+  geom_point(position = position_dodge(width = 0.25), size = 3) +
+  facet_wrap(~high_RS*condition.f, scales = "free_x") +
+  labs(x = "Phase", y = "Mean Rating", color = "Rating Type") +
+  scale_color_manual(values = c("rating_coop" = "lightblue", "rating_forgiving" = "#FF6961", "rating_playAgain" = "#77DD77"),
+                     labels = c("Cooperative", "Forgiving", "Play again")) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+# temporarily paste code from Figure 2 to compare styles: 
+# Plot
+ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
+  stat_summary(fun = mean, geom = "line") +
+  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
+               alpha = 0.3, linetype = 0) +
+  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
+  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
+  labs(x = "Round",
+       y = "Percentage Return",
+       color = "Trust Game Phase") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        legend.box = "vertical", # This will stack the legends vertically
+        legend.key = element_blank(),
+        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
+        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10)) +
+  scale_color_manual(values = c("darkblue", "darkgreen"),
+                     labels = c("Pre-manipulation", "Post-manipulation")) +
+  scale_fill_manual(values = c("darkblue", "darkgreen"),
+                    guide = "none") + # Hide the fill legend
+  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
+  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
+         linetype = guide_legend(title = "Defection round")) +
+  facet_wrap(~condition.f)
+
+```
+
+```{r}
+library(ggplot2)
+
+ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f))) +
+  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem, color = condition.f), 
+                width = .1, position = position_dodge(width = 0.25)) +
+  geom_line(aes(color = condition.f), linetype = "dashed", position = position_dodge(width = 0.25)) +
+  geom_point(aes(color = condition.f), position = position_dodge(width = 0.25), size = 3) +
+  facet_grid(high_RS ~ rating_type, scales = "free_x") +
+  scale_color_manual(values = c("Control" = "blue", "Manipulation" = "red")) +
+  labs(x = "Phase", y = "Mean Rating", color = "Condition") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
+### Comparing pre and post ratings
+<!-- Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
+<!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
+Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
+-->
+
+Participants high on RS in the Manipulation condition rated the co-players in the "post" phase similarly on cooperation, lower on 
+forgiveness `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
+and lower on willingness to face them again `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`. 
+Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
+(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
+Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
+Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).
+
+For participants low on RS, those in the Manipulation condition rated the co-players in the "post" phase lower on all three attributes 
+(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
+Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
+Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`). 
+Those in the Control condition did not differ in their ratings of the "pre" and "post" phase players.
+
+In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the control condition.
+In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the control condition even though the player continues to use the same strategy.
+After exposure to the forgiving HMM, both groups rate the "post" co-player worse than the "pre" even though they are the same.
+
+```{r}
+d_finished %>% 
+  dplyr::select(id,Turing.choice) %>% 
+  unique() %>% 
+  group_by(Turing.choice) %>%
+  summarise(Count = n()) %>%
+  mutate(Percentage = (Count / sum(Count)) * 100)
+```
 
@@
 
-This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of Hidden Markov Models (HMMs) as artificial agents in economic games.
-The use of HMM-based artificial agents in economic games led to similar investment and returns to those recorded in human dyadic interactions.
-Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.
-This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control.
-
-Following the exposure manipulation, participants reduced their returns overall whilst the returns of those in the control group did not change between the pre and post phase of the experiment.
+We used a randomized control online experiment where participants played a RTG with AI agents designed to simulate human-like trust-building scenarios.
+Participants were then exposed to either forgiving or human-like AI agents before playing another RTG.
+Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation.
+Instead, participants reduced their returns overall whilst the returns of those in the control group did not change between the pre and post phase of the experiment.
+Neither did participants show more forgiving behavior in the manipulation condition after the one-off defection by the agent.
 Why did participants reduce their returns even though they were repeatedly exposed to a more cooperative and more forgiving AI?
-A look at how the participants rated their co-players might shed some light on what is driving this reduction in returns for those exposed to the forgiving AI.
-
-Those exposed to the forgiving AI rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure.
+A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the manipulation condition.
+
+Those exposed to the forgiving agent rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure.
 One possible explanation for this drop in rating is that participants exhibited a negative contrast effect.
@@
 
-It is worth highlighting that Rejection Sensitivity was not correlated to overall performance in the task.
+It is worth highlighting that RS was not correlated to overall performance in the task.
 Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
-However, in examining the player ratings by rejection sensitivity (RS) group more closely, we observe that individuals with high rejection sensitivity demonstrate a heightened attunement to changes in the behavior of their AI co-players.
+However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their AI co-players.
 Specifically, when exposed to a more forgiving AI agent, these participants accurately increased their ratings of the agent, indicating a sensitive and appropriate response to the behavioral manipulation.
@@
 
-Importantly, our approach utilizing HMMs to simulate interactive partners in social dilemmas offers a promising avenue for future research, especially in understanding and potentially mitigating trust deficits in individuals with high rejection sensitivity.
+This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of generative models of human behavior to design artificial agents that can interact in economic games.
+The use of these agents in the RTG led to similar investment and returns to those recorded in human dyadic interactions.
+Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.
+This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control.
+Importantly, our approach utilizing HMMs to simulate interactive partners in social dilemmas offers a promising avenue for future research, especially in understanding and potentially mitigating trust deficits in individuals with high RS.
 The nuanced behaviors elicited through the interaction with HMM agents highlight the complexity of trust dynamics and the potential for computational models to offer novel insights into human social behaviors.
@@
 
-<!-- The study found that RSQ (Rejection Sensitivity Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game.
+<!-- The study found that RSQ (RS Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game.
 This gap highlights a need for further research, despite theoretical connections suggested in the literature review.

### 2024-06-05 supplement.Rmd
version 3

@@
 
-We fit a linear mixed effects model to participant returns as a proportion of the multiplied investment received. The results of the model are presented in attached table.
+We fit a linear mixed effects model to participant returns as a proportion of the multiplied investment received as described below. The results of the model are presented in Table \ref{tab:myTable}
+
+\[
+\begin{split}
+\text{R}_{ij} = & \, \beta_0 + \beta_1 \text{ Phase}_i + \beta_2 \text{ Condition}_i + \beta_3 \text{ Investment}_i + \beta_4 \text{ RS}_i + \\
+& \beta_5 (\text{Phase} \times \text{Condition})_i + \beta_6 (\text{Phase} \times \text{Investment})_i + \beta_7 (\text{Phase} \times \text{RS})_i + \\
+& \beta_8 (\text{Condition} \times \text{Investment})_i + \beta_9 (\text{Condition} \times \text{RS})_i + \beta_{10} (\text{Investment} \times \text{RS})_i + \\
+& \beta_{11} (\text{Phase} \times \text{Condition} \times \text{Investment})_i + \beta_{12} (\text{Phase} \times \text{Condition} \times \text{RS})_i + \\
+& \beta_{13} (\text{Phase} \times \text{Investment} \times \text{RS})_i + \beta_{14} (\text{Condition} \times \text{Investment} \times \text{RS})_i + \\
+& \beta_{15} (\text{Phase} \times \text{Condition} \times \text{Investment} \times \text{RS})_i + \\
+& b_{0j} + b_{1j} \text{ (Phase)}_i+ \epsilon_{ij}
+\end{split}
+\]
+
+where:
+
+- \( \text{R}_{ij} \): percentage of tripled investment returned to investor for participant \( j \) in observation \( i \)
+- \( \beta_0 \): intercept
+- \( \beta_1 \) to \( \beta_4 \): main effects of Phase (RTG game pre vs. post-manipulation), Condition (manipulation vs. control), Investment, and RS (High vs Low RS), respectively
+- \( \beta_5 \) to \( \beta_{10} \): interaction effects between each pair of the four factors, showing how the relationship between one factor and the return percentage not available changes depending on the level of another factor
+- \( \beta_{11} \) to \( \beta_{14} \): three-way interaction effects among the four factors, indicating how the interaction between two factors is further modified by the third factor
+- \( \beta_{15} \): four-way interaction effect between Phase, Condition, Investment, and RS, describing how the interaction among three factors is modified by the fourth factor
+- \( b_{0j} \): player-wise random intercept for player \( j \)
+- \( b_{1j} \): player-wise random slope for Phase for player \( j \)
+- \( \epsilon_{ij} \): error term for player \( j \) in observation \( i \)
 

### 2024-06-26 HMMfit.Rmd
trstee HMM fit & joint state analytics

---
title: "HMM fitting to data of participants returns in the Exposure experiment"
author: "Ismail Guennouni"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include =F)
```

```{r}
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(grid)
library(gridExtra)
library(forcats)
library(ggsignif)
library(magick)
library(doParallel)
library(factoextra)

```

## Fitting HMM model to participants data

Here we fit an HMM model to data generated by participants in both RTGs (pre and post manipulation) for both conditions. We specify that the transition between latent states depends on the investment received. In order to distinguish transitons by condition and Phase, we add a contrast variabel that has three levels (Pre-manipulation, post-control and post exposure). We compare models with different number of states and choose the one with the lowest BIC.

```{r}
load("data/d_finished.RData")

# select columns of interest and also solve issue with the participant with two RS_scores by averaging them. 
data_HMM <- d_finished %>%
  dplyr::select(playerId, roundType, investment, returns, roundNum, gameNum.f, condition.f, RS_score, LPFS_score, investorState.f) %>% 
  filter(roundType == "trust", !is.na(roundNum), gameNum.f %in% c("pre", "post")) %>% 
  dplyr::select(-roundType) %>%
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment == 0, 0, returns/(3*investment))) %>%
  group_by(playerId, investment, returns, roundNum, gameNum.f, condition.f, LPFS_score, investorState.f, inv_pct, ret_pct_0) %>%
  summarize(RS_score = mean(RS_score, na.rm = TRUE)) %>%
  ungroup()

# Data for both pre and post games
data_HMM_all <- data_HMM %>% 
  arrange(playerId, gameNum.f, roundNum) %>%
  group_by(playerId,gameNum.f) %>% 
  dplyr::mutate(next_investment = lead(investment, default=0),
                full_ctrst = factor(ifelse(gameNum.f=="pre", 0,ifelse(condition.f=="Control",1,2)))) %>%
  ungroup() %>% 
  unique()

# FOCUS ON POST GAMES ONLY 
data_HMM_post <- data_HMM %>% 
  filter(gameNum.f == "post") %>% 
  group_by(playerId) %>% 
  dplyr::mutate(next_investment = lead(investment, default=0)) %>%
  ungroup() %>%
  unique()

```

<!-- Here we extend the depmixS4 package functionality for analyzing data using hidden Markov models (HMMs). It involves defining new response classes and their associated methods. -->
```{r}

# Define a response class representing Gaussian distributions with discrete support
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

# Define a generic function for creating instances of the discgaus class
setGeneric("discgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("discgaus"))

# Define a method for creating instances of the discgaus class
setMethod("discgaus", 
          signature(y="ANY"), 
          function(y, pstart=NULL, fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), ...) {
            # Convert y to a matrix
            y <- matrix(y, length(y))
            # Create x matrix
            x <- matrix(1)
            # Initialize parameters list
            parameters <- list()
            # Number of parameters
            npar <- 2
            # Check if fixed parameters are provided, else initialize as FALSE
            if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
            # Check if initial parameters are provided, else initialize default values
            if(!is.null(pstart)) {
              if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu = 10, sigma = 3)
            }
            # Create a new object of class discgaus
            mod <- new("discgaus", parameters=parameters, fixed=fixed, x=x, y=y, npar=npar, breaks=breaks)
            mod
          }
)

# Define method for displaying parameters of discgaus class
setMethod("show", "discgaus",
          function(object) {
            cat("Gaussian with discrete support\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

# Define method for computing density of discgaus class
setMethod("dens", "discgaus",
          function(object, log=FALSE) {
            p <- pnorm(object@breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - 
                 pnorm(object@breaks[-length(object@breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) 
            else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

# Define method for setting parameters of discgaus class
setMethod("setpars", "discgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values) != npar) stop("length of 'values' must be", npar)
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

# Define method for getting parameters of discgaus class
setMethod("getpars", "discgaus",
          function(object, which="pars", ...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

# Define method for fitting discgaus class
setMethod("fit", "discgaus",
          function(object, w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object, c(pars[1], exp(pars[2])))
            object
          }
)

# Define a response class representing truncated Gaussian distributions with discrete support
setClass("truncdiscgaus", contains="discgaus", slots=c(min="numeric", max="numeric"))

# Define method for computing density of truncdiscgaus class
setMethod("dens", "truncdiscgaus",
          function(object, log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            prec <- pnorm(object@max, mean = object@parameters$mu, sd = object@parameters$sigma) - 
                    pnorm(object@min, mean = object@parameters$mu, sd = object@parameters$sigma)
            if(is.na(prec) || prec < 1e-12) {
              p <- rep(1/(length(breaks)-1), length(object@y))
              if(log) return(log(p)) else return(p)
            } else {
              p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - 
                   pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
              p <- p/sum(p)
              if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) 
              else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
            }
          }
)

# Define a generic function for creating instances of the truncdiscgaus class
setGeneric("truncdiscgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("truncdiscgaus"))

# Define a method for creating instances of the truncdiscgaus class
setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y, pstart=NULL, fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), min=-0.5, max=20.5, ...) {
            y <- matrix(y, length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
            if(!is.null(pstart)) {
              if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus", parameters=parameters, fixed=fixed, x=x, y=y, npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)

# Define a response class representing Gaussian distributions with variable discrete support
setClass("vtdgaus", contains="response", slots=c(yield="numeric"))

# Define a generic function for creating instances of the vtdgaus class
setGeneric("vtdgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("vtdgaus"))

# Define a method for creating instances of the vtdgaus class
setMethod("vtdgaus", 
          signature(y="ANY"), 
          function(y, yield, pstart=NULL, fixed=NULL, ...) {
            y <- matrix(y, length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
            if(!is.null(pstart)) {
              if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu=.5, sigma=1)
            }
            mod <- new("vtdgaus", parameters=parameters, fixed=fixed, x=x, y=y, npar=npar, yield=yield)
            mod
          }
)

# Define method for displaying parameters of vtdgaus class
setMethod("show", "vtdgaus",
          function(object) {
            cat("Gaussian with variable discrete support for percentage responses\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

# Define method for computing density of vtdgaus class
setMethod("dens", "vtdgaus",
          function(object, log=FALSE) {
            prec <- pnorm(1 + .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma) - 
                    pnorm(0 - .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma)
            if(prec < 1e-12) {
              p <- 1/(object@yield + 1)
            } else {
              p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - 
                   pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
              norm <- pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - 
                      pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
              p <- p/norm
            }
            p[object@yield == 0] <- 1
            if(log) return(log(p)) else return(p)
          }
)

# Define method for setting parameters of vtdgaus class
setMethod("setpars", "vtdgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values) != npar) stop("length of 'values' must be", npar)
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

# Define method for getting parameters of vtdgaus class
setMethod("getpars", "vtdgaus",
          function(object, which="pars", ...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

# Define method for fitting vtdgaus class
setMethod("fit", "vtdgaus",
          function(object, w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object, c(pars[1], exp(pars[2])))
            object
          }
)

# Function for reordering states in an HMM based on labels
label_switch <- function(mod, labels) {
  if(!is(mod, "depmix") || !is(mod, "depmix.fitted")) stop("this function is for depmix models")
  n_states <- mod@nstates
  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
    stop("labels needs to be a vector of unique integers between 1 and", n_states)
  }
  inv_labels <- sapply(1:n_states, function(x) which(labels == x))
  tmp <- mod
  # Relabel prior
  ppars <- getpars(mod@prior)
  fpars <- getpars(mod@prior, which="fixed")
  out_pars <- as.numeric(t(matrix(ppars, nrow=length(ppars)/n_states, byrow=TRUE)[, inv_labels]))
  out_fixed <- as.logical(t(matrix(fpars, nrow=length(fpars)/n_states, byrow=TRUE)[, inv_labels]))
  if(!tmp@prior@family$link == "identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
  # Relabel transition
  for(i in 1:n_states) {
    ppars <- getpars(mod@transition[[inv_labels[i]]])
    fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
    out_pars <- c(out_pars, as.numeric(t(matrix(ppars, nrow=length(ppars)/n_states, byrow=TRUE)[, inv_labels])))
    out_fixed <- c(out_fixed, as.logical(t(matrix(fpars, nrow=length(fpars)/n_states, byrow=TRUE)[, inv_labels])))
    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
    if(!tmp@transition[[i]]@family$link == "identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
  }
  # Relabel response
  for(i in 1:n_states) {
    out_pars <- c(out_pars, unlist(lapply(mod@response[[inv_labels[i]]], getpars)))
    out_fixed <- c(out_fixed, unlist(lapply(mod@response[[inv_labels[i]]], getpars, which="fixed")))
  }
  tmp <- setpars(tmp, out_fixed, which="fixed")
  tmp <- setpars(tmp, out_pars)
  if(is(tmp, "depmix.fitted")) tmp@posterior <- viterbi(tmp)
  return(tmp)
}

# Function for ordering states in a truncated discretised Gaussian model
order_mod_truncdiscgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0, ns)
  for(i in 1:ns) {
    # Calculate expected return in each state
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- truncdiscgaus(seq(0, 20), pstart=tpars, min=-0.5, max=20.5)
    sum[i] <- sum(seq(0, 20) * dens(dmod))
  }
  cat("Expected values for states: ", sum, "\n")
  # Reorder the states
  mod <- label_switch(mod, rank(sum, ties.method="first"))
  return(mod)
}

# Function for ordering states in a variable response truncated discretised Gaussian model
order_mod_vtdgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0, ns)
  for(i in 1:ns) {
    # Calculate expected return in each state
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- vtdgaus(seq(0, 1, length=61), pstart=tpars, yield=rep(60, 61))
    sum[i] <- sum(seq(0, 1, length=61) * dens(dmod))
  }
  # Reorder the states
  mod <- label_switch(mod, rank(sum, ties.method="first"))
  return(mod)
}

```

# Fitting HMM to ALL Participants data (Pre and Post)

```{r}
# 
# #| label: setup-parallel-3
# 
# # Set up the parallel backend
# num_cores <- 8
# #create the cluster
# my.cluster <- parallel::makeCluster(
#   num_cores,
#   type = "FORK"
#   )
# 
# #register it to be used by %dopar%
# doParallel::registerDoParallel(cl = my.cluster)
```

```{r}
#| label: fit-HMM-trustee-models
#| cache: true

# # Depmix simple Models
# priordat_all <- data_HMM_all |>
#   filter(roundNum==1)
# 
# tr_fdmod_all <- list()
# 
# max_ns <- 8
# 
# 
# set.seed(20240513)
# my_seeds <- sample(1000:50000, max_ns)
# 
# 
# # Fitting HMM models to data
# tr_fdmod_all <- foreach(state = 2:max_ns, .packages = c("depmixS4"), .combine = "c") %dopar% {
#   # Define depmix model with simple transition only depending on next_inv
#   print(state)
#   tr_simple <- depmix(returns ~ 1, data=data_HMM_all, nstates=state, transition=~ next_investment*full_ctrst, prior=~ investment, initdata=priordat_all, family=multinomial("mlogit"), ntimes=rep(15,206*2))
# 
#   mus <- (1:state)*(1/(state+1))
#   sigmas <- rep(1/(state+1), state)
#   if(state > 8) sigmas <- sigmas + .1
#   rModels_tr <- list()
#   for(s in 1:state) {
#     rModels_tr[[s]] <- list(vtdgaus(y=data_HMM_all$ret_pct_0, yield = 3*data_HMM_all$investment, pstart=c(mus[s],sigmas[s])))
#   }
# 
#   tr_dmod <- makeDepmix(response=rModels_tr, transition=tr_simple@transition, prior=tr_simple@prior, ntimes = rep(15,206*2), homogeneous=FALSE)
# 
#   tmp1 <- fit(tr_dmod, emcontrol=em.control(random.start = FALSE))
#   set.seed(my_seeds[state])
#   tmp2 <- multistart(tr_dmod, nstart=20)
#   if(!is.na(logLik(tmp2)) && logLik(tmp2) > logLik(tmp1)) {tmp1 <- tmp2}
#   tmp1
# }
# 
# # Stop the parallel backend
# stopImplicitCluster()
# 
# 
# save(tr_fdmod_all, file="data/tr_fdmod_all.RData", compression_level = 9)

```

```{r}
#|  label: HMM-trustee-model-comparison
#|  include: false

load("data/tr_fdmod_all.RData")

df_bics_all <- data.frame(nstate = 2:max_ns,
                          logLik = unlist(lapply(tr_fdmod_all, logLik)),
                          num_params= unlist(lapply(tr_fdmod_all, function(x) attr(logLik(x), which="df"))),
                          AIC = unlist(lapply(tr_fdmod_all, AIC)),
                          BIC = unlist(lapply(tr_fdmod_all, BIC)))
df_bics_all

# Here we choose BIC as best m
nstate_tr <- which.min(df_bics_all[,5]) + 1
nstate_tr

best_mod_all <- order_mod_vtdgaus(tr_fdmod_all[[nstate_tr-1]])
summary(best_mod_all)
```

```{r}
#| label: tbl-BICs-tr
#| warning: false
#| echo: false
#| tbl-cap: Model comparison of HMMs of trustee actions with different numbers of latent states.

df_bics_all %>%
  mutate_if(is.numeric, format, digits=4, nsmall = 0, big.mark = ",") %>%
knitr::kable(booktabs = TRUE,
  col.names = c("#states", "log(lik)", "# pars", "AIC", "BIC"))
```

<!-- Function to get transiton probabilities from fitted HMM models -->
```{r}
get_trans_probs <- function(fmod, numCovIn, numCovTr) {
  # --- Parameters & Setup ---
  ns <- nstates(fmod) # Number of states in the HMM
  investment_values <- seq(0, 20, length = 100)
  # Create a data frame to store the transition probabilities.
  # Each row represents a combination of "from" state, "to" state, and "investment" value.
  trans_prob_all <- expand.grid(
    from = paste0("from State ", 1:ns),
    to = paste0("State ", 1:ns),
    investment = investment_values
  )
  trans_prob_all$prob_pre <- 0
  trans_prob_all$prob_ctrl <- 0
  trans_prob_all$prob_expo <- 0

  # --- Calculate Transition Probabilities ---
  for (from in 1:ns) { 
    start_idx <- ns * numCovIn + 1 + (from - 1) * numCovTr * ns
    end_idx <- start_idx + numCovTr * ns - 1
    pars <- matrix(getpars(fmod)[start_idx:end_idx], ncol = numCovTr)
    print(pars)

    # Initialize matrices to store transition probabilities before normalization
    y0 <- matrix(0, nrow = length(investment_values), ncol = ns) # Pre-intervention group
    y1 <- matrix(0, nrow = length(investment_values), ncol = ns) # Control group
    y2 <- matrix(0, nrow = length(investment_values), ncol = ns) # Exposed group

    for (to in 1:ns) {
      y0[, to] <- exp(pars[to, 1] + pars[to, 2] * investment_values)
      y2[, to] <- exp(pars[to, 1] + pars[to, 2] * investment_values + pars[to, 3] + pars[to, 5] * investment_values) # Exposed: Intercept + investment + group effect + interaction
      y1[, to] <- exp(pars[to, 1] + pars[to, 2] * investment_values + pars[to, 4] + pars[to, 6] * investment_values) # Control: Intercept + investment + group effect + interaction
    }

    # Normalize transition probabilities for each group to ensure they sum to 1 across all "to" states
    y0 <- y0 / rowSums(y0)
    y1 <- y1 / rowSums(y1)
    y2 <- y2 / rowSums(y2)

    for (to in 1:ns) {
      idx <- trans_prob_all$from == paste0("from State ", from) & 
             trans_prob_all$to == paste0("State ", to)
      trans_prob_all$prob_pre[idx] <- y0[, to]
      trans_prob_all$prob_ctrl[idx] <- y1[, to]
      trans_prob_all$prob_expo[idx] <- y2[, to]
    }
  }

  saveRDS(trans_prob_all, "data/trans_prob_all.RDS")
  return(trans_prob_all)
}
```

```{r}

# Plotting the trustee HMM policy by state 
tr_policy <- list()
for(i in 1:nstates(best_mod_all)) {
  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(best_mod_all@response[[i]][[1]]@parameters))))  
}

perc_return <- seq(0,1, length=61)

respplot <- data.frame(state = rep(1:nstates(best_mod_all), each=length(perc_return)),
                            perc_return = perc_return,
                            probability = unlist(tr_policy)) |>
  mutate(state = factor(paste0("s",state))) |> 
  ggplot(                 
       aes(x = perc_return,
           y = probability,
           fill = state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  # labs(fill='Latent trustee state') +
  xlab("Percentage return") +
  theme_minimal() + 
  theme(legend.position = "right") + ggtitle("State conditional strategies")

##################################################################################
# Plotting Prior (initial) state probabilities (new)

# Function to calculate softmax probabilities
softmax <- function(x) exp(x) / sum(exp(x))

# Get parameters from model
params <- getpars(getmodel(best_mod_all, "prior"))

# Extract intercepts and investment coefficients, and set initial investment value
intercepts <- params[1:nstate_tr]
investments <- params[(nstate_tr + 1):(2 * nstate_tr)]
# set average investment of trustee in neutral state as value to calculate initial state probabilities in linear model
investment_value <- 9

# Calculate logits and initial probabilities
logits <- intercepts + investments * investment_value
initial_probs <- softmax(logits)

# Create a data frame for plotting
priordf <- data.frame(
  state = factor(paste0("s", 1:nstate_tr)),
  prob = initial_probs
)

# Plot the initial state probabilities
priorplot <- ggplot(priordf, aes(x = state, y = prob, fill = state)) +
  geom_bar(stat = "identity", width = 0.5) +
  xlab("State") +
  ylab("Prior Probability") +
  ggtitle("Initial State Probability for average investment") +
  theme_minimal()

```

```{r,include=F}

trans_prob_all <- get_trans_probs(best_mod_all, numCovIn=2, numCovTr=6)

ctrl_prePost_plot <- ggplot(trans_prob_all,aes(x=investment)) + 
  geom_line(aes(y=prob_ctrl, colour = as.factor(to), linetype="Post-control")) + 
  geom_line(aes(y=prob_pre, colour = as.factor(to), linetype="Pre-control")) +
  facet_wrap(~from) + 
  ylim(c(0,1)) + 
  ggtitle("Transition function pre- and post-control group") + 
  labs(x = "Investment", y = "Transition probability", color='Transition to:') + 
  scale_linetype_manual(values = c("Post-control" = "solid", "Pre-control" = "dotted")) +
  theme_bw()

expo_prePost_plot  <- ggplot(trans_prob_all,aes(x=investment)) + 
  geom_line(aes(y=prob_expo, colour = as.factor(to), linetype="Post-exposure")) + 
  geom_line(aes(y=prob_pre, colour = as.factor(to), linetype="Pre-exposure")) +
  facet_wrap(~from) + 
  ylim(c(0,1)) + 
  ggtitle("Transition function pre- and post-exposure manipulation") + 
  labs(x = "Investment", y = "Transition probability", color='Transition to:', linetype = "Line Type") + 
  scale_linetype_manual(values = c("Post-exposure" = "solid", "Pre-exposure" = "dotted")) +
  theme_bw()

expo_prePost_plot 
ctrl_prePost_plot
```

```{r}
##############################################################################

#| label: fig-trustee-policy-plot
#| echo: false
#| fig.cap: HMM model of the trustees 
#| fig.align: center
#| fig.width: 12
#| fig.height: 6

# transplot <- plot_transitions(best_mod_all,numCovIn=2,numCovTr=2) + labs(x = "Investment", y = "Probability", color='Transition to') + theme_minimal() + theme(legend.position = "bottom")

gridExtra::grid.arrange(respplot, priorplot, expo_prePost_plot, ncol=3, layout_matrix=cbind(c(1,2),c(3,3),c(3,3)))
```

<!-- Get Trustee posterior states from model and add to data table -->

```{r, include = FALSE}
set.seed(20221010)
# mention we are using local decoding (refer to the book). 
predTrStatesAll <- posterior(best_mod_all, type="local")

#############
data_HMM_all$postTrState <-  factor(predTrStatesAll, levels= c(1,2,3,4,5),labels=c("State 1","State 2","State 3","State 4","State 5"))
############
```

```{r}

#| label: fig-trustee-posterior-plot
#| echo: false
#| fig.cap: POsterior state of the trustee in eahc round by Condition and Phase
#| fig.align: center
#| fig.width: 8
#| fig.height: 12
postHMM <- ggplot(data_HMM_all) +
  geom_bar(aes(x = roundNum, group = postTrState, fill = postTrState),  position = position_stack(reverse = TRUE)) + 
  # facet_wrap(~factor(condition.f, levels = rev(levels(as.factor(condition.f))))) +  
  facet_wrap(~condition.f*gameNum.f)+
  labs(x = "Round", fill='Posterior trustee state') +
  theme_bw() +
  theme(legend.position = "bottom")

postHMM
```

<!-- Cluster joint latent states: -->

```{r}
# Load necessary libraries
library(factoextra)
library(hexbin)

# Assuming data_HMM_post is your dataframe

# Rename the levels of investorState and postTrState
data_HMM_all <- data_HMM_all %>%
  mutate(investorState = recode(investorState.f, 
                                "unhappy" = "1", 
                                "neutral" = "2", 
                                "happy" = "3"),
         postTrState = gsub("State ", "", postTrState))

# Define the desired order of joint states
desired_order <- c(
  "1-1", "1-2", "1-3", "1-4", "1-5",
  "2-1", "2-2", "2-3", "2-4", "2-5",
  "3-1", "3-2", "3-3", "3-4", "3-5"
)

# Ensure Joint_State and Next_Joint_State are factors with the specified order
data_HMM_all <- data_HMM_all %>%
  mutate(Joint_State = factor(paste(investorState, postTrState, sep = "-"), 
                              levels = desired_order, ordered = TRUE))

# Plot the frequency of each Joint_State
# Define a color palette based on the desired order
color_palette <- colorRampPalette(c("red","orange", "green", "cyan", "blue"))(length(desired_order))
state_pair_to_color <- setNames(color_palette, desired_order)

# Plot the frequency of each Joint_State with the assigned colors
ggplot(data_HMM_all, aes(x = Joint_State)) +
  geom_bar(aes(fill = Joint_State)) +
  scale_fill_manual(values = state_pair_to_color) +
  labs(title = "Frequency of Joint States",
       x = "Joint State",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Plot joint states distribution by phase and condition 
ggplot(data_HMM_all, aes(x = Joint_State, fill = gameNum.f)) +
  geom_bar(position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("lightblue", "blue"), # Example colors for different conditions
                    name = "Condition") +
  labs(title = "Frequency of Joint States by Condition",
       x = "Joint State",
       y = "Frequency") +
  facet_wrap(~condition.f)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Create next joint state for transition analysis
data_HMM_all <- data_HMM_all %>%
  arrange(playerId, gameNum.f, roundNum) %>%
  group_by(playerId, gameNum.f) %>%
  mutate(Next_Joint_State = lead(Joint_State, order_by = roundNum)) %>%
  ungroup() %>%
  mutate(Next_Joint_State = factor(Next_Joint_State, levels = desired_order, ordered = TRUE))

# Write to CSV if needed (for reference or debugging)
write.csv(data_HMM_all, "data/data_HMM_all.csv")

# Calculate transition counts
transition_counts <- data_HMM_all %>%
  filter(!is.na(Next_Joint_State)) %>%
  count(Joint_State, Next_Joint_State)

# Convert counts to probabilities and reshape to wide format
transition_matrix <- transition_counts %>%
  group_by(Joint_State) %>%
  mutate(transition_prob = n / sum(n)) %>%
  ungroup() %>%
  dplyr::select(Joint_State, Next_Joint_State, transition_prob) %>%
  pivot_wider(names_from = Next_Joint_State, values_from = transition_prob, values_fill = list(transition_prob = 0))

# Ensure the column order in transition_matrix
transition_matrix <- transition_matrix %>%
  dplyr::select(Joint_State, all_of(desired_order))

# Visualize the transition matrix
transition_matrix_long <- transition_matrix %>%
  pivot_longer(cols = -Joint_State, names_to = "Next_Joint_State", values_to = "probability")

ggplot(transition_matrix_long, aes(x = Next_Joint_State , y = Joint_State, fill = probability)) +
  geom_tile() +
  scale_fill_gradient2(low = "white", high = "blue", mid = "lightblue", midpoint = 0.5, limit = c(0, 1), space = "Lab", name="Transition Probability") +
  theme_minimal() +
  labs(title = "Transition Matrix of Joint States", x = "Current Joint State", y = "Next Joint State") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
```{r}
frequency_table <- data_HMM_all %>%
  count(Joint_State, condition.f, gameNum.f) %>%
  spread(gameNum.f, n, fill = 0) %>% 
  mutate(Difference = `post` - `pre`)

# Plot the differences
ggplot(frequency_table, aes(x = Joint_State, y = Difference, fill = condition.f)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("lightblue", "blue"), # Example colors for different conditions
                    name = "Condition") +
  labs(title = "Difference in Frequency of Joint States by Condition (post - pre)",
       x = "Joint State",
       y = "Difference in Frequency") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<!-- Clustering of whole time series using K-means  -->
```{r}

# Reshape the data
df_wide <- data_HMM_all %>%
  dplyr::select(playerId, gameNum.f,investorState, postTrState) %>%
  mutate(across(c(investorState, postTrState), as.numeric)) %>%
  group_by(playerId, gameNum.f) %>%
  mutate(id = row_number()) %>%
  pivot_wider(names_from = id, values_from = c(investorState, postTrState)) %>%
  ungroup()

# Separate playerId and gameNum.f columns for later use
playerIds <- df_wide$playerId
gameNums <- df_wide$gameNum.f

# Convert the rest of the data frame to a matrix
data_matrix <- df_wide %>% dplyr::select(-playerId, -gameNum.f) %>% as.matrix()

# Determine the optimal number of clusters using the elbow method
wss <- vector("numeric", length = 10)

for (k in 1:10) {
  set.seed(123)  # For reproducibility
  kmeans_result <- kmeans(data_matrix, centers = k, nstart = 25)
  wss[k] <- kmeans_result$tot.withinss
}

# Plot the elbow plot
plot(1:10, wss, type = "b", pch = 19, frame = FALSE,
     xlab = "Number of Clusters K",
     ylab = "Total Within-Clusters Sum of Squares",
     main = "Elbow Plot for Determining Optimal K")

```

<!-- # Perform k-means clustering with the optimal number of clusters -->
```{r}

set.seed(123)
final_kmeans <- kmeans(data_matrix, centers = 3, nstart = 25)

# Add cluster assignments to the original data
df_wide$cluster <- factor(final_kmeans$cluster)

# Combine playerId and gameNum.f with cluster assignments
df_wide_with_id <- df_wide %>% 
  mutate(playerId = playerIds, gameNum.f = gameNums)

# Join the cluster assignments back to the original data
df_clustered <- data_HMM_all %>% 
  left_join(df_wide_with_id %>% dplyr::select(playerId, gameNum.f, cluster), 
            by = c("playerId", "gameNum.f"))

df_clustered %>%
  group_by(cluster, Joint_State) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(cluster) %>%
  mutate(proportion = count / sum(count)) %>%
  ggplot(aes(x = factor(cluster), y = proportion, fill = factor(Joint_State))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = state_pair_to_color) +
  theme_minimal() +
  labs(title = "Proportion of Joint States by Cluster",
       x = "Cluster",
       y = "Proportion",
       fill = "Joint State") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

df_clustered <- df_clustered %>%
  mutate(cluster_mood = case_when(
    cluster == 2 ~ "Joint Unhappy",
    cluster == 1 ~ "Joint Happy",
    cluster == 3 ~ "Joint Neutral",
    TRUE ~ "Other"  # This catches any other values, if they exist
  ) %>% factor(levels = c("Joint Unhappy", "Joint Neutral", "Joint Happy")))

# Save the results to a CSV file (optional)
write.csv(df_clustered, "clustered_data.csv", row.names = FALSE)

```

```{r}
# Calculate the mean path of joint states for each cluster over time
df_mean_path <- df_clustered %>%
  group_by(cluster_mood, roundNum) %>%
  summarize(meanInvestorState = mean(as.numeric(investorState), na.rm = TRUE),
            meanPostTrState = mean(as.numeric(postTrState), na.rm = TRUE)) %>%
  ungroup()

# Combine both states in a single plot using facets
df_mean_path_long <- df_mean_path %>%
  pivot_longer(cols = c(meanInvestorState, meanPostTrState), 
               names_to = "StateType", 
               values_to = "MeanState")

 # Define custom colors for each cluster
cluster_colors <- c("Joint Unhappy" = "red", "Joint Neutral" = "darkgreen", "Joint Happy" = "blue")

ggplot(df_mean_path_long, aes(x = roundNum, y = MeanState, color = cluster_mood)) +
  geom_line(size = 1.5) +
  geom_point(size = 2) +
  scale_color_manual(values = cluster_colors) +
  facet_wrap(~ StateType, scales = "free_y") +
  labs(title = "Mean Path of States by Cluster Over Time",
       x = "Time",
       y = "Mean State",
       color = "Cluster") +
  theme_minimal()

```

```{r}
cluster_means <- df_clustered %>%
  group_by(cluster_mood, roundNum) %>%
  summarise(across(c(investment, ret_pct_0), mean))

ggplot(cluster_means, aes(x = roundNum, y = investment, color = factor(cluster_mood))) +
  geom_line() +
  scale_color_manual(values = cluster_colors) +
  labs(x = "Round Number", y = "Average investment by cluster", color = "Cluster")

ggplot(cluster_means, aes(x = roundNum, y = ret_pct_0, color = factor(cluster_mood))) +
  geom_line() +
  scale_color_manual(values = cluster_colors) +
  labs(x = "Round Number", y = "Average % return by cluster", color = "Cluster")

# Load the patchwork library
library(patchwork)

# Assuming your cluster_means dataframe and cluster_colors vector are already defined
# First plot for investment
p1 <- ggplot(cluster_means, aes(x = roundNum, y = investment, color = factor(cluster_mood))) +
  geom_line() +
  scale_color_manual(values = cluster_colors) +
  labs(x = "Round Number", y = "Average investment by cluster", color = "Cluster") +
  theme_minimal()

# Second plot for % return
p2 <- ggplot(cluster_means, aes(x = roundNum, y = ret_pct_0, color = factor(cluster_mood))) +
  geom_line() +
  scale_color_manual(values = cluster_colors) +
  labs(x = "Round Number", y = "Average % return by cluster", color = "Cluster") +
  theme_minimal()

# Combine the two plots
combined_plot <- p1 / p2

# Print the combined plot
print(combined_plot)

```
```{r}

# Calculate cumulative rewards and investments for each game
cumulative_summary <- df_clustered %>%
  group_by(playerId, cluster_mood, gameNum.f) %>%
  summarise(
    cum_investor_reward = sum(returns - investment, na.rm = TRUE),
    cum_trustee_reward = sum(3*investment - returns, na.rm = TRUE),
    .groups = 'drop'
  )  %>%
  group_by(cluster_mood) %>%
  summarise(
    avg_investor_cum = mean(cum_investor_reward),
    avg_trustee_cum = mean(cum_trustee_reward)
  ) %>%
  ungroup()

# Print summary
print(cumulative_summary)

```

<!-- Correlate the Cluster with initial state of the trustee:  -->

```{r}
library(dplyr)
library(nnet)  # for multinomial logistic regression
library(broom)  # for tidy output
library(ggplot2)

# First, let's filter for the initial round and select relevant variables
df_initial <- df_clustered %>%
  filter(roundNum == 1) %>%
  dplyr::select(playerId,gameNum.f, condition.f, cluster_mood, investment, postTrState)

# Create a dataframe with the pre and post cluster moods for each player and condition
pre_post_data <- df_initial %>%
  dplyr::select(playerId, gameNum.f, condition.f, cluster_mood) %>%
  spread(key = gameNum.f, value = cluster_mood) %>%
  filter(!is.na(pre) & !is.na(post))

# Calculate the distribution of people between clusters for pre and post
distribution_pre <- pre_post_data %>%
  group_by(condition.f, pre) %>%
  summarise(count = n()) %>%
  mutate(time = "pre") %>%
  rename(cluster_mood = pre)

distribution_post <- pre_post_data %>%
  group_by(condition.f, post) %>%
  summarise(count = n()) %>%
  mutate(time = "post") %>%
  rename(cluster_mood = post)

# Combine the pre and post data
combined_distribution <- bind_rows(distribution_pre, distribution_post)

# Ensure the time column is a factor with levels in the correct order
combined_distribution$time <- factor(combined_distribution$time, levels = c("pre", "post"))

# Plot the evolution between pre and post
ggplot(combined_distribution, aes(x = time, y = count, group = cluster_mood, color = cluster_mood)) +
  geom_line(aes(linetype = cluster_mood)) +
  geom_point() +
  facet_wrap(~ condition.f, scales = "fixed") +
  theme_minimal() +
  labs(title = "Evolution of Cluster Moods Pre to Post by Condition",
       x = "Time",
       y = "Count",
       color = "Cluster Mood",
       linetype = "Cluster Mood") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
<!-- #Testing whether differences between pre and post cluster frequencies/proportions are statistically significant -->
```{r}
# Create a contingency table
contingency_table <- combined_distribution %>%
  filter(time == "pre") %>%
  inner_join(combined_distribution %>% filter(time == "post"), by = c("condition.f", "cluster_mood")) %>%
  dplyr::select(condition.f, cluster_mood, count.x, count.y) %>%
  pivot_wider(names_from = condition.f, values_from = c(count.x, count.y))

# Rename columns for clarity
colnames(contingency_table) <- c("cluster_mood", "Control_Pre", "Manipulation_Pre", "Control_Post", "Manipulation_Post")

# Print contingency table
print(contingency_table)

# Function to perform proportion test for each class
perform_prop_test <- function(pre_counts, post_counts, total_pre, total_post) {
  counts <- c(pre_counts, post_counts)
  totals <- c(total_pre, total_post)
  test_result <- prop.test(counts, totals)
  return(test_result)
}

# Total counts for Control and Manipulation groups
total_control_pre <- sum(contingency_table$Control_Pre)
total_control_post <- sum(contingency_table$Control_Post)
total_manipulation_pre <- sum(contingency_table$Manipulation_Pre)
total_manipulation_post <- sum(contingency_table$Manipulation_Post)

# Perform tests for Control group
control_results <- lapply(1:nrow(contingency_table), function(i) {
  perform_prop_test(contingency_table$Control_Pre[i], contingency_table$Control_Post[i], total_control_pre, total_control_post)
})

# Perform tests for Manipulation group
manipulation_results <- lapply(1:nrow(contingency_table), function(i) {
  perform_prop_test(contingency_table$Manipulation_Pre[i], contingency_table$Manipulation_Post[i], total_manipulation_pre, total_manipulation_post)
})

# Print results for Control group
cat("Control Group Proportion Tests:\n")
for (i in 1:length(control_results)) {
  cat(paste("\nCluster Mood:", contingency_table$cluster_mood[i]))
  print(control_results[[i]])
}

# Print results for Manipulation group
cat("\n\nManipulation Group Proportion Tests:\n")
for (i in 1:length(manipulation_results)) {
  cat(paste("\nCluster Mood:", contingency_table$cluster_mood[i]))
  print(manipulation_results[[i]])
}

```

```{r}
# Ensure cluster_mood is a factor
df_initial$cluster_mood <- factor(df_initial$cluster_mood)

# Fit the multinomial logistic regression model
model <- multinom(cluster_mood ~ investment + postTrState, data = df_initial)

# Summary of the model
summary(model)

# Get tidy output of the model
tidy_model <- tidy(model)
print(tidy_model)

# Visualize the effects
ggplot(df_initial, aes(x = investment, fill = cluster_mood)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~postTrState) +
  labs(title = "Distribution of Initial Investment by Cluster Mood and Trustee State",
       x = "Initial Investment Percentage",
       y = "Density") +
  theme_minimal()
```

### Key takeaways:

Higher initial investments slightly decrease the likelihood of being in Joint Neutral or Joint Happy clusters compared to Joint Unhappy.
Higher trustee states, particularly 3 and above, strongly increase the odds of being in Joint Neutral or Joint Happy clusters compared to Joint Unhappy.
The effect of trustee state is more pronounced for the Joint Happy cluster than for Joint Neutral.

These results suggest that the trustee's initial state is a stronger predictor of cluster membership than the initial investment amount, with higher trustee states associated with more positive joint outcomes.

<!-- Check associations with RS -->
```{r}
#########################
# Perform ANOVA to test for significant differences in RS_score between clusters
df_aov_RS <-  df_clustered %>% dplyr::select(playerId,gameNum.f, RS_score,cluster_mood) %>% unique()

# Calculate the mean and standard error of RS_score for each cluster
mean_rs_score <- df_clustered %>%
  group_by(cluster_mood,condition.f) %>%
  summarize(mean_RS_score = mean(RS_score, na.rm = TRUE),
            se_RS_score = sd(RS_score, na.rm = TRUE) / sqrt(n()))

# Print the mean RS score for each cluster
print(mean_rs_score)

anova_result_rs <- aov(RS_score ~ cluster_mood*gameNum.f, data=df_aov_RS)

# Print the summary of the ANOVA result
summary(anova_result_rs)

# Plot the mean RS score for each cluster with error bars using geom_point and geom_errorbar
ggplot(mean_rs_score, aes(x = cluster_mood, y = mean_RS_score, color = cluster_mood)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean_RS_score - se_RS_score, ymax = mean_RS_score + se_RS_score),
                width = 0.2, size = 1.5) +
  facet_wrap(~condition.f)+
  scale_color_manual(values = cluster_colors) +
  labs(title = "Average RS Score for Each Cluster",
       x = "Cluster",
       y = "Average RS Score",
       color = "Cluster") +
  theme_minimal()

```

<!-- Check association WITH LPFS -->
```{r}
# Perform ANOVA to test for significant differences in LPFS_score between clusters
df_aov_lpfs <-  df_clustered %>% dplyr::select(playerId,LPFS_score,cluster_mood) %>% unique()

anova_result_lpfs <- aov(LPFS_score ~ cluster_mood, data = df_aov_lpfs)

# Print the summary of the ANOVA result
summary(anova_result_lpfs)

# Calculate the mean LPFS_score for each cluster
mean_lpfs_score <- df_clustered %>%
  group_by(cluster_mood) %>%
  summarize(mean_LPFS_score = mean(LPFS_score, na.rm = TRUE))

# Print the mean LPFS score for each cluster
print(mean_lpfs_score)

```
Association between RS and variability in joitn states 
```{r}

# Calculate the number of joint state changes for each player
result_vol <- df_clustered %>%
  arrange(playerId, roundNum) %>%
  group_by(playerId) %>%
  summarize(
    RS_score = first(RS_score),
    joint_state_changes = sum(Joint_State != Next_Joint_State, na.rm = TRUE)) %>%
  ungroup()

# Perform correlation analysis
correlation <- cor.test(result_vol$RS_score, result_vol$joint_state_changes)

# Print correlation results
print(correlation)

# Create a scatter plot
ggplot(result_vol, aes(x = RS_score, y = joint_state_changes)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Relationship between RS_score and Joint State Changes",
       x = "RS_score",
       y = "Number of Joint State Changes")
```

```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(forcats)

# Assuming the data is already loaded into a dataframe called 'df'

# Categorize RS_score into high and low groups
df_categorized <- df_clustered %>%
  mutate(RS_group = case_when(
    RS_score > 15 ~ "High RS",
    RS_score <= 10 ~ "Low RS",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(RS_group))

# Calculate the distribution of joint states for each RS group
joint_state_distribution <- df_categorized %>%
  group_by(RS_group, Joint_State) %>%
  summarise(count = n()) %>%
  group_by(RS_group) %>%
  mutate(proportion = count / sum(count))

# Visualization: Stacked bar plot
ggplot(joint_state_distribution, aes(x = RS_group, y = proportion, fill = Joint_State)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Joint States by Rejection Sensitivity Group",
       x = "Rejection Sensitivity Group",
       y = "Proportion",
       fill = "Joint State") +
  theme_minimal()

# Chi-square test of independence
chi_square_result <- chisq.test(table(df_categorized$RS_group, df_categorized$Joint_State))
print(chi_square_result)

# Post-hoc analysis: Standardized residuals
residuals <- chi_square_result$residuals
print("Standardized Residuals:")
print(residuals)

# Visualization: Mosaic plot
library(vcd)
mosaic(~ RS_group + Joint_State, data = df_categorized, 
       main = "Mosaic Plot of Joint States by RS Group")

# Effect size: Cramer's V
library(effectsize)
cramer_v <- cramers_v(df_categorized$RS_group, df_categorized$Joint_State)
print(paste("Cramer's V:", cramer_v))

# Optional: Analyze transitions between joint states
transitions <- df_categorized %>%
  arrange(playerId, roundNum) %>%
  group_by(playerId) %>%
  mutate(next_state = lead(Joint_State)) %>%
  filter(!is.na(next_state)) %>%
  group_by(RS_group, Joint_State, next_state) %>%
  summarise(transition_count = n()) %>%
  ungroup()

# Visualization: Transition heatmap
ggplot(transitions, aes(x = Joint_State, y = next_state, fill = transition_count)) +
  geom_tile() +
  facet_wrap(~ RS_group) +
  scale_fill_viridis_c() +
  labs(title = "Joint State Transitions by RS Group",
       x = "Current State", y = "Next State", fill = "Transition Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Fitting HMM to Participants data in the post Phase with manipulation vs Control factor in transition function:

```{r}

#| label: setup-parallel-4

# # Set up the parallel backend
# num_cores <- 8
# #create the cluster
# my.cluster <- parallel::makeCluster(
#   num_cores,
#   type = "FORK"
#   )
# 
# #register it to be used by %dopar%
# doParallel::registerDoParallel(cl = my.cluster)
```

```{r}
#| label: fit-HMM-trustee-models
#| cache: true

# # Depmix simple Models 
# priordat <- data_HMM_post |>
#   filter(roundNum==1)
# 
# tr_fdmod_cond <- list()
# 
# max_ns <- 8
# 
# 
# set.seed(20240513)
# my_seeds <- sample(1000:50000, max_ns)
# 
# 
# # Fitting HMM models to data 
# tr_fdmod_cond <- foreach(state = 2:max_ns, .packages = c("depmixS4"), .combine = "c") %dopar% {
#   print(state)
#   # Define depmix model with simple transition only depending on next_inv
#   tr_cond <- depmix(returns ~ 1, data=data_HMM_post, nstates=state, transition=~ next_investment*condition.f, prior=~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes=rep(15,207))
#   
#   mus <- (1:state)*(1/(state+1))
#   sigmas <- rep(1/(state+1), state)
#   if(state > 8) sigmas <- sigmas + .1
#   rModels_tr <- list()
#   for(s in 1:state) {
#     rModels_tr[[s]] <- list(vtdgaus(y=data_HMM_post$ret_pct_0, yield = 3*data_HMM_post$investment, pstart=c(mus[s],sigmas[s])))
#   }
#   
#   tr_dmod <- makeDepmix(response=rModels_tr, transition=tr_cond@transition, prior=tr_cond@prior, ntimes = rep(15,207), homogeneous=FALSE)
#   
#   tmp1 <- fit(tr_dmod, emcontrol=em.control(random.start = FALSE))
#   set.seed(my_seeds[state])
#   tmp2 <- multistart(tr_dmod, nstart=20)
#   if(!is.na(logLik(tmp2)) && logLik(tmp2) > logLik(tmp1)) {tmp1 <- tmp2}
#   tmp1
# }
# 
# # Stop the parallel backend
# stopImplicitCluster()
# 
# 
# save(tr_fdmod_cond, file="data/tr_fdmod_cond.RData", compression_level = 9)

```

```{r}
#|  label: HMM-trustee-cond-comparison
#|  include: false

#############################
load("data/tr_fdmod_cond.RData")
############################

max_ns <- length(tr_fdmod_cond) +1 
df_bics_cond <- data.frame(nstate = 2:max_ns,
                          logLik = unlist(lapply(tr_fdmod_cond, logLik)),
                          unlist(lapply(tr_fdmod_cond, function(x) attr(logLik(x), which="df"))),
                          AIC = unlist(lapply(tr_fdmod_cond, AIC)),
                          BIC = unlist(lapply(tr_fdmod_cond, BIC)))
df_bics_cond

# Here we choose BIC as best m
nstate_tr <- which.min(df_bics_cond[,5]) + 1

############################
best_mod_cond <- order_mod_vtdgaus(tr_fdmod_cond[[nstate_tr-1]])
###########################
summary(best_mod_cond)
```

```{r}
#| label: tbl-BICs-tr-cond
#| warning: false
#| echo: false
#| tbl-cap: Model comparison of HMMs of trustee actions with different numbers of latent states.

df_bics_cond %>%
  mutate_if(is.numeric, format, digits=4, nsmall = 0, big.mark = ",") %>%
knitr::kable(booktabs = TRUE,
  col.names = c("#states", "log(lik)", "# par", "AIC", "BIC"))
```

```{r}
#| label: fig-trustee-policy-plot
#| echo: false
#| fig.cap: Probability mass function of the trustee's policy conditional on its latent state as an output of the best fitting XX state HMM using a truncated discretised Gaussian as a response function
#| fig.align: center
#| fig.width: 12
#| fig.height: 6

# Plotting the trustee HMM policy by state 
tr_policy <- list()
for(i in 1:nstates(best_mod_cond)) {
  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(best_mod_cond@response[[i]][[1]]@parameters))))  
}

perc_return <- seq(0,1, length=61)

respplot <- data.frame(state = rep(1:nstates(best_mod_cond), each=length(perc_return)),
                            perc_return = perc_return,
                            probability = unlist(tr_policy)) |>
  mutate(state = factor(paste0("s",state))) |> 
  ggplot(                 
       aes(x = perc_return,
           y = probability,
           fill = state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  # labs(fill='Latent trustee state') +
  xlab("Percentage return") +
  theme_minimal() + 
  theme(legend.position = "right") + ggtitle("State conditional strategies")

# Function to calculate softmax probabilities
softmax <- function(x) exp(x) / sum(exp(x))

# Assuming best_mod_cond is your fitted model
params <- getpars(getmodel(best_mod_cond, "prior"))

# Extract intercepts and investment coefficients, and set initial investment value
intercepts <- params[1:nstate_tr]
investments <- params[(nstate_tr + 1):(2 * nstate_tr)]
investment_value <- 9

# Calculate logits and initial probabilities
logits <- intercepts + investments * investment_value
initial_probs <- softmax(logits)

# Create a data frame for plotting
priordf <- data.frame(
  state = factor(paste0("s", 1:nstate_tr)),
  prob = initial_probs
)

# Plot the initial state probabilities
priorplot <- ggplot(priordf, aes(x = state, y = prob, fill = state)) +
  geom_bar(stat = "identity", width = 0.5) +
  xlab("State") +
  ylab("Prior Probability") +
  ggtitle("Initial State Probability") +
  theme_minimal()

# Plotting Transition function between states
plot_transitions_cond <- function(fmod, numCovIn, numCovTr) {

  ns <- nstates(fmod)
  conditions <- c(0, 1)  # 0 for control, 1 for manipulation

  trans_prob <- data.frame(
    from = rep(1:ns, each=80*ns*2),
    to = rep(1:ns, each=80*2),
    inv = rep(seq(0, 20, length=80), times=ns*2),
    condition = rep(rep(conditions, each=80), times=ns),
    prob = 0
  )

  y <- matrix(0.0, ncol=ns, nrow=80)

  for (from in 1:ns) {
    pars <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)], ncol=numCovTr)
    print(pars)
    for (cond in conditions) {
      for (to in 1:ns) {
        x <- trans_prob[trans_prob$from == from & trans_prob$to == to & trans_prob$condition == cond, "inv"]
        y[, to] <- exp(pars[to, 1] + pars[to, 2]*x + pars[to, 3]*cond + pars[to, 4]*x*cond)
      }
      y <- y / rowSums(y)
      for (to in 1:ns) {
        trans_prob$prob[trans_prob$from == from & trans_prob$to == to & trans_prob$condition == cond] <- y[, to]
      }
    }
  }

  trans_prob <- trans_prob %>%
    mutate(from = paste0("s", from),
           to = paste0("s", to))

  ggplot(trans_prob, aes(x=inv, y=prob, colour = as.factor(to), linetype = as.factor(condition))) + 
    geom_line() + 
    facet_wrap(~from) + 
    ylim(c(0, 1)) +
    scale_linetype_manual(values=c("0" = "dashed", "1" = "solid")) +
    labs(linetype = "Condition")
}
transplot <- plot_transitions_cond(best_mod_cond,numCovIn=2,numCovTr=4) + labs(x = "Investment", y = "Probability", color='Transition to') + theme_minimal() + theme(legend.position = "bottom")

gridExtra::grid.arrange(respplot, priorplot, transplot, ncol=3, layout_matrix=cbind(c(1,2),c(3,3),c(3,3)))

```

<!-------------------------       LATENT STATE POSTERIOR PLOTS BY CONDITIONS        ------------------------------>

```{r, include = FALSE}
set.seed(20221010)
# Get investor posterior states from model and add to data table

# mention we are using local decoding (refer to the book). 
predTrStates <- posterior(best_mod_cond, type="local")

#############
data_HMM_post$postTrState <-  factor(predTrStates, levels= c(1,2,3,4,5),labels=c("State 1","State 2","State 3","State 4","State 5"))
############
```

```{r}
postHMM <- ggplot(data_HMM_post %>% dplyr::filter(gameNum.f=="post")) +
  geom_bar(aes(x = roundNum, group = postTrState, fill = postTrState),  position = position_stack(reverse = TRUE)) + 
  facet_wrap(~factor(condition.f, levels = rev(levels(as.factor(condition.f))))) +  
  labs(x = "Round", fill='Posterior trustee state') +
  theme_bw() +
  theme(legend.position = "bottom")

postHMM
```

<!-------------------------       CLUSTERING PAIRS OF LATENT STATES      ------------------------------>

```{r}
# Load required libraries
library(reshape2)
library(ggplot2)
library(dplyr)

# Example structure of data_hmm_post:
# playerId | roundNum | PosttrState | investorState
# 1        | 1        | 2           | 1
# 1        | 2        | 3           | 2
# ...

# Create a combined state pair column
data_HMM_new <- data_HMM_post %>%
  distinct(playerId, roundNum, .keep_all = TRUE) %>%
  mutate(StatePair = paste(investorState.f, postTrState, sep = "-"))

# Define the order of state pairs from most uncooperative to most cooperative
state_order <- c("unhappy-State 1", "unhappy-State 2", "unhappy-State 3", "unhappy-State 4", "unhappy-State 5",
                 "neutral-State 1", "neutral-State 2", "neutral-State 3", "neutral-State 4", "neutral-State 5",
                 "happy-State 1", "happy-State 2", "happy-State 3", "happy-State 4", "happy-State 5")

# Convert StatePair to a factor with the defined order
data_HMM_new$StatePair <- factor(data_HMM_new$StatePair, levels = state_order, ordered = TRUE)

# Convert roundNum to a numeric type and then to factor with levels in the correct order
data_HMM_new$roundNum <- as.numeric(as.character(data_HMM_new$roundNum))
data_HMM_new$roundNum <- factor(data_HMM_new$roundNum)

# Pivot the data to have players as rows and rounds as columns
heatmap_data <- data_HMM_new %>%
  dplyr::select(playerId, roundNum, StatePair) %>%
  pivot_wider(names_from = roundNum, values_from = StatePair)

# Convert the data to long format for ggplot2
heatmap_data_long <- heatmap_data %>%
  pivot_longer(cols = -playerId, names_to = "roundNum", values_to = "StatePair")

# Ensure roundNum, playerId, and StatePair are treated as characters and then as factors
heatmap_data_long$roundNum <- as.character(heatmap_data_long$roundNum)
heatmap_data_long$playerId <- as.character(heatmap_data_long$playerId)
heatmap_data_long$StatePair <- as.character(heatmap_data_long$StatePair)

heatmap_data_long$roundNum <- as.factor(heatmap_data_long$roundNum)
heatmap_data_long$playerId <- as.factor(heatmap_data_long$playerId)
heatmap_data_long$StatePair <- as.factor(heatmap_data_long$StatePair)

#Get unique state pairs and assign a color to each
unique_state_pairs <- unique(heatmap_data_long$StatePair)
color_palette <- scales::hue_pal()(length(unique_state_pairs))
state_pair_to_color <- setNames(color_palette, unique_state_pairs)

# Assign colors: cold for uncooperative and warm for cooperative
color_palette <- colorRampPalette(c("blue", "cyan", "yellow", "orange", "red"))(length(state_order))
state_pair_to_color <- setNames(color_palette, state_order)

# Plot the heatmap
ggplot(heatmap_data_long, aes(x = factor(roundNum, levels = unique(roundNum)), y = playerId)) +
  geom_tile(aes(fill = factor(StatePair, levels=unique(StatePair))), color = "white") +
  scale_fill_manual(values = state_pair_to_color, 
                    breaks = unique_state_pairs,
                    labels = unique_state_pairs) +
  theme_minimal() +
  labs(title = "Sequence of Latent State Pairs (Investor-Trustee) for Each Participant",
       x = "Round",
       y = "Participant ID",
       fill = "State Pair") +
  theme(axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())

# Visual of frequencies of joint states by phase and condition .

```

 <!-- # 3D representation of joint latent space -->

```{r}

# Load required libraries
library(plotly)
library(dplyr)

# Define the order of investor states
investor_state_order <- c("unhappy", "neutral", "happy")

# Function to plot evolution of a specific player in 3D
plot_player_evolution_3d <- function(player_index) {
  player_data <- data_HMM_new %>% filter(playerId == player_index)

  # Ensure postTrState is converted to numeric for y-axis
  player_data <- player_data %>%
    mutate(
      TrusteeState = as.numeric(gsub("State ", "", postTrState)),
      InvestorStateNum = as.numeric(factor(investorState.f, levels = investor_state_order))
    )

  # Create 3D plot
  p <- plot_ly(player_data, x = ~InvestorStateNum, y = ~TrusteeState, z = ~roundNum, type = 'scatter3d', mode = 'lines+markers',
               marker = list(size = 5, color = ~roundNum, colorscale = 'Viridis', showscale = TRUE),
               line = list(width = 4, color = ~roundNum, colorscale = 'Viridis', showscale = FALSE)) %>%
    add_trace(
      x = ~InvestorStateNum, y = ~TrusteeState, z = ~roundNum, 
      marker = list(size = 8, symbol = I("circle")),
      text = ~paste("Round:", roundNum),
      hoverinfo = 'text'
    ) %>%
    add_trace(
      x = player_data$InvestorStateNum[1], y = player_data$TrusteeState[1], z = player_data$roundNum[1], 
      marker = list(size = 10, symbol = I("diamond"), color = "red"),
      text = "Start",
      hoverinfo = 'text'
    ) %>%
    add_trace(
      x = player_data$InvestorStateNum[nrow(player_data)], y = player_data$TrusteeState[nrow(player_data)], z = player_data$roundNum[nrow(player_data)], 
      marker = list(size = 10, symbol = I("cross"), color = "green"),
      text = "End",
      hoverinfo = 'text'
    ) %>%
    layout(
      title = paste("Evolution of State Pairs Over Time for Player", player_index),
      scene = list(
        xaxis = list(title = 'Investor State', tickvals = 1:3, ticktext = investor_state_order),
        yaxis = list(title = 'Trustee State', tickvals = 1:5),
        zaxis = list(title = 'Round')
      )
    )

  # Print the plot
  print(p)
}

# Example usage
plot_player_evolution_3d("YEpcoSejRCLubbkPp")
plot_player_evolution_3d("EZYk8rfDNxcuwSpEr")
plot_player_evolution_3d("6SYmH8njCALTjKyp2")

```

<!-- Cluster joint latent states: -->

```{r}
# Load necessary libraries
library(factoextra)
library(hexbin)

# Assuming data_HMM_post is your dataframe

# Rename the levels of investorState and postTrState
data_HMM_clust <- data_HMM_post %>%
  distinct(playerId, roundNum, .keep_all = TRUE) %>%
  mutate(investorState = recode(investorState.f, 
                                "unhappy" = "1", 
                                "neutral" = "2", 
                                "happy" = "3"),
         postTrState = gsub("State ", "", postTrState))

# Define the desired order of joint states
desired_order <- c(
  "1-1", "1-2", "1-3", "1-4", "1-5",
  "2-1", "2-2", "2-3", "2-4", "2-5",
  "3-1", "3-2", "3-3", "3-4", "3-5"
)

# Ensure Joint_State and Next_Joint_State are factors with the specified order
data_HMM_clust <- data_HMM_clust %>%
  mutate(Joint_State = factor(paste(investorState, postTrState, sep = "-"), 
                              levels = desired_order, ordered = TRUE))

# Plot the frequency of each Joint_State
# Define a color palette based on the desired order
color_palette <- colorRampPalette(c("blue", "cyan", "yellow", "orange", "red"))(length(desired_order))
state_pair_to_color <- setNames(color_palette, desired_order)

# Plot the frequency of each Joint_State with the assigned colors
ggplot(data_HMM_clust, aes(x = Joint_State)) +
  geom_bar(aes(fill = Joint_State)) +
  scale_fill_manual(values = state_pair_to_color) +
  labs(title = "Frequency of Joint States",
       x = "Joint State",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data_HMM_clust, aes(x = Joint_State, fill = condition.f)) +
  geom_bar(position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("grey40", "grey80"), # Example colors for different conditions
                    name = "Condition") +
  labs(title = "Frequency of Joint States by Condition",
       x = "Joint State",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create next joint state for transition analysis
data_HMM_clust <- data_HMM_clust %>%
  arrange(playerId, roundNum) %>%
  group_by(playerId) %>%
  mutate(Next_Joint_State = lead(Joint_State, order_by = roundNum)) %>%
  ungroup() %>%
  mutate(Next_Joint_State = factor(Next_Joint_State, levels = desired_order, ordered = TRUE))

# Write to CSV if needed (for reference or debugging)
write.csv(data_HMM_clust, "data/data_HMM_clust.csv")

# Calculate transition counts
transition_counts <- data_HMM_clust %>%
  filter(!is.na(Next_Joint_State)) %>%
  count(Joint_State, Next_Joint_State)

# Convert counts to probabilities and reshape to wide format
transition_matrix <- transition_counts %>%
  group_by(Joint_State) %>%
  mutate(transition_prob = n / sum(n)) %>%
  ungroup() %>%
  select(Joint_State, Next_Joint_State, transition_prob) %>%
  pivot_wider(names_from = Next_Joint_State, values_from = transition_prob, values_fill = list(transition_prob = 0))

# Ensure the column order in transition_matrix
transition_matrix <- transition_matrix %>%
  select(Joint_State, all_of(desired_order))

# Visualize the transition matrix
transition_matrix_long <- transition_matrix %>%
  pivot_longer(cols = -Joint_State, names_to = "Next_Joint_State", values_to = "probability")

ggplot(transition_matrix_long, aes(x = Next_Joint_State , y = Joint_State, fill = probability)) +
  geom_tile() +
  scale_fill_gradient2(low = "white", high = "blue", mid = "lightblue", midpoint = 0.5, limit = c(0, 1), space = "Lab", name="Transition Probability") +
  theme_minimal() +
  labs(title = "Transition Matrix of Joint States", x = "Current Joint State", y = "Next Joint State") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
```{r}
library(dplyr)
library(ggplot2)

# Function to compute the stationary distribution
stationary_distribution <- function(transition_matrix) {
  eigen_result <- eigen(t(transition_matrix))
  # Find the eigenvector associated with the eigenvalue 1
  stationary_vector <- Re(eigen_result$vectors[, which.max(Re(eigen_result$values))])
  # Normalize the vector to sum to 1
  stationary_vector <- stationary_vector / sum(stationary_vector)
  return(stationary_vector)
}

# Compute the stationary distribution
transition_matrix_numeric <- transition_matrix %>%
  dplyr::select(-Joint_State) %>%  # Remove the 'Joint_State' column
  as.matrix()

# Print the resulting transition matrix
print(transition_matrix_numeric)
stationary_dist <- stationary_distribution(transition_matrix_numeric)

print("Stationary Distribution:")
print(stationary_dist)

# Create a data frame for plotting with initial names
plot_data <- data.frame(
  State = desired_order,
  Probability = stationary_dist
)

# Plot the stationary distribution
ggplot(plot_data, aes(x = factor(State, levels = desired_order), y = Probability)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Stationary Distribution", x = "State", y = "Probability") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

<!-- Visual of times series of joint states by initial trustee latent state (investor always starts neutral) -->
```{r}
# # Extract the trustee state at the first trial for each player
# first_trial_Trstate <- data_HMM_clust %>%
#   filter(roundNum == 1) %>%
#   dplyr::select(playerId, firstPostTrState = postTrState)
# 
# # Merge with the main data to add the first trial state as a column
# data_HMM_clust <- data_HMM_clust %>%
#   left_join(first_trial_Trstate, by = "playerId")
# 
# # Visualize the joint states over rounds with color based on the first trial trustee state
# ggplot(data_HMM_clust, aes(x = roundNum, y = Joint_State, group=playerId, color = as.factor(firstPostTrState))) +
#   geom_line(size = 1) +
#   geom_point(size = 2) +
#   facet_wrap(~ as.factor(firstPostTrState), scales = "free_y") +
#   labs(title = "Joint States Over Rounds", x = "Round", y = "Joint State", color = "First Trial Trustee State") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

<!-- Clustering of whole time series using K-means  -->
```{r}

# Reshape the data
df_wide <- data_HMM_clust %>%
  dplyr::select(playerId, investorState, postTrState) %>%
  mutate(across(c(investorState, postTrState), as.numeric)) %>%
  group_by(playerId) %>%
  mutate(id = row_number()) %>%
  pivot_wider(names_from = id, values_from = c(investorState, postTrState)) %>%
  ungroup()

# Separate playerId column for later use
playerIds <- df_wide$playerId

# Convert the rest of the data frame to a matrix
data_matrix <- df_wide %>% select(-playerId) %>% as.matrix()

# Determine the optimal number of clusters using the elbow method
wss <- vector("numeric", length = 10)

for (k in 1:10) {
  set.seed(123)  # For reproducibility
  kmeans_result <- kmeans(data_matrix, centers = k, nstart = 25)
  wss[k] <- kmeans_result$tot.withinss
}

# Plot the elbow plot
plot(1:10, wss, type = "b", pch = 19, frame = FALSE,
     xlab = "Number of Clusters K",
     ylab = "Total Within-Clusters Sum of Squares",
     main = "Elbow Plot for Determining Optimal K")

# Perform k-means clustering with the optimal number of clusters
set.seed(123)
final_kmeans <- kmeans(data_matrix, centers = 3, nstart = 25)

# Add cluster assignments to the original data
df_wide$cluster <- factor(final_kmeans$cluster)

# Combine playerId with cluster assignments
df_wide_with_id <- df_wide %>% 
  mutate(playerId = playerIds)

# Join the cluster assignments back to the original data
df_clustered <- data_HMM_clust %>% 
  left_join(df_wide_with_id %>% select(playerId, cluster), by = "playerId") %>% 
  mutate(cluster = recode(cluster, `1` = "2", `2` = "1"))  %>% # renaming clusters so that they are in increasing order of cooperation (1 least cooeprative and 3 most) 
  mutate(cluster = factor(cluster, levels = c("1", "2", "3")))

# Save the results to a CSV file (optional)
# write.csv(df_clustered, "clustered_data.csv", row.names = FALSE)

# Display the clustering results
print(df_clustered)

# Frenquencies of clusters 
 df_clustered %>%
  group_by(cluster, condition.f) %>%
  summarize(frequency = n()/15) %>%
  ungroup()

```

```{r}
# Calculate the mean path of joint states for each cluster over time
df_mean_path <- df_clustered %>%
  group_by(cluster, roundNum) %>%
  summarize(meanInvestorState = mean(as.numeric(investorState), na.rm = TRUE),
            meanPostTrState = mean(as.numeric(postTrState), na.rm = TRUE)) %>%
  ungroup()

# Combine both states in a single plot using facets
df_mean_path_long <- df_mean_path %>%
  pivot_longer(cols = c(meanInvestorState, meanPostTrState), 
               names_to = "StateType", 
               values_to = "MeanState")

 # Define custom colors for each cluster
cluster_colors <- c("1" = "red", "2" = "green", "3" = "blue")

ggplot(df_mean_path_long, aes(x = roundNum, y = MeanState, color = cluster)) +
  geom_line(size = 1.5) +
  geom_point(size = 2) +
  scale_color_manual(values = cluster_colors) +
  facet_wrap(~ StateType, scales = "free_y") +
  labs(title = "Mean Path of States by Cluster Over Time",
       x = "Time",
       y = "Mean State",
       color = "Cluster") +
  theme_minimal()

# frequencies of clusters  what is specific about these people 
# whether peop
```

```{r}
cluster_means <- df_clustered %>%
  group_by(cluster, roundNum) %>%
  summarise(across(c(investment, ret_pct_0), mean))

 # Define custom colors for each cluster
cluster_colors <- c("1" = "red", "2" = "green", "3" = "blue")

ggplot(cluster_means, aes(x = roundNum, y = investment, color = factor(cluster))) +
  geom_line() +
  scale_color_manual(values = cluster_colors) +
  labs(x = "Round Number", y = "Average investment by cluster", color = "Cluster")

ggplot(cluster_means, aes(x = roundNum, y = ret_pct_0, color = factor(cluster))) +
  geom_line() +
  scale_color_manual(values = cluster_colors) +
  labs(x = "Round Number", y = "Average % return by cluster", color = "Cluster")
```

<!-- Check association WITH LPFS -->
```{r}
# Perform ANOVA to test for significant differences in LPFS_score between clusters
anova_result_lpfs <- aov(LPFS_score ~ cluster, data = df_clustered)

# Print the summary of the ANOVA result
summary(anova_result_lpfs)

# Calculate the mean LPFS_score for each cluster
mean_lpfs_score <- df_clustered %>%
  group_by(cluster) %>%
  summarize(mean_LPFS_score = mean(LPFS_score, na.rm = TRUE))

# Print the mean LPFS score for each cluster
print(mean_lpfs_score)

```

<!-- Check associations with RS -->
```{r}
#########################
# Perform ANOVA to test for significant differences in RS_score between clusters

anova_result_rs <- aov(RS_score ~ cluster, data = df_clustered %>% dplyr::select(playerId,RS_score,cluster) %>% unique())

# Print the summary of the ANOVA result
summary(anova_result_rs)

# Calculate the mean and standard error of RS_score for each cluster
mean_rs_score <- df_clustered %>%
  group_by(cluster) %>%
  summarize(mean_RS_score = mean(RS_score, na.rm = TRUE),
            se_RS_score = sd(RS_score, na.rm = TRUE) / sqrt(n()))

# Print the mean RS score for each cluster
print(mean_rs_score)

# Define custom colors for each cluster
cluster_colors <- c("1" = "red", "2" = "green", "3" = "blue")

# Plot the mean RS score for each cluster with error bars using geom_point and geom_errorbar
ggplot(mean_rs_score, aes(x = cluster, y = mean_RS_score, color = cluster)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean_RS_score - se_RS_score, ymax = mean_RS_score + se_RS_score),
                width = 0.2, size = 1.5) +
  scale_color_manual(values = cluster_colors) +
  labs(title = "Average RS Score for Each Cluster",
       x = "Cluster",
       y = "Average RS Score",
       color = "Cluster") +
  theme_minimal()

```
<!-- Check associations with firs trustee state and first trustee return %  -->

```{r}

```

<!-- ######## OLD WORK  -->
```{r}
# Load necessary libraries
# library(DirichletReg)
# 
# 
# # Assuming merged_data is your dataframe
# 
# # Prepare the data for Dirichlet regression
# # Create a matrix of the cluster probabilities
# probability_matrix <- as.matrix(merged_data %>%
#   select(starts_with("Cluster_")))
# 
# # Add a small constant to avoid zeros, which can cause issues in Dirichlet regression
# probability_matrix <- probability_matrix + 1e-6
# probability_matrix <- DR_data(probability_matrix)
# 
# # Fit a Dirichlet regression model
# dirichlet_model <- DirichReg(probability_matrix ~ RS_score + LPFS_score, data = merged_data)
# 
# # Summary of the model
# summary(dirichlet_model)
# 
# # Predict the probabilities for each cluster
# predicted_probabilities <- predict(dirichlet_model, merged_data)
# 
# # Add predicted probabilities to the dataframe
# predicted_probabilities_df <- as.data.frame(predicted_probabilities)
# names(predicted_probabilities_df) <- paste0("Predicted_Cluster_", 1:5)
# 
# merged_data <- cbind(merged_data, predicted_probabilities_df)
# 
# # Evaluate the model by comparing actual and predicted probabilities
# # Calculate the Mean Squared Error (MSE) as an example
# mse <- function(actual, predicted) {
#   mean((actual - predicted)^2)
# }
# 
# mse_values <- sapply(1:5, function(i) {
#   actual_col <- paste0("Cluster_", i)
#   predicted_col <- paste0("Predicted_Cluster_", i)
#   mse(merged_data[[actual_col]], merged_data[[predicted_col]])
# })
# 
# print(mse_values)
# 
# # Visualize the predicted vs actual probabilities for each cluster
# predicted_actual_df <- merged_data %>%
#   select(playerId, starts_with("Cluster_"), starts_with("Predicted_Cluster_")) %>%
#   pivot_longer(cols = -playerId, names_to = "Cluster", values_to = "Probability") %>%
#   mutate(Type = ifelse(grepl("Predicted", Cluster), "Predicted", "Actual"),
#          Cluster = gsub("Predicted_", "", Cluster))
# 
# ggplot(predicted_actual_df, aes(x = Cluster, y = Probability, fill = Type)) +
#   geom_boxplot() +
#   facet_wrap(~ Cluster) +
#   theme_minimal() +
#   labs(title = "Actual vs Predicted Cluster Probabilities",
#        x = "Cluster",
#        y = "Probability",
#        fill = "Type")

```

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)

# Create a transition matrix
transition_counts <- data_HMM_clust %>%
  arrange(playerId, roundNum) %>%
  group_by(playerId) %>%
  mutate(next_cluster = lead(Cluster, order_by = roundNum)) %>%
  filter(!is.na(next_cluster)) %>%  # Remove rows with NA next_cluster
  ungroup() %>%
  count(Cluster, next_cluster) %>%
  spread(key = next_cluster, value = n, fill = 0)

# Normalize the transition counts to get probabilities
transition_matrix <- as.matrix(transition_counts[,-1])
row_sums <- rowSums(transition_matrix)
transition_matrix <- transition_matrix / row_sums

print("Transition Matrix:")
print(transition_matrix)

```

<!----------------------------- Compute stationary distribution of Clusters:  ---------------------->

```{r}
# Function to compute the stationary distribution of a transition matrix
stationary_distribution <- function(transition_matrix) {
  eigen_result <- eigen(t(transition_matrix))
  # Find the eigenvector associated with the eigenvalue 1
  stationary_vector <- Re(eigen_result$vectors[, which.max(Re(eigen_result$values))])
  # Normalize the vector to sum to 1
  stationary_vector <- stationary_vector / sum(stationary_vector)
  return(stationary_vector)
}

# Compute the stationary distribution
stationary_dist <- stationary_distribution(transition_matrix)

print("Stationary Distribution:")
print(stationary_dist)

```

<!------------------- ########## TRYING NEW CLUSTERING WITH FEATURES OF THE DYAD ################ ------------------->

```{r}
# Ensure joint states are numeric for clustering
data_HMM_clust <- data_HMM_clust %>%
  mutate(Joint_State_numeric = as.numeric(factor(Joint_State)))

# Extract features
features <- data_HMM_clust %>%
  group_by(playerId) %>%
  mutate(
    cumulative_invPayoff = cumsum(returns - investment+20),
    cumulative_trPayoff = cumsum(3*investment -returns)) %>%
  summarize(
    avg_joint_state_duration = mean(rle(Joint_State_numeric)$lengths),
    avg_investment = mean(investment),
    avg_return = mean(returns)
    # avg_invPayoff = mean(cumulative_invPayoff),
    # avg_trPayoff = mean(cumulative_trPayoff) 
  ) %>%
  ungroup()

# Prepare the feature matrix for clustering
feature_matrix <- features %>%
  dplyr::select(-playerId) %>%
  scale()

# Determine the optimal number of clusters using the elbow method
fviz_nbclust(feature_matrix, kmeans, method = "wss")

# Perform k-means clustering
k <- 3  # Choose the number of clusters based on the elbow plot
kmeans_result <- kmeans(feature_matrix, centers = k, nstart = 25)

# Add cluster assignments to the features dataframe
features$cluster <- kmeans_result$cluster

# Visualize the clusters
fviz_cluster(kmeans_result, data = feature_matrix)

pca_result <- prcomp(feature_matrix, scale. = TRUE)
print(pca_result)
summary(pca_result)

# Load required library (if not already loaded)
library(ggbiplot)

# Assuming your PCA result is in pca_result
g <- ggbiplot(pca_result, 
              obs.scale = 1, 
              var.scale = 1, 
              ellipse = TRUE,
              circle = TRUE,
              varname.size = 4
              )
g <- g + scale_color_discrete(name = '')
g <- g + theme(legend.direction = 'horizontal', 
               legend.position = 'top')
g <- g + theme_minimal()
print(g)

```
```{r}
# Interpret the clusters
cluster_summary <- features %>%
  group_by(cluster) %>%
  summarize_all(mean)

print(cluster_summary)

```

<!-- NEW APPROACH  -->

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(cluster)
library(ggplot2)
library(factoextra)

# Example data frame structure (assuming you have the actual data in a similar structure)
data <- data_HMM_clust %>% dplyr::select(playerId, roundNum, investorState,postTrState)

# Number of categories in each time series
n_cat1 <- 3
n_cat2 <- 5

# Function to calculate frequency distribution
frequency_distribution <- function(series, n_cat) {
  table(factor(series, levels = 1:n_cat)) / length(series)
}

# Function to calculate transition matrix
transition_matrix <- function(series, n_cat) {
  series <- factor(series, levels = 1:n_cat)
  trans_mat <- matrix(0, nrow = n_cat, ncol = n_cat)
  for (i in 1:(length(series) - 1)) {
    trans_mat[as.numeric(series[i]), as.numeric(series[i + 1])] <- trans_mat[as.numeric(series[i]), as.numeric(series[i + 1])] + 1
  }
  trans_mat <- trans_mat / rowSums(trans_mat)
  trans_mat[is.nan(trans_mat)] <- 0  # Replace NaN with 0
  return(trans_mat)
}

# Function to calculate co-occurrence matrix
co_occurrence_matrix <- function(series1, series2, n_cat1, n_cat2) {
  series1 <- factor(series1, levels = 1:n_cat1)
  series2 <- factor(series2, levels = 1:n_cat2)
  co_mat <- matrix(0, nrow = n_cat1, ncol = n_cat2)
  for (i in 1:length(series1)) {
    co_mat[as.numeric(series1[i]), as.numeric(series2[i])] <- co_mat[as.numeric(series1[i]), as.numeric(series2[i])] + 1
  }
  co_mat <- co_mat / sum(co_mat)
  print(co_mat)
  return(co_mat)
}

# Extract unique player IDs
player_ids <- unique(data$playerId)

# Initialize an empty list to store feature vectors
feature_list <- list()

# Loop over each playerId to extract features for each dyad
for (player_id in player_ids) {
  dyad_data <- data %>% filter(playerId == player_id)

  time_series1 <- dyad_data$investorState 
  time_series2 <- dyad_data$postTrState

  # Calculate frequency distributions
  freq_dist1 <- frequency_distribution(time_series1, n_cat1)
  freq_dist2 <- frequency_distribution(time_series2, n_cat2)

  # Calculate transition matrices
  trans_matrix1 <- as.vector(transition_matrix(time_series1, n_cat1))
  trans_matrix2 <- as.vector(transition_matrix(time_series2, n_cat2))

  # Calculate co-occurrence matrix
  co_matrix <- as.vector(co_occurrence_matrix(time_series1, time_series2, n_cat1, n_cat2))

  # Construct the feature vector for the current dyad
  features <- c(freq_dist1, freq_dist2, trans_matrix1, trans_matrix2, co_matrix)

  # Append the feature vector to the list
  feature_list[[as.character(player_id)]] <- features
}

# Convert the list of feature vectors into a dataframe
features_df <- do.call(rbind, feature_list)

# Perform PCA for dimensionality reduction
pca_result <- prcomp(features_df, scale. = TRUE)
pca_data <- data.frame(pca_result$x)

# Clustering using K-means (adjust the number of clusters as needed)
set.seed(123)  # For reproducibility
kmeans_result <- kmeans(pca_data, centers = 2)

# Add cluster labels to the PCA data
pca_data$cluster <- as.factor(kmeans_result$cluster)

# Visualize the clusters
ggplot(pca_data, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(size = 3) +
  ggtitle('Clustering of Dyadic Categorical Time Series') +
  xlab('Principal Component 1') +
  ylab('Principal Component 2') +
  theme_minimal()

# Print the clustering results
kmeans_result

```

### 2024-06-26 article.Rmd
trstee HMM fit & joint state analytics

@@
 
-# Find IDs for those who participated in experiment and add RS_score and high_RS as new columsn to d_finished.
+# Find IDs for those who participated in experiment and add RS_score and high_RS as new columns to d_finished.
 # Merging d_finished with relevant columns from RS_data based on matching ids
@@
   inner_join(RS_data %>% dplyr::select(id, RS_score, high_RS), by = "id")
-
-save(d_finished, file = "data/d_finished.RData")
 
@@
 print(temp)
+
+###################### Extract latent Investor State in each round ####################
+
+# Select the desired columns
+d_selected <- d_finished %>%
+  dplyr::select(roundNum, gameNumber, playerId, starts_with("State_"))
+
+# Reshape the data
+d_reshaped <- d_selected %>%
+  pivot_longer(cols = starts_with("State_"), 
+               names_to = "State_name", 
+               values_to = "investorState") %>%
+  mutate(roundNum = as.numeric(str_extract(State_name, "(?<=r)\\d+")),
+         gameNumber = as.numeric(str_extract(State_name, "(?<=game_)\\d+"))) %>%
+  dplyr::select(playerId, gameNumber, roundNum, investorState) %>%
+  arrange(playerId, gameNumber, roundNum) %>%
+  filter(!is.na(investorState)) %>% unique()
+
+# Join the reshaped data back to the original dataset
+d_finished <- d_finished %>%
+  dplyr::select(-investorState) %>% 
+  left_join(d_reshaped, by = c("playerId", "gameNumber", "roundNum"))
 ```
@@
                                       phase.f = factor(ifelse(gameNumber %in% c("1", "2"), 
-                                                       ifelse(gameNumber == "1", "pre", "post"), NA), levels = c("pre", "post"))) %>% 
+                                                       ifelse(gameNumber == "1", "pre", "post"), NA), levels = c("pre", "post")),
+                                      investorState.f = factor(investorState, labels= c("unhappy", "neutral", "happy"), levels =c("unhappy", "neutral", "happy"))) %>% 
                               rename(# Before Exposure
@@
                             )
+
 ```
@@
   )
+save(d_finished, file = "data/d_finished.RData")
 
@@
 avg_ret_df <- d_finished %>%
-  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score) %>% 
+  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score, investorState.f) %>% 
   filter(roundType=="trust",!is.na(roundNum)) %>% 
@@
 ```{r modAllReturns, include=FALSE, cache=TRUE}
-
+##############  Filter only pre and post games
 prepost_df <- avg_ret_df %>% filter(gameNum.f %in% c("pre","post"))
-
+##############
 mod_returns_pct_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
@@
 contrasts(prepost_df$high_RS)
-
-```
-```{r}
-#RS_model <- lm( RS_score ~ ret_pct_na*inv_scaled + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+contrasts(prepost_df$investorState.f)
 ```
@@
 # Compute EMMs for the interaction of interest
-emms <- emmeans(mod_returns_pct_RS, ~ phase.f * condition.f | high_RS, pbkrtest.limit = 6300)
+emms <- emmeans(mod_returns_pct_RS, ~ phase.f * condition.f | high_RS, pbkrtest.limit = 6300, adjust="tukey")
 
@@
 effectsize::eta_squared(mod_returns_pct_RS)
+```
+
+<!---------------- TRYING A MODEL WITH LATENT HMM STATE AS COVARIATE INSTEAD OF INVESTMENT  ------------------->
+
+```{r}
+mod_returns_latent <- mixed( ret_pct_na ~ phase.f*condition.f*investorState.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+
+summary(mod_returns_latent)
+```
+
+```{r, include= F}
+# Load necessary packages
+library(lme4)
+library(emmeans)
+library(ggplot2)
+library(dplyr)
+
+# Assuming `mod_returns_latent` is your fitted model and `prepost_df` is your data frame
+# Calculate estimated marginal means
+emm <- emmeans(mod_returns_latent, ~ high_RS*condition.f*phase.f   | investorState.f, adjust = "tukey")
+
+# Calculate pairwise comparisons
+contrast_results <- contrast(emm, method = "pairwise")
+
+# Extract significant contrasts
+significant_contrasts <- summary(contrast_results, infer = c(TRUE, TRUE))
+
+# Convert emmeans object to data frame for plotting
+emm_df <- as.data.frame(emm)
+
+# Plot the marginal means using ggplot2
+p <- ggplot(emm_df, aes(x = investorState.f, y = emmean, color = high_RS, group = high_RS)) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(width = 0.5)) +
+  geom_line(position = position_dodge(width = 0.5), aes(linetype = high_RS)) +
+  labs(title = "Marginal Means of Returns by Investor State and RS Group",
+       x = "Investor State",
+       y = "Estimated Marginal Mean of Returns") +
+  theme_minimal() +
+  scale_color_manual(name = "RS Group", values = c("low RS" = "blue", "high RS" = "red"))
+
+# Add significance annotations
+for(i in 1:nrow(significant_contrasts)) {
+  row <- significant_contrasts[i, ]
+  if(row$p.value < 0.05) {
+    p <- p + annotate("text", 
+                      label = "*", 
+                      x = which(levels(emm_df$investorState.f) == row$investorState.f),
+                      y = max(emm_df$emmean) + 0.05 * i,
+                      color = "black", 
+                      size = 5, 
+                      vjust = -0.5)
+  }
+}
+
+# Print the plot
+print(p)
+
+```
+```{r}
+# Load necessary packages
+library(lme4)
+library(emmeans)
+library(ggplot2)
+library(dplyr)
+library(ggsignif)
+
+# Assuming `mod_returns_latent` is your fitted model and `prepost_df` is your data frame
+# Calculate estimated marginal means
+emm <- emmeans(mod_returns_latent, ~  high_RS | investorState.f  )
+
+# Calculate pairwise comparisons
+contrast_results <- pairs(emm)
+
+print(contrast_results)
+
+# Convert emmeans object to data frame for plotting
+emm_df <- as.data.frame(emm)
+
+# Convert contrast results to data frame for annotation
+contrast_df <- as.data.frame(contrast_results)
+significant_contrasts <- contrast_df %>%
+  filter(p.value < 0.05) %>%
+  mutate(significance = ifelse(p.value < 0.001, "***", ifelse(p.value < 0.01, "**", "*")))
+
+print(significant_contrasts)
+
+# Plot the marginal means using ggplot2
+p <- ggplot(emm_df, aes(x = investorState.f, y = emmean, color = high_RS, group = high_RS)) +
+  geom_point(position = position_dodge(width = 0.8), size = 3) +
+  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2, position = position_dodge(width = 0.8)) +
+  geom_line(position = position_dodge(width = 0.8), aes(linetype = high_RS)) +
+  labs(title = "Marginal Means of Returns by Investor State and RS Group",
+       x = "Investor State",
+       y = "Estimated Marginal Mean of Returns") +
+  theme_minimal() +
+  scale_color_manual(name = "RS Group", values = c("low RS" = "blue", "high RS" = "red")) +
+  theme(plot.title = element_text(size = 16), legend.position = "bottom")
+
+# Print the plot
+print(p)
+
 ```
@@
                      factor_levels = list(condition.f = c("Control", "Manipulation")),
-                     legend_title = "Phase")
+                     legend_title = "Phase",
+                     emmeans_arg = list(p.adjust.method = "tukey"))
 
@@
 d_long <- d_finished %>%
-  gather(key = "state_col", value = "investorState", starts_with("State_")) %>%
-  mutate(
-    round_extracted = as.numeric(str_extract(state_col, "(?<=State_r)\\d+")),
-    game_extracted = as.numeric(str_extract(state_col, "(?<=game_)\\d+"))
-  ) %>%
-  # Filter to keep rows where extracted round/game numbers match the actual round/game numbers
-  filter(round_extracted == roundNum, game_extracted == gameNumber) %>%
-  # Select and rename the desired columns
-  dplyr::select(id, condition.f, gameNum.f, gameNumber, roundNum, investorState)
-
+  dplyr::select(id, condition.f, gameNum.f, gameNumber, roundNum, investorState.f) %>% 
+  drop_na()
 # View the resulting dataset
@@
 # Filter for only games 1 and 2
-filtered_data <- d_long %>%
+filtered_prepost <- d_long %>%
   filter(gameNum.f %in% c("pre", "post"))
@@
 # Create the histogram with stacked bars and specified colors
-ggplot(filtered_data, aes(x = as.factor(roundNum), fill = as.factor(investorState))) +
-  geom_bar(aes(group = investorState), position = position_stack(reverse = FALSE), stat = "count") +
+ggplot(filtered_prepost, aes(x = as.factor(roundNum), fill = investorState.f)) +
+  geom_bar(aes(group = investorState.f), position = position_stack(reverse = FALSE), stat = "count") +
+  scale_fill_manual(values = state_colors) +
+  facet_wrap(~condition.f*gameNum.f, scales = "free_x", labeller = label_both) +
+  labs(x = "Round", fill = "Investor State", title = "Histogram of Latent HMM Investor States per Game and Round") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+```
+```{r}
+# Filter for expo games only
+filtered_expo <- d_long %>%
+  filter(gameNum.f %in% c("expo1", "expo2", "expo3"))
+
+# Create the histogram with stacked bars and specified colors
+ggplot(filtered_expo, aes(x = as.factor(roundNum), fill = investorState.f)) +
+  geom_bar(aes(group = investorState.f), position = position_stack(reverse = FALSE), stat = "count") +
   scale_fill_manual(values = state_colors) +
@@
   theme(legend.position = "bottom")
-
-```
-
-```{r}
-counts <- filtered_data %>%
-  group_by(condition.f, gameNum.f, investorState,id) %>%
+```
+
+<!------------------------ AVERAGING LATENT STATE COUNT ACROSS TRIALS ----------------------------->
+
+```{r}
+counts <- filtered_prepost %>%
+  group_by(condition.f, gameNum.f, investorState.f,id) %>%
   summarise(count = n()) %>%
@@
 # Calculate the average count per state per condition
-total_rounds <- filtered_data  %>%
+total_rounds <- filtered_prepost  %>%
   group_by(condition.f, gameNum.f) %>%
@@
 
-ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = as.factor(investorState))) +
+ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = investorState.f)) +
   geom_bar(stat = "identity") +
@@
 # Create the histogram with stacked bars and specified colors
-ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = as.factor(investorState))) +
-  geom_bar(aes(group = investorState), position = position_stack(reverse = FALSE), stat = "identity") +
+ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = investorState.f)) +
+  geom_bar(aes(group = investorState.f), position = position_stack(reverse = FALSE), stat = "identity") +
   scale_fill_manual(values = state_colors) +

### 2024-06-26 exploratory.Rmd
trstee HMM fit & joint state analytics

--- 2024-06-26 HMMfit.Rmd
@@
 ---
-title: "HMM fitting to data of participants returns in the Exposure experiment"
+title: "exploratory"
 author: "Ismail Guennouni"
@@
 
+Here we fit an HMM to all trustee data in the exposure experiment, both pre/post and control/manipulation.
+
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, include =F)
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 library(doParallel)
-library(factoextra)
-
-```
-
-## Fitting HMM model to participants data
-
-Here we fit an HMM model to data generated by participants in both RTGs (pre and post manipulation) for both conditions.
-We specify that the transition between latent states depends on the investment received.
-In order to distinguish transitons by condition and Phase, we add a contrast variabel that has three levels (Pre-manipulation, post-control and post exposure).
-We compare models with different number of states and choose the one with the lowest BIC.
+
+```
+
+## Exploratory analyses to model individual games
 
@@
 
-# select columns of interest and also solve issue with the participant with two RS_scores by averaging them. 
 data_HMM <- d_finished %>%
-  dplyr::select(playerId, roundType, investment, returns, roundNum, gameNum.f, condition.f, RS_score, LPFS_score, investorState.f) %>% 
-  filter(roundType == "trust", !is.na(roundNum), gameNum.f %in% c("pre", "post")) %>% 
-  dplyr::select(-roundType) %>%
+  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score) %>% 
+  filter(roundType=="trust",!is.na(roundNum),gameNum.f %in% c("pre","post")) %>% 
   mutate(roundNum = as.numeric(as.character(roundNum))) %>%
-  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment == 0, 0, returns/(3*investment))) %>%
-  group_by(playerId, investment, returns, roundNum, gameNum.f, condition.f, LPFS_score, investorState.f, inv_pct, ret_pct_0) %>%
-  summarize(RS_score = mean(RS_score, na.rm = TRUE)) %>%
-  ungroup()
-
-# Data for both pre and post games
-data_HMM_all <- data_HMM %>% 
-  arrange(playerId, gameNum.f, roundNum) %>%
-  group_by(playerId,gameNum.f) %>% 
-  dplyr::mutate(next_investment = lead(investment, default=0),
-                full_ctrst = factor(ifelse(gameNum.f=="pre", 0,ifelse(condition.f=="Control",1,2)))) %>%
-  ungroup() %>% 
-  unique()
-
-# FOCUS ON POST GAMES ONLY 
-data_HMM_post <- data_HMM %>% 
-  filter(gameNum.f == "post") %>% 
-  group_by(playerId) %>% 
-  dplyr::mutate(next_investment = lead(investment, default=0)) %>%
-  ungroup() %>%
-  unique()
-
-```
-
-<!-- Here we extend the depmixS4 package functionality for analyzing data using hidden Markov models (HMMs). It involves defining new response classes and their associated methods. -->
+  mutate(inv_scaled = as.vector(scale(investment))) %>% 
+  mutate(next_investment = lead(investment, default=0)) %>%
+  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)),ret_pct_na = ifelse(investment==0,NA,returns/(3*investment)))
+```
+
 ```{r}
@@
 
-# Fitting HMM to ALL Participants data (Pre and Post)
-
-```{r}
-# 
-# #| label: setup-parallel-3
-# 
-# # Set up the parallel backend
-# num_cores <- 8
-# #create the cluster
-# my.cluster <- parallel::makeCluster(
-#   num_cores,
-#   type = "FORK"
-#   )
-# 
-# #register it to be used by %dopar%
-# doParallel::registerDoParallel(cl = my.cluster)
+```{r}
+
+#| label: setup-parallel-3
+
+# Set up the parallel backend
+num_cores <- 8
+#create the cluster
+my.cluster <- parallel::makeCluster(
+  num_cores,
+  type = "FORK"
+  )
+
+#register it to be used by %dopar%
+doParallel::registerDoParallel(cl = my.cluster)
 ```
@@
 
-# # Depmix simple Models
-# priordat_all <- data_HMM_all |>
-#   filter(roundNum==1)
-# 
-# tr_fdmod_all <- list()
-# 
-# max_ns <- 8
-# 
-# 
-# set.seed(20240513)
-# my_seeds <- sample(1000:50000, max_ns)
-# 
-# 
-# # Fitting HMM models to data
-# tr_fdmod_all <- foreach(state = 2:max_ns, .packages = c("depmixS4"), .combine = "c") %dopar% {
-#   # Define depmix model with simple transition only depending on next_inv
-#   print(state)
-#   tr_simple <- depmix(returns ~ 1, data=data_HMM_all, nstates=state, transition=~ next_investment*full_ctrst, prior=~ investment, initdata=priordat_all, family=multinomial("mlogit"), ntimes=rep(15,206*2))
-# 
-#   mus <- (1:state)*(1/(state+1))
-#   sigmas <- rep(1/(state+1), state)
-#   if(state > 8) sigmas <- sigmas + .1
-#   rModels_tr <- list()
-#   for(s in 1:state) {
-#     rModels_tr[[s]] <- list(vtdgaus(y=data_HMM_all$ret_pct_0, yield = 3*data_HMM_all$investment, pstart=c(mus[s],sigmas[s])))
-#   }
-# 
-#   tr_dmod <- makeDepmix(response=rModels_tr, transition=tr_simple@transition, prior=tr_simple@prior, ntimes = rep(15,206*2), homogeneous=FALSE)
-# 
-#   tmp1 <- fit(tr_dmod, emcontrol=em.control(random.start = FALSE))
-#   set.seed(my_seeds[state])
-#   tmp2 <- multistart(tr_dmod, nstart=20)
-#   if(!is.na(logLik(tmp2)) && logLik(tmp2) > logLik(tmp1)) {tmp1 <- tmp2}
-#   tmp1
-# }
-# 
-# # Stop the parallel backend
-# stopImplicitCluster()
-# 
-# 
-# save(tr_fdmod_all, file="data/tr_fdmod_all.RData", compression_level = 9)
+# Depmix simple Models 
+priordat <- data_HMM |>
+  filter(roundNum==1)
+
+tr_fdmod <- list()
+
+max_ns <- 10
+
+set.seed(20240513)
+my_seeds <- sample(1000:50000, max_ns)
+
+# Fitting HMM models to data 
+tr_fdmod <- foreach(state = 2:max_ns, .packages = c("depmixS4"), .combine = "c") %dopar% {
+  # Define depmix model with simple transition only depending on next_inv
+  tr_simple <- depmix(returns ~ 1, data=data_HMM, nstates=state, transition=~ next_investment, prior=~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes=rep(15,414))
+
+  mus <- (1:state)*(1/(state+1))
+  sigmas <- rep(1/(state+1), state)
+  if(state > 8) sigmas <- sigmas + .1
+  rModels_tr <- list()
+  for(s in 1:state) {
+    rModels_tr[[s]] <- list(vtdgaus(y=data_HMM$ret_pct_0, yield = 3*data_HMM$investment, pstart=c(mus[s],sigmas[s])))
+  }
+
+  tr_dmod <- makeDepmix(response=rModels_tr, transition=tr_simple@transition, prior=tr_simple@prior, ntimes = rep(15,414), homogeneous=FALSE)
+
+  tmp1 <- fit(tr_dmod, emcontrol=em.control(random.start = FALSE))
+  set.seed(my_seeds[state])
+  tmp2 <- multistart(tr_dmod, nstart=20)
+  if(!is.na(logLik(tmp2)) && logLik(tmp2) > logLik(tmp1)) {tmp1 <- tmp2}
+  tmp1
+}
+
+# Stop the parallel backend
+stopImplicitCluster()
+
+save(tr_fdmod, file="data/tr_fdmod.RData", compression_level = 9)
 
@@
 
-load("data/tr_fdmod_all.RData")
-
-df_bics_all <- data.frame(nstate = 2:max_ns,
-                          logLik = unlist(lapply(tr_fdmod_all, logLik)),
-                          num_params= unlist(lapply(tr_fdmod_all, function(x) attr(logLik(x), which="df"))),
-                          AIC = unlist(lapply(tr_fdmod_all, AIC)),
-                          BIC = unlist(lapply(tr_fdmod_all, BIC)))
-df_bics_all
+df_bics_tr <- data.frame(nstate = 2:max_ns,
+                          logLik = unlist(lapply(tr_fdmod, logLik)),
+                          unlist(lapply(tr_fdmod, function(x) attr(logLik(x), which="df"))),
+                          AIC = unlist(lapply(tr_fdmod, AIC)),
+                          BIC = unlist(lapply(tr_fdmod, BIC)))
+df_bics_tr
 
 # Here we choose BIC as best m
-nstate_tr <- which.min(df_bics_all[,5]) + 1
-nstate_tr
-
-best_mod_all <- order_mod_vtdgaus(tr_fdmod_all[[nstate_tr-1]])
-summary(best_mod_all)
+nstate_tr <- which.min(df_bics_tr[,5]) + 1
+
+best_mod_tr_all <- order_mod_vtdgaus(tr_fdmod[[nstate_tr-1]])
+summary(best_mod_tr_all)
 ```
@@
 
-df_bics_all %>%
+df_bics_tr %>%
   mutate_if(is.numeric, format, digits=4, nsmall = 0, big.mark = ",") %>%
 knitr::kable(booktabs = TRUE,
-  col.names = c("#states", "log(lik)", "# pars", "AIC", "BIC"))
-```
-
-<!-- Function to get transiton probabilities from fitted HMM models -->
-```{r}
-get_trans_probs <- function(fmod, numCovIn, numCovTr) {
-  # --- Parameters & Setup ---
-  ns <- nstates(fmod) # Number of states in the HMM
-  investment_values <- seq(0, 20, length = 100)
-  # Create a data frame to store the transition probabilities.
-  # Each row represents a combination of "from" state, "to" state, and "investment" value.
-  trans_prob_all <- expand.grid(
-    from = paste0("from State ", 1:ns),
-    to = paste0("State ", 1:ns),
-    investment = investment_values
-  )
-  trans_prob_all$prob_pre <- 0
-  trans_prob_all$prob_ctrl <- 0
-  trans_prob_all$prob_expo <- 0
-
-  # --- Calculate Transition Probabilities ---
-  for (from in 1:ns) { 
-    start_idx <- ns * numCovIn + 1 + (from - 1) * numCovTr * ns
-    end_idx <- start_idx + numCovTr * ns - 1
-    pars <- matrix(getpars(fmod)[start_idx:end_idx], ncol = numCovTr)
-    print(pars)
-
-    # Initialize matrices to store transition probabilities before normalization
-    y0 <- matrix(0, nrow = length(investment_values), ncol = ns) # Pre-intervention group
-    y1 <- matrix(0, nrow = length(investment_values), ncol = ns) # Control group
-    y2 <- matrix(0, nrow = length(investment_values), ncol = ns) # Exposed group
-
-    for (to in 1:ns) {
-      y0[, to] <- exp(pars[to, 1] + pars[to, 2] * investment_values)
-      y2[, to] <- exp(pars[to, 1] + pars[to, 2] * investment_values + pars[to, 3] + pars[to, 5] * investment_values) # Exposed: Intercept + investment + group effect + interaction
-      y1[, to] <- exp(pars[to, 1] + pars[to, 2] * investment_values + pars[to, 4] + pars[to, 6] * investment_values) # Control: Intercept + investment + group effect + interaction
-    }
-
-    # Normalize transition probabilities for each group to ensure they sum to 1 across all "to" states
-    y0 <- y0 / rowSums(y0)
-    y1 <- y1 / rowSums(y1)
-    y2 <- y2 / rowSums(y2)
-
-    for (to in 1:ns) {
-      idx <- trans_prob_all$from == paste0("from State ", from) & 
-             trans_prob_all$to == paste0("State ", to)
-      trans_prob_all$prob_pre[idx] <- y0[, to]
-      trans_prob_all$prob_ctrl[idx] <- y1[, to]
-      trans_prob_all$prob_expo[idx] <- y2[, to]
-    }
-  }
-
-  saveRDS(trans_prob_all, "data/trans_prob_all.RDS")
-  return(trans_prob_all)
-}
-```
-
-```{r}
+  col.names = c("#states", "log(lik)", "# par", "AIC", "BIC"))
+```
+
+```{r}
+#| label: fig-trustee-policy-plot
+#| echo: false
+#| fig.cap: Probability mass function of the trustee's policy conditional on its latent state as an output of the best fitting XX state HMM using a truncated discretised Gaussian as a response function
+#| fig.align: center
+#| fig.width: 12
+#| fig.height: 6
 
@@
 tr_policy <- list()
-for(i in 1:nstates(best_mod_all)) {
-  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(best_mod_all@response[[i]][[1]]@parameters))))  
+for(i in 1:nstates(best_mod_tr_all)) {
+  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(best_mod_tr_all@response[[i]][[1]]@parameters))))  
 }
@@
 
-respplot <- data.frame(state = rep(1:nstates(best_mod_all), each=length(perc_return)),
+respplot <- data.frame(state = rep(1:nstates(best_mod_tr_all), each=length(perc_return)),
                             perc_return = perc_return,
@@
 
-##################################################################################
-# Plotting Prior (initial) state probabilities (new)
+# Plotting Prior (initial) state probabilities
+# priorplot <- data.frame(state=factor(paste0("s",1:nstate_tr)), prob=getpars(getmodel(best_mod_tr_all, "prior"))) |>
+#   ggplot(aes(x = state, y = prob, fill=state)) +
+#   geom_bar(stat = "identity", width = 0.5) +
+#   xlab("state") +
+#   ylab("prior probability") +
+#   ggtitle("Initial state probability") +
+#   theme_minimal()
 
@@
 
-# Get parameters from model
-params <- getpars(getmodel(best_mod_all, "prior"))
+# Assuming best_mod_tr_all is your fitted model
+params <- getpars(getmodel(best_mod_tr_all, "prior"))
 
@@
 investments <- params[(nstate_tr + 1):(2 * nstate_tr)]
-# set average investment of trustee in neutral state as value to calculate initial state probabilities in linear model
 investment_value <- 9
@@
 
-```
-
-```{r,include=F}
-
-trans_prob_all <- get_trans_probs(best_mod_all, numCovIn=2, numCovTr=6)
-
-ctrl_prePost_plot <- ggplot(trans_prob_all,aes(x=investment)) + 
-  geom_line(aes(y=prob_ctrl, colour = as.factor(to), linetype="Post-control")) + 
-  geom_line(aes(y=prob_pre, colour = as.factor(to), linetype="Pre-control")) +
-  facet_wrap(~from) + 
-  ylim(c(0,1)) + 
-  ggtitle("Transition function pre- and post-control group") + 
-  labs(x = "Investment", y = "Transition probability", color='Transition to:') + 
-  scale_linetype_manual(values = c("Post-control" = "solid", "Pre-control" = "dotted")) +
-  theme_bw()
-
-expo_prePost_plot  <- ggplot(trans_prob_all,aes(x=investment)) + 
-  geom_line(aes(y=prob_expo, colour = as.factor(to), linetype="Post-exposure")) + 
-  geom_line(aes(y=prob_pre, colour = as.factor(to), linetype="Pre-exposure")) +
-  facet_wrap(~from) + 
-  ylim(c(0,1)) + 
-  ggtitle("Transition function pre- and post-exposure manipulation") + 
-  labs(x = "Investment", y = "Transition probability", color='Transition to:', linetype = "Line Type") + 
-  scale_linetype_manual(values = c("Post-exposure" = "solid", "Pre-exposure" = "dotted")) +
-  theme_bw()
-
-expo_prePost_plot 
-ctrl_prePost_plot
-```
-
-```{r}
-##############################################################################
-
-#| label: fig-trustee-policy-plot
-#| echo: false
-#| fig.cap: HMM model of the trustees 
-#| fig.align: center
-#| fig.width: 12
-#| fig.height: 6
-
-# transplot <- plot_transitions(best_mod_all,numCovIn=2,numCovTr=2) + labs(x = "Investment", y = "Probability", color='Transition to') + theme_minimal() + theme(legend.position = "bottom")
-
-gridExtra::grid.arrange(respplot, priorplot, expo_prePost_plot, ncol=3, layout_matrix=cbind(c(1,2),c(3,3),c(3,3)))
-```
-
-<!-- Get Trustee posterior states from model and add to data table -->
-
-```{r, include = FALSE}
-set.seed(20221010)
-# mention we are using local decoding (refer to the book). 
-predTrStatesAll <- posterior(best_mod_all, type="local")
-
-#############
-data_HMM_all$postTrState <-  factor(predTrStatesAll, levels= c(1,2,3,4,5),labels=c("State 1","State 2","State 3","State 4","State 5"))
-############
-```
-
-```{r}
-
-#| label: fig-trustee-posterior-plot
-#| echo: false
-#| fig.cap: POsterior state of the trustee in eahc round by Condition and Phase
-#| fig.align: center
-#| fig.width: 8
-#| fig.height: 12
-postHMM <- ggplot(data_HMM_all) +
-  geom_bar(aes(x = roundNum, group = postTrState, fill = postTrState),  position = position_stack(reverse = TRUE)) + 
-  # facet_wrap(~factor(condition.f, levels = rev(levels(as.factor(condition.f))))) +  
-  facet_wrap(~condition.f*gameNum.f)+
-  labs(x = "Round", fill='Posterior trustee state') +
-  theme_bw() +
-  theme(legend.position = "bottom")
-
-postHMM
-```
-
-<!-- Cluster joint latent states: -->
-
-```{r}
-# Load necessary libraries
-library(factoextra)
-library(hexbin)
-
-# Assuming data_HMM_post is your dataframe
-
-# Rename the levels of investorState and postTrState
-data_HMM_all <- data_HMM_all %>%
-  mutate(investorState = recode(investorState.f, 
-                                "unhappy" = "1", 
-                                "neutral" = "2", 
-                                "happy" = "3"),
-         postTrState = gsub("State ", "", postTrState))
-
-# Define the desired order of joint states
-desired_order <- c(
-  "1-1", "1-2", "1-3", "1-4", "1-5",
-  "2-1", "2-2", "2-3", "2-4", "2-5",
-  "3-1", "3-2", "3-3", "3-4", "3-5"
-)
-
-# Ensure Joint_State and Next_Joint_State are factors with the specified order
-data_HMM_all <- data_HMM_all %>%
-  mutate(Joint_State = factor(paste(investorState, postTrState, sep = "-"), 
-                              levels = desired_order, ordered = TRUE))
-
-# Plot the frequency of each Joint_State
-# Define a color palette based on the desired order
-color_palette <- colorRampPalette(c("red","orange", "green", "cyan", "blue"))(length(desired_order))
-state_pair_to_color <- setNames(color_palette, desired_order)
-
-# Plot the frequency of each Joint_State with the assigned colors
-ggplot(data_HMM_all, aes(x = Joint_State)) +
-  geom_bar(aes(fill = Joint_State)) +
-  scale_fill_manual(values = state_pair_to_color) +
-  labs(title = "Frequency of Joint States",
-       x = "Joint State",
-       y = "Frequency") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-#Plot joint states distribution by phase and condition 
-ggplot(data_HMM_all, aes(x = Joint_State, fill = gameNum.f)) +
-  geom_bar(position = position_dodge(width = 0.9)) +
-  scale_fill_manual(values = c("lightblue", "blue"), # Example colors for different conditions
-                    name = "Condition") +
-  labs(title = "Frequency of Joint States by Condition",
-       x = "Joint State",
-       y = "Frequency") +
-  facet_wrap(~condition.f)+
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-# Create next joint state for transition analysis
-data_HMM_all <- data_HMM_all %>%
-  arrange(playerId, gameNum.f, roundNum) %>%
-  group_by(playerId, gameNum.f) %>%
-  mutate(Next_Joint_State = lead(Joint_State, order_by = roundNum)) %>%
-  ungroup() %>%
-  mutate(Next_Joint_State = factor(Next_Joint_State, levels = desired_order, ordered = TRUE))
-
-# Write to CSV if needed (for reference or debugging)
-write.csv(data_HMM_all, "data/data_HMM_all.csv")
-
-# Calculate transition counts
-transition_counts <- data_HMM_all %>%
-  filter(!is.na(Next_Joint_State)) %>%
-  count(Joint_State, Next_Joint_State)
-
-# Convert counts to probabilities and reshape to wide format
-transition_matrix <- transition_counts %>%
-  group_by(Joint_State) %>%
-  mutate(transition_prob = n / sum(n)) %>%
-  ungroup() %>%
-  dplyr::select(Joint_State, Next_Joint_State, transition_prob) %>%
-  pivot_wider(names_from = Next_Joint_State, values_from = transition_prob, values_fill = list(transition_prob = 0))
-
-# Ensure the column order in transition_matrix
-transition_matrix <- transition_matrix %>%
-  dplyr::select(Joint_State, all_of(desired_order))
-
-# Visualize the transition matrix
-transition_matrix_long <- transition_matrix %>%
-  pivot_longer(cols = -Joint_State, names_to = "Next_Joint_State", values_to = "probability")
-
-ggplot(transition_matrix_long, aes(x = Next_Joint_State , y = Joint_State, fill = probability)) +
-  geom_tile() +
-  scale_fill_gradient2(low = "white", high = "blue", mid = "lightblue", midpoint = 0.5, limit = c(0, 1), space = "Lab", name="Transition Probability") +
-  theme_minimal() +
-  labs(title = "Transition Matrix of Joint States", x = "Current Joint State", y = "Next Joint State") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-```{r}
-frequency_table <- data_HMM_all %>%
-  count(Joint_State, condition.f, gameNum.f) %>%
-  spread(gameNum.f, n, fill = 0) %>% 
-  mutate(Difference = `post` - `pre`)
-
-# Plot the differences
-ggplot(frequency_table, aes(x = Joint_State, y = Difference, fill = condition.f)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
-  scale_fill_manual(values = c("lightblue", "blue"), # Example colors for different conditions
-                    name = "Condition") +
-  labs(title = "Difference in Frequency of Joint States by Condition (post - pre)",
-       x = "Joint State",
-       y = "Difference in Frequency") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
-
-<!-- Clustering of whole time series using K-means  -->
-```{r}
-
-# Reshape the data
-df_wide <- data_HMM_all %>%
-  dplyr::select(playerId, gameNum.f,investorState, postTrState) %>%
-  mutate(across(c(investorState, postTrState), as.numeric)) %>%
-  group_by(playerId, gameNum.f) %>%
-  mutate(id = row_number()) %>%
-  pivot_wider(names_from = id, values_from = c(investorState, postTrState)) %>%
-  ungroup()
-
-# Separate playerId and gameNum.f columns for later use
-playerIds <- df_wide$playerId
-gameNums <- df_wide$gameNum.f
-
-# Convert the rest of the data frame to a matrix
-data_matrix <- df_wide %>% dplyr::select(-playerId, -gameNum.f) %>% as.matrix()
-
-# Determine the optimal number of clusters using the elbow method
-wss <- vector("numeric", length = 10)
-
-for (k in 1:10) {
-  set.seed(123)  # For reproducibility
-  kmeans_result <- kmeans(data_matrix, centers = k, nstart = 25)
-  wss[k] <- kmeans_result$tot.withinss
-}
-
-# Plot the elbow plot
-plot(1:10, wss, type = "b", pch = 19, frame = FALSE,
-     xlab = "Number of Clusters K",
-     ylab = "Total Within-Clusters Sum of Squares",
-     main = "Elbow Plot for Determining Optimal K")
-
-```
-
-<!-- # Perform k-means clustering with the optimal number of clusters -->
-```{r}
-
-set.seed(123)
-final_kmeans <- kmeans(data_matrix, centers = 3, nstart = 25)
-
-# Add cluster assignments to the original data
-df_wide$cluster <- factor(final_kmeans$cluster)
-
-# Combine playerId and gameNum.f with cluster assignments
-df_wide_with_id <- df_wide %>% 
-  mutate(playerId = playerIds, gameNum.f = gameNums)
-
-# Join the cluster assignments back to the original data
-df_clustered <- data_HMM_all %>% 
-  left_join(df_wide_with_id %>% dplyr::select(playerId, gameNum.f, cluster), 
-            by = c("playerId", "gameNum.f"))
-
-df_clustered %>%
-  group_by(cluster, Joint_State) %>%
-  summarise(count = n(), .groups = 'drop') %>%
-  group_by(cluster) %>%
-  mutate(proportion = count / sum(count)) %>%
-  ggplot(aes(x = factor(cluster), y = proportion, fill = factor(Joint_State))) +
-  geom_bar(stat = "identity", position = "stack") +
-  scale_fill_manual(values = state_pair_to_color) +
-  theme_minimal() +
-  labs(title = "Proportion of Joint States by Cluster",
-       x = "Cluster",
-       y = "Proportion",
-       fill = "Joint State") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-df_clustered <- df_clustered %>%
-  mutate(cluster_mood = case_when(
-    cluster == 2 ~ "Joint Unhappy",
-    cluster == 1 ~ "Joint Happy",
-    cluster == 3 ~ "Joint Neutral",
-    TRUE ~ "Other"  # This catches any other values, if they exist
-  ) %>% factor(levels = c("Joint Unhappy", "Joint Neutral", "Joint Happy")))
-
-# Save the results to a CSV file (optional)
-write.csv(df_clustered, "clustered_data.csv", row.names = FALSE)
-
-```
-
-```{r}
-# Calculate the mean path of joint states for each cluster over time
-df_mean_path <- df_clustered %>%
-  group_by(cluster_mood, roundNum) %>%
-  summarize(meanInvestorState = mean(as.numeric(investorState), na.rm = TRUE),
-            meanPostTrState = mean(as.numeric(postTrState), na.rm = TRUE)) %>%
-  ungroup()
-
-# Combine both states in a single plot using facets
-df_mean_path_long <- df_mean_path %>%
-  pivot_longer(cols = c(meanInvestorState, meanPostTrState), 
-               names_to = "StateType", 
-               values_to = "MeanState")
-
- # Define custom colors for each cluster
-cluster_colors <- c("Joint Unhappy" = "red", "Joint Neutral" = "darkgreen", "Joint Happy" = "blue")
-
-ggplot(df_mean_path_long, aes(x = roundNum, y = MeanState, color = cluster_mood)) +
-  geom_line(size = 1.5) +
-  geom_point(size = 2) +
-  scale_color_manual(values = cluster_colors) +
-  facet_wrap(~ StateType, scales = "free_y") +
-  labs(title = "Mean Path of States by Cluster Over Time",
-       x = "Time",
-       y = "Mean State",
-       color = "Cluster") +
-  theme_minimal()
-
-```
-
-```{r}
-cluster_means <- df_clustered %>%
-  group_by(cluster_mood, roundNum) %>%
-  summarise(across(c(investment, ret_pct_0), mean))
-
-ggplot(cluster_means, aes(x = roundNum, y = investment, color = factor(cluster_mood))) +
-  geom_line() +
-  scale_color_manual(values = cluster_colors) +
-  labs(x = "Round Number", y = "Average investment by cluster", color = "Cluster")
-
-ggplot(cluster_means, aes(x = roundNum, y = ret_pct_0, color = factor(cluster_mood))) +
-  geom_line() +
-  scale_color_manual(values = cluster_colors) +
-  labs(x = "Round Number", y = "Average % return by cluster", color = "Cluster")
-
-# Load the patchwork library
-library(patchwork)
-
-# Assuming your cluster_means dataframe and cluster_colors vector are already defined
-# First plot for investment
-p1 <- ggplot(cluster_means, aes(x = roundNum, y = investment, color = factor(cluster_mood))) +
-  geom_line() +
-  scale_color_manual(values = cluster_colors) +
-  labs(x = "Round Number", y = "Average investment by cluster", color = "Cluster") +
-  theme_minimal()
-
-# Second plot for % return
-p2 <- ggplot(cluster_means, aes(x = roundNum, y = ret_pct_0, color = factor(cluster_mood))) +
-  geom_line() +
-  scale_color_manual(values = cluster_colors) +
-  labs(x = "Round Number", y = "Average % return by cluster", color = "Cluster") +
-  theme_minimal()
-
-# Combine the two plots
-combined_plot <- p1 / p2
-
-# Print the combined plot
-print(combined_plot)
-
-```
-```{r}
-
-# Calculate cumulative rewards and investments for each game
-cumulative_summary <- df_clustered %>%
-  group_by(playerId, cluster_mood, gameNum.f) %>%
-  summarise(
-    cum_investor_reward = sum(returns - investment, na.rm = TRUE),
-    cum_trustee_reward = sum(3*investment - returns, na.rm = TRUE),
-    .groups = 'drop'
-  )  %>%
-  group_by(cluster_mood) %>%
-  summarise(
-    avg_investor_cum = mean(cum_investor_reward),
-    avg_trustee_cum = mean(cum_trustee_reward)
-  ) %>%
-  ungroup()
-
-# Print summary
-print(cumulative_summary)
-
-```
-
-<!-- Correlate the Cluster with initial state of the trustee:  -->
-
-```{r}
-library(dplyr)
-library(nnet)  # for multinomial logistic regression
-library(broom)  # for tidy output
-library(ggplot2)
-
-# First, let's filter for the initial round and select relevant variables
-df_initial <- df_clustered %>%
-  filter(roundNum == 1) %>%
-  dplyr::select(playerId,gameNum.f, condition.f, cluster_mood, investment, postTrState)
-
-# Create a dataframe with the pre and post cluster moods for each player and condition
-pre_post_data <- df_initial %>%
-  dplyr::select(playerId, gameNum.f, condition.f, cluster_mood) %>%
-  spread(key = gameNum.f, value = cluster_mood) %>%
-  filter(!is.na(pre) & !is.na(post))
-
-# Calculate the distribution of people between clusters for pre and post
-distribution_pre <- pre_post_data %>%
-  group_by(condition.f, pre) %>%
-  summarise(count = n()) %>%
-  mutate(time = "pre") %>%
-  rename(cluster_mood = pre)
-
-distribution_post <- pre_post_data %>%
-  group_by(condition.f, post) %>%
-  summarise(count = n()) %>%
-  mutate(time = "post") %>%
-  rename(cluster_mood = post)
-
-# Combine the pre and post data
-combined_distribution <- bind_rows(distribution_pre, distribution_post)
-
-# Ensure the time column is a factor with levels in the correct order
-combined_distribution$time <- factor(combined_distribution$time, levels = c("pre", "post"))
-
-# Plot the evolution between pre and post
-ggplot(combined_distribution, aes(x = time, y = count, group = cluster_mood, color = cluster_mood)) +
-  geom_line(aes(linetype = cluster_mood)) +
-  geom_point() +
-  facet_wrap(~ condition.f, scales = "fixed") +
-  theme_minimal() +
-  labs(title = "Evolution of Cluster Moods Pre to Post by Condition",
-       x = "Time",
-       y = "Count",
-       color = "Cluster Mood",
-       linetype = "Cluster Mood") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-<!-- #Testing whether differences between pre and post cluster frequencies/proportions are statistically significant -->
-```{r}
-# Create a contingency table
-contingency_table <- combined_distribution %>%
-  filter(time == "pre") %>%
-  inner_join(combined_distribution %>% filter(time == "post"), by = c("condition.f", "cluster_mood")) %>%
-  dplyr::select(condition.f, cluster_mood, count.x, count.y) %>%
-  pivot_wider(names_from = condition.f, values_from = c(count.x, count.y))
-
-# Rename columns for clarity
-colnames(contingency_table) <- c("cluster_mood", "Control_Pre", "Manipulation_Pre", "Control_Post", "Manipulation_Post")
-
-# Print contingency table
-print(contingency_table)
-
-# Function to perform proportion test for each class
-perform_prop_test <- function(pre_counts, post_counts, total_pre, total_post) {
-  counts <- c(pre_counts, post_counts)
-  totals <- c(total_pre, total_post)
-  test_result <- prop.test(counts, totals)
-  return(test_result)
-}
-
-# Total counts for Control and Manipulation groups
-total_control_pre <- sum(contingency_table$Control_Pre)
-total_control_post <- sum(contingency_table$Control_Post)
-total_manipulation_pre <- sum(contingency_table$Manipulation_Pre)
-total_manipulation_post <- sum(contingency_table$Manipulation_Post)
-
-# Perform tests for Control group
-control_results <- lapply(1:nrow(contingency_table), function(i) {
-  perform_prop_test(contingency_table$Control_Pre[i], contingency_table$Control_Post[i], total_control_pre, total_control_post)
-})
-
-# Perform tests for Manipulation group
-manipulation_results <- lapply(1:nrow(contingency_table), function(i) {
-  perform_prop_test(contingency_table$Manipulation_Pre[i], contingency_table$Manipulation_Post[i], total_manipulation_pre, total_manipulation_post)
-})
-
-# Print results for Control group
-cat("Control Group Proportion Tests:\n")
-for (i in 1:length(control_results)) {
-  cat(paste("\nCluster Mood:", contingency_table$cluster_mood[i]))
-  print(control_results[[i]])
-}
-
-# Print results for Manipulation group
-cat("\n\nManipulation Group Proportion Tests:\n")
-for (i in 1:length(manipulation_results)) {
-  cat(paste("\nCluster Mood:", contingency_table$cluster_mood[i]))
-  print(manipulation_results[[i]])
-}
-
-```
-
-```{r}
-# Ensure cluster_mood is a factor
-df_initial$cluster_mood <- factor(df_initial$cluster_mood)
-
-# Fit the multinomial logistic regression model
-model <- multinom(cluster_mood ~ investment + postTrState, data = df_initial)
-
-# Summary of the model
-summary(model)
-
-# Get tidy output of the model
-tidy_model <- tidy(model)
-print(tidy_model)
-
-# Visualize the effects
-ggplot(df_initial, aes(x = investment, fill = cluster_mood)) +
-  geom_density(alpha = 0.5) +
-  facet_wrap(~postTrState) +
-  labs(title = "Distribution of Initial Investment by Cluster Mood and Trustee State",
-       x = "Initial Investment Percentage",
-       y = "Density") +
-  theme_minimal()
-```
-
-### Key takeaways:
-
-Higher initial investments slightly decrease the likelihood of being in Joint Neutral or Joint Happy clusters compared to Joint Unhappy.
-Higher trustee states, particularly 3 and above, strongly increase the odds of being in Joint Neutral or Joint Happy clusters compared to Joint Unhappy.
-The effect of trustee state is more pronounced for the Joint Happy cluster than for Joint Neutral.
-
-These results suggest that the trustee's initial state is a stronger predictor of cluster membership than the initial investment amount, with higher trustee states associated with more positive joint outcomes.
-
-<!-- Check associations with RS -->
-```{r}
-#########################
-# Perform ANOVA to test for significant differences in RS_score between clusters
-df_aov_RS <-  df_clustered %>% dplyr::select(playerId,gameNum.f, RS_score,cluster_mood) %>% unique()
-
-# Calculate the mean and standard error of RS_score for each cluster
-mean_rs_score <- df_clustered %>%
-  group_by(cluster_mood,condition.f) %>%
-  summarize(mean_RS_score = mean(RS_score, na.rm = TRUE),
-            se_RS_score = sd(RS_score, na.rm = TRUE) / sqrt(n()))
-
-# Print the mean RS score for each cluster
-print(mean_rs_score)
-
-anova_result_rs <- aov(RS_score ~ cluster_mood*gameNum.f, data=df_aov_RS)
-
-# Print the summary of the ANOVA result
-summary(anova_result_rs)
-
-# Plot the mean RS score for each cluster with error bars using geom_point and geom_errorbar
-ggplot(mean_rs_score, aes(x = cluster_mood, y = mean_RS_score, color = cluster_mood)) +
-  geom_point(size = 5) +
-  geom_errorbar(aes(ymin = mean_RS_score - se_RS_score, ymax = mean_RS_score + se_RS_score),
-                width = 0.2, size = 1.5) +
-  facet_wrap(~condition.f)+
-  scale_color_manual(values = cluster_colors) +
-  labs(title = "Average RS Score for Each Cluster",
-       x = "Cluster",
-       y = "Average RS Score",
-       color = "Cluster") +
-  theme_minimal()
-
-```
-
-<!-- Check association WITH LPFS -->
-```{r}
-# Perform ANOVA to test for significant differences in LPFS_score between clusters
-df_aov_lpfs <-  df_clustered %>% dplyr::select(playerId,LPFS_score,cluster_mood) %>% unique()
-
-anova_result_lpfs <- aov(LPFS_score ~ cluster_mood, data = df_aov_lpfs)
-
-# Print the summary of the ANOVA result
-summary(anova_result_lpfs)
-
-# Calculate the mean LPFS_score for each cluster
-mean_lpfs_score <- df_clustered %>%
-  group_by(cluster_mood) %>%
-  summarize(mean_LPFS_score = mean(LPFS_score, na.rm = TRUE))
-
-# Print the mean LPFS score for each cluster
-print(mean_lpfs_score)
-
-```
-Association between RS and variability in joitn states 
-```{r}
-
-# Calculate the number of joint state changes for each player
-result_vol <- df_clustered %>%
-  arrange(playerId, roundNum) %>%
-  group_by(playerId) %>%
-  summarize(
-    RS_score = first(RS_score),
-    joint_state_changes = sum(Joint_State != Next_Joint_State, na.rm = TRUE)) %>%
-  ungroup()
-
-# Perform correlation analysis
-correlation <- cor.test(result_vol$RS_score, result_vol$joint_state_changes)
-
-# Print correlation results
-print(correlation)
-
-# Create a scatter plot
-ggplot(result_vol, aes(x = RS_score, y = joint_state_changes)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE, color = "red") +
-  labs(title = "Relationship between RS_score and Joint State Changes",
-       x = "RS_score",
-       y = "Number of Joint State Changes")
-```
-
-```{r}
-# Load required libraries
-library(dplyr)
-library(ggplot2)
-library(tidyr)
-library(forcats)
-
-# Assuming the data is already loaded into a dataframe called 'df'
-
-# Categorize RS_score into high and low groups
-df_categorized <- df_clustered %>%
-  mutate(RS_group = case_when(
-    RS_score > 15 ~ "High RS",
-    RS_score <= 10 ~ "Low RS",
-    TRUE ~ NA_character_
-  )) %>%
-  filter(!is.na(RS_group))
-
-# Calculate the distribution of joint states for each RS group
-joint_state_distribution <- df_categorized %>%
-  group_by(RS_group, Joint_State) %>%
-  summarise(count = n()) %>%
-  group_by(RS_group) %>%
-  mutate(proportion = count / sum(count))
-
-# Visualization: Stacked bar plot
-ggplot(joint_state_distribution, aes(x = RS_group, y = proportion, fill = Joint_State)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Distribution of Joint States by Rejection Sensitivity Group",
-       x = "Rejection Sensitivity Group",
-       y = "Proportion",
-       fill = "Joint State") +
-  theme_minimal()
-
-# Chi-square test of independence
-chi_square_result <- chisq.test(table(df_categorized$RS_group, df_categorized$Joint_State))
-print(chi_square_result)
-
-# Post-hoc analysis: Standardized residuals
-residuals <- chi_square_result$residuals
-print("Standardized Residuals:")
-print(residuals)
-
-# Visualization: Mosaic plot
-library(vcd)
-mosaic(~ RS_group + Joint_State, data = df_categorized, 
-       main = "Mosaic Plot of Joint States by RS Group")
-
-# Effect size: Cramer's V
-library(effectsize)
-cramer_v <- cramers_v(df_categorized$RS_group, df_categorized$Joint_State)
-print(paste("Cramer's V:", cramer_v))
-
-# Optional: Analyze transitions between joint states
-transitions <- df_categorized %>%
-  arrange(playerId, roundNum) %>%
-  group_by(playerId) %>%
-  mutate(next_state = lead(Joint_State)) %>%
-  filter(!is.na(next_state)) %>%
-  group_by(RS_group, Joint_State, next_state) %>%
-  summarise(transition_count = n()) %>%
-  ungroup()
-
-# Visualization: Transition heatmap
-ggplot(transitions, aes(x = Joint_State, y = next_state, fill = transition_count)) +
-  geom_tile() +
-  facet_wrap(~ RS_group) +
-  scale_fill_viridis_c() +
-  labs(title = "Joint State Transitions by RS Group",
-       x = "Current State", y = "Next State", fill = "Transition Count") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-
-# Fitting HMM to Participants data in the post Phase with manipulation vs Control factor in transition function:
-
-```{r}
-
-#| label: setup-parallel-4
-
-# # Set up the parallel backend
-# num_cores <- 8
-# #create the cluster
-# my.cluster <- parallel::makeCluster(
-#   num_cores,
-#   type = "FORK"
-#   )
-# 
-# #register it to be used by %dopar%
-# doParallel::registerDoParallel(cl = my.cluster)
-```
-
-```{r}
-#| label: fit-HMM-trustee-models
-#| cache: true
-
-# # Depmix simple Models 
-# priordat <- data_HMM_post |>
-#   filter(roundNum==1)
-# 
-# tr_fdmod_cond <- list()
-# 
-# max_ns <- 8
-# 
-# 
-# set.seed(20240513)
-# my_seeds <- sample(1000:50000, max_ns)
-# 
-# 
-# # Fitting HMM models to data 
-# tr_fdmod_cond <- foreach(state = 2:max_ns, .packages = c("depmixS4"), .combine = "c") %dopar% {
-#   print(state)
-#   # Define depmix model with simple transition only depending on next_inv
-#   tr_cond <- depmix(returns ~ 1, data=data_HMM_post, nstates=state, transition=~ next_investment*condition.f, prior=~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes=rep(15,207))
-#   
-#   mus <- (1:state)*(1/(state+1))
-#   sigmas <- rep(1/(state+1), state)
-#   if(state > 8) sigmas <- sigmas + .1
-#   rModels_tr <- list()
-#   for(s in 1:state) {
-#     rModels_tr[[s]] <- list(vtdgaus(y=data_HMM_post$ret_pct_0, yield = 3*data_HMM_post$investment, pstart=c(mus[s],sigmas[s])))
-#   }
-#   
-#   tr_dmod <- makeDepmix(response=rModels_tr, transition=tr_cond@transition, prior=tr_cond@prior, ntimes = rep(15,207), homogeneous=FALSE)
-#   
-#   tmp1 <- fit(tr_dmod, emcontrol=em.control(random.start = FALSE))
-#   set.seed(my_seeds[state])
-#   tmp2 <- multistart(tr_dmod, nstart=20)
-#   if(!is.na(logLik(tmp2)) && logLik(tmp2) > logLik(tmp1)) {tmp1 <- tmp2}
-#   tmp1
-# }
-# 
-# # Stop the parallel backend
-# stopImplicitCluster()
-# 
-# 
-# save(tr_fdmod_cond, file="data/tr_fdmod_cond.RData", compression_level = 9)
-
-```
-
-```{r}
-#|  label: HMM-trustee-cond-comparison
-#|  include: false
-
-#############################
-load("data/tr_fdmod_cond.RData")
-############################
-
-max_ns <- length(tr_fdmod_cond) +1 
-df_bics_cond <- data.frame(nstate = 2:max_ns,
-                          logLik = unlist(lapply(tr_fdmod_cond, logLik)),
-                          unlist(lapply(tr_fdmod_cond, function(x) attr(logLik(x), which="df"))),
-                          AIC = unlist(lapply(tr_fdmod_cond, AIC)),
-                          BIC = unlist(lapply(tr_fdmod_cond, BIC)))
-df_bics_cond
-
-# Here we choose BIC as best m
-nstate_tr <- which.min(df_bics_cond[,5]) + 1
-
-############################
-best_mod_cond <- order_mod_vtdgaus(tr_fdmod_cond[[nstate_tr-1]])
-###########################
-summary(best_mod_cond)
-```
-
-```{r}
-#| label: tbl-BICs-tr-cond
-#| warning: false
-#| echo: false
-#| tbl-cap: Model comparison of HMMs of trustee actions with different numbers of latent states.
-
-df_bics_cond %>%
-  mutate_if(is.numeric, format, digits=4, nsmall = 0, big.mark = ",") %>%
-knitr::kable(booktabs = TRUE,
-  col.names = c("#states", "log(lik)", "# par", "AIC", "BIC"))
-```
-
-```{r}
-#| label: fig-trustee-policy-plot
-#| echo: false
-#| fig.cap: Probability mass function of the trustee's policy conditional on its latent state as an output of the best fitting XX state HMM using a truncated discretised Gaussian as a response function
-#| fig.align: center
-#| fig.width: 12
-#| fig.height: 6
-
-# Plotting the trustee HMM policy by state 
-tr_policy <- list()
-for(i in 1:nstates(best_mod_cond)) {
-  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(best_mod_cond@response[[i]][[1]]@parameters))))  
-}
-
-perc_return <- seq(0,1, length=61)
-
-respplot <- data.frame(state = rep(1:nstates(best_mod_cond), each=length(perc_return)),
-                            perc_return = perc_return,
-                            probability = unlist(tr_policy)) |>
-  mutate(state = factor(paste0("s",state))) |> 
-  ggplot(                 
-       aes(x = perc_return,
-           y = probability,
-           fill = state)) +
-  geom_bar(stat = "identity",
-           position = "dodge") + 
-  # labs(fill='Latent trustee state') +
-  xlab("Percentage return") +
-  theme_minimal() + 
-  theme(legend.position = "right") + ggtitle("State conditional strategies")
-
-# Function to calculate softmax probabilities
-softmax <- function(x) exp(x) / sum(exp(x))
-
-# Assuming best_mod_cond is your fitted model
-params <- getpars(getmodel(best_mod_cond, "prior"))
-
-# Extract intercepts and investment coefficients, and set initial investment value
-intercepts <- params[1:nstate_tr]
-investments <- params[(nstate_tr + 1):(2 * nstate_tr)]
-investment_value <- 9
-
-# Calculate logits and initial probabilities
-logits <- intercepts + investments * investment_value
-initial_probs <- softmax(logits)
-
-# Create a data frame for plotting
-priordf <- data.frame(
-  state = factor(paste0("s", 1:nstate_tr)),
-  prob = initial_probs
-)
-
-# Plot the initial state probabilities
-priorplot <- ggplot(priordf, aes(x = state, y = prob, fill = state)) +
-  geom_bar(stat = "identity", width = 0.5) +
-  xlab("State") +
-  ylab("Prior Probability") +
-  ggtitle("Initial State Probability") +
-  theme_minimal()
-
 # Plotting Transition function between states
-plot_transitions_cond <- function(fmod, numCovIn, numCovTr) {
+plot_transitions <- function(fmod,numCovIn,numCovTr) {
 
   ns <- nstates(fmod)
-  conditions <- c(0, 1)  # 0 for control, 1 for manipulation
 
   trans_prob <- data.frame(
-    from = rep(1:ns, each=80*ns*2),
-    to = rep(1:ns, each=80*2),
-    inv = rep(seq(0, 20, length=80), times=ns*2),
-    condition = rep(rep(conditions, each=80), times=ns),
+    from = rep(1:ns, each=80*ns),
+    to = rep(1:ns, each=80),
+    inv = seq(0,20,length=80),
     prob = 0
@@
 
-  y <- matrix(0.0, ncol=ns, nrow=80)
-
-  for (from in 1:ns) {
+  y <- matrix(0.0,ncol=ns, nrow=80)
+
+  for(from in 1:ns) {
+    #pars <- matrix(getpars(fmod)[seq(ns + 1 + (from - 1)*2*ns, ns + 1 + (from - 1)*2*ns + 2*ns - 1)],ncol=2)
     pars <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)], ncol=numCovTr)
+    paste0("from state: ", from)
     print(pars)
-    for (cond in conditions) {
-      for (to in 1:ns) {
-        x <- trans_prob[trans_prob$from == from & trans_prob$to == to & trans_prob$condition == cond, "inv"]
-        y[, to] <- exp(pars[to, 1] + pars[to, 2]*x + pars[to, 3]*cond + pars[to, 4]*x*cond)
-      }
-      y <- y / rowSums(y)
-      for (to in 1:ns) {
-        trans_prob$prob[trans_prob$from == from & trans_prob$to == to & trans_prob$condition == cond] <- y[, to]
-      }
+    for(to in 1:ns) {
+      x <- trans_prob[trans_prob$from == from & trans_prob$to == to, "inv"]
+      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
+    }
+    y <- y/rowSums(y)
+    for(to in 1:ns) {
+      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
     }
@@
 
-  ggplot(trans_prob, aes(x=inv, y=prob, colour = as.factor(to), linetype = as.factor(condition))) + 
-    geom_line() + 
-    facet_wrap(~from) + 
-    ylim(c(0, 1)) +
-    scale_linetype_manual(values=c("0" = "dashed", "1" = "solid")) +
-    labs(linetype = "Condition")
-}
-transplot <- plot_transitions_cond(best_mod_cond,numCovIn=2,numCovTr=4) + labs(x = "Investment", y = "Probability", color='Transition to') + theme_minimal() + theme(legend.position = "bottom")
+  ggplot(trans_prob,aes(x=inv,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))
+}
+
+transplot <- plot_transitions(best_mod_tr_all,numCovIn=2,numCovTr=2) + labs(x = "Investment", y = "Probability", color='Transition to') + theme_minimal() + theme(legend.position = "bottom")
 
@@
 ```
-
-<!-------------------------       LATENT STATE POSTERIOR PLOTS BY CONDITIONS        ------------------------------>
-
-```{r, include = FALSE}
-set.seed(20221010)
-# Get investor posterior states from model and add to data table
-
-# mention we are using local decoding (refer to the book). 
-predTrStates <- posterior(best_mod_cond, type="local")
-
-#############
-data_HMM_post$postTrState <-  factor(predTrStates, levels= c(1,2,3,4,5),labels=c("State 1","State 2","State 3","State 4","State 5"))
-############
-```
-
-```{r}
-postHMM <- ggplot(data_HMM_post %>% dplyr::filter(gameNum.f=="post")) +
-  geom_bar(aes(x = roundNum, group = postTrState, fill = postTrState),  position = position_stack(reverse = TRUE)) + 
-  facet_wrap(~factor(condition.f, levels = rev(levels(as.factor(condition.f))))) +  
-  labs(x = "Round", fill='Posterior trustee state') +
-  theme_bw() +
-  theme(legend.position = "bottom")
-
-postHMM
-```
-
-<!-------------------------       CLUSTERING PAIRS OF LATENT STATES      ------------------------------>
-
-```{r}
-# Load required libraries
-library(reshape2)
-library(ggplot2)
-library(dplyr)
-
-# Example structure of data_hmm_post:
-# playerId | roundNum | PosttrState | investorState
-# 1        | 1        | 2           | 1
-# 1        | 2        | 3           | 2
-# ...
-
-# Create a combined state pair column
-data_HMM_new <- data_HMM_post %>%
-  distinct(playerId, roundNum, .keep_all = TRUE) %>%
-  mutate(StatePair = paste(investorState.f, postTrState, sep = "-"))
-
-# Define the order of state pairs from most uncooperative to most cooperative
-state_order <- c("unhappy-State 1", "unhappy-State 2", "unhappy-State 3", "unhappy-State 4", "unhappy-State 5",
-                 "neutral-State 1", "neutral-State 2", "neutral-State 3", "neutral-State 4", "neutral-State 5",
-                 "happy-State 1", "happy-State 2", "happy-State 3", "happy-State 4", "happy-State 5")
-
-# Convert StatePair to a factor with the defined order
-data_HMM_new$StatePair <- factor(data_HMM_new$StatePair, levels = state_order, ordered = TRUE)
-
-# Convert roundNum to a numeric type and then to factor with levels in the correct order
-data_HMM_new$roundNum <- as.numeric(as.character(data_HMM_new$roundNum))
-data_HMM_new$roundNum <- factor(data_HMM_new$roundNum)
-
-# Pivot the data to have players as rows and rounds as columns
-heatmap_data <- data_HMM_new %>%
-  dplyr::select(playerId, roundNum, StatePair) %>%
-  pivot_wider(names_from = roundNum, values_from = StatePair)
-
-# Convert the data to long format for ggplot2
-heatmap_data_long <- heatmap_data %>%
-  pivot_longer(cols = -playerId, names_to = "roundNum", values_to = "StatePair")
-
-# Ensure roundNum, playerId, and StatePair are treated as characters and then as factors
-heatmap_data_long$roundNum <- as.character(heatmap_data_long$roundNum)
-heatmap_data_long$playerId <- as.character(heatmap_data_long$playerId)
-heatmap_data_long$StatePair <- as.character(heatmap_data_long$StatePair)
-
-heatmap_data_long$roundNum <- as.factor(heatmap_data_long$roundNum)
-heatmap_data_long$playerId <- as.factor(heatmap_data_long$playerId)
-heatmap_data_long$StatePair <- as.factor(heatmap_data_long$StatePair)
-
-#Get unique state pairs and assign a color to each
-unique_state_pairs <- unique(heatmap_data_long$StatePair)
-color_palette <- scales::hue_pal()(length(unique_state_pairs))
-state_pair_to_color <- setNames(color_palette, unique_state_pairs)
-
-# Assign colors: cold for uncooperative and warm for cooperative
-color_palette <- colorRampPalette(c("blue", "cyan", "yellow", "orange", "red"))(length(state_order))
-state_pair_to_color <- setNames(color_palette, state_order)
-
-# Plot the heatmap
-ggplot(heatmap_data_long, aes(x = factor(roundNum, levels = unique(roundNum)), y = playerId)) +
-  geom_tile(aes(fill = factor(StatePair, levels=unique(StatePair))), color = "white") +
-  scale_fill_manual(values = state_pair_to_color, 
-                    breaks = unique_state_pairs,
-                    labels = unique_state_pairs) +
-  theme_minimal() +
-  labs(title = "Sequence of Latent State Pairs (Investor-Trustee) for Each Participant",
-       x = "Round",
-       y = "Participant ID",
-       fill = "State Pair") +
-  theme(axis.text.y = element_blank(), 
-        axis.ticks.y = element_blank())
-
-# Visual of frequencies of joint states by phase and condition .
-
-```
-
- <!-- # 3D representation of joint latent space -->
-
-```{r}
-
-# Load required libraries
-library(plotly)
-library(dplyr)
-
-# Define the order of investor states
-investor_state_order <- c("unhappy", "neutral", "happy")
-
-# Function to plot evolution of a specific player in 3D
-plot_player_evolution_3d <- function(player_index) {
-  player_data <- data_HMM_new %>% filter(playerId == player_index)
-
-  # Ensure postTrState is converted to numeric for y-axis
-  player_data <- player_data %>%
-    mutate(
-      TrusteeState = as.numeric(gsub("State ", "", postTrState)),
-      InvestorStateNum = as.numeric(factor(investorState.f, levels = investor_state_order))
-    )
-
-  # Create 3D plot
-  p <- plot_ly(player_data, x = ~InvestorStateNum, y = ~TrusteeState, z = ~roundNum, type = 'scatter3d', mode = 'lines+markers',
-               marker = list(size = 5, color = ~roundNum, colorscale = 'Viridis', showscale = TRUE),
-               line = list(width = 4, color = ~roundNum, colorscale = 'Viridis', showscale = FALSE)) %>%
-    add_trace(
-      x = ~InvestorStateNum, y = ~TrusteeState, z = ~roundNum, 
-      marker = list(size = 8, symbol = I("circle")),
-      text = ~paste("Round:", roundNum),
-      hoverinfo = 'text'
-    ) %>%
-    add_trace(
-      x = player_data$InvestorStateNum[1], y = player_data$TrusteeState[1], z = player_data$roundNum[1], 
-      marker = list(size = 10, symbol = I("diamond"), color = "red"),
-      text = "Start",
-      hoverinfo = 'text'
-    ) %>%
-    add_trace(
-      x = player_data$InvestorStateNum[nrow(player_data)], y = player_data$TrusteeState[nrow(player_data)], z = player_data$roundNum[nrow(player_data)], 
-      marker = list(size = 10, symbol = I("cross"), color = "green"),
-      text = "End",
-      hoverinfo = 'text'
-    ) %>%
-    layout(
-      title = paste("Evolution of State Pairs Over Time for Player", player_index),
-      scene = list(
-        xaxis = list(title = 'Investor State', tickvals = 1:3, ticktext = investor_state_order),
-        yaxis = list(title = 'Trustee State', tickvals = 1:5),
-        zaxis = list(title = 'Round')
-      )
-    )
-
-  # Print the plot
-  print(p)
-}
-
-# Example usage
-plot_player_evolution_3d("YEpcoSejRCLubbkPp")
-plot_player_evolution_3d("EZYk8rfDNxcuwSpEr")
-plot_player_evolution_3d("6SYmH8njCALTjKyp2")
-
-```
-
-<!-- Cluster joint latent states: -->
-
-```{r}
-# Load necessary libraries
-library(factoextra)
-library(hexbin)
-
-# Assuming data_HMM_post is your dataframe
-
-# Rename the levels of investorState and postTrState
-data_HMM_clust <- data_HMM_post %>%
-  distinct(playerId, roundNum, .keep_all = TRUE) %>%
-  mutate(investorState = recode(investorState.f, 
-                                "unhappy" = "1", 
-                                "neutral" = "2", 
-                                "happy" = "3"),
-         postTrState = gsub("State ", "", postTrState))
-
-# Define the desired order of joint states
-desired_order <- c(
-  "1-1", "1-2", "1-3", "1-4", "1-5",
-  "2-1", "2-2", "2-3", "2-4", "2-5",
-  "3-1", "3-2", "3-3", "3-4", "3-5"
-)
-
-# Ensure Joint_State and Next_Joint_State are factors with the specified order
-data_HMM_clust <- data_HMM_clust %>%
-  mutate(Joint_State = factor(paste(investorState, postTrState, sep = "-"), 
-                              levels = desired_order, ordered = TRUE))
-
-# Plot the frequency of each Joint_State
-# Define a color palette based on the desired order
-color_palette <- colorRampPalette(c("blue", "cyan", "yellow", "orange", "red"))(length(desired_order))
-state_pair_to_color <- setNames(color_palette, desired_order)
-
-# Plot the frequency of each Joint_State with the assigned colors
-ggplot(data_HMM_clust, aes(x = Joint_State)) +
-  geom_bar(aes(fill = Joint_State)) +
-  scale_fill_manual(values = state_pair_to_color) +
-  labs(title = "Frequency of Joint States",
-       x = "Joint State",
-       y = "Frequency") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(data_HMM_clust, aes(x = Joint_State, fill = condition.f)) +
-  geom_bar(position = position_dodge(width = 0.9)) +
-  scale_fill_manual(values = c("grey40", "grey80"), # Example colors for different conditions
-                    name = "Condition") +
-  labs(title = "Frequency of Joint States by Condition",
-       x = "Joint State",
-       y = "Frequency") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Create next joint state for transition analysis
-data_HMM_clust <- data_HMM_clust %>%
-  arrange(playerId, roundNum) %>%
-  group_by(playerId) %>%
-  mutate(Next_Joint_State = lead(Joint_State, order_by = roundNum)) %>%
-  ungroup() %>%
-  mutate(Next_Joint_State = factor(Next_Joint_State, levels = desired_order, ordered = TRUE))
-
-# Write to CSV if needed (for reference or debugging)
-write.csv(data_HMM_clust, "data/data_HMM_clust.csv")
-
-# Calculate transition counts
-transition_counts <- data_HMM_clust %>%
-  filter(!is.na(Next_Joint_State)) %>%
-  count(Joint_State, Next_Joint_State)
-
-# Convert counts to probabilities and reshape to wide format
-transition_matrix <- transition_counts %>%
-  group_by(Joint_State) %>%
-  mutate(transition_prob = n / sum(n)) %>%
-  ungroup() %>%
-  select(Joint_State, Next_Joint_State, transition_prob) %>%
-  pivot_wider(names_from = Next_Joint_State, values_from = transition_prob, values_fill = list(transition_prob = 0))
-
-# Ensure the column order in transition_matrix
-transition_matrix <- transition_matrix %>%
-  select(Joint_State, all_of(desired_order))
-
-# Visualize the transition matrix
-transition_matrix_long <- transition_matrix %>%
-  pivot_longer(cols = -Joint_State, names_to = "Next_Joint_State", values_to = "probability")
-
-ggplot(transition_matrix_long, aes(x = Next_Joint_State , y = Joint_State, fill = probability)) +
-  geom_tile() +
-  scale_fill_gradient2(low = "white", high = "blue", mid = "lightblue", midpoint = 0.5, limit = c(0, 1), space = "Lab", name="Transition Probability") +
-  theme_minimal() +
-  labs(title = "Transition Matrix of Joint States", x = "Current Joint State", y = "Next Joint State") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-```{r}
-library(dplyr)
-library(ggplot2)
-
-# Function to compute the stationary distribution
-stationary_distribution <- function(transition_matrix) {
-  eigen_result <- eigen(t(transition_matrix))
-  # Find the eigenvector associated with the eigenvalue 1
-  stationary_vector <- Re(eigen_result$vectors[, which.max(Re(eigen_result$values))])
-  # Normalize the vector to sum to 1
-  stationary_vector <- stationary_vector / sum(stationary_vector)
-  return(stationary_vector)
-}
-
-# Compute the stationary distribution
-transition_matrix_numeric <- transition_matrix %>%
-  dplyr::select(-Joint_State) %>%  # Remove the 'Joint_State' column
-  as.matrix()
-
-# Print the resulting transition matrix
-print(transition_matrix_numeric)
-stationary_dist <- stationary_distribution(transition_matrix_numeric)
-
-print("Stationary Distribution:")
-print(stationary_dist)
-
-# Create a data frame for plotting with initial names
-plot_data <- data.frame(
-  State = desired_order,
-  Probability = stationary_dist
-)
-
-# Plot the stationary distribution
-ggplot(plot_data, aes(x = factor(State, levels = desired_order), y = Probability)) +
-  geom_bar(stat = "identity", fill = "steelblue") +
-  labs(title = "Stationary Distribution", x = "State", y = "Probability") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-
-<!-- Visual of times series of joint states by initial trustee latent state (investor always starts neutral) -->
-```{r}
-# # Extract the trustee state at the first trial for each player
-# first_trial_Trstate <- data_HMM_clust %>%
-#   filter(roundNum == 1) %>%
-#   dplyr::select(playerId, firstPostTrState = postTrState)
-# 
-# # Merge with the main data to add the first trial state as a column
-# data_HMM_clust <- data_HMM_clust %>%
-#   left_join(first_trial_Trstate, by = "playerId")
-# 
-# # Visualize the joint states over rounds with color based on the first trial trustee state
-# ggplot(data_HMM_clust, aes(x = roundNum, y = Joint_State, group=playerId, color = as.factor(firstPostTrState))) +
-#   geom_line(size = 1) +
-#   geom_point(size = 2) +
-#   facet_wrap(~ as.factor(firstPostTrState), scales = "free_y") +
-#   labs(title = "Joint States Over Rounds", x = "Round", y = "Joint State", color = "First Trial Trustee State") +
-#   theme_minimal() +
-#   theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
-
-<!-- Clustering of whole time series using K-means  -->
-```{r}
-
-# Reshape the data
-df_wide <- data_HMM_clust %>%
-  dplyr::select(playerId, investorState, postTrState) %>%
-  mutate(across(c(investorState, postTrState), as.numeric)) %>%
-  group_by(playerId) %>%
-  mutate(id = row_number()) %>%
-  pivot_wider(names_from = id, values_from = c(investorState, postTrState)) %>%
-  ungroup()
-
-# Separate playerId column for later use
-playerIds <- df_wide$playerId
-
-# Convert the rest of the data frame to a matrix
-data_matrix <- df_wide %>% select(-playerId) %>% as.matrix()
-
-# Determine the optimal number of clusters using the elbow method
-wss <- vector("numeric", length = 10)
-
-for (k in 1:10) {
-  set.seed(123)  # For reproducibility
-  kmeans_result <- kmeans(data_matrix, centers = k, nstart = 25)
-  wss[k] <- kmeans_result$tot.withinss
-}
-
-# Plot the elbow plot
-plot(1:10, wss, type = "b", pch = 19, frame = FALSE,
-     xlab = "Number of Clusters K",
-     ylab = "Total Within-Clusters Sum of Squares",
-     main = "Elbow Plot for Determining Optimal K")
-
-# Perform k-means clustering with the optimal number of clusters
-set.seed(123)
-final_kmeans <- kmeans(data_matrix, centers = 3, nstart = 25)
-
-# Add cluster assignments to the original data
-df_wide$cluster <- factor(final_kmeans$cluster)
-
-# Combine playerId with cluster assignments
-df_wide_with_id <- df_wide %>% 
-  mutate(playerId = playerIds)
-
-# Join the cluster assignments back to the original data
-df_clustered <- data_HMM_clust %>% 
-  left_join(df_wide_with_id %>% select(playerId, cluster), by = "playerId") %>% 
-  mutate(cluster = recode(cluster, `1` = "2", `2` = "1"))  %>% # renaming clusters so that they are in increasing order of cooperation (1 least cooeprative and 3 most) 
-  mutate(cluster = factor(cluster, levels = c("1", "2", "3")))
-
-# Save the results to a CSV file (optional)
-# write.csv(df_clustered, "clustered_data.csv", row.names = FALSE)
-
-# Display the clustering results
-print(df_clustered)
-
-# Frenquencies of clusters 
- df_clustered %>%
-  group_by(cluster, condition.f) %>%
-  summarize(frequency = n()/15) %>%
-  ungroup()
-
-```
-
-```{r}
-# Calculate the mean path of joint states for each cluster over time
-df_mean_path <- df_clustered %>%
-  group_by(cluster, roundNum) %>%
-  summarize(meanInvestorState = mean(as.numeric(investorState), na.rm = TRUE),
-            meanPostTrState = mean(as.numeric(postTrState), na.rm = TRUE)) %>%
-  ungroup()
-
-# Combine both states in a single plot using facets
-df_mean_path_long <- df_mean_path %>%
-  pivot_longer(cols = c(meanInvestorState, meanPostTrState), 
-               names_to = "StateType", 
-               values_to = "MeanState")
-
- # Define custom colors for each cluster
-cluster_colors <- c("1" = "red", "2" = "green", "3" = "blue")
-
-ggplot(df_mean_path_long, aes(x = roundNum, y = MeanState, color = cluster)) +
-  geom_line(size = 1.5) +
-  geom_point(size = 2) +
-  scale_color_manual(values = cluster_colors) +
-  facet_wrap(~ StateType, scales = "free_y") +
-  labs(title = "Mean Path of States by Cluster Over Time",
-       x = "Time",
-       y = "Mean State",
-       color = "Cluster") +
-  theme_minimal()
-
-# frequencies of clusters  what is specific about these people 
-# whether peop
-```
-
-```{r}
-cluster_means <- df_clustered %>%
-  group_by(cluster, roundNum) %>%
-  summarise(across(c(investment, ret_pct_0), mean))
-
- # Define custom colors for each cluster
-cluster_colors <- c("1" = "red", "2" = "green", "3" = "blue")
-
-ggplot(cluster_means, aes(x = roundNum, y = investment, color = factor(cluster))) +
-  geom_line() +
-  scale_color_manual(values = cluster_colors) +
-  labs(x = "Round Number", y = "Average investment by cluster", color = "Cluster")
-
-ggplot(cluster_means, aes(x = roundNum, y = ret_pct_0, color = factor(cluster))) +
-  geom_line() +
-  scale_color_manual(values = cluster_colors) +
-  labs(x = "Round Number", y = "Average % return by cluster", color = "Cluster")
-```
-
-<!-- Check association WITH LPFS -->
-```{r}
-# Perform ANOVA to test for significant differences in LPFS_score between clusters
-anova_result_lpfs <- aov(LPFS_score ~ cluster, data = df_clustered)
-
-# Print the summary of the ANOVA result
-summary(anova_result_lpfs)
-
-# Calculate the mean LPFS_score for each cluster
-mean_lpfs_score <- df_clustered %>%
-  group_by(cluster) %>%
-  summarize(mean_LPFS_score = mean(LPFS_score, na.rm = TRUE))
-
-# Print the mean LPFS score for each cluster
-print(mean_lpfs_score)
-
-```
-
-<!-- Check associations with RS -->
-```{r}
-#########################
-# Perform ANOVA to test for significant differences in RS_score between clusters
-
-anova_result_rs <- aov(RS_score ~ cluster, data = df_clustered %>% dplyr::select(playerId,RS_score,cluster) %>% unique())
-
-# Print the summary of the ANOVA result
-summary(anova_result_rs)
-
-# Calculate the mean and standard error of RS_score for each cluster
-mean_rs_score <- df_clustered %>%
-  group_by(cluster) %>%
-  summarize(mean_RS_score = mean(RS_score, na.rm = TRUE),
-            se_RS_score = sd(RS_score, na.rm = TRUE) / sqrt(n()))
-
-# Print the mean RS score for each cluster
-print(mean_rs_score)
-
-# Define custom colors for each cluster
-cluster_colors <- c("1" = "red", "2" = "green", "3" = "blue")
-
-# Plot the mean RS score for each cluster with error bars using geom_point and geom_errorbar
-ggplot(mean_rs_score, aes(x = cluster, y = mean_RS_score, color = cluster)) +
-  geom_point(size = 5) +
-  geom_errorbar(aes(ymin = mean_RS_score - se_RS_score, ymax = mean_RS_score + se_RS_score),
-                width = 0.2, size = 1.5) +
-  scale_color_manual(values = cluster_colors) +
-  labs(title = "Average RS Score for Each Cluster",
-       x = "Cluster",
-       y = "Average RS Score",
-       color = "Cluster") +
-  theme_minimal()
-
-```
-<!-- Check associations with firs trustee state and first trustee return %  -->
-
-```{r}
-
-```
-
-<!-- ######## OLD WORK  -->
-```{r}
-# Load necessary libraries
-# library(DirichletReg)
-# 
-# 
-# # Assuming merged_data is your dataframe
-# 
-# # Prepare the data for Dirichlet regression
-# # Create a matrix of the cluster probabilities
-# probability_matrix <- as.matrix(merged_data %>%
-#   select(starts_with("Cluster_")))
-# 
-# # Add a small constant to avoid zeros, which can cause issues in Dirichlet regression
-# probability_matrix <- probability_matrix + 1e-6
-# probability_matrix <- DR_data(probability_matrix)
-# 
-# # Fit a Dirichlet regression model
-# dirichlet_model <- DirichReg(probability_matrix ~ RS_score + LPFS_score, data = merged_data)
-# 
-# # Summary of the model
-# summary(dirichlet_model)
-# 
-# # Predict the probabilities for each cluster
-# predicted_probabilities <- predict(dirichlet_model, merged_data)
-# 
-# # Add predicted probabilities to the dataframe
-# predicted_probabilities_df <- as.data.frame(predicted_probabilities)
-# names(predicted_probabilities_df) <- paste0("Predicted_Cluster_", 1:5)
-# 
-# merged_data <- cbind(merged_data, predicted_probabilities_df)
-# 
-# # Evaluate the model by comparing actual and predicted probabilities
-# # Calculate the Mean Squared Error (MSE) as an example
-# mse <- function(actual, predicted) {
-#   mean((actual - predicted)^2)
-# }
-# 
-# mse_values <- sapply(1:5, function(i) {
-#   actual_col <- paste0("Cluster_", i)
-#   predicted_col <- paste0("Predicted_Cluster_", i)
-#   mse(merged_data[[actual_col]], merged_data[[predicted_col]])
-# })
-# 
-# print(mse_values)
-# 
-# # Visualize the predicted vs actual probabilities for each cluster
-# predicted_actual_df <- merged_data %>%
-#   select(playerId, starts_with("Cluster_"), starts_with("Predicted_Cluster_")) %>%
-#   pivot_longer(cols = -playerId, names_to = "Cluster", values_to = "Probability") %>%
-#   mutate(Type = ifelse(grepl("Predicted", Cluster), "Predicted", "Actual"),
-#          Cluster = gsub("Predicted_", "", Cluster))
-# 
-# ggplot(predicted_actual_df, aes(x = Cluster, y = Probability, fill = Type)) +
-#   geom_boxplot() +
-#   facet_wrap(~ Cluster) +
-#   theme_minimal() +
-#   labs(title = "Actual vs Predicted Cluster Probabilities",
-#        x = "Cluster",
-#        y = "Probability",
-#        fill = "Type")
-
-```
-
-```{r}
-# Load necessary libraries
-library(dplyr)
-library(tidyr)
-
-# Create a transition matrix
-transition_counts <- data_HMM_clust %>%
-  arrange(playerId, roundNum) %>%
-  group_by(playerId) %>%
-  mutate(next_cluster = lead(Cluster, order_by = roundNum)) %>%
-  filter(!is.na(next_cluster)) %>%  # Remove rows with NA next_cluster
-  ungroup() %>%
-  count(Cluster, next_cluster) %>%
-  spread(key = next_cluster, value = n, fill = 0)
-
-# Normalize the transition counts to get probabilities
-transition_matrix <- as.matrix(transition_counts[,-1])
-row_sums <- rowSums(transition_matrix)
-transition_matrix <- transition_matrix / row_sums
-
-print("Transition Matrix:")
-print(transition_matrix)
-
-```
-
-<!----------------------------- Compute stationary distribution of Clusters:  ---------------------->
-
-```{r}
-# Function to compute the stationary distribution of a transition matrix
-stationary_distribution <- function(transition_matrix) {
-  eigen_result <- eigen(t(transition_matrix))
-  # Find the eigenvector associated with the eigenvalue 1
-  stationary_vector <- Re(eigen_result$vectors[, which.max(Re(eigen_result$values))])
-  # Normalize the vector to sum to 1
-  stationary_vector <- stationary_vector / sum(stationary_vector)
-  return(stationary_vector)
-}
-
-# Compute the stationary distribution
-stationary_dist <- stationary_distribution(transition_matrix)
-
-print("Stationary Distribution:")
-print(stationary_dist)
-
-```
-
-<!------------------- ########## TRYING NEW CLUSTERING WITH FEATURES OF THE DYAD ################ ------------------->
-
-```{r}
-# Ensure joint states are numeric for clustering
-data_HMM_clust <- data_HMM_clust %>%
-  mutate(Joint_State_numeric = as.numeric(factor(Joint_State)))
-
-# Extract features
-features <- data_HMM_clust %>%
-  group_by(playerId) %>%
-  mutate(
-    cumulative_invPayoff = cumsum(returns - investment+20),
-    cumulative_trPayoff = cumsum(3*investment -returns)) %>%
-  summarize(
-    avg_joint_state_duration = mean(rle(Joint_State_numeric)$lengths),
-    avg_investment = mean(investment),
-    avg_return = mean(returns)
-    # avg_invPayoff = mean(cumulative_invPayoff),
-    # avg_trPayoff = mean(cumulative_trPayoff) 
-  ) %>%
-  ungroup()
-
-# Prepare the feature matrix for clustering
-feature_matrix <- features %>%
-  dplyr::select(-playerId) %>%
-  scale()
-
-# Determine the optimal number of clusters using the elbow method
-fviz_nbclust(feature_matrix, kmeans, method = "wss")
-
-# Perform k-means clustering
-k <- 3  # Choose the number of clusters based on the elbow plot
-kmeans_result <- kmeans(feature_matrix, centers = k, nstart = 25)
-
-# Add cluster assignments to the features dataframe
-features$cluster <- kmeans_result$cluster
-
-# Visualize the clusters
-fviz_cluster(kmeans_result, data = feature_matrix)
-
-pca_result <- prcomp(feature_matrix, scale. = TRUE)
-print(pca_result)
-summary(pca_result)
-
-# Load required library (if not already loaded)
-library(ggbiplot)
-
-# Assuming your PCA result is in pca_result
-g <- ggbiplot(pca_result, 
-              obs.scale = 1, 
-              var.scale = 1, 
-              ellipse = TRUE,
-              circle = TRUE,
-              varname.size = 4
-              )
-g <- g + scale_color_discrete(name = '')
-g <- g + theme(legend.direction = 'horizontal', 
-               legend.position = 'top')
-g <- g + theme_minimal()
-print(g)
-
-```
-```{r}
-# Interpret the clusters
-cluster_summary <- features %>%
-  group_by(cluster) %>%
-  summarize_all(mean)
-
-print(cluster_summary)
-
-```
-
-<!-- NEW APPROACH  -->
-
-```{r}
-# Load necessary libraries
-library(dplyr)
-library(tidyr)
-library(cluster)
-library(ggplot2)
-library(factoextra)
-
-# Example data frame structure (assuming you have the actual data in a similar structure)
-data <- data_HMM_clust %>% dplyr::select(playerId, roundNum, investorState,postTrState)
-
-# Number of categories in each time series
-n_cat1 <- 3
-n_cat2 <- 5
-
-# Function to calculate frequency distribution
-frequency_distribution <- function(series, n_cat) {
-  table(factor(series, levels = 1:n_cat)) / length(series)
-}
-
-# Function to calculate transition matrix
-transition_matrix <- function(series, n_cat) {
-  series <- factor(series, levels = 1:n_cat)
-  trans_mat <- matrix(0, nrow = n_cat, ncol = n_cat)
-  for (i in 1:(length(series) - 1)) {
-    trans_mat[as.numeric(series[i]), as.numeric(series[i + 1])] <- trans_mat[as.numeric(series[i]), as.numeric(series[i + 1])] + 1
-  }
-  trans_mat <- trans_mat / rowSums(trans_mat)
-  trans_mat[is.nan(trans_mat)] <- 0  # Replace NaN with 0
-  return(trans_mat)
-}
-
-# Function to calculate co-occurrence matrix
-co_occurrence_matrix <- function(series1, series2, n_cat1, n_cat2) {
-  series1 <- factor(series1, levels = 1:n_cat1)
-  series2 <- factor(series2, levels = 1:n_cat2)
-  co_mat <- matrix(0, nrow = n_cat1, ncol = n_cat2)
-  for (i in 1:length(series1)) {
-    co_mat[as.numeric(series1[i]), as.numeric(series2[i])] <- co_mat[as.numeric(series1[i]), as.numeric(series2[i])] + 1
-  }
-  co_mat <- co_mat / sum(co_mat)
-  print(co_mat)
-  return(co_mat)
-}
-
-# Extract unique player IDs
-player_ids <- unique(data$playerId)
-
-# Initialize an empty list to store feature vectors
-feature_list <- list()
-
-# Loop over each playerId to extract features for each dyad
-for (player_id in player_ids) {
-  dyad_data <- data %>% filter(playerId == player_id)
-
-  time_series1 <- dyad_data$investorState 
-  time_series2 <- dyad_data$postTrState
-
-  # Calculate frequency distributions
-  freq_dist1 <- frequency_distribution(time_series1, n_cat1)
-  freq_dist2 <- frequency_distribution(time_series2, n_cat2)
-
-  # Calculate transition matrices
-  trans_matrix1 <- as.vector(transition_matrix(time_series1, n_cat1))
-  trans_matrix2 <- as.vector(transition_matrix(time_series2, n_cat2))
-
-  # Calculate co-occurrence matrix
-  co_matrix <- as.vector(co_occurrence_matrix(time_series1, time_series2, n_cat1, n_cat2))
-
-  # Construct the feature vector for the current dyad
-  features <- c(freq_dist1, freq_dist2, trans_matrix1, trans_matrix2, co_matrix)
-
-  # Append the feature vector to the list
-  feature_list[[as.character(player_id)]] <- features
-}
-
-# Convert the list of feature vectors into a dataframe
-features_df <- do.call(rbind, feature_list)
-
-# Perform PCA for dimensionality reduction
-pca_result <- prcomp(features_df, scale. = TRUE)
-pca_data <- data.frame(pca_result$x)
-
-# Clustering using K-means (adjust the number of clusters as needed)
-set.seed(123)  # For reproducibility
-kmeans_result <- kmeans(pca_data, centers = 2)
-
-# Add cluster labels to the PCA data
-pca_data$cluster <- as.factor(kmeans_result$cluster)
-
-# Visualize the clusters
-ggplot(pca_data, aes(x = PC1, y = PC2, color = cluster)) +
-  geom_point(size = 3) +
-  ggtitle('Clustering of Dyadic Categorical Time Series') +
-  xlab('Principal Component 1') +
-  ylab('Principal Component 2') +
-  theme_minimal()
-
-# Print the clustering results
-kmeans_result
-
-```

### 2024-07-03 article.Rmd
improved intro and discussion

@@
 ---
-title: "Can Exposure To Forgiving AI Foster Cooperative Play?"
+title: "When Forgiveness Backfires: an AI-Mediated Randomised Controlled Experiment on Rejection Sensitivity and Cooperation"
 author:
@@
 abstract: |
+ Can exposure to forgiving behavior improve cooperative outcomes for individuals prone to rejection sensitivity?
+This randomized controlled experiment explores this question using a novel approach with Hidden Markov Model (HMM)-based AI agents in repeated trust games (RTGs).
+Participants (N = 206), pre-screened for high or low rejection sensitivity (RS), interacted with either consistently behaving or more forgiving co-players, which were in reality HMM agents.
+Unexpectedly, exposure to forgiving agents led to reduced cooperation in subsequent interactions.
+Control group participants maintained consistent behavior.
+RS levels influenced perceptions of the agent's cooperativeness but did not significantly affect behavioral outcomes, revealing a perception-behavior dissociation.
+These findings challenge assumptions about fostering cooperation through simple exposure to positive interactions.
+They also demonstrate the potential of HMM-based AI agents for studying interactive social dynamics, offering a methodological advancement for future research.
+Our results suggest the need for nuanced interventions that gradually shape expectations and behaviors to foster cooperative outcomes, particularly for individuals with high rejection sensitivity.
 
@@
 
-<!-- Trust-based cooperation is an important element of human social interactions, facilitating seamless interpersonal and intergroup relations -->
-
-<!-- <!-- ### Lack of trust and social dysfunction  -->
-
-<!-- Research into the determinants of psychopathology has also linked trust-based constructs to the emergence of mental health disorders [@fonagy_mentalizing_2017].
-Many studies have drawn attention to difficulties in social cognition as a crucial concern within personality disorders (See review in @herpertz_social-cognitive_2014).
-Individuals with these disorders often find it challenging to initiate and sustain social interactions, leading to behaviors that undermine their relationships.
-Specifically, individuals facing such social cognitive challenges tend to act uncooperatively, which has been identified as a key marker for the severity and presence of  a cluster of Personality Disorder (PD) symptoms [@mulder_relationship_1999].
--->
-
-<!-- <!-- ### Could lack of trust be due to sampling bias for uncaring early caregivers ? -->
-
-<!-- One perspective of how these difficulties in the social domain emerge is through experience with early caregivers.
-Attachment theory [@bowlby_attachment_1978] posits that the quality of early caregiver relationships shapes an individual's future ability to form secure attachments and trust others.
-Research strongly suggests a connection between insecure attachment styles (characterized by anxiety or avoidance) and lower levels of trust in relationships.
-Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998].
-Similarly, @fonagy_role_2014 posit that when a young learner grows up surrounded by unreliable communicators, they may adapt by becoming mistrustful of social knowledge.
-This mistrust serves as a protective mechanism, particularly within environments where caregivers lack benign intention.
--->
-
-<!-- <!-- In their model, drawn on Bowlby’s (1973) attachment theory, Downey and colleagues (1996, 1998) proposed that RS may evolve through early experiences of repeated rejection from primary caregivers, which lead individuals to form the expectation that they are likely to be rejected by others.
-This expectation is associated with hypervigilance for rejection-relevant cues, which in turn leads to negative emotional responses and maladaptive behavioral reactions.
-Even if RS is more strongly activated in threatening situations, it is assumed to be a stable disposition, a sort of implicit theory about the self (i.e., rejectable) and others (i.e., rejecters).
-As such, it can lead to long-term negative effects on individual functioning (e.g., self-regulation, mental health, per- sonality, social cognition) and undermine relationship quality (Gao, Assink, Cipriani, & Lin, 2017).
--->
-
-<!-- <!-- ### Exposure as a way to correct the sampling bias.  -->
-
-<!-- If this adaptive mistrust is the source of social dysfunction, we can ask whether exposing those who exhibit it to cooperative and forgiving interaction partners might correct this bias.
-Research in the fields of behavioral economics and psychology has explored how positive social interactions influence trust and cooperation.
-The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995].
-In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors and trust levels, even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005].
-@fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior.
-Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior.
-These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust.
--->
-
-<!-- <!-- ### Describe experiment -->
-
-<!-- In this study, using a randomised control trial design, we assess the effectiveness of a manipulation aimed at repairing the potential breakdown of cooperation in the RTG from a low investment sent by a computerized investor.
-The manipulation focuses on repeatedly exposing participants to agents that have been designed to have a limited propensity to retaliate to a non-cooperative action by the participants.
-In order to select a sample that might exhibit difficulties in building relationships based on Trust dues to adverse early experiences, we focused on stratifying participants based on Rejection Sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection.
-It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997].
-Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017).
-However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas.
-By startifying participants based on RS, we can directly examine whether the manipulation is particularly effective  for those most vulnerable to trust-related issues.
-Additionally, comparing outcomes for those with high and low RS will help us determine if the exposure-based manipulation has differential impacts depending on an individual's underlying sensitivity to rejection.
-The computerized investor in the RTG was programmed to play according to a hidden Markov model estimated from real players' data in prior research.
-A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
-Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
--->
-
 Trust is fundamental to human social interactions, facilitating seamless relations at both interpersonal and intergroup levels.
 The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders [@fonagy_mentalizing_2017].
-Individuals with personality disorders often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors – a marker for the severity of PD symptoms [@herpertz_social-cognitive_2014; @mulder_relationship_1999].
+Individuals with personality disorders (PD) often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors – a marker for the severity of PD symptoms [@herpertz_social-cognitive_2014; @mulder_relationship_1999].
 
@@
 Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998].
-These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection [@downey_early_1997].
+These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection [@downey_implications_1996; @downey_early_1997].
 RS has been linked to the development of various mental health conditions, including depression, anxiety, personality disorders, and self-harm [@gao_associations_2017].
-Given the potential influence of RS on trust-related dynamics, it becomes particularly relevant to our understanding of social dysfunction.
-Could an ingrained sensitivity to rejection contribute to a mistrust bias?
-Similarly, learners exposed to unreliable communicators could develop mistrust of social knowledge as a protective strategy [@fonagy_role_2014].
-
-If this adaptive mistrust is the source of social dysfunction, we can ask whether exposing those who exhibit it to cooperative and forgiving interaction partners might correct this bias.
-Research in the fields of behavioral economics and psychology has explored how positive social interactions influence trust and cooperation.
+A recent meta-analysis revealed prosocial behavior and interpersonal trust as two key processes of interpersonal functioning that are markedly impaired in PDs and which are likely to contribute to interpersonal dysfunction in this population [@hepp_prosociality_2022].
+The interaction of rejection sensitivity and trust-based constructs has been explored, particularly in Borderline Personality Disorder (BPD).
+@miano_rejection_2013 and @richetin_emotional_2018 found that rejection sensitivity mediates the relationship between BPD features and lower trust appraisal.
+However, the interaction between trustworthiness and rejection sensitivity hasn't been studied as extensively in other personality disorders or in non-clinical populations, leaving a gap in our understanding of how these constructs might interplay.
+Could an ingrained sensitivity to rejection contribute to reduced pro-social behavior such as a failure to show trustworthiness?
+
+If indeed RS contributes to interpersonal dysfunction through failure to show trustworthiness, could exposing those who exhibit it to cooperative and forgiving interaction partners lead to improved interpersonal outcomes?
+Research in the fields of behavioral economics and psychology has explored how positive social interactions influence cooperation and pro-social behavior.
 The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995].
-In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors and trust levels, even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005].
+In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005].
 @fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior.
 Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior.
-These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust.
-
-<!-- In this study, we use a randomized control trial to test whether exposing participants to forgiving and more cooperative co-players can produce more trustworthy behavior and repair potential breakdowns in RTG cooperation.
-We stratify the sample of participants based on rejection sensitivity (RS).
-In order to simulate realistic social interaction while maintaining a high degree of experimental control, the computerized investor in the RTG was designed to play using a generative model of how humans play.
-We fit Hidden markov Models (HMMs) to real players' data from prior studies to build a responsive yet consistent co-player.
-A key aspect of this agent is that its actions depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
-An advantage of having a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation.
-Such generative  model of behavior, informed by empirical data and integrating a responsive trust mechanism, represent a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.
-In this study, we leverage this flexibility to expose participants in between trust games to artificial agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
-In the control condition, the participants play the human like agents throughout.
-We would expect those in the manipulation condition to behave more cooperatively and have a lower propensity to retaliate to the co-player's defection.
-We would also expect those exhibiting higher rejection sensitivity to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
--->
-
-In this study, we use a randomized control online experiment to test whether exposing participants with varying levels of RS to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation.
-In order to simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players data.
-A key aspect of these agents is that their actions depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios.
-An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation.
-In this study, participants were given the role of the trustee.
-After playing a 15 rounds RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition.
+These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust and potentially fostering trustworthy behavior.
+
+In this study, we use a randomized controlled online experiment to test whether exposing participants with varying RS levels to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation.
+To simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players' data.
+A key aspect of these agents is that their actions depend on a latent "trust state" which reacts dynamically to the trustees' returns, simulating real-life trust-building scenarios.
+An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively, or the propensity to retaliate after breakdowns of cooperation.
+
+We pre-screened participants for high or low rejection sensitivity (RS) using a validated questionnaire, then assigned them exclusively to the trustee role in a series of trust games.
+After playing a 15-round RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition.
 In the Manipulation condition, participants were exposed over three RTGs to HMM investors designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
-In the Control condition, participants, played three RTGs against the same human-like HMM.
-After this exposure phase, all participants played another 15 round RTG with a human-like HMM investor.
-We hypothesise that those in the manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase.
-We also expect those exhibiting higher RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
+In the Control condition, participants played three RTGs against the same human-like HMM.
+After this exposure phase, all participants played another 15-round RTG with a human-like HMM investor.
+We expect those exhibiting high RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
+We also hypothesize that those in the Manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase, with those in the higher RS group showing a bigger effect.
 
@@
 
-```{r loadCSV}
-
-library(dplyr)
-d_finished <- readRDS("../scriptMongo/data/d_finished.csv")
+```{r loadData}
+
+load("data/d_finished.RData")
 
@@
 
-###### get whether high or low RS and score from df_screening
-
-# Load the "df_screening.RData" file
-load("data/df_screening.RData")
-
-RS_data <- df_screening %>%  mutate(high_RS = factor(ifelse(RS_score > 15, 1, 0), 
+# Step 3:  create binary : high or low RS
+d_finished <- d_finished %>%
+  mutate(high_RS = factor(ifelse(RS_score > 15, 1, 0), 
                                     levels = c(0, 1), 
-                                    labels = c("low RS", "high RS")))  %>% 
-                             dplyr::select(id, MH.f, RS_score, high_RS)
-
-# Find IDs for those who participated in experiment and add RS_score and high_RS as new columns to d_finished.
-# Merging d_finished with relevant columns from RS_data based on matching ids
-d_finished <- d_finished %>%
-  inner_join(RS_data %>% dplyr::select(id, RS_score, high_RS), by = "id")
+                                    labels = c("low RS", "high RS")))
 
@@
 A total of 206 participants (56% female) were recruited on the Prolific Academic platform (prolific.co).
+The required sample size was determined based on an a priori power analysis.
 The mean age of participants was 34.6 years, with an 11.9 years standard deviation.
 Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.
-Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire (RSQ) to form two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10).
+A total of 1195 participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10).
 
@@
 The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : high or low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A).
-Participants within each pre-screened group were randomly assigned to one of the two levels of the first factor.
+Participants within each pre-screened group were randomly assigned to one of the two levels of the Condition factor.
 The games were designed and implemented online using Empirica v1 [@almaatouq_empirica_2021].
+The planned experiment received approval from the University of Heidelberg's Medical Faculty ethics commission (ID:S-708/2023) and the experiment was performed in accordance with the ethics board guidelines and regulations.
+All participants provided informed consent prior to their participation.
 
@@
 
+The computerised investor consisted of a hidden Markov model (HMM) trained on an independent existing behavioral RTG data set of human investors.
+This data-driven approach thus sought to learn an investor strategy that mimics human-like interactions.
+The data set used for training consists of 388 ten round games with the same player (full details can be found in the Supplementary Information).
+On this data set, the HMM was inferred with three latent states that could be interpreted as reflecting a “low-trust”, a “medium-trust”, and a “high-trust” state.
+A separate output distribution, that maps each HMM state onto possible investments from 0 to 20 separately, is learned (Figure \@ref(fig:HMMPanels).B).
+In analogy to the latent states, these distributions can be interpreted as reflecting “low-trust”, “medium-trust”, or “high-trust” dispositions.
+Finally, the HMM is specified by transition probabilities that describe the transition between states.
+The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C)).
+The initial state for the HMM investor in each instance of the game was set to the “mid-trust” state.
+Details on how the HMM state conditional probabilities and transition functions are specified can be found in the supplement.
+
+In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair it, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) and round 13 (post-manipulation).
+On all other rounds, the investor’s actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
+
+## Manipulation
+
+In all phases of the RTG other than the ‘Exposure phase’ (Figure \ref{fig:HMMPanels}.A), participants interacted with this human-like HMM.
+In the ‘Manipulation’ Condition of the exposure phase, however, the parameters of this HMM were adjusted to design a ‘forgiving’ and ultimately more cooperative agent.
+To achieve this, we changed the state transition probabilities of the HMM such that it becomes impossible for it to remain in a low trust state, effectively setting the transition probability for remaining in a “low-trust” state to 0.
+The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
+The policies conditional on the latent states and the transition function in the other latent states remain unchanged.
+
 \begin{landscape}
@@
 Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
-Panel D shows the transition probabilities of the forgiving HMM agent which are identical to those of the human-like agent, except in the low trust state.
-Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
+Panel D shows the transition probabilities of the forgiving HMM agent in the low trust state.
+Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state, as a proxy for coaxing behavior.
+Transitions in the medium and high trust states were identical for both agents.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
 
@@
 \end{landscape}
-
-The strategy of the computerised investor was modelled on the behavior of human investors in the RTG using an existing dataset consisting of 388  ten round games with the same co-player.
-Full detail on the datasets used in the Supplementary Information.
-Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states.
-The HMM consists of three outputs.
-First, a distribution over the possible investments from 0 to 20 for each latent state of the investor (Figure \@ref(fig:HMMPanels).B).
-These distributions can be interpreted as reflecting "low-trust", "medium-trust", or "high-trust" dispositions.
-The HMM also specifies how  the investor can move between these latent states.
-The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C).
-Finally, the HMM also provides as an output the initial distribution of latent states for the investor.
-In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) or round 13 (post-manipulation).
-On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
-The initial state for the HMM investor in each instance of the game was the "mid-trust" state.
-
-## Manipulation
-
-The HMM resulting from fitting the model to human investor play in the RTG (Human-like HMM) isused as a basis to design a forgiving and ultimately more cooperative agent.
-To achieve that, we changed the transition function of the investor HMM to make it impossible to remain in a low trust state once the agent transitions there.
-This is done by selecting  parameters of the transition function to make the probability of remaining in the "low-trust" state, when the agent is in that state, effectively nil.
-The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
-The policies conditional on the latent states and the transition function in the other latent states remain unchanged.
 
@@
 On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round.
-After completing 15 rounds of the RTG, participants rated how cooperative, forgiving they perceived the investor to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
-
 Phase 2 ("exposure") consisted of three 7-round RTGs.
@@
 Those in the Control condition faced the same human-like HMM agent as in the "pre" phase and rated each co-player on the same attributes.
-To keep the experience similar to the "pre" phase, the agent in the control condition was also designed to send a very low investment in round 5 of each of the three games.
-In the post-manipulation phase ("post"), participants in both conditions were told they would face a new player, and faced the same human-like HMM as in "pre" phase.
-Participants then completed the Levels of Personality Functioning Scale (LPFS) questionnaire (see the supplement for details).
+To keep the experience similar to the "pre" phase, the agent in the Control condition was also designed to send a very low investment in round 5 of each of the three games.
+In the post-manipulation phase ("post"), participants in both conditions faced the same human-like HMM as in "pre" phase.
+
+At the beginning of each game in all three phases, participants were told they would face a new player and had to wait to be paired with an available co-player.
+This simulated the waiting time in real social interaction tasks.
+After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-playerr to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
+After completing the three game phases, participants then completed the Levels of Personality Functioning Scale (LPFS) questionnaire (see the supplement for details).
 Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.
 
-## Statistical analysis
-
-To explore whether participants behaved differently in the RTG after the manipulation compared to the control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.
+## Statistical Analysis
+
+To explore whether participants behaved differently in the RTG after the manipulation compared to the Control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.
 post-manipulation), Condition (manipulation vs.
-control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase.
+Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase.
+The full specification of the statistical model can be found in the supplement.
+To test whether participants behaved differently between conditions in the Exposure phase, we fit a linear mixed effects model to participants returns only in the exposure phase, with Condition (manipulation vs.
+Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts.
+Finally, to test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
 
@@
 A similar process was used to establish the random effects structures of linear mixed-effects models used to analyse the HMM agent investments as well as the participants' ratings of the co-players.
+There is no agreed upon way to calculate effect sizes for mixed effects models.
+Instead, we will report on testing differences in marginal means.
 For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex".
 We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
-
-```{r ratingsProcess}
-
-# d_finished now has the 'Condition' column with the value equal to expo_opponent for all rounds and rows per playerId
-
-d_finished <- d_finished  %>% mutate( gameNum.f = factor(gameNumber,labels = c("pre", "expo1", "expo2", "expo3","post"),
-                                      levels=c("1", "11", "12", "13", "2")),
-                                      phase.f = factor(ifelse(gameNumber %in% c("1", "2"), 
-                                                       ifelse(gameNumber == "1", "pre", "post"), NA), levels = c("pre", "post")),
-                                      investorState.f = factor(investorState, labels= c("unhappy", "neutral", "happy"), levels =c("unhappy", "neutral", "happy"))) %>% 
-                              rename(# Before Exposure
-                            rating_cooperative_1 = `ratePLayer_trustGame_1.How cooperative do you think the other player is?`,
-                            rating_forgiving_1 = `ratePLayer_trustGame_1.How forgiving do you think the other player is?`,
-                            rating_playAgain_1 = `ratePLayer_trustGame_1.If given the choice, would you like to play again with the other player?`,
-
-                            # During Exposure
-                            rating_cooperative_11 = `ratePLayer_trustGame_11.How cooperative do you think the other player is?`,
-                            rating_forgiving_11 = `ratePLayer_trustGame_11.How forgiving do you think the other player is?`,
-                            rating_playAgain_11 = `ratePLayer_trustGame_11.If given the choice, would you like to play again with the other player?`,
-                            rating_cooperative_12 = `ratePLayer_trustGame_12.How cooperative do you think the other player is?`,
-                            rating_forgiving_12 = `ratePLayer_trustGame_12.How forgiving do you think the other player is?`,
-                            rating_playAgain_12 = `ratePLayer_trustGame_12.If given the choice, would you like to play again with the other player?`,
-                            rating_cooperative_13 = `ratePLayer_trustGame_13.How cooperative do you think the other player is?`,
-                            rating_forgiving_13 = `ratePLayer_trustGame_13.How forgiving do you think the other player is?`,
-                            rating_playAgain_13 = `ratePLayer_trustGame_13.If given the choice, would you like to play again with the other player?`,
-
-                            # After Exposure
-                            rating_cooperative_2 = `ratePLayer_trustGame_2.How cooperative do you think the other player is?`,
-                            rating_forgiving_2 = `ratePLayer_trustGame_2.How forgiving do you think the other player is?`,
-                            rating_playAgain_2 = `ratePLayer_trustGame_2.If given the choice, would you like to play again with the other player?`,
-
-                            )
-
-```
 
@@
 mu
-
-# ggplot(mu,aes(x=gameNum.f)) + 
-#   geom_bar(aes(y=mean_forgiv, stat = "identity", position = "dodge", color="blue")) + 
-#   geom_bar(aes(y=mean_coop, stat = "identity", position = "dodge", color="darkgreen")) +
-#   geom_bar(aes(y=mean_playAgain, stat = "identity", position = "dodge", color="red")) +
-#   ggtitle("Player ratings after each game") +
-#   xlab("Game Number") +
-#   theme_minimal() + 
-#   theme(legend.position = "bottom")
-#   
-```
-
-```{r LPFS}
-library(dplyr)
-
-# Step 1: Convert LPFS columns to numeric
-d_finished <- d_finished %>%
-  mutate(across(starts_with("LPFS"), as.numeric))
-
-# Step 2: Calculate means
-# Identify LPFS columns
-lpfs_cols = grep("^LPFS", names(d_finished), value = TRUE)
-lpfs_1_to_6_cols = grep("^LPFS\\.[1-6]$", names(d_finished), value = TRUE)
-lpfs_7_to_12_cols = grep("^LPFS\\.(7|8|9|10|11|12)$", names(d_finished), value = TRUE)
-
-# Calculate means using vectorized operations
-d_finished <- d_finished %>%
-  mutate(
-    LPFS_score = rowMeans(dplyr::select(., all_of(lpfs_cols)), na.rm = TRUE),
-    LPFS_self_score = rowMeans(dplyr::select(., all_of(lpfs_1_to_6_cols)), na.rm = TRUE),
-    LPFS_other_score = rowMeans(dplyr::select(., all_of(lpfs_7_to_12_cols)), na.rm = TRUE)
-  )
-save(d_finished, file = "data/d_finished.RData")
-
-```
-
-```{r dfCorr}
-
-library(dplyr)
-library(GGally)
-
-df_corr <- d_finished %>% dplyr::select(id, RS_score, LPFS_score, LPFS_other_score,payoffTrust1,payoffTrust2) %>%
-  unique()
-
-# df_corr %>% summarise(
-#   Pearson_correlation = cor(LPFS_score, RS_score, method = "pearson", use = "complete.obs"),
-#   Spearman_correlation = cor(LPFS_score, RS_score, method = "spearman", use = "complete.obs")
-# )
-# 
-# # View the correlations
-# print(correlations)
-# 
-# # Now plot the scatter plot for the entire dataset (not by individual id)
-# ggplot(df_corr, aes(x = LPFS_score, y = RS_score)) +
-#   geom_point() +
-#   geom_smooth(method = "lm", se = FALSE) +
-#   labs(x = "LPFS Score", y = "RS Score", title = "Scatterplot of LPFS Score vs. RS Score") +
-#   theme_minimal()
-
-# Select only the numerical columns for the correlation matrix (excluding the 'id' column)
-numerical_data <- df_corr %>% dplyr::select(-id)
-
-# Calculate the correlation matrix
-correlation_matrix <- cor(numerical_data, use = "complete.obs")
-
-# Print the correlation matrix
-print(correlation_matrix)
-
-# Now create a scatterplot matrix
-# Install the package if you haven't already
-if (!require("GGally")) install.packages("GGally")
-
-ggpairs(numerical_data)
-
-```
-
-```{r corrPayoffs}
-
-temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
-  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f, RS_score, LPFS_score) %>%
-  mutate(perf = payoffTrust1 + payoffTrust2) %>% 
-  unique()
-
-# #                            Bonus Calculation
-# avg_perf <- mean(temp2$perf)
-# lowRS_bonus <- temp2 %>% filter(high_RS== "low RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
-# write.csv(lowRS_bonus,"data/bonusLow.csv")
-# 
-# highRS_bonus <- temp2 %>% filter(high_RS== "high RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
-# write.csv(highRS_bonus,"data/bonusHigh.csv")
-
-ggplot(data = temp2, aes(x = payoffTrust1, y = payoffTrust2)) +
-  geom_point() +  # Color points by id
-  labs(x = "Payoff Trustee 1", y = "Payoff Trustee 2", color = "ID") +
-  theme_minimal() +
-  theme(legend.position = "right")
-
-```
-
-```{r}
-# Sanity checks
-
-# prolific_highRS <- read.csv(file="/Users/mbp17/Downloads/prolific_export_highRS.csv")
-# 
-# # Which Ids in highRS file but not in Prolific? 
-# 
-# # Find IDs in highRS_bonus that are not in prolific_highRS
-# ids_in_highRS_not_in_prolific <- setdiff(highRS_bonus$id, prolific_highRS$Participant.id)
-# 
-# # Find IDs in prolific_highRS that are not in highRS_bonus
-# ids_in_prolific_not_in_highRS <- setdiff(prolific_highRS$Participant.id, highRS_bonus$id)
-# 
-# # ids_in_highRS_not_in_prolific now contains IDs that are in highRS_bonus but not in prolific_highRS
-# # ids_in_prolific_not_in_highRS now contains IDs that are in prolific_highRS but not in highRS_bonus
-# 
-# # print these variables to see the IDs
-# print(ids_in_highRS_not_in_prolific)
-# print(ids_in_prolific_not_in_highRS)
-```
-
-```{r}
-library(ggplot2)
-library(dplyr)
-
-# First, reshape data into a longer format.
-long_data_temp <- d_finished %>%
-  dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
-  dplyr::select(id, payoffTrust1, payoffTrust2) %>%
-  gather(key = "game", value = "payoff", -id) %>%
-  mutate(game = factor(game, levels = c("payoffTrust1", "payoffTrust2"))) %>% 
-  unique()
-
-# Now, create the dumbbell plot
-ggplot(long_data_temp, aes(x = game, y = payoff, group = id)) +
-  geom_line(aes(group = id), color = "gray") + # Connect payoffs with a line
-  geom_point(aes(color = id), size = 3, alpha = 0.6) + # Payoff points
-  theme_minimal() +
-  theme(legend.position = "none") + # Hide legend if too many playerids
-  labs(x = "Game", y = "Payoff", title = "Payoff Comparison between Game 1 and Game 2")
 
@@
 
-```{r}
-# Plot Invs and Rets in absolute values for pre and post
-ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=roundNum)) + 
-  stat_summary(aes(y=returns, color="Returns"), fun = "mean", geom = "line", group = 1) +
-  stat_summary(aes(y=investment, color="Investment"), fun = "mean", geom = "line", group = 1) +
-  facet_wrap(~gameNum.f) +
-  scale_color_manual(values = c("Returns" = "red", "Investment" = "blue")) +
-  labs(color = "") +
-  xlab("Round Number")+
-  theme_bw() +
-  theme(legend.position = "bottom",  # Move legend to bottom
-        legend.title.align = 0.5,  # Center-align the legend title
-        legend.key = element_blank(),
-        plot.title = element_text(hjust = 0.5),  # Center-align the plot title
-        axis.text = element_text(color = "black"),  # Customize axis text
-        axis.title = element_text(color = "black")) + # Customize axis title
-  facet_wrap(~condition.f*gameNum.f)
-
-# # Plot Invs and Rets in percentages for pre and post
-# ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")),aes(x=roundNum)) + 
-#   stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=ret_pct_0)) + 
-#   stat_summary(fun = "mean", geom = "line",group=1, color = "blue", alpha =0.6,position = position_dodge2(width=0.8),aes(y=inv_pct)) + 
-#   geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
-#   facet_wrap(~gameNum.f)
-
-ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=roundNum)) + 
-  stat_summary(aes(y=ret_pct_0, color="Returns"), fun = "mean", geom = "line", group = 1) +
-  stat_summary(aes(y=inv_pct, color="Investment"), fun = "mean", geom = "line", group = 1) +
-  geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
-  facet_wrap(~gameNum.f) +
-  scale_color_manual(values = c("Returns" = "red", "Investment" = "blue")) +
-  labs(color = "") +
-  xlab("Round Number")+
-  ylab("Investment / Returns (%)")+
-  theme_bw() +
-  theme(legend.position = "bottom",  # Move legend to bottom
-        legend.title.align = 0.5,  # Center-align the legend title
-        plot.title = element_text(hjust = 0.5),  # Center-align the plot title
-        axis.text = element_text(color = "black"),  # Customize axis text
-        axis.title = element_text(color = "black"))+ # Customize axis title
-  facet_wrap(~condition.f*gameNum.f)
-
-```
-
-```{r}
-# Plot Invs and Rets in absolute values for exposure
-ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")),aes(x=roundNum)) + 
-  stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=returns)) + 
-  #stat_summary(fun.data = "mean_ci", geom = "errorbar",alpha =0.4, width = 0.2, aes(y=ret_pct) ) + 
-  stat_summary(fun = "mean", geom = "line",group=1, color = "blue",aes(y=investment)) + 
-  facet_wrap(~condition.f*gameNum.f)
-
-# Plot Invs and Rets in percentages for exposure
-ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")),aes(x=roundNum)) + 
-  stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=ret_pct_0)) + 
-  #stat_summary(fun.data = "mean_ci", geom = "errorbar",alpha =0.4, width = 0.2, aes(y=ret_pct) ) + 
-  stat_summary(fun = "mean", geom = "line",group=1, color = "blue", alpha =0.6,position = position_dodge2(width=0.8),aes(y=inv_pct)) + 
-  geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
-  facet_wrap(~condition.f*gameNum.f)
-
-```
-
 <!-- Linear mixed effects model for all returns  -->
@@
 
-## Analysis of participant returns
+## Analysis of Participant Returns
 
@@
 
-<!---------------- TRYING A MODEL WITH LATENT HMM STATE AS COVARIATE INSTEAD OF INVESTMENT  ------------------->
-
-```{r}
-mod_returns_latent <- mixed( ret_pct_na ~ phase.f*condition.f*investorState.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
-
-summary(mod_returns_latent)
-```
-
-```{r, include= F}
-# Load necessary packages
-library(lme4)
-library(emmeans)
-library(ggplot2)
-library(dplyr)
-
-# Assuming `mod_returns_latent` is your fitted model and `prepost_df` is your data frame
-# Calculate estimated marginal means
-emm <- emmeans(mod_returns_latent, ~ high_RS*condition.f*phase.f   | investorState.f, adjust = "tukey")
-
-# Calculate pairwise comparisons
-contrast_results <- contrast(emm, method = "pairwise")
-
-# Extract significant contrasts
-significant_contrasts <- summary(contrast_results, infer = c(TRUE, TRUE))
-
-# Convert emmeans object to data frame for plotting
-emm_df <- as.data.frame(emm)
-
-# Plot the marginal means using ggplot2
-p <- ggplot(emm_df, aes(x = investorState.f, y = emmean, color = high_RS, group = high_RS)) +
-  geom_point(position = position_dodge(width = 0.5), size = 3) +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(width = 0.5)) +
-  geom_line(position = position_dodge(width = 0.5), aes(linetype = high_RS)) +
-  labs(title = "Marginal Means of Returns by Investor State and RS Group",
-       x = "Investor State",
-       y = "Estimated Marginal Mean of Returns") +
-  theme_minimal() +
-  scale_color_manual(name = "RS Group", values = c("low RS" = "blue", "high RS" = "red"))
-
-# Add significance annotations
-for(i in 1:nrow(significant_contrasts)) {
-  row <- significant_contrasts[i, ]
-  if(row$p.value < 0.05) {
-    p <- p + annotate("text", 
-                      label = "*", 
-                      x = which(levels(emm_df$investorState.f) == row$investorState.f),
-                      y = max(emm_df$emmean) + 0.05 * i,
-                      color = "black", 
-                      size = 5, 
-                      vjust = -0.5)
-  }
-}
-
-# Print the plot
-print(p)
-
-```
-```{r}
-# Load necessary packages
-library(lme4)
-library(emmeans)
-library(ggplot2)
-library(dplyr)
-library(ggsignif)
-
-# Assuming `mod_returns_latent` is your fitted model and `prepost_df` is your data frame
-# Calculate estimated marginal means
-emm <- emmeans(mod_returns_latent, ~  high_RS | investorState.f  )
-
-# Calculate pairwise comparisons
-contrast_results <- pairs(emm)
-
-print(contrast_results)
-
-# Convert emmeans object to data frame for plotting
-emm_df <- as.data.frame(emm)
-
-# Convert contrast results to data frame for annotation
-contrast_df <- as.data.frame(contrast_results)
-significant_contrasts <- contrast_df %>%
-  filter(p.value < 0.05) %>%
-  mutate(significance = ifelse(p.value < 0.001, "***", ifelse(p.value < 0.01, "**", "*")))
-
-print(significant_contrasts)
-
-# Plot the marginal means using ggplot2
-p <- ggplot(emm_df, aes(x = investorState.f, y = emmean, color = high_RS, group = high_RS)) +
-  geom_point(position = position_dodge(width = 0.8), size = 3) +
-  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2, position = position_dodge(width = 0.8)) +
-  geom_line(position = position_dodge(width = 0.8), aes(linetype = high_RS)) +
-  labs(title = "Marginal Means of Returns by Investor State and RS Group",
-       x = "Investor State",
-       y = "Estimated Marginal Mean of Returns") +
-  theme_minimal() +
-  scale_color_manual(name = "RS Group", values = c("low RS" = "blue", "high RS" = "red")) +
-  theme(plot.title = element_text(size = 16), legend.position = "bottom")
-
-# Print the plot
-print(p)
-
-```
-
 On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60% of the endowment for investments and 35-50% of the total yield for returns, as reported in previous studies [@charness_investment_2008; @fiedler_social_2011].
 
-```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received (y-axis) by Condition, Phase, and game round (x-axis).
+```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received (y-axis) by Condition, Phase, and game round (x-axis) averaged across RS groups.
 The blue line shows the returns in the Pre phase and the green line those in the Post phase.
@@
 The dotted lines identify the rounds where the pre-programmed one-off low investment occurs.
-We note lower average returns post vs pre in the manipulation condition, whilst returns in the control condition are similar between the two phases.",fig.align="center", fig.width=6, fig.height = 4}
-
-# Create a data frame for the vertical linesr
+We note lower average returns post vs pre in the Manipulation condition, whilst returns in the Control condition are similar between the two phases.",fig.align="center", fig.width=6, fig.height = 4}
+
+# Create a data frame for the vertical lines
 vline_data <- data.frame(
@@
 post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
-As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the manipulation condition, pre - post,
-`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.
+As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the Manipulation condition, pre - post,
+`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the Control condition.
 We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.
@@
 such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, reflected  that returns were more affected by investments in the control condition.
-We also find a three way interaction between Phase,  Investment and RS
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation).
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, reflected that returns were more affected by investments in the Control condition.
+We also find a three way interaction between Phase, Investment and RS, `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_inv_scaled_high_RS)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation).
 Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
-For completeness, these effects are discussed in more detail the supplement.
+For completeness, these effects are discussed in more detail in the supplement.
 
@@
 ```
-
-<!-- To interpret this interaction we plot the effect of the Investment on returns for each level of Condition, Phase and RS.
-Figure \ref{fig:fourwayInteractionPlot} shows that while the effect of investment on returns is similar across RS groups and across phases in the Control condition, the Manipulation condition shows a higher effect of investment in the post phase compared to pre in the low RS group, but a lower effect of investment compared to before for the high RS group.
-This suggests that high RS participants based their return less on the investment after being exposed to the forgiving AI compared to the low RS group.
--->
 
@@
 
-Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor?.
+Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor?
 To explore this question, we restrict the analysis to the trials following the pre-programmed defection by the HMM agent in both the "pre" (trials 12 to 15) and the "post" phases (trials 13 to 15).
@@
 Finally, we still find an Investment by Condition interaction 
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the manipulation condition compared to the control condition in post-defection trials.
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the Manipulation condition compared to the Control condition in post-defection trials.
 However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.
@@
 
-```{r}
-library(ggplot2)
-
-# Plotting
-ggplot(prepost_df %>% filter(condition.f=="Manipulation", high_RS == "high RS", roundNum < 12), aes(x = roundNum, y = investment, group = playerId, color = as.factor(gameNum.f))) +
-  geom_line(alpha = 0.5) +  # Set alpha for transparency to handle overlap
-  facet_wrap(~ playerId, scales = 'free_y') + # Faceting by playerID, each gets its own y-axis scale
-  theme_minimal() +
-  labs(color = 'Game Number') +
-  theme(
-    strip.text = element_blank(), # Remove facet labels since there will be too many
-    axis.text.x = element_text(angle = 90, hjust = 1) # Rotate x-axis labels for readability
-  )
-
-```
-
-### HMM investor
+### HMM Investor in Pre and Post Phases
 <!-- Linear mixed effects model for investments -->
-To test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
-As seen in Figure \@ref(fig:boxPlots), we find no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.
-
-When asked during debrief whether they thought the investors they faced were Human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player.
+Was the HMM's strategy similar between pre and post phases in the control condition?
+Was participants' behavior post exposure differentiated enough to induce a different reaction from the HMM?
+To answer these questions, we test for differences in the HMM agent's investment by Phase, Condition and RS using a linear mixed-effects model as described in the methods section.
+As seen in Figure \@ref(fig:boxPlots), we find no main or interaction effects, indicating the HMM's behavior was on aggregate similar across levels of Phase, Condition and RS.
+This consistency in the investor's behavior is a desirable feature of the HMM agent when the participants behavior is largely simialr between phases.
+More importantly, it indicates that the lower returns of participants in the post phase of the manipulation condition were not differentiated enough to make the HMM react by transitioning to lower latent trust states.
+It is also noteworthy that the HMM agent was relatively successful in imitating human behavior in this paradigm: When asked during debrief whether they thought the investors they faced were human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player.
 When asked to justify their choice, many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.
 
-### Questionnaire scores and performance
-
-<!-- We found no correlation between either the RS score and task performance, as defined by the total points accumulated in both the "pre" and "post phase of the RTG.
-The LPFS score was also not correlated with task performance.
-These two scores exhibited a positive correlation  (Rho = 0.52, p < 0.001) -->
-
-```{r}
+### Exposure Phase Trials
+
+```{r}
+ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
+  stat_summary(fun = mean, geom = "line") +
+  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
+               alpha = 0.3, linetype = 0) +
+  # geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
+  # scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
+  labs(x = "Round",
+       y = "Percentage Return",
+       color = "Trust Game Phase") +
+  theme_bw() +
+  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
+  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
+         linetype = guide_legend(title = "Defection round")) +
+  facet_wrap(~condition.f)
+```
+
+```{r}
+
+expo_df <- avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3"))
+
+mod_returns_expo <- mixed( ret_pct_na ~ condition.f*inv_scaled*high_RS + (1 | playerId), expo_df, REML= TRUE, method="KR")
+mod_inv_summary <- summary(mod_returns_expo)
+mod_inv_summary
+
+em_ret_cond <- emmeans::emmeans(mod_returns_expo , c("condition.f"))
+pairs(em_ret_cond)
+
+# Calculate the slopes (coefficients) of inv_scaled for each condition
+inv_slopes <- emtrends(mod_returns_expo, ~ condition.f, var = "inv_scaled")
+
+# To see pairwise comparisons of the slopes
+print(inv_slopes)
+pairs(inv_slopes)
+
+##### Three way interaction
+inv_slopes3 <- emtrends(mod_returns_expo, ~ condition.f * high_RS, var = "inv_scaled")
+summary(inv_slopes3)
+
+pairwise_comparisons3 <- pairs(inv_slopes3, by = "high_RS")
+summary(pairwise_comparisons3)
+
+######### HMM investment in EXPO PHASE 
+mod_invs_expo <- mixed(investment ~ condition.f*high_RS + (1 | playerId), expo_df %>% filter(roundNum >= 5), REML= TRUE, method="KR")
+summary(mod_invs_expo)
+
+emcheck_inv <- pairs(emmeans::emmeans(mod_invs_expo, c("condition.f"), pbkrtest.limit = 6300))
+emcheck_inv
+```
+So far we focused on analysing behavior for the pre and post phases.
+Here, we look at returns and investments in the exposure phase.
+The linear mixed effects model of participants' returns in the exposure phase does not show a main effect of Condition on returns.
+There was a main effect of Investment, `r papaja::apa_print(mod_returns_expo)$full_result$inv_scaled`, with participants positively reciprocating higher investments, an interaction effect between Condition and Investment `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2",papaja::apa_print(mod_returns_expo)$full_result$condition_f_inv_scaled)`, showing a stronger positive reciprocity in the Control condition, and finally a three way interaction between the RS group, Condition and Investment `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2",papaja::apa_print(mod_returns_expo)$full_result$condition_f_inv_scaled_high_RS)`, showing that this stronger positive reciprocity to investment in the Control condition is higher for participants with high RS.
+The linear mixed effects model of the HMM investments shows a main effect of Condition `r papaja::apa_print(mod_invs_expo)$full_result$condition_f`, suggesting higher overall investments for the forgiving HMM  compared to the human-like HMM, but no difference in investments when facing low and high RS groups.
+
+In summary, despite the forgiving HMM sending overall higher investments in the exposure phase, participants returned similar proportions of the multiplied investments as those facing the human-like HMM.
+The positive reciprocity of returns to investments was higher in the Control condition with this relationship stronger for the high RS group.
+
+### Questionnaire Scores and Performance
+
+```{r}
+
+temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
+  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f, RS_score, LPFS_score) %>%
+  mutate(perf = payoffTrust1 + payoffTrust2) %>% 
+  unique()
 cor.test(temp2$LPFS_score, temp2$perf, method = "spearman")
@@
 
-Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity score (RSQ), Spearman's `r spearman_rho`, $p < 0.001$, there was no correlation between these questionnaire scores and participant's return or overall task performance.
+Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity  Questionnaire score (RSQ), Spearman's `r spearman_rho`, $p < 0.001$, there was no correlation between these questionnaire scores and participant's return or overall task performance.
 
@@
 
-## Player ratings
+## Player Ratings
 
@@
 
-### Comparing pre and exposure ratings
+### Comparing Pre and Exposure Ratings
 
@@
 There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
-Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
+Those in the Control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
 
@@
 Cooperation, forgiveness, and willingness to play again ratings remained relatively stable for the low RS group except post-manipulation, where they were lower.
-In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the manipulation condition, but decreased over time in the control group,.
+In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the Manipulation condition, but decreased over time in the Control group.
 Ratings were also markedly lower in the post-manipulation game for this group.",fig.align="center", fig.width=6, fig.height = 6}
@@
 
-### Comparing pre and post ratings
-<!-- Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
+### Comparing Pre and Post Ratings
+<!-- Comparing post-exposure to exposure ratings, participants in the Manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
 <!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
-Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
+Ratings for those in the Control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
 -->
@@
 Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
-(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
+(Cooperation: `r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
 Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
@@
 
-In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the control condition.
-In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the control condition even though the player continues to use the same strategy.
+In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the Control condition.
+In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the Control condition even though the player continues to use the same strategy.
 After exposure to the forgiving HMM, both groups rate the "post" co-player worse than the "pre" even though they are the same.
@@
 
-We used a randomized control online experiment where participants played a RTG with AI agents designed to simulate human-like trust-building scenarios.
+We used a randomized controlled online experiment where participants played a RTG with AI agents designed to simulate human-like trust-building scenarios.
 Participants were then exposed to either forgiving or human-like AI agents before playing another RTG.
 Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation.
-Instead, participants reduced their returns overall whilst the returns of those in the control group did not change between the pre and post phase of the experiment.
-Neither did participants show more forgiving behavior in the manipulation condition after the one-off defection by the agent.
+Instead, participants reduced their returns overall whilst the returns of those in the Control group did not change between the pre and post phase of the experiment.
+Neither did participants show more forgiving behavior in the Manipulation condition after the one-off defection by the agent.
 Why did participants reduce their returns even though they were repeatedly exposed to a more cooperative and more forgiving AI?
-A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the manipulation condition.
+A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the Manipulation condition.
 
@@
 
-<!-- Participants also perceived the forgiving AI during the exposure phase as more cooperative, but not more forgiving.
-The latter might be the result of the participants not having many opportunities to test the co-player's forgiveness propensity since most decided to continue cooperating.
--->
-
 It is worth highlighting that RS was not correlated to overall performance in the task.
 Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
-However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their AI co-players.
-Specifically, when exposed to a more forgiving AI agent, these participants accurately increased their ratings of the agent, indicating a sensitive and appropriate response to the behavioral manipulation.
-This adjustment reflects a nuanced perception of social cues and a capacity to modify judgments based on the behavior of interaction partners, potentially indicating a perceived alignment or support from the co-player that mitigates their heightened sensitivity to potential social rebuffs.
-High RS participants' lower ratings of essentially similar co-players as they continued to face them may indicate that the occasional pre-programmed defections of the AI agents failed to mitigate concerns over rejection and possibly exacerbated perceptions of uncooperativeness and unforgiveness, leading to a stronger preference against future interactions.
+However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their AI co-players: They accurately perceived increased cooperativeness in the forgiving AI agent, showing an ability to adjust their judgments based on interaction patterns.
+This may indicate that perceived support from cooperative partners can mitigate their rejection concerns.
+However, their declining ratings of similar co-players over time, despite consistent AI behavior, suggests that occasional pre-programmed defections may have amplified their perceptions of untrustworthiness and reluctance for future interactions.
+These nuanced perceptions align with the heightened social awareness often associated with high rejection sensitivity [@downey_implications_1996].
 Individuals with low RS were less affected by AI behavioral variations, showing a less volatile baseline of social perception.
-They still exhibit behavior consistent with a contrast effect in the manipulation condition.
-This dichotomy highlights the potential of tailored interactions, mediated by advanced AI agents, to address and modulate specific social and psychological predispositions in human participants.
-
-<!-- Interestingly, for participants with low RS, the manipulation did not significantly alter their perceptions of the AI's cooperativeness, forgiveness, or their willingness to interact in the future.
-This group's ratings remained largely unchanged between the pre- and post-exposure phases, regardless of the AI's behavior.
-This stability suggests that individuals with lower sensitivity to rejection may base their ratings more on consistent behavioral patterns or other criteria not directly influenced by the forgiving nature of the AI agents.
--->
-
-This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of generative models of human behavior to design artificial agents that can interact in economic games.
+They still exhibit behavior consistent with a contrast effect in the Manipulation condition.
+This stability in perception might reflect a lower propensity to scrutinize social cues or a more robust internal model of social interactions that is less easily perturbed by short-term variations in partner behavior.
+
+Despite these differences in perception, the actual cooperative behavior – as measured by returns in the trust game – did not significantly differ between RS groups.
+This suggests a potential dissociation between the cognitive-perceptual processes involved in evaluating social partners and the behavioral output in economic games.
+There are multiple potential explanations for this: First, the ratings may reflect explicit, conscious evaluations of social partners, while behavior in the trust game might be guided more by implicit, automatic processes that are less differentiated between RS groups.
+This interpretation aligns with dual-process theories of social cognition [@lieberman_social_2007].
+Second, the structured nature of the trust game might provide a context where the behavioral expression of rejection sensitivity is attenuated.
+In contrast, the more open-ended nature of providing ratings might allow for greater expression of RS-related perceptual biases.
+
+These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged.
+Conversely, therapies that target these perceptual processes might not necessarily translate into immediate behavioral changes.
+For future research, it would be valuable to explore whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods.
+Additionally, investigating whether there are specific conditions or thresholds under which perceptual differences do begin to manifest in behavioral divergences could provide important insights into the dynamics of rejection sensitivity.
+These findings also suggest that simply exposing individuals to highly cooperative or forgiving partners may not be sufficient to induce lasting changes in behavior.
+Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time, perhaps by combining periods of forgiveness with clear communication about mutual expectations for cooperation [@balliet_reward_2011].
+Future research could explore the temporal dynamics of these effects, investigating whether longer exposure periods or different patterns of forgiving behavior might lead to more positive outcomes.
+Additionally, examining the role of explicit communication and meta-cognitive reflection on the cooperative process could provide insights into how to more effectively leverage positive interactions to promote lasting cooperative tendencies.
+
+<!-- This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of generative models of human behavior to design artificial agents that can interact in economic games.
 The use of these agents in the RTG led to similar investment and returns to those recorded in human dyadic interactions.
@@
 This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control.
-Importantly, our approach utilizing HMMs to simulate interactive partners in social dilemmas offers a promising avenue for future research, especially in understanding and potentially mitigating trust deficits in individuals with high RS.
-The nuanced behaviors elicited through the interaction with HMM agents highlight the complexity of trust dynamics and the potential for computational models to offer novel insights into human social behaviors.
-
-<!-- In conclusion, this study demonstrates the value of integrating HMM-based artificial agents in the study of trust and cooperation.
-By showcasing the potential of these agents to both simulate realistic human interactions and influence real human behavior, our research paves the way for future investigations into the mechanisms of trust repair and the broader implications of artificial agents in social psychology.
 -->
 
-<!-- ## Other notable points: -->
-
-<!-- The study found that RSQ (RS Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game.
-This gap highlights a need for further research, despite theoretical connections suggested in the literature review.
--->
+# Conclusion
+
+This randomised controlled experiment enabled us to uncover unexpected effects of exposure to forgiving behavior on subsequent cooperation, particularly in relation to rejection sensitivity.
+These findings challenge existing assumptions about fostering cooperative behavior and suggest the need for more nuanced interventions.
+Importantly, the use of HMM-based AI agents in this study represents a significant methodological advancement.
+By providing a balance between experimental control and realistic, adaptive behavior, these agents allowed for a nuanced exploration of trust dynamics that would be challenging to achieve with human confederates or simplistic computer algorithms.
+This approach opens up new possibilities for studying complex social interactions in controlled settings, potentially bridging the gap between laboratory experiments and real-world social dynamics.
 

### 2024-07-03 supplement.Rmd
improved intro and discussion

@@
 ---
-title: "Can Exposure To Forgiving AI Foster Cooperative Play?"
-subtitle: "Supplementary Information"
+title: "Supplementary Information"
+subtitle: ""
 
@@
 
-We estimated a three-state model for investor's behaviour, using maximum likelihood estimation via the Expectation-Maximisation algorithm as implemented in the depmixS4 (Visser & Speekenbrink) package for R.
+We estimated a three-state model for investor's behaviour, using maximum likelihood estimation via the Expectation-Maximisation algorithm as implemented in the depmixS4 package for R [@visser_depmixs4_2021].
 The model was estimated using investments from existing datasets of human dyads playing 10 rounds of the RTG with the same trustee.
 The dataset consisted of a total of 381 games from two data sources: First, a total of 93 repeated trust games with healthy investors and a mix of healthy trustees and trustees diagnosed with Borderline Personality Disorder (BPD) [@king-casas_rupture_2008].
-The second source was from data collected as part of a project investigating social exchanges at the Montague lab at Virginia Tech and consists of 288 games.
+The second source was from data collected as part of a project investigating social exchanges in BPD and antisocial personality disorder reported on elsewhere [@euler_interpersonal_2021; @huang_multidirectional_2020; @rifkin-zybutz_impaired_2021] and consists of 288 games.
+In both datasets, the investor on which we modelled the HMM's strategy was always selected from a healthy population and the trustees were a mix of healthy participants and those with personality disorders allowing for a diversified interaction behavior.
 
-## Mixed-effects model results for participant returns
+## Mixed-effects models for participant returns
 

### 2024-07-09 article.Rmd
version 3 finalised

@@
 ---
-title: "When Forgiveness Backfires: an AI-Mediated Randomised Controlled Experiment on Rejection Sensitivity and Cooperation"
+title: "When Forgiveness Backfires: Rejection Sensitivity and Cooperative Behavior Following Exposure to Adaptive Forgiving Agents"
+
 author:
-  # "Ismail Guennouni, Georgia Koppe, Christoph Korn"
-
-keywords:
-- Social Trust
-- Repair of Cooperation
-- Cognitive Manipulation 
-- Economic Games 
-- Hidden Markov Models
+  "Ismail Guennouni^1^*, Georgia Koppe^2^, Christoph Korn^3^"
 
 abstract: |
- Can exposure to forgiving behavior improve cooperative outcomes for individuals prone to rejection sensitivity?
-This randomized controlled experiment explores this question using a novel approach with Hidden Markov Model (HMM)-based AI agents in repeated trust games (RTGs).
-Participants (N = 206), pre-screened for high or low rejection sensitivity (RS), interacted with either consistently behaving or more forgiving co-players, which were in reality HMM agents.
+  Can exposure to forgiving behavior improve cooperative outcomes, and does Rejection Sensitivity (RS) moderate this effect?
+This randomized controlled experiment explores these questions using a novel approach with Hidden Markov Model (HMM)-based artificial agents in repeated trust games (RTGs).
+Participants (N = 206), pre-screened for high or low RS, interacted with either consistently behaving or more forgiving co-players, which were in reality HMM agents.
 Unexpectedly, exposure to forgiving agents led to reduced cooperation in subsequent interactions.
@@
 RS levels influenced perceptions of the agent's cooperativeness but did not significantly affect behavioral outcomes, revealing a perception-behavior dissociation.
-These findings challenge assumptions about fostering cooperation through simple exposure to positive interactions.
-They also demonstrate the potential of HMM-based AI agents for studying interactive social dynamics, offering a methodological advancement for future research.
+These findings challenge assumptions about fostering cooperation through simple exposure to positive interactions and highlight the importance of accounting for negative contrast effects when designing such interventions.
+They also demonstrate the potential of HMM-based artificial agents for studying interactive social dynamics, offering a methodological advancement for future research.
 Our results suggest the need for nuanced interventions that gradually shape expectations and behaviors to foster cooperative outcomes, particularly for individuals with high rejection sensitivity.
@@
 biblio-style: spphys
-
 output:
   bookdown::pdf_document2:
+    keep_tex: true
     toc: false
@@
 ---
+\small
+^1^ *Central Institute of Mental Health, Medical Faculty Mannheim/Heidelberg University, Germany*
+
+^2^ *Faculty of Mathematics and Computer Science, Heidelberg University, Germany*
+
+^3^ *Department of General Psychiatry, Section Social Neuroscience, Heidelberg University, Germany*
+
+^\*^ *Corresponding author. Address: Central institute of Mental Health, J5, Mannheim, Germany. Email: ismail.guennouni@zi-mannheim.de* <br>
+
+\vspace{1cm}
+
+**Keywords** : Social Trust; Forgiveness Intervention; Rejection Sensitivity; Interpersonal Dynamics; Hidden Markov Models
+
+\vspace{1cm}
+
+**Highlights:**
+
+- Exposure to forgiving artificial agents unexpectedly led to reduced cooperation in subsequent interactions, challenging assumptions about fostering trust through positive experiences
+
+- Rejection sensitivity influenced perceptions of artificial agents' cooperativeness but not behavioral outcomes, revealing a dissociation between perception and behavior
+
+- Novel Hidden Markov Model-based artificial agents provide adaptive, human-like behavior in trust games, allowing for precise control of "cooperative" and "forgiving" strategies while maintaining experimental control of co-player strategies
+
+\pagebreak
 
@@
 The interaction of rejection sensitivity and trust-based constructs has been explored, particularly in Borderline Personality Disorder (BPD).
-@miano_rejection_2013 and @richetin_emotional_2018 found that rejection sensitivity mediates the relationship between BPD features and lower trust appraisal.
+@miano_rejection_2013 and @richetin_emotional_2018 found that rejection sensitivity mediated the relationship between BPD features and lower trust appraisal.
+@abramov_influence_2022-1 found that higher baseline feelings of rejection in individuals with BPD predict slower trust formation and less pronounced declines in trust following trust violations during the trust game.
 However, the interaction between trustworthiness and rejection sensitivity hasn't been studied as extensively in other personality disorders or in non-clinical populations, leaving a gap in our understanding of how these constructs might interplay.
@@
 
-A total of 206 participants (56% female) were recruited on the Prolific Academic platform (prolific.co).
-The required sample size was determined based on an a priori power analysis.
+To have participants with large differences in RS, a total of 1195 participants were pre-screened on the Prolific Academic platform (prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10) totalling 206 participants (56% female).
+These were then invited through prolific to take part in the main experiment study.
+The required sample size was determined using an a priori power analysis to have an 80% probability to detect a small effect size with a 5% type I error rate in a repeated measures ANOVA with between factors.
 The mean age of participants was 34.6 years, with an 11.9 years standard deviation.
 Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.
-A total of 1195 participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10).
 
@@
 
-The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : high or low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A).
+The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : High or Low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A).
 Participants within each pre-screened group were randomly assigned to one of the two levels of the Condition factor.
@@
 Participants were not told they were facing computerised co-players.
+We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a varying time lapse (randomly chosen between 5 and 10 seconds).
 
@@
 After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-playerr to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
-After completing the three game phases, participants then completed the Levels of Personality Functioning Scale (LPFS) questionnaire (see the supplement for details).
+After completing the three game phases, participants then completed the Levels of Personality Functioning Scale Brief-Form (LPFS-BF) questionnaire [@weekers_level_2019].
+This is a self-report measure designed to assess core elements of personality functioning as defined in the Alternative Model for Personality Disorders in the DSM-5 [@american_psychiatric_association_diagnostic_2013], and provides a dimensional assessment of personality functioning, which complements the categorical approach of RS.
 Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.
@@
 summary(mod_invs_RS)
-
+```
+
+```{r}
+saveRDS(summary(mod_invs_RS), "data/mod_invs_RS.RDS")
 ```
@@
 
-We used a randomized controlled online experiment where participants played a RTG with AI agents designed to simulate human-like trust-building scenarios.
-Participants were then exposed to either forgiving or human-like AI agents before playing another RTG.
+We used a randomized controlled online experiment where participants played a RTG with artificial agents designed to simulate human-like trust-building scenarios.
+Participants were then exposed to either forgiving or human-like HMM agents before playing another RTG.
 Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation.
@@
 Neither did participants show more forgiving behavior in the Manipulation condition after the one-off defection by the agent.
-Why did participants reduce their returns even though they were repeatedly exposed to a more cooperative and more forgiving AI?
+Why did participants reduce their returns even though they were repeatedly exposed to more cooperative and more forgiving artificial agents?
 A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the Manipulation condition.
@@
 If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favourable or even negative by comparison [@kobre_negative_1972].
-As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving AI [@zentall_within-trial_2005].
+As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving agents [@zentall_within-trial_2005].
 If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.
@@
 Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
-However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their AI co-players: They accurately perceived increased cooperativeness in the forgiving AI agent, showing an ability to adjust their judgments based on interaction patterns.
+However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their HMM co-players: They accurately perceived increased cooperativeness in the forgiving agent, showing an ability to adjust their judgments based on interaction patterns.
 This may indicate that perceived support from cooperative partners can mitigate their rejection concerns.
-However, their declining ratings of similar co-players over time, despite consistent AI behavior, suggests that occasional pre-programmed defections may have amplified their perceptions of untrustworthiness and reluctance for future interactions.
+However, their declining ratings of similar co-players over time, despite consistent agent behavior, suggests that occasional pre-programmed defections may have amplified their perceptions of untrustworthiness and reluctance for future interactions.
 These nuanced perceptions align with the heightened social awareness often associated with high rejection sensitivity [@downey_implications_1996].
-Individuals with low RS were less affected by AI behavioral variations, showing a less volatile baseline of social perception.
+Individuals with low RS were less affected by the agent's behavioral variations, showing a less volatile baseline of social perception.
 They still exhibit behavior consistent with a contrast effect in the Manipulation condition.
@@
 These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged.
-Conversely, therapies that target these perceptual processes might not necessarily translate into immediate behavioral changes.
-For future research, it would be valuable to explore whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods.
-Additionally, investigating whether there are specific conditions or thresholds under which perceptual differences do begin to manifest in behavioral divergences could provide important insights into the dynamics of rejection sensitivity.
-These findings also suggest that simply exposing individuals to highly cooperative or forgiving partners may not be sufficient to induce lasting changes in behavior.
-Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time, perhaps by combining periods of forgiveness with clear communication about mutual expectations for cooperation [@balliet_reward_2011].
-Future research could explore the temporal dynamics of these effects, investigating whether longer exposure periods or different patterns of forgiving behavior might lead to more positive outcomes.
-Additionally, examining the role of explicit communication and meta-cognitive reflection on the cooperative process could provide insights into how to more effectively leverage positive interactions to promote lasting cooperative tendencies.
+Conversely, therapies targeting these perceptual processes might not necessarily translate into immediate behavioral changes.
+Future research should investigate whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods.
+Additionally, exploring specific conditions or thresholds under which perceptual differences begin to manifest in behavioral divergences could provide valuable insights into the dynamics of rejection sensitivity.
+Simply exposing individuals to highly cooperative or forgiving partners may not suffice to induce lasting changes in behavior.
+Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time.
+Combining periods of forgiveness with clear communication about mutual expectations for cooperation might be more effective [@balliet_reward_2011].
+Future studies should also consider the temporal dynamics of these effects, examining whether longer exposure periods or different patterns of forgiving behavior could lead to more positive outcomes.
+Moreover, investigating the role of explicit communication and meta-cognitive reflection on the cooperative process could offer insights into leveraging positive interactions to promote enduring cooperative tendencies.
 
@@
 These findings challenge existing assumptions about fostering cooperative behavior and suggest the need for more nuanced interventions.
-Importantly, the use of HMM-based AI agents in this study represents a significant methodological advancement.
+Importantly, the use of HMM-based artificial agents in this study represents a significant methodological advancement.
 By providing a balance between experimental control and realistic, adaptive behavior, these agents allowed for a nuanced exploration of trust dynamics that would be challenging to achieve with human confederates or simplistic computer algorithms.
@@
 
+# Author contributions statement {.unnumbered}
+I.
+Guennouni, G.
+Koppe and C.
+Korn.
+designed and developed the study concept.
+Experiment design, testing and data collection were performed by I.
+Guennouni.
+I.
+Guennouni analysed and interpreted the data under the supervision of G.
+Koppe and C.
+Korn.
+All authors jointly wrote and approved the final version of the manuscript for submission.
+
+# Funding {.unnumbered}
+This publication was supported through state funds approved by the State Parliament of Baden-Württemberg for the Innovation Campus Health + Life Science alliance Heidelberg Mannheim.
+I.
+Guennouni was supported by a research fellowship from the AI Health Innovation Cluster
+
+# Competing interests statement {.unnumbered}
+The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+# Additional information  {.unnumbered}
+
+## Correspondence {.unnumbered}
+
+All correspondence and requests for materials should be addressed to I. Guennouni.
+
+## Transparency and data availability {.unnumbered}
+
+Preregistration: The hypotheses and methods were not preregistered.The analysis plan was not preregistered.
+Materials: All study materials are publicly available (https://github.com/ismailg/exposure-public).
+Data: All primary data are publicly available (https://github.com/ismailg/exposure-public).
+Analysis scripts: All analysis scripts are publicly available (https://github.com/ismailg/exposure-public).
+
 # References

### 2024-07-09 article.tex
version 3 finalised

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lscape}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={When Forgiveness Backfires: Rejection Sensitivity and Cooperative Behavior Following Exposure to Adaptive Forgiving Agents},
  pdfauthor={Ismail Guennouni1*, Georgia Koppe2, Christoph Korn3},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{When Forgiveness Backfires: Rejection Sensitivity and Cooperative Behavior Following Exposure to Adaptive Forgiving Agents}
\author{Ismail Guennouni\textsuperscript{1}*, Georgia Koppe\textsuperscript{2}, Christoph Korn\textsuperscript{3}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Can exposure to forgiving behavior improve cooperative outcomes, and does Rejection Sensitivity (RS) moderate this effect? This randomized controlled experiment explores these questions using a novel approach with Hidden Markov Model (HMM)-based artificial agents in repeated trust games (RTGs). Participants (N = 206), pre-screened for high or low RS, interacted with either consistently behaving or more forgiving co-players, which were in reality HMM agents. Unexpectedly, exposure to forgiving agents led to reduced cooperation in subsequent interactions. Control group participants maintained consistent behavior. RS levels influenced perceptions of the agent's cooperativeness but did not significantly affect behavioral outcomes, revealing a perception-behavior dissociation. These findings challenge assumptions about fostering cooperation through simple exposure to positive interactions and highlight the importance of accounting for negative contrast effects when designing such interventions. They also demonstrate the potential of HMM-based artificial agents for studying interactive social dynamics, offering a methodological advancement for future research. Our results suggest the need for nuanced interventions that gradually shape expectations and behaviors to foster cooperative outcomes, particularly for individuals with high rejection sensitivity.
\end{abstract}

\small

\textsuperscript{1} \emph{Central Institute of Mental Health, Medical Faculty Mannheim/Heidelberg University, Germany}

\textsuperscript{2} \emph{Faculty of Mathematics and Computer Science, Heidelberg University, Germany}

\textsuperscript{3} \emph{Department of General Psychiatry, Section Social Neuroscience, Heidelberg University, Germany}

\textsuperscript{*} \emph{Corresponding author. Address: Central institute of Mental Health, J5, Mannheim, Germany. Email: \href{mailto:ismail.guennouni@zi-mannheim.de}{\nolinkurl{ismail.guennouni@zi-mannheim.de}}}

\vspace{1cm}

\textbf{Keywords} : Social Trust; Forgiveness Intervention; Rejection Sensitivity; Interpersonal Dynamics; Hidden Markov Models

\vspace{1cm}

\textbf{Highlights:}

\begin{itemize}
\item
  Exposure to forgiving artificial agents unexpectedly led to reduced cooperation in subsequent interactions, challenging assumptions about fostering trust through positive experiences
\item
  Rejection sensitivity influenced perceptions of artificial agents' cooperativeness but not behavioral outcomes, revealing a dissociation between perception and behavior
\item
  Novel Hidden Markov Model-based artificial agents provide adaptive, human-like behavior in trust games, allowing for precise control of ``cooperative'' and ``forgiving'' strategies while maintaining experimental control of co-player strategies
\end{itemize}

\pagebreak

\section{Introduction}\label{introduction}

Trust is fundamental to human social interactions, facilitating seamless relations at both interpersonal and intergroup levels. The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders (Fonagy and Campbell 2017). Individuals with personality disorders (PD) often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors -- a marker for the severity of PD symptoms (Herpertz and Bertsch 2014; Mulder et al. 1999).

One explanation for such social challenges lies in early caregiver experiences. Attachment theory (Bowlby 1978) suggests that the quality of these relationships shapes our capacity for secure attachments and trust. Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken (Mikulincer 1998). These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection (Downey and Feldman 1996; Downey, Khouri, and Feldman 1997). RS has been linked to the development of various mental health conditions, including depression, anxiety, personality disorders, and self-harm (Gao et al. 2017). A recent meta-analysis revealed prosocial behavior and interpersonal trust as two key processes of interpersonal functioning that are markedly impaired in PDs and which are likely to contribute to interpersonal dysfunction in this population (Hepp and Niedtfeld 2022). The interaction of rejection sensitivity and trust-based constructs has been explored, particularly in Borderline Personality Disorder (BPD). Miano et al. (2013) and Richetin et al. (2018) found that rejection sensitivity mediated the relationship between BPD features and lower trust appraisal. Abramov et al. (2022) found that higher baseline feelings of rejection in individuals with BPD predict slower trust formation and less pronounced declines in trust following trust violations during the trust game. However, the interaction between trustworthiness and rejection sensitivity hasn't been studied as extensively in other personality disorders or in non-clinical populations, leaving a gap in our understanding of how these constructs might interplay. Could an ingrained sensitivity to rejection contribute to reduced pro-social behavior such as a failure to show trustworthiness?

If indeed RS contributes to interpersonal dysfunction through failure to show trustworthiness, could exposing those who exhibit it to cooperative and forgiving interaction partners lead to improved interpersonal outcomes? Research in the fields of behavioral economics and psychology has explored how positive social interactions influence cooperation and pro-social behavior. The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions (Joyce, Dickhaut, and McCabe 1995). In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors even among individuals who are initially inclined to be distrustful (King-Casas et al. 2005). Fowler and Christakis (2010) studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior. Similarly, research on social learning theory (Bandura 1977) has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior. These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust and potentially fostering trustworthy behavior.

In this study, we use a randomized controlled online experiment to test whether exposing participants with varying RS levels to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation. To simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players' data. A key aspect of these agents is that their actions depend on a latent ``trust state'' which reacts dynamically to the trustees' returns, simulating real-life trust-building scenarios. An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively, or the propensity to retaliate after breakdowns of cooperation.

We pre-screened participants for high or low rejection sensitivity (RS) using a validated questionnaire, then assigned them exclusively to the trustee role in a series of trust games. After playing a 15-round RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition. In the Manipulation condition, participants were exposed over three RTGs to HMM investors designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust. In the Control condition, participants played three RTGs against the same human-like HMM. After this exposure phase, all participants played another 15-round RTG with a human-like HMM investor. We expect those exhibiting high RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity. We also hypothesize that those in the Manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase, with those in the higher RS group showing a bigger effect.

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

To have participants with large differences in RS, a total of 1195 participants were pre-screened on the Prolific Academic platform (prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score \textgreater{} 15) and the other with low RS (RSQ score \textless{} 10) totalling 206 participants (56\% female). These were then invited through prolific to take part in the main experiment study. The required sample size was determined using an a priori power analysis to have an 80\% probability to detect a small effect size with a 5\% type I error rate in a repeated measures ANOVA with between factors. The mean age of participants was 34.6 years, with an 11.9 years standard deviation. Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.

\subsection{Design and Procedure}\label{design-and-procedure}

The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : High or Low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \ref{fig:HMMPanels}.A). Participants within each pre-screened group were randomly assigned to one of the two levels of the Condition factor. The games were designed and implemented online using Empirica v1 (Almaatouq et al. 2021). The planned experiment received approval from the University of Heidelberg's Medical Faculty ethics commission (ID:S-708/2023) and the experiment was performed in accordance with the ethics board guidelines and regulations. All participants provided informed consent prior to their participation.

\subsection{Tasks and Measures}\label{tasks-and-measures}

\subsubsection{Repeated Trust Game}\label{repeated-trust-game}

Participants played a 15-round RTG (Joyce, Dickhaut, and McCabe 1995) in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right. The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised co-players. We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a varying time lapse (randomly chosen between 5 and 10 seconds).

The computerised investor consisted of a hidden Markov model (HMM) trained on an independent existing behavioral RTG data set of human investors. This data-driven approach thus sought to learn an investor strategy that mimics human-like interactions. The data set used for training consists of 388 ten round games with the same player (full details can be found in the Supplementary Information). On this data set, the HMM was inferred with three latent states that could be interpreted as reflecting a ``low-trust'', a ``medium-trust'', and a ``high-trust'' state. A separate output distribution, that maps each HMM state onto possible investments from 0 to 20 separately, is learned (Figure \ref{fig:HMMPanels}.B). In analogy to the latent states, these distributions can be interpreted as reflecting ``low-trust'', ``medium-trust'', or ``high-trust'' dispositions. Finally, the HMM is specified by transition probabilities that describe the transition between states. The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.C)). The initial state for the HMM investor in each instance of the game was set to the ``mid-trust'' state. Details on how the HMM state conditional probabilities and transition functions are specified can be found in the supplement.

In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair it, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) and round 13 (post-manipulation). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).

\subsection{Manipulation}\label{manipulation}

In all phases of the RTG other than the `Exposure phase' (Figure \ref{fig:HMMPanels}.A), participants interacted with this human-like HMM. In the `Manipulation' Condition of the exposure phase, however, the parameters of this HMM were adjusted to design a `forgiving' and ultimately more cooperative agent. To achieve this, we changed the state transition probabilities of the HMM such that it becomes impossible for it to remain in a low trust state, effectively setting the transition probability for remaining in a ``low-trust'' state to 0. The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D. The policies conditional on the latent states and the transition function in the other latent states remain unchanged.

\begin{landscape}
\begin{figure}

{\centering \includegraphics[width=1\linewidth,height=1\textheight]{article_files/figure-latex/HMMPanels-1}

}

\caption{\small{A: A timeline of the experiment for both conditions. The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor. The difference between conditions is the type of agents participants are exposed to in the exposure phase. Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG against human trustees. From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C. Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1. Panel D shows the transition probabilities of the forgiving HMM agent in the low trust state. Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state, as a proxy for coaxing behavior. Transitions in the medium and high trust states were identical for both agents.}}\label{fig:HMMPanels}
\end{figure}
\end{landscape}

\subsection{Procedure}\label{procedure}

At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Participants were randomly assigned to either a Control or Manipulation condition. The timeline of the experiment is shown in Figure \ref{fig:HMMPanels}.A. Phase one (``pre'') consisted of a 15 round RTG in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round. Phase 2 (``exposure'') consisted of three 7-round RTGs. Participants in the Manipulation condition faced the forgiving HMM investor and rated the agent on the same attributes as in the pre-manipulation phase. Those in the Control condition faced the same human-like HMM agent as in the ``pre'' phase and rated each co-player on the same attributes. To keep the experience similar to the ``pre'' phase, the agent in the Control condition was also designed to send a very low investment in round 5 of each of the three games. In the post-manipulation phase (``post''), participants in both conditions faced the same human-like HMM as in ``pre'' phase.

At the beginning of each game in all three phases, participants were told they would face a new player and had to wait to be paired with an available co-player. This simulated the waiting time in real social interaction tasks. After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-playerr to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating). After completing the three game phases, participants then completed the Levels of Personality Functioning Scale Brief-Form (LPFS-BF) questionnaire (Weekers, Hutsebaut, and Kamphuis 2019). This is a self-report measure designed to assess core elements of personality functioning as defined in the Alternative Model for Personality Disorders in the DSM-5 (American Psychiatric Association 2013), and provides a dimensional assessment of personality functioning, which complements the categorical approach of RS. Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.

\subsection{Statistical Analysis}\label{statistical-analysis}

To explore whether participants behaved differently in the RTG after the manipulation compared to the Control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.~post-manipulation), Condition (manipulation vs.~Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase. The full specification of the statistical model can be found in the supplement. To test whether participants behaved differently between conditions in the Exposure phase, we fit a linear mixed effects model to participants returns only in the exposure phase, with Condition (manipulation vs.~Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts. Finally, to test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.

The model was estimated using the \texttt{afex} package (Singmann et al. 2022) in R. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure (Matuschek et al. 2017). A similar process was used to establish the random effects structures of linear mixed-effects models used to analyse the HMM agent investments as well as the participants' ratings of the co-players. There is no agreed upon way to calculate effect sizes for mixed effects models. Instead, we will report on testing differences in marginal means. For the \(F\)-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package ``afex''. We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

\section{Behavioral Results}\label{behavioral-results}

\subsection{Analysis of Participant Returns}\label{analysis-of-participant-returns}

On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60\% of the endowment for investments and 35-50\% of the total yield for returns, as reported in previous studies (Charness, Cobo-Reyes, and Jiménez 2008; Fiedler, Haruvy, and Li 2011).

\begin{figure}

{\centering \includegraphics[width=\textwidth]{article_files/figure-latex/gamesPlot-1}

}

\caption{Averages and standard errors of the trustee's return as a percentage of the multiplied investment received (y-axis) by Condition, Phase, and game round (x-axis) averaged across RS groups. The blue line shows the returns in the Pre phase and the green line those in the Post phase. The left Panel shows returns in the Control condition and the right one those in the Manipulation condition. The dotted lines identify the rounds where the pre-programmed one-off low investment occurs. We note lower average returns post vs pre in the Manipulation condition, whilst returns in the Control condition are similar between the two phases.}\label{fig:gamesPlot}
\end{figure}

Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs.~Post RTG game), \(F(1, 201.63) = 5.81\), \(p = .017\), with higher percentage returns in the Pre RTG compared to the Post. Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.~post-manipulation), \(F(1, 201.63) = 4.38\), \(p = .038\). As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the Manipulation condition, pre - post,
\(\Delta M = 0.03\), 95\% CI \([0.01, 0.05]\), \(t(201.50) = 3.15\), \(p = .002\), but no change in the Control condition. We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.

There was also a significant main effect of Investment,
\(F(1, 5955.67) = 325.35\), \(p < .001\),
such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction,
\(F(1, 5955.67) = 13.92\), \(p < .001\), reflected that returns were more affected by investments in the Control condition. We also find a three way interaction between Phase, Investment and RS, \(F(1, 5864.62) = 7.84\), \(p = .005\), showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation). Finally, we find a four-way interaction between Condition, Phase, Investment and RS \(F(1, 5864.62) = 9.24\), \(p = .002\). For completeness, these effects are discussed in more detail in the supplement.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{article_files/figure-latex/boxPlots-1}

}

\caption{Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition. The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns. The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.}\label{fig:boxPlots}
\end{figure}

\subsubsection{Post Defection Trials}\label{post-defection-trials}

Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor? To explore this question, we restrict the analysis to the trials following the pre-programmed defection by the HMM agent in both the ``pre'' (trials 12 to 15) and the ``post'' phases (trials 13 to 15). We fit the same mixed effects model as for all the trials with the exception of the RS variable. This is because RS did not show main or interaction effects in the main model, and also due to the necessity of running a simpler model to accommodate the low number of trials. We find a significant main effect of Phase \(F(1, 227.47) = 4.76\), \(p = .030\) with returns lower in the second game post defection trials compared to the first. We also find a main effect of Investment
\(F(1, 1296.14) = 156.41\), \(p < .001\)
where participants continued to return higher proportions when receiving higher investments.
Finally, we still find an Investment by Condition interaction
\(F(1, 5955.67) = 13.92\), \(p < .001\) showing a lower effects of investment on the Manipulation condition compared to the Control condition in post-defection trials. However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.

\subsubsection{HMM Investor in Pre and Post Phases}\label{hmm-investor-in-pre-and-post-phases}

Was the HMM's strategy similar between pre and post phases in the control condition? Was participants' behavior post exposure differentiated enough to induce a different reaction from the HMM? To answer these questions, we test for differences in the HMM agent's investment by Phase, Condition and RS using a linear mixed-effects model as described in the methods section. As seen in Figure \ref{fig:boxPlots}, we find no main or interaction effects, indicating the HMM's behavior was on aggregate similar across levels of Phase, Condition and RS. This consistency in the investor's behavior is a desirable feature of the HMM agent when the participants behavior is largely simialr between phases. More importantly, it indicates that the lower returns of participants in the post phase of the manipulation condition were not differentiated enough to make the HMM react by transitioning to lower latent trust states. It is also noteworthy that the HMM agent was relatively successful in imitating human behavior in this paradigm: When asked during debrief whether they thought the investors they faced were human or not, \(41\)\% of participants thought they were either facing a human or were not sure of the nature of the co-player. When asked to justify their choice, many answers reflected participants projecting human traits such as ``spitefulness'' or ``greed'' onto the artificial co-player's behavior.

\subsubsection{Exposure Phase Trials}\label{exposure-phase-trials}

So far we focused on analysing behavior for the pre and post phases. Here, we look at returns and investments in the exposure phase. The linear mixed effects model of participants' returns in the exposure phase does not show a main effect of Condition on returns. There was a main effect of Investment, \(F(1, 4117.20) = 233.19\), \(p < .001\), with participants positively reciprocating higher investments, an interaction effect between Condition and Investment \(F(1, 4117.20) = 45.93\), \(p < .001\), showing a stronger positive reciprocity in the Control condition, and finally a three way interaction between the RS group, Condition and Investment \(F(1, 4117.20) = 4.21\), \(p = .040\), showing that this stronger positive reciprocity to investment in the Control condition is higher for participants with high RS. The linear mixed effects model of the HMM investments shows a main effect of Condition \(F(1, 202) = 197.64\), \(p < .001\), suggesting higher overall investments for the forgiving HMM compared to the human-like HMM, but no difference in investments when facing low and high RS groups.

In summary, despite the forgiving HMM sending overall higher investments in the exposure phase, participants returned similar proportions of the multiplied investments as those facing the human-like HMM. The positive reciprocity of returns to investments was higher in the Control condition with this relationship stronger for the high RS group.

\subsubsection{Questionnaire Scores and Performance}\label{questionnaire-scores-and-performance}

Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity Questionnaire score (RSQ), Spearman's \(r_{\mathrm{s}} = .52\), \(p < 0.001\), there was no correlation between these questionnaire scores and participant's return or overall task performance.

\subsection{Player Ratings}\label{player-ratings}

Figure \ref{fig:plotRatings} shows participants' ratings of each player they faced by condition and RS group. We will focus on two contrasts to analyse the ratings by Condition and RS group. The first is between the rating in the first phase (``pre'') when participants phase the human-like HMM, and the average rating during the exposure phase (average of ``expo1'', ``expo2'' and ``expo3'') where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition). The second contrast is between the ``pre'' and ``post'' phases of the experiment where in both conditions participants face the same human-like HMM.

\subsubsection{Comparing Pre and Exposure Ratings}\label{comparing-pre-and-exposure-ratings}

For those with high RS, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative \(\Delta M = 2.57\), 95\% CI \([0.84, 4.30]\), \(t(808) = 2.91\), \(p = .004\). There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
Those in the Control condition rated the investors faced in the exposure group (same HMM) as less cooperative \(\Delta M = -2.65\), 95\% CI \([-4.37, -0.94]\), \(t(808) = -3.04\), \(p = .002\), less forgiving \(\Delta M = -2.19\), 95\% CI \([-4.00, -0.39]\), \(t(808) = -2.38\), \(p = .017\), and were less keen on facing them again \(\Delta M = -3.62\), 95\% CI \([-5.73, -1.50]\), \(t(808) = -3.36\), \(p < .001\).

For those with low RS, there was no difference in any of the ratings between the ``pre'' and ``exposure'' phases for the Manipulation condition. For the Control condition, participants indicated less willingness to face the exposure co-player compared to the ``pre'' player, \(\Delta M = -2.43\), 95\% CI \([-4.52, -0.34]\), \(t(808) = -2.29\), \(p = .023\) but also did not differ on cooperation and forgiveness ratings. In summary, those with low RS had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving. In contrast, we see that exposing participants with high RS to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{article_files/figure-latex/plotRatings-1}

}

\caption{Averages and standard errors of the participants ratings of the opponent (y-axis) by each game and condition for each phase (x-axis). Pre and post are the 15 round RTGs before and after the exposure phase respectively. The games titled expo 1 to 3 are the three 7 round games during the exposure phase. The blue line represents participants' perception of co-player cooperation, the red line indicates perceived co-player forgiveness, and the green line shows the participants' willingness to play again with the same co-player. Cooperation, forgiveness, and willingness to play again ratings remained relatively stable for the low RS group except post-manipulation, where they were lower. In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the Manipulation condition, but decreased over time in the Control group. Ratings were also markedly lower in the post-manipulation game for this group.}\label{fig:plotRatings}
\end{figure}

\subsubsection{Comparing Pre and Post Ratings}\label{comparing-pre-and-post-ratings}

Participants high on RS in the Manipulation condition rated the co-players in the ``post'' phase similarly on cooperation, lower on
forgiveness \(\Delta M = -0.88\), 95\% CI \([-1.63, -0.14]\), \(t(808) = -2.33\), \(p = .020\),
and lower on willingness to face them again \(\Delta M = -1.29\), 95\% CI \([-2.16, -0.42]\), \(t(808) = -2.92\), \(p = .004\).
Those in the Control condition rated the investors post the exposure phase lower on all three attributes
(Cooperation: \(\Delta M = -0.96\), 95\% CI \([-1.66, -0.26]\), \(t(808) = -2.70\), \(p = .007\),
Forgiveness: \(\Delta M = -1.00\), 95\% CI \([-1.74, -0.26]\), \(t(808) = -2.66\), \(p = .008\),
Play again: \(\Delta M = -1.21\), 95\% CI \([-2.07, -0.35]\), \(t(808) = -2.76\), \(p = .006\)).

For participants low on RS, those in the Manipulation condition rated the co-players in the ``post'' phase lower on all three attributes
(Cooperation:\(\Delta M = -0.74\), 95\% CI \([-1.45, -0.03]\), \(t(808) = -2.03\), \(p = .042\),
Forgiveness: \(\Delta M = -1.14\), 95\% CI \([-1.89, -0.39]\), \(t(808) = -2.98\), \(p = .003\),
Play again: \(\Delta M = -1.30\), 95\% CI \([-2.18, -0.42]\), \(t(808) = -2.90\), \(p = .004\)).
Those in the Control condition did not differ in their ratings of the ``pre'' and ``post'' phase players.

In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the Control condition. In contrast, the high RS group shows a negative bias towards the co-players after the ``pre'' phase in the Control condition even though the player continues to use the same strategy. After exposure to the forgiving HMM, both groups rate the ``post'' co-player worse than the ``pre'' even though they are the same.

\section{Discussion}\label{discussion}

We used a randomized controlled online experiment where participants played a RTG with artificial agents designed to simulate human-like trust-building scenarios. Participants were then exposed to either forgiving or human-like HMM agents before playing another RTG. Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation. Instead, participants reduced their returns overall whilst the returns of those in the Control group did not change between the pre and post phase of the experiment. Neither did participants show more forgiving behavior in the Manipulation condition after the one-off defection by the agent. Why did participants reduce their returns even though they were repeatedly exposed to more cooperative and more forgiving artificial agents? A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the Manipulation condition.

Those exposed to the forgiving agent rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure. One possible explanation for this drop in rating is that participants exhibited a negative contrast effect. This occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people. If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favourable or even negative by comparison (Kobre and Lipsitt 1972). As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving agents (Zentall 2005). If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.

It is worth highlighting that RS was not correlated to overall performance in the task. Neither did it moderate the change in returns between Conditions or have an effect on participant returns. However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their HMM co-players: They accurately perceived increased cooperativeness in the forgiving agent, showing an ability to adjust their judgments based on interaction patterns. This may indicate that perceived support from cooperative partners can mitigate their rejection concerns. However, their declining ratings of similar co-players over time, despite consistent agent behavior, suggests that occasional pre-programmed defections may have amplified their perceptions of untrustworthiness and reluctance for future interactions. These nuanced perceptions align with the heightened social awareness often associated with high rejection sensitivity (Downey and Feldman 1996). Individuals with low RS were less affected by the agent's behavioral variations, showing a less volatile baseline of social perception. They still exhibit behavior consistent with a contrast effect in the Manipulation condition. This stability in perception might reflect a lower propensity to scrutinize social cues or a more robust internal model of social interactions that is less easily perturbed by short-term variations in partner behavior.

Despite these differences in perception, the actual cooperative behavior -- as measured by returns in the trust game -- did not significantly differ between RS groups. This suggests a potential dissociation between the cognitive-perceptual processes involved in evaluating social partners and the behavioral output in economic games. There are multiple potential explanations for this: First, the ratings may reflect explicit, conscious evaluations of social partners, while behavior in the trust game might be guided more by implicit, automatic processes that are less differentiated between RS groups. This interpretation aligns with dual-process theories of social cognition (Lieberman 2007). Second, the structured nature of the trust game might provide a context where the behavioral expression of rejection sensitivity is attenuated. In contrast, the more open-ended nature of providing ratings might allow for greater expression of RS-related perceptual biases.

These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged. Conversely, therapies targeting these perceptual processes might not necessarily translate into immediate behavioral changes. Future research should investigate whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods. Additionally, exploring specific conditions or thresholds under which perceptual differences begin to manifest in behavioral divergences could provide valuable insights into the dynamics of rejection sensitivity. Simply exposing individuals to highly cooperative or forgiving partners may not suffice to induce lasting changes in behavior. Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time. Combining periods of forgiveness with clear communication about mutual expectations for cooperation might be more effective (Balliet, Mulder, and Van Lange 2011). Future studies should also consider the temporal dynamics of these effects, examining whether longer exposure periods or different patterns of forgiving behavior could lead to more positive outcomes. Moreover, investigating the role of explicit communication and meta-cognitive reflection on the cooperative process could offer insights into leveraging positive interactions to promote enduring cooperative tendencies.

\section{Conclusion}\label{conclusion}

This randomised controlled experiment enabled us to uncover unexpected effects of exposure to forgiving behavior on subsequent cooperation, particularly in relation to rejection sensitivity. These findings challenge existing assumptions about fostering cooperative behavior and suggest the need for more nuanced interventions. Importantly, the use of HMM-based artificial agents in this study represents a significant methodological advancement. By providing a balance between experimental control and realistic, adaptive behavior, these agents allowed for a nuanced exploration of trust dynamics that would be challenging to achieve with human confederates or simplistic computer algorithms. This approach opens up new possibilities for studying complex social interactions in controlled settings, potentially bridging the gap between laboratory experiments and real-world social dynamics.

\pagebreak

\section*{Author contributions statement}\label{author-contributions-statement}
\addcontentsline{toc}{section}{Author contributions statement}

I. Guennouni, G. Koppe and C. Korn. designed and developed the study concept. Experiment design, testing and data collection were performed by I. Guennouni. I. Guennouni analysed and interpreted the data under the supervision of G. Koppe and C. Korn. All authors jointly wrote and approved the final version of the manuscript for submission.

\section*{Funding}\label{funding}
\addcontentsline{toc}{section}{Funding}

This publication was supported through state funds approved by the State Parliament of Baden-Württemberg for the Innovation Campus Health + Life Science alliance Heidelberg Mannheim. I. Guennouni was supported by a research fellowship from the AI Health Innovation Cluster

\section*{Competing interests statement}\label{competing-interests-statement}
\addcontentsline{toc}{section}{Competing interests statement}

The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.

\section*{Additional information}\label{additional-information}
\addcontentsline{toc}{section}{Additional information}

\subsection*{Correspondence}\label{correspondence}
\addcontentsline{toc}{subsection}{Correspondence}

All correspondence and requests for materials should be addressed to I. Guennouni.

\subsection*{Transparency and data availability}\label{transparency-and-data-availability}
\addcontentsline{toc}{subsection}{Transparency and data availability}

Preregistration: The hypotheses and methods were not preregistered.The analysis plan was not preregistered. Materials: All study materials are publicly available (\url{https://github.com/ismailg/exposure-public}). Data: All primary data are publicly available (\url{https://github.com/ismailg/exposure-public}). Analysis scripts: All analysis scripts are publicly available (\url{https://github.com/ismailg/exposure-public}).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-abramov_influence_2022-1}
Abramov, Gamze, Jason Kautz, Sebastien Miellet, and Frank P. Deane. 2022. {``The {Influence} of {Attachment Style}, {Self-protective Beliefs}, and {Feelings} of {Rejection} on the {Decline} and {Growth} of {Trust} as a {Function} of {Borderline Personality Disorder Trait Count}.''} \emph{Journal of Psychopathology and Behavioral Assessment} 44 (3): 773--86. \url{https://doi.org/10.1007/s10862-022-09965-9}.

\bibitem[\citeproctext]{ref-almaatouq_empirica_2021}
Almaatouq, Abdullah, Joshua Becker, James P. Houghton, Nicolas Paton, Duncan J. Watts, and Mark E. Whiting. 2021. {``Empirica: A Virtual Lab for High-Throughput Macro-Level Experiments.''} \emph{Behavior Research Methods} 53 (5): 2158--71. \url{https://doi.org/10.3758/s13428-020-01535-9}.

\bibitem[\citeproctext]{ref-american_psychiatric_association_diagnostic_2013}
American Psychiatric Association. 2013. \emph{Diagnostic and {Statistical Manual} of {Mental Disorders}}. Fifth Edition. American Psychiatric Association. \url{https://doi.org/10.1176/appi.books.9780890425596}.

\bibitem[\citeproctext]{ref-balliet_reward_2011}
Balliet, Daniel, Laetitia B. Mulder, and Paul A. M. Van Lange. 2011. {``Reward, Punishment, and Cooperation: {A} Meta-Analysis.''} \emph{Psychological Bulletin} 137 (4): 594--615. \url{https://doi.org/10.1037/a0023489}.

\bibitem[\citeproctext]{ref-bandura_social_1977}
Bandura, Albert. 1977. \emph{Social {Learning Theory}}. Prentice Hall.

\bibitem[\citeproctext]{ref-bowlby_attachment_1978}
Bowlby, John. 1978. {``Attachment Theory and Its Therapeutic Implications.''} \emph{Adolescent Psychiatry} 6: 5--33.

\bibitem[\citeproctext]{ref-charness_investment_2008}
Charness, Gary, Ramón Cobo-Reyes, and Natalia Jiménez. 2008. {``An Investment Game with Third-Party Intervention.''} \emph{Journal of Economic Behavior \& Organization} 68 (1): 18--28. \url{https://doi.org/10.1016/j.jebo.2008.02.006}.

\bibitem[\citeproctext]{ref-downey_implications_1996}
Downey, Geraldine, and Scott I. Feldman. 1996. {``Implications of Rejection Sensitivity for Intimate Relationships.''} \emph{Journal of Personality and Social Psychology} 70 (6): 1327--43. \url{https://doi.org/10.1037/0022-3514.70.6.1327}.

\bibitem[\citeproctext]{ref-downey_early_1997}
Downey, Geraldine, Hala Khouri, and Scott I. Feldman. 1997. {``Early Interpersonal Trauma and Later Adjustment: {The} Mediational Role of Rejection Sensitivity.''} In \emph{Developmental Perspectives on Trauma: {Theory}, Research, and Intervention}, 85--114. Rochester Symposium on Developmental Psychology, {Vol}. 8. Rochester, NY, US: University of Rochester Press.

\bibitem[\citeproctext]{ref-fiedler_social_2011}
Fiedler, Marina, Ernan Haruvy, and Sherry Xin Li. 2011. {``Social Distance in a Virtual World Experiment.''} \emph{Games and Economic Behavior} 72 (2): 400--426. \url{https://doi.org/10.1016/j.geb.2010.09.004}.

\bibitem[\citeproctext]{ref-fonagy_mentalizing_2017}
Fonagy, Peter, and Chloe Campbell. 2017. {``\href{https://www.ncbi.nlm.nih.gov/pubmed/29135441}{Mentalizing, Attachment and Epistemic Trust: How Psychotherapy Can Promote Resilience}.''} \emph{Psychiatria Hungarica: A Magyar Pszichiatriai Tarsasag Tudomanyos Folyoirata} 32 (3): 283--87.

\bibitem[\citeproctext]{ref-fowler_cooperative_2010}
Fowler, James H., and Nicholas A. Christakis. 2010. {``Cooperative Behavior Cascades in Human Social Networks.''} \emph{Proceedings of the National Academy of Sciences} 107 (12): 5334--38. \url{https://doi.org/10.1073/pnas.0913149107}.

\bibitem[\citeproctext]{ref-gao_associations_2017}
Gao, Shuling, Mark Assink, Andrea Cipriani, and Kangguang Lin. 2017. {``Associations Between Rejection Sensitivity and Mental Health Outcomes: {A} Meta-Analytic Review.''} \emph{Clinical Psychology Review} 57 (November): 59--74. \url{https://doi.org/10.1016/j.cpr.2017.08.007}.

\bibitem[\citeproctext]{ref-hepp_prosociality_2022}
Hepp, Johanna, and Inga Niedtfeld. 2022. {``Prosociality in Personality Disorders: {Status} Quo and Research Agenda.''} \emph{Current Opinion in Psychology} 44 (April): 208--14. \url{https://doi.org/10.1016/j.copsyc.2021.09.013}.

\bibitem[\citeproctext]{ref-herpertz_social-cognitive_2014}
Herpertz, Sabine C., and Katja Bertsch. 2014. {``The Social-Cognitive Basis of Personality Disorders.''} \emph{Current Opinion in Psychiatry} 27 (1): 73--77. \url{https://doi.org/10.1097/YCO.0000000000000026}.

\bibitem[\citeproctext]{ref-joyce_trust_1995}
Joyce, Berg, John Dickhaut, and Kevin McCabe. 1995. {``Trust, {Reciprocity}, and {Social History}.''} \emph{Games and Economic Behavior} 10 (1): 122--42.

\bibitem[\citeproctext]{ref-king-casas_getting_2005}
King-Casas, Brooks, Damon Tomlin, Cedric Anen, Colin F. Camerer, Steven R. Quartz, and P. Read Montague. 2005. {``Getting to {Know You}: {Reputation} and {Trust} in a {Two-Person Economic Exchange}.''} \emph{Science} 308 (5718): 78--83. \url{https://doi.org/10.1126/science.1108062}.

\bibitem[\citeproctext]{ref-kobre_negative_1972}
Kobre, Kenneth R, and Lewis P Lipsitt. 1972. {``A Negative Contrast Effect in Newborns.''} \emph{Journal of Experimental Child Psychology} 14 (1): 81--91. \url{https://doi.org/10.1016/0022-0965(72)90033-1}.

\bibitem[\citeproctext]{ref-lieberman_social_2007}
Lieberman, Matthew D. 2007. {``Social {Cognitive Neuroscience}: {A Review} of {Core Processes}.''} \emph{Annual Review of Psychology} 58 (1): 259--89. \url{https://doi.org/10.1146/annurev.psych.58.110405.085654}.

\bibitem[\citeproctext]{ref-matuschek_balancing_2017}
Matuschek, Hannes, Reinhold Kliegl, Shravan Vasishth, Harald Baayen, and Douglas Bates. 2017. {``Balancing {Type I} Error and Power in Linear Mixed Models.''} \emph{Journal of Memory and Language} 94 (June): 305--15. \url{https://doi.org/10.1016/j.jml.2017.01.001}.

\bibitem[\citeproctext]{ref-miano_rejection_2013}
Miano, Annemarie, Eric A. Fertuck, Arnoud Arntz, and Barbara Stanley. 2013. {``Rejection {Sensitivity Is} a {Mediator Between Borderline Personality Disorder Features} and {Facial Trust Appraisal}.''} \emph{Journal of Personality Disorders} 27 (4): 442--56. \url{https://doi.org/10.1521/pedi_2013_27_096}.

\bibitem[\citeproctext]{ref-mikulincer_attachment_1998}
Mikulincer, Mario. 1998. {``Attachment Working Models and the Sense of Trust: {An} Exploration of Interaction Goals and Affect Regulation.''} \emph{Journal of Personality and Social Psychology} 74 (5): 1209--24. \url{https://doi.org/10.1037/0022-3514.74.5.1209}.

\bibitem[\citeproctext]{ref-mulder_relationship_1999}
Mulder, R. T., P. R. Joyce, P. F. Sullivan, C. M. Bulik, and F. A. Carter. 1999. {``The Relationship Among Three Models of Personality Psychopathology: {DSM-III-R} Personality Disorder, {TCI} Scores and {DSQ} Defences.''} \emph{Psychological Medicine} 29 (4): 943--51. \url{https://doi.org/10.1017/S0033291799008533}.

\bibitem[\citeproctext]{ref-richetin_emotional_2018}
Richetin, Juliette, Anita Poggi, Paola Ricciardelli, Eric A. Fertuck, and Emanuele Preti. 2018. {``The Emotional Components of Rejection Sensitivity as a Mediator Between {Borderline Personality Disorder} and Biased Appraisal of Trust in Faces.''} \emph{Clinical Neuropsychiatry: Journal of Treatment Evaluation} 15 (4): 200--205.

\bibitem[\citeproctext]{ref-singmann_afex_2022}
Singmann, Henrik, Ben Bolker, Jake Westfall, Frederik Aust, Mattan S. Ben-Shachar, Søren Højsgaard, John Fox, et al. 2022. {``Afex: {Analysis} of {Factorial Experiments}.''}

\bibitem[\citeproctext]{ref-weekers_level_2019}
Weekers, Laura C., Joost Hutsebaut, and Jan H. Kamphuis. 2019. {``The {Level} of {Personality Functioning Scale}-{Brief Form} 2.0: {Update} of a Brief Instrument for Assessing Level of Personality Functioning.''} \emph{Personality and Mental Health} 13 (1): 3--14. \url{https://doi.org/10.1002/pmh.1434}.

\bibitem[\citeproctext]{ref-zentall_within-trial_2005}
Zentall, Thomas R. 2005. {``A {Within-trial Contrast Effect} and Its {Implications} for {Several Social Psychological Phenomena}.''} \emph{International Journal of Comparative Psychology} 18 (4). \url{https://doi.org/10.46867/ijcp.2005.18.04.08}.

\end{CSLReferences}

\end{document}

### 2024-07-09 supplement.Rmd
version 3 finalised

@@
 
-# Repeated Trust Game
+# A Snapshot of the Repeated Trust Game as Seen by Participants {-}
 
-Participants played a 15-round repeated trust game [@joyce_trust_1995] in the trustee role against a computer-programmed investor.
-Each player was represented with an icon with the participant always on the left of the screen and the opponent on the right.
-The participants were able to choose the icon that represents them at the start of the experiment.
-The icon representing the opponent changed at the start of each new game, to simulate a new interaction partner.
-Participants were not told they were facing computerised opponents.
-We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a random time lapse (randomly chosen between 5 and 10 seconds).
+Figure \@ref(fig:trustGameScreenshot) shows a screenshot of the repeated Trust Game at the moment the participant is required to make a decision of how much to send back to the Investor.
 
-```{r trustGameScreenshot, include=F, fig.cap = "Screenshot of the repeated Trust Game.
-The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee.
-The investor is endowed with 20 units at the start of each round.
-They need to decide how much of that endowment they want to invest with the trustee.
-The investment is then multiplied by a factor of 3 and sent to the trustee.
-The trustee in turn needs to decide how much of the multiplied investment they want to send back to the investor.
-Shown here is the stage at which the trustee makes a decision of how much to send back the investor", out.width="70%", out.height="70%", fig.align='center'}
+```{r trustGameScreenshot, include=T, fig.cap = "Screenshot of the RTG as seen by participants at the decision phase.", out.width="70%", out.height="70%", fig.align='center'}
 
@@
 
-# Hidden markov Model used to simulate the Investor's actions
+# Hidden Markov Model Used to Simulate the Investor's Actions {-}
 
@@
 
-## Mixed-effects models for participant returns
+# Mixed-effects Models for Participant Returns {-}
 
-We fit a linear mixed effects model to participant returns as a proportion of the multiplied investment received as described below. The results of the model are presented in Table \ref{tab:myTable}
+We fit a linear mixed effects model to participant returns as a proportion of the multiplied investment received as described below.
+The results of the model are presented in Table \@ref(tab:myTableRet)
 
@@
          Term = gsub("phase.f1", "Phase", Term),
-         Term = gsub("high_RS1", "high_RS", Term),
-         Term = gsub("inv_scaled", "Investment", Term))
-
-# Create a summary table using apa_table
-summary_table <- papaja::apa_table(
-  mod_df,
-  caption = "Summary of Mixed-Effects Model of participant returns over all rounds",
-  align = c("l", "c", "c", "c", "c", "c")
-)
+         Term = gsub("high_RS1", "RS_group", Term),
+         Term = gsub("inv_scaled", "Investment", Term)) %>%
+  mutate(across(where(is.numeric), ~round(., 2)))  # Round all numeric columns to 2 decimal places
 
@@
 
-```{r myTable, results='asis'}
-summary_table
+```{r myTableRet, results='asis'}
+# Create table using kable and kableExtra
+kable(mod_df, 
+      caption = "Summary of Mixed-Effects Model of participant returns across all rounds",
+      booktabs = TRUE,
+      align = c("l", "c", "c", "c", "c", "c")) %>%
+  kable_styling(latex_options = c("striped", "hold_position"))
+```
+
+# Mixed-effects Models for HMM Investments  {-}
+
+We fit a linear mixed effects model to the investments sent by the HMM with Phase, Condition and RS groups as fixed effects, and player-wise intercepts.
+The results of the model are presented in Table \ref{tab:myTableInv}.
+
+```{r, include=F}
+# Create a summary object
+model_inv <- readRDS("data/mod_invs_RS.RDS")
+
+# Extract the fixed effects as a data frame
+mod_df_inv <- as.data.frame(model_inv$coefficients)
+
+mod_df_inv <- mod_df_inv %>%
+  rownames_to_column(var = "Term") %>% 
+  mutate(Term = gsub("condition.f1", "Condition", Term),
+         Term = gsub("phase.f1", "Phase", Term),
+         Term = gsub("high_RS1", "RS_group", Term)) %>%
+  mutate(across(where(is.numeric), ~round(., 2)))  # Round all numeric columns to 2 decimal places
+
+```
+
+```{r myTableInv, results='asis'}
+# Create table using kable and kableExtra
+kable(mod_df_inv, 
+      caption = "Summary of Mixed-Effects Model of HMM investments across all rounds",
+      booktabs = TRUE,
+      align = c("l", "c", "c", "c", "c", "c")) %>%
+  kable_styling(latex_options = c("striped", "hold_position"))
 ```
@@
 
-# References
+# References {-}

### 2024-10-10 article.tex
format changes JPCS

@@
   pdftitle={When Forgiveness Backfires: Rejection Sensitivity and Cooperative Behavior Following Exposure to Adaptive Forgiving Agents},
-  pdfauthor={Ismail Guennouni1*, Georgia Koppe2, Christoph Korn3},
+  pdfauthor={Ismail Guennouni1,2,3,4*, Georgia Koppe1,2,3\^{}\textbackslash dagger, Christoph Korn4\^{}\textbackslash dagger},
   hidelinks,
@@
 \title{When Forgiveness Backfires: Rejection Sensitivity and Cooperative Behavior Following Exposure to Adaptive Forgiving Agents}
-\author{Ismail Guennouni\textsuperscript{1}*, Georgia Koppe\textsuperscript{2}, Christoph Korn\textsuperscript{3}}
+\author{Ismail Guennouni\textsuperscript{1,2,3,4}*, Georgia Koppe\textsuperscript{1,2,3}\(^\dagger\), Christoph Korn\textsuperscript{4}\(^\dagger\)}
 \date{}
@@
 \maketitle
-\begin{abstract}
-Can exposure to forgiving behavior improve cooperative outcomes, and does Rejection Sensitivity (RS) moderate this effect?
-This randomized controlled experiment explores these questions using a novel approach with Hidden Markov Model (HMM)-based artificial agents in repeated trust games (RTGs).
-Participants (N = 206), pre-screened for high or low RS, interacted with either consistently behaving or more forgiving co-players, which were in reality HMM agents.
-Unexpectedly, exposure to forgiving agents led to reduced cooperation in subsequent interactions.
-Control group participants maintained consistent behavior.
-RS levels influenced perceptions of the agent's cooperativeness but did not significantly affect behavioral outcomes, revealing a perception-behavior dissociation.
-These findings challenge assumptions about fostering cooperation through simple exposure to positive interactions and highlight the importance of accounting for negative contrast effects when designing such interventions.
-They also demonstrate the potential of HMM-based artificial agents for studying interactive social dynamics, offering a methodological advancement for future research.
-Our results suggest the need for nuanced interventions that gradually shape expectations and behaviors to foster cooperative outcomes, particularly for individuals with high rejection sensitivity.
-\end{abstract}
 
@@
 
-\textsuperscript{1} \emph{Central Institute of Mental Health, Medical Faculty Mannheim/Heidelberg University, Germany}
+\textsuperscript{1} \emph{Department of Psychiatry and Psychotherapy, Central Institute of
+Mental Health, Medical Faculty Mannheim, Heidelberg University,
+Mannheim, Germany}
 
-\textsuperscript{2} \emph{Faculty of Mathematics and Computer Science, Heidelberg University, Germany}
+\textsuperscript{2} \emph{Interdisciplinary Center for Scientific Computing, Faculty of
+Mathematics and Computer Science, Heidelberg University, Heidelberg,
+Germany}
 
-\textsuperscript{3} \emph{Department of General Psychiatry, Section Social Neuroscience, Heidelberg University, Germany}
+\textsuperscript{3} \emph{Hector Institute for AI in Psychiatry, Central Institute of Mental
+Health, Medical Faculty Mannheim, Heidelberg University, Mannheim
+Germany}
 
-\textsuperscript{*} \emph{Corresponding author.
-Address: Central institute of Mental Health, J5, Mannheim, Germany.
-Email: \href{mailto:ismail.guennouni@zi-mannheim.de}{\nolinkurl{ismail.guennouni@zi-mannheim.de}}}
+\textsuperscript{4} \emph{Department of General Psychiatry, Section Social Neuroscience,
+Heidelberg University, Germany}
+
+\(^\dagger\) \emph{Joint last author}
+
+\textsuperscript{*} \emph{Corresponding author. Address: Central institute of Mental Health,
+J5, Mannheim, Germany. Email:
+\href{mailto:ismail.guennouni@zi-mannheim.de}{\nolinkurl{ismail.guennouni@zi-mannheim.de}}}
+
+\pagebreak
+
+\textbf{Abstract:}
+
+Can exposure to forgiving behavior improve cooperative outcomes, and
+does Rejection Sensitivity (RS) moderate this effect? This randomized
+controlled experiment explores these questions using a novel approach
+with Hidden Markov Model (HMM)-based artificial agents in repeated trust
+games (RTGs). Participants (N = 206), pre-screened for high or low RS,
+interacted with either consistently behaving or more forgiving
+co-players, which were in reality HMM agents. Unexpectedly, exposure to
+forgiving agents led to reduced cooperation in subsequent interactions.
+Control group participants maintained consistent behavior. RS levels
+influenced perceptions of the agent's cooperativeness but did not
+significantly affect behavioral outcomes, revealing a
+perception-behavior dissociation. These findings challenge assumptions
+about fostering cooperation through simple exposure to positive
+interactions and highlight the importance of accounting for negative
+contrast effects when designing such interventions. They also
+demonstrate the potential of HMM-based artificial agents for studying
+interactive social dynamics, offering a methodological advancement for
+future research. Our results suggest the need for nuanced interventions
+that gradually shape expectations and behaviors to foster cooperative
+outcomes, particularly addressing the discrepancy between perceptions
+and behavior in individuals with high rejection sensitivity.
 
@@
 
-\textbf{Keywords} : Social Trust; Forgiveness Intervention; Rejection Sensitivity; Interpersonal Dynamics; Hidden Markov Models
+\textbf{Keywords} : Interpersonal functioning; Rejection Sensitivity;
+Forgiveness Intervention; Trust-based Cooperation; Hidden Markov Models
 
@@
 
-\textbf{Highlights:}
+\textbf{General Scientific Summary:}
 
-\begin{itemize}
-\item
-  Exposure to forgiving artificial agents unexpectedly led to reduced cooperation in subsequent interactions, challenging assumptions about fostering trust through positive experiences
-\item
-  Rejection sensitivity influenced perceptions of artificial agents' cooperativeness but not behavioral outcomes, revealing a dissociation between perception and behavior
-\item
-  Novel Hidden Markov Model-based artificial agents provide adaptive, human-like behavior in trust games, allowing for precise control of ``cooperative'' and ``forgiving'' strategies while maintaining experimental control of co-player strategies
-\end{itemize}
+This study reveals that exposing individuals to forgiving behavior in
+economic games can unexpectedly decrease their subsequent cooperation.
+Interestingly, while people with high rejection sensitivity more
+accurately perceived their partners' cooperativeness, this didn't
+translate into more cooperative behavior. These findings challenge
+assumptions about fostering cooperation and highlight the need for more
+nuanced interventions to promote reciprocity, especially for individuals
+prone to expecting rejection.
 
@@
 
-Trust is fundamental to human social interactions, facilitating seamless relations at both interpersonal and intergroup levels.
-The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders (Fonagy and Campbell 2017).
-Individuals with personality disorders (PD) often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors -- a marker for the severity of PD symptoms (Herpertz and Bertsch 2014; Mulder et al.
-1999).
+Trust is fundamental to human social interactions, facilitating seamless
+relations at both interpersonal and intergroup levels. The study of
+psychopathology has linked deficits in trust-based constructs to the
+development of mental health disorders (Fonagy \& Campbell, 2017).
+Individuals with personality disorders (PD) often struggle to form and
+maintain social connections, a difficulty reflected in uncooperative
+behaviors -- a marker for the severity of PD symptoms
+(Herpertz \& Bertsch, 2014; Mulder et al., 1999).
 
-One explanation for such social challenges lies in early caregiver experiences.
-Attachment theory (Bowlby 1978) suggests that the quality of these relationships shapes our capacity for secure attachments and trust.
-Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken (Mikulincer 1998).
-These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection (Downey and Feldman 1996; Downey, Khouri, and Feldman 1997).
-RS has been linked to the development of various mental health conditions, including depression, anxiety, personality disorders, and self-harm (Gao et al.
-2017).
-A recent meta-analysis revealed prosocial behavior and interpersonal trust as two key processes of interpersonal functioning that are markedly impaired in PDs and which are likely to contribute to interpersonal dysfunction in this population (Hepp and Niedtfeld 2022).
-The interaction of rejection sensitivity and trust-based constructs has been explored, particularly in Borderline Personality Disorder (BPD).
-Miano et al.
-(2013) and Richetin et al.
-(2018) found that rejection sensitivity mediated the relationship between BPD features and lower trust appraisal.
-Abramov et al.
-(2022) found that higher baseline feelings of rejection in individuals with BPD predict slower trust formation and less pronounced declines in trust following trust violations during the trust game.
-However, the interaction between trustworthiness and rejection sensitivity hasn't been studied as extensively in other personality disorders or in non-clinical populations, leaving a gap in our understanding of how these constructs might interplay.
-Could an ingrained sensitivity to rejection contribute to reduced pro-social behavior such as a failure to show trustworthiness?
+One explanation for such social challenges lies in early caregiver
+experiences. Attachment theory (Bowlby, 1978) suggests that
+the quality of these relationships shapes our capacity for secure
+attachments and trust. Individuals with higher levels of insecure
+attachment may recall negative trust-related experiences more easily,
+report fewer positive trust experiences, and use less constructive
+coping strategies when trust is broken (Mikulincer, 1998).
+These insecure attachment patterns are often associated with heightened
+rejection sensitivity (RS), a tendency to anxiously expect, readily
+perceive, and intensely react to rejection (Downey et al., 1997; Downey \& Feldman, 1996). RS has been linked to the development of various
+mental health conditions, including depression, anxiety, personality
+disorders, and self-harm (Gao et al., 2017). Individuals high in
+RS show attentional biases towards social threat cues, which may
+contribute to difficulties in social interactions
+(Berenson et al., 2009). A recent meta-analysis revealed prosocial
+behavior and interpersonal trust as two key processes of interpersonal
+functioning that are markedly impaired in PDs and which are likely to
+contribute to interpersonal dysfunction in this population
+(Hepp \& Niedtfeld, 2022). The interaction of RS and trust-based
+constructs has been explored, particularly in Borderline Personality
+Disorder (BPD). Miano et al. (2013) and Richetin et al. (2018) found
+that RS mediated the relationship between BPD features and lower trust
+appraisal. Abramov et al. (2022) found that higher baseline feelings
+of rejection in individuals with BPD predict slower trust formation and
+less pronounced declines in trust following trust violations during the
+trust game. However, the interaction between \emph{reciprocity} and RS hasn't
+been studied as extensively, leaving a gap in our understanding of how
+these constructs might interplay. Could an ingrained sensitivity to
+rejection contribute to reduced pro-social behavior such as a failure to
+show reciprocity as a result of perceived rejection by the interaction
+partner?
 
-If indeed RS contributes to interpersonal dysfunction through failure to show trustworthiness, could exposing those who exhibit it to cooperative and forgiving interaction partners lead to improved interpersonal outcomes?
-Research in the fields of behavioral economics and psychology has explored how positive social interactions influence cooperation and pro-social behavior.
-The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions (Joyce, Dickhaut, and McCabe 1995).
-In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors even among individuals who are initially inclined to be distrustful (King-Casas et al.
-2005).
-Fowler and Christakis (2010) studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior.
-Similarly, research on social learning theory (Bandura 1977) has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior.
-These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust and potentially fostering trustworthy behavior.
+Given that RS may be a manifestation of maladaptive attachment styles,
+it's crucial to explore whether exposure to consistently forgiving
+interaction partners could help reshape these interpersonal expectations
+and behaviors. If RS indeed influences reciprocity in repeated
+interactions, could such exposure serve as a corrective experience,
+potentially improving interpersonal outcomes? This approach aligns with
+attachment theory, which suggests that new, positive relational
+experiences can modify internal working models of relationships
+(Bowlby, 1988). Research in the fields of behavioral economics
+and psychology has also explored how positive social interactions
+influence cooperation and pro-social behavior. The use of the repeated
+trust game (RTG), a well-established experimental approach, has allowed
+for the analysis of the development of trust through ongoing
+interactions (Joyce et al., 1995). In this paradigm, cycles of mutual
+trust, where each party's trust is reciprocated with trustworthiness,
+have the effect of enhancing cooperative behaviors even among
+individuals who are initially inclined to be distrustful
+(King-Casas et al., 2005). Fowler \& Christakis (2010) studied behavior in
+social networks interacting in a public goods game and found that
+cooperative behavior tends to cluster, suggesting that exposure to
+cooperative peers can lead to more cooperative behavior. Similarly,
+research on social learning theory (Bandura, 1977) has long
+demonstrated that individuals learn and model the behavior of those
+around them, indicating that if someone is consistently exposed to
+cooperative and positive individuals, they're likely to emulate this
+behavior. These insights highlight that engaging with compassionate and
+forgiving others can be an effective method for mitigating deeply
+ingrained mistrust and potentially fostering trustworthy behavior.
 
-In this study, we use a randomized controlled online experiment to test whether exposing participants with varying RS levels to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation.
-To simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players' data.
-A key aspect of these agents is that their actions depend on a latent ``trust state'' which reacts dynamically to the trustees' returns, simulating real-life trust-building scenarios.
-An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively, or the propensity to retaliate after breakdowns of cooperation.
+In this study, we use a randomized controlled online experiment to test
+whether exposing participants with varying RS levels to forgiving and
+more cooperative co-players results in more trustworthy behavior and a
+repair of potential breakdowns in RTG cooperation. To simulate realistic
+social interaction while maintaining a high degree of experimental
+control, we take a novel paradigmatic approach: We use generative models
+of how humans play the RTG to design an agent that plays the role of the
+investor, based on Hidden Markov Models (HMMs) fitted to real players'
+data. A key aspect of these agents is that their actions depend on a
+latent ``trust state'' which reacts dynamically to the trustees' returns,
+simulating real-life trust-building scenarios. An advantage of having
+such a generative model of behavior is the possibility of controlling
+different aspects of the agent's strategy such as its general policy,
+the propensity to cooperate actively, or the propensity to trust again
+after breakdowns of cooperation. To further mimic real-world
+interactions and examine participants' responses to one-off breakdowns
+of cooperation, we incorporate occasional pre-programmed low investments
+by the agent.
 
-We pre-screened participants for high or low rejection sensitivity (RS) using a validated questionnaire, then assigned them exclusively to the trustee role in a series of trust games.
-After playing a 15-round RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition.
-In the Manipulation condition, participants were exposed over three RTGs to HMM investors designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
-In the Control condition, participants played three RTGs against the same human-like HMM.
-After this exposure phase, all participants played another 15-round RTG with a human-like HMM investor.
-We expect those exhibiting high RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
-We also hypothesize that those in the Manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase, with those in the higher RS group showing a bigger effect.
+We pre-screened participants for high or low RS using a validated
+questionnaire, then assigned them exclusively to the trustee role in a
+series of trust games. After playing a 15-round RTG with a human-like
+HMM investor, they were randomly assigned to either a Control or
+Manipulation condition. In the Manipulation condition, participants were
+exposed over three RTGs to HMM investors designed with a limited
+propensity for retaliation, potentially mitigating ingrained mistrust.
+In the Control condition, participants played three RTGs against the
+same human-like HMM. After this exposure phase, all participants played
+another 15-round RTG with a human-like HMM investor, similar to the one
+in the pre-exposure phase. We expect those exhibiting high RS to be more
+responsive to the pre-programmed change in the agent's cooperativeness
+and retaliation propensity. We also hypothesize that those in the
+Manipulation condition would behave more cooperatively and have a lower
+propensity to retaliate to the co-player's defection after the exposure
+phase, with those in the higher RS group showing a bigger effect.
 
@@
 
-To have participants with large differences in RS, a total of 1195 participants were pre-screened on the Prolific Academic platform (prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score \textgreater{} 15) and the other with low RS (RSQ score \textless{} 10) totalling 206 participants (56\% female).
-These were then invited through prolific to take part in the main experiment study.
-The required sample size was determined using an a priori power analysis to have an 80\% probability to detect a small effect size with a 5\% type I error rate in a repeated measures ANOVA with between factors.
-The mean age of participants was 34.6 years, with an 11.9 years standard deviation.
-Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.
+To have participants with large differences in RS, a total of 1195
+participants were pre-screened on the Prolific Academic platform
+(prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to
+finally select two similarly sized groups: One with high RS (RSQ score
+\textgreater{} 15, N=103) and the other with low RS (RSQ score \textless{} 10, N=103)
+totalling 206 participants (56\% female). These were then invited through
+prolific to take part in the main experiment. The required sample size
+was determined using an \emph{a priori} power analysis to have an 80\%
+probability to detect a small effect size (Cohen's f = 0.10) for a
+within-between interaction with a 5\% type I error rate in a repeated
+measures ANOVA. The sample size calculation assumed 2 groups, 2
+measurement per group and was performed using the G*Power software
+(Faul et al., 2009). The mean age of participants was 34.6 years,
+with an 11.9 years standard deviation. The majority of participants
+identified ethnically as White (\(80\)\%). The online cohort registered 30
+unique countries of birth with the most frequent being the U.K (\(33\)\%)
+followed by Poland (\(10\)\%) and Portugal (\(10\)\%). Participants were paid
+a fixed fee of £6 plus a bonus payment dependent on their performance
+that averaged £0.5. Data was collected over multiple sessions between
+December 2023 and February 2024.
 
@@
 
-The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : High or Low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \ref{fig:HMMPanels}.A).
-Participants within each pre-screened group were randomly assigned to one of the two levels of the Condition factor.
-The games were designed and implemented online using Empirica v1 (Almaatouq et al.
-2021).
-The planned experiment received approval from the University of Heidelberg's Medical Faculty ethics commission (ID:S-708/2023) and the experiment was performed in accordance with the ethics board guidelines and regulations.
-All participants provided informed consent prior to their participation.
+The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS :
+High or Low) by 2 (Phase: Trust-Game Pre Manipulation, Trust-Game Post
+Manipulation) design, with repeated measures on the Phase factor (Figure
+\ref{fig:HMMPanels}.A). Participants within each pre-screened RS group
+were randomly assigned to one of the two levels of the Condition factor,
+resulting in 101 participants in the Manipulation condition and 105 in
+the Control condition. The games were designed and implemented online
+using Empirica v1 (Almaatouq et al., 2021). The planned experiment
+received approval from the University of Heidelberg's Medical Faculty
+ethics commission (ID:S-708/2023) and the experiment was performed in
+accordance with the ethics board guidelines and regulations. All
+participants provided informed consent prior to their participation.
 
@@
 
-\subsubsection{Repeated Trust Game}\label{repeated-trust-game}
+\subsubsection{Repeated Trust Game and HMM Investor}\label{repeated-trust-game-and-hmm-investor}
 
-Participants played a 15-round RTG (Joyce, Dickhaut, and McCabe 1995) in the trustee role against a computer-programmed investor.
-On each round the investor is endowed with 20 units and decides how much of that endowment to invest.
-This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor.
-If the trustee returns more than one third of the amount, the investor makes a gain.
-Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right.
-The participants were able to choose the icon that represents them at the start of the experiment.
-The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner.
-Participants were not told they were facing computerised co-players.
-We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a varying time lapse (randomly chosen between 5 and 10 seconds).
+Participants played a 15-round RTG (Joyce et al., 1995) in the trustee
+role against a computer-programmed investor. On each round the investor
+is endowed with 20 units and decides how much of that endowment to
+invest. This investment is tripled and the trustee then decides how to
+split this tripled amount between them and the investor. If the trustee
+returns more than one third of the amount, the investor makes a gain.
+Each player was represented with an icon with the participant always on
+the left of the screen and the co-player on the right. The participants
+were able to choose the icon that represents them at the start of the
+experiment. The icon representing the co-player changed at the start of
+each new game, to simulate a new interaction partner. Participants were
+not told they were facing computerised co-players. We chose to simulate
+the behavior of a human interaction partner through allowing for a delay
+whilst pairing with new opponents as the start of each game as well as
+programming the agents to respond during each round after a varying time
+lapse (randomly chosen between 5 and 10 seconds).
 
-The computerised investor consisted of a hidden Markov model (HMM) trained on an independent existing behavioral RTG data set of human investors.
-This data-driven approach thus sought to learn an investor strategy that mimics human-like interactions.
-The data set used for training consists of 388 ten round games with the same player (full details can be found in the Supplementary Information).
-On this data set, the HMM was inferred with three latent states that could be interpreted as reflecting a ``low-trust'', a ``medium-trust'', and a ``high-trust'' state.
-A separate output distribution, that maps each HMM state onto possible investments from 0 to 20 separately, is learned (Figure \ref{fig:HMMPanels}.B).
-In analogy to the latent states, these distributions can be interpreted as reflecting ``low-trust'', ``medium-trust'', or ``high-trust'' dispositions.
-Finally, the HMM is specified by transition probabilities that describe the transition between states.
-The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \ref{fig:HMMPanels}.C)).
-The initial state for the HMM investor in each instance of the game was set to the ``mid-trust'' state.
-Details on how the HMM state conditional probabilities and transition functions are specified can be found in the supplement.
+The computerised investor consisted of a hidden Markov model (HMM)
+trained on an independent existing behavioral RTG data set of human
+investors. This data-driven approach thus sought to learn an investor
+strategy that mimics human-like interactions. The data set used for
+training consists of 388 ten round games with the same player (full
+details can be found in the Supplementary Information). On this data
+set, the HMM was inferred with three latent states that could be
+interpreted as reflecting a ``low-trust'', a ``medium-trust'', and a
+``high-trust'' state. A separate output distribution, that maps each HMM
+state onto possible investments from 0 to 20 separately, is learned
+(Figure \ref{fig:HMMPanels}.B). In analogy to the latent states, these
+distributions can be interpreted as reflecting ``low-trust'',
+``medium-trust'', or ``high-trust'' dispositions. Finally, the HMM is
+specified by transition probabilities that describe the transition
+between states. The probability of these transitions was modelled as a
+function of their net return (i.e return - investment) in the previous
+round (see Figure \ref{fig:HMMPanels}.C)). The initial state for the
+HMM investor in each instance of the game was set to the ``mid-trust''
+state. Details on how the HMM state conditional probabilities and
+transition functions are specified can be found in the supplement.
 
-In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair it, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) and round 13 (post-manipulation).
-On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
+In order to instigate a potential breakdown of trust, thereby allowing
+us to probe efforts to repair it, the computerised agent was programmed
+to provide a low investment on round 12 (pre-manipulation) and round 13
+(post-manipulation). On all other rounds, the investor's actions were
+determined by randomly drawing an investment from the state-conditional
+distribution, with the state over rounds determined by randomly drawing
+the next state from the state-transition distribution as determined from
+the net return on the previous round (disregarding the net return
+immediately after the pre-programmed low investment rounds).
 
@@
 
-In all phases of the RTG other than the `Exposure phase' (Figure \ref{fig:HMMPanels}.A), participants interacted with this human-like HMM.
-In the `Manipulation' Condition of the exposure phase, however, the parameters of this HMM were adjusted to design a `forgiving' and ultimately more cooperative agent.
-To achieve this, we changed the state transition probabilities of the HMM such that it becomes impossible for it to remain in a low trust state, effectively setting the transition probability for remaining in a ``low-trust'' state to 0.
-The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
-The policies conditional on the latent states and the transition function in the other latent states remain unchanged.
+In all phases of the RTG other than the `Exposure phase' (Figure
+\ref{fig:HMMPanels}.A), participants interacted with this human-like
+HMM. In the `Manipulation' Condition of the exposure phase, however, the
+parameters of this HMM were adjusted to design a `forgiving' and
+ultimately more cooperative agent. To achieve this, we changed the state
+transition probabilities of the HMM such that it becomes impossible for
+it to remain in a low trust state, effectively setting the transition
+probability for remaining in a ``low-trust'' state to 0. The resulting
+transition function is shown in Figure \ref{fig:HMMPanels}.D. The
+policies conditional on the latent states and the transition function in
+the other latent states remain unchanged.
+
+\subsection{Procedure}\label{procedure}
+
+At the start of the experiment, participants provided informed consent
+and were instructed the study would consist of three phases in which
+they would face a different other player. Participants were told their
+goal was to maximise the number of points in all phases. They were not
+told the number of rounds of each phase. Participants were randomly
+assigned to either a Control or Manipulation condition. The timeline of
+the experiment is shown in Figure \ref{fig:HMMPanels}.A. Phase one
+(``pre'') consisted of a 15 round RTG in which participants took the role
+of trustee, facing the same investor over all 15 rounds. On each round,
+after being informed about the amount sent by the investor participants
+decided how much of the tripled investment to return to the investor,
+before continuing to the next round. Phase 2 (``exposure'') consisted of
+three 7-round RTGs. Participants in the Manipulation condition faced the
+forgiving HMM investor and rated the agent on the same attributes as in
+the pre-manipulation phase. Those in the Control condition faced the
+same human-like HMM agent as in the ``pre'' phase and rated each co-player
+on the same attributes. To keep the experience similar to the ``pre''
+phase, the agent in the Control condition was also designed to send a
+very low investment in round 5 of each of the three games. In the
+post-manipulation phase (``post''), participants in both conditions faced
+the same human-like HMM as in ``pre'' phase.
+
+At the beginning of each game in all three phases, participants were
+told they would face a new player and had to wait to be paired with an
+available co-player. This simulated the waiting time in real social
+interaction tasks. After completing each RTG in each phase, participants
+rated how cooperative and forgiving they perceived the co-player to be,
+and whether they would like to play with them again (all on a scale from
+1 to 10 with 10 being the most positive rating). After completing the
+three game phases, participants then completed the Levels of Personality
+Functioning Scale Brief-Form (LPFS-BF) questionnaire
+(Weekers et al., 2019). This is a self-report measure designed to assess
+core elements of personality functioning as defined in the Alternative
+Model for Personality Disorders in the DSM-5
+(American Psychiatric Association, 2013), and provides a
+dimensional assessment of personality functioning, which complements the
+categorical approach of RS. Finally, participants were asked whether
+they thought the other players were human or computer agents, to probe
+how well the agent can mimic human behavior, then debriefed and thanked
+for their participation.
 
@@
 The difference between conditions is the type of agents participants are exposed to in the exposure phase.
-Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG against human trustees.
+Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG.
 From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B.
@@
 
-\subsection{Procedure}\label{procedure}
-
-At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player.
-Participants were told their goal was to maximise the number of points in all phases.
-They were not told the number of rounds of each phase.
-Participants were randomly assigned to either a Control or Manipulation condition.
-The timeline of the experiment is shown in Figure \ref{fig:HMMPanels}.A.
-Phase one (``pre'') consisted of a 15 round RTG in which participants took the role of trustee, facing the same investor over all 15 rounds.
-On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round.
-Phase 2 (``exposure'') consisted of three 7-round RTGs.
-Participants in the Manipulation condition faced the forgiving HMM investor and rated the agent on the same attributes as in the pre-manipulation phase.
-Those in the Control condition faced the same human-like HMM agent as in the ``pre'' phase and rated each co-player on the same attributes.
-To keep the experience similar to the ``pre'' phase, the agent in the Control condition was also designed to send a very low investment in round 5 of each of the three games.
-In the post-manipulation phase (``post''), participants in both conditions faced the same human-like HMM as in ``pre'' phase.
-
-At the beginning of each game in all three phases, participants were told they would face a new player and had to wait to be paired with an available co-player.
-This simulated the waiting time in real social interaction tasks.
-After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-playerr to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
-After completing the three game phases, participants then completed the Levels of Personality Functioning Scale Brief-Form (LPFS-BF) questionnaire (Weekers, Hutsebaut, and Kamphuis 2019).
-This is a self-report measure designed to assess core elements of personality functioning as defined in the Alternative Model for Personality Disorders in the DSM-5 (American Psychiatric Association 2013), and provides a dimensional assessment of personality functioning, which complements the categorical approach of RS.
-Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.
-
 \subsection{Statistical Analysis}\label{statistical-analysis}
 
-To explore whether participants behaved differently in the RTG after the manipulation compared to the Control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.~post-manipulation), Condition (manipulation vs.~Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase.
-The full specification of the statistical model can be found in the supplement.
-To test whether participants behaved differently between conditions in the Exposure phase, we fit a linear mixed effects model to participants returns only in the exposure phase, with Condition (manipulation vs.~Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts.
-Finally, to test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
+To test whether participants behaved differently in the RTG after the
+manipulation compared to the Control group, we model the percentage
+return (percentage of tripled investment returned to investor) using a
+linear mixed-effects model with Phase (RTG game pre vs.
+post-manipulation), Condition (manipulation vs.~Control), Investment,
+and RS (High vs Low RS group) as well as their interactions as fixed
+effects, and player-wise random intercepts and slopes for Phase. The
+full specification of the statistical model can be found in the
+supplement. To test whether participants behaved differently between
+conditions in the Exposure phase, we fit a linear mixed effects model to
+participants returns only in the exposure phase, with Condition
+(manipulation vs.~Control), Investment, and RS (High vs Low RS group) as
+well as their interactions as fixed effects, and player-wise random
+intercepts. Finally, to test whether the HMM agent's behavior differed
+between Phases, Conditions and RS groups, we estimate a linear
+mixed-effects model of investments sent by the computerised HMM agent
+with Condition, Phase and RS and their interaction as fixed effects, and
+a similar random effects structure to the returns model.
 
-The model was estimated using the \texttt{afex} package (Singmann et al.
-2022) in R.
-More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure (Matuschek et al.
-2017).
-A similar process was used to establish the random effects structures of linear mixed-effects models used to analyse the HMM agent investments as well as the participants' ratings of the co-players.
-There is no agreed upon way to calculate effect sizes for mixed effects models.
-Instead, we will report on testing differences in marginal means.
-For the \(F\)-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package ``afex''.
-We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
+The model was estimated using the \texttt{afex} package (Singmann et al., 2022)
+in R. More complex models with additional random effects could not be
+estimated reliably, and as such the estimated model can be considered to
+include the optimal random effects structure
+(Matuschek et al., 2017). A similar process was used to establish the
+random effects structures of linear mixed-effects models used to analyse
+the HMM agent investments as well as the participants' ratings of the
+co-players. There is no agreed upon way to calculate effect sizes for
+mixed effects models. Instead, we will report on testing differences in
+marginal means. For the \(F\)-tests, we used the Kenward-Roger
+approximation to the degrees of freedom, as implemented in the R package
+``afex''. We Z-transform the Investment variable (subtract the overall
+investment mean and divide by overall standard deviation) as centering
+is beneficial to interpreting the main effects more easily in the
+presence of interactions. To probe significant interactions, we
+conducted planned contrasts using the \texttt{emmeans} package in R. Given that
+we were testing multiple pre-planned comparisons, we applied the ``Sidak''
+correction to control for familywise error rate while maintaining
+reasonable statistical power. This approach allowed us to investigate
+specific hypotheses about the differential effects of our manipulation
+across phases and RS groups, while protecting against inflated Type I
+error rates.
 
@@
 
-On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60\% of the endowment for investments and 35-50\% of the total yield for returns, as reported in previous studies (Charness, Cobo-Reyes, and Jiménez 2008; Fiedler, Haruvy, and Li 2011).
+On average, investments and returns, as shown in Figure
+\ref{fig:gamesPlot}, fell within the documented range of 40-60\% of the
+endowment for investments and 35-50\% of the total yield for returns, as
+reported in previous studies (Charness et al., 2008; Fiedler et al., 2011).
 
@@
 
-Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs.~Post RTG game), \(F(1, 201.63) = 5.81\), \(p = .017\), with higher percentage returns in the Pre RTG compared to the Post.
-Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.~post-manipulation), \(F(1, 201.63) = 4.38\), \(p = .038\).
-As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the Manipulation condition, pre - post,
-\(\Delta M = 0.03\), 95\% CI \([0.01, 0.05]\), \(t(201.50) = 3.15\), \(p = .002\), but no change in the Control condition.
-We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.
+Mixed-effects analysis on the percentage returns shows a significant
+main effect of Phase (Pre vs.~Post RTG game),
+\(F(1, 201.63) = 5.81\), \(p = .017\), with
+higher percentage returns in the Pre RTG compared to the Post.
+Importantly, we also find an interaction between Condition and Phase
+(RTG pre- vs.~post-manipulation),
+\(F(1, 201.63) = 4.38\), \(p = .038\).
+As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that,
+contrary to our expectations, there was a decrease in the percentage
+returned only in the Manipulation condition, pre - post,
+\(\Delta M = 0.03\), 95\% CI \([0.01, 0.05]\), \(t(201.50) = 3.15\), \(p = .002\),
+but no change in the Control condition. We find no interaction between
+Phase, Condition and RS, suggesting that there was no difference between
+RS groups for this interaction.
 
@@
 \(F(1, 5955.67) = 325.35\), \(p < .001\),
-such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction,
-\(F(1, 5955.67) = 13.92\), \(p < .001\), reflected that returns were more affected by investments in the Control condition.
-We also find a three way interaction between Phase, Investment and RS, \(F(1, 5864.62) = 7.84\), \(p = .005\), showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation).
-Finally, we find a four-way interaction between Condition, Phase, Investment and RS \(F(1, 5864.62) = 9.24\), \(p = .002\).
-For completeness, these effects are discussed in more detail in the supplement.
+such that higher investments were associated with higher percentage
+returns indicating positive reciprocity. An Investment by Condition
+interaction,
+\(F(1, 5955.67) = 13.92\), \(p < .001\),
+reflected that returns were more affected by investments in the Control
+condition. We also find a three way interaction between Phase,
+Investment and RS,
+\(F(1, 5864.62) = 7.84\), \(p = .005\),
+showing that the differentiated effect of the investment on the
+proportion returned by RS group is itself moderated by the Phase (pre-
+vs post manipulation). Finally, we find a four-way interaction between
+Condition, Phase, Investment and RS
+\(F(1, 5864.62) = 9.24\), \(p = .002\).
+
+To examine this four-way interaction, we conducted a contrast analysis
+of how the effect of investment on returns changed from pre- to
+post-manipulation for different RS groups in both conditions. Starting
+with the Manipulation condition, for participants with low RS, the
+effect of investment on returns increased significantly from pre- to
+post-manipulation,
+\(\Delta M = 0.03\), \(95\%\ \mathrm{CI}_\mathrm{\scriptstyle Sidak(3)}\) \([0.01, 0.05]\), \(t(5881.28) = 3.15\), \(p_\mathrm{\scriptstyle Sidak(3)} = .005\).
+This suggests that after the manipulation, low RS participants became
+more responsive to their co-player's investments, returning
+proportionally more as investments increased. In contrast, for
+participants with high RS, the effect of investment on returns decreased
+significantly from pre- to post-manipulation,
+\(\Delta M = -0.02\), \(95\%\ \mathrm{CI}_\mathrm{\scriptstyle Sidak(3)}\) \([-0.04, 0.00]\), \(t(5891.81) = -2.67\), \(p_\mathrm{\scriptstyle Sidak(3)} = .023\).
+This indicates that high RS participants became less responsive to their
+co-player's investments after the manipulation, with smaller increases
+in returns as investments increased. The difference in these pre-post
+changes between high and low RS groups was significant,
+\(\Delta M = -0.05\), \(95\%\ \mathrm{CI}_\mathrm{\scriptstyle Sidak(3)}\) \([-0.08, -0.02]\), \(t(5887.33) = -4.11\), \(p_\mathrm{\scriptstyle Sidak(3)} < .001\).
+This result suggests that the manipulation had significantly different
+effects on how low and high RS participants responded to their
+co-player's investments.
+
+In the Control condition, we observed no significant changes in how
+participants responded to their co-player's investments between the pre
+and post phases, regardless of their RS level.
 
@@
 The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns.
-The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.}\label{fig:boxPlots}
+The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.
+RS did not affect returns and investments by Phase and Condition.}\label{fig:boxPlots}
 \end{figure}
@@
 
-Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor?
-To explore this question, we restrict the analysis to the trials following the pre-programmed defection by the HMM agent in both the ``pre'' (trials 12 to 15) and the ``post'' phases (trials 13 to 15).
-We fit the same mixed effects model as for all the trials with the exception of the RS variable.
-This is because RS did not show main or interaction effects in the main model, and also due to the necessity of running a simpler model to accommodate the low number of trials.
-We find a significant main effect of Phase \(F(1, 227.47) = 4.76\), \(p = .030\) with returns lower in the second game post defection trials compared to the first.
-We also find a main effect of Investment
+Did participants learn to be more forgiving and cooperative after
+witnessing the pre-programmed defection by the HMM investor? To explore
+this question, we restrict the analysis to the trials following the
+pre-programmed defection by the HMM agent in both the ``pre'' (trials 12
+to 15) and the ``post'' phases (trials 13 to 15). We fit the same mixed
+effects model as for all the trials with the exception of the RS
+variable. This is because RS did not show main effects in the previous
+model, and also due to the necessity of running a simpler model to
+accommodate the low number of trials. We find a significant main effect
+of Phase \(F(1, 227.47) = 4.76\), \(p = .030\)
+with returns lower in the second game post defection trials compared to
+the first. We also find a main effect of Investment
 \(F(1, 1296.14) = 156.41\), \(p < .001\)
-where participants continued to return higher proportions when receiving higher investments.
-Finally, we still find an Investment by Condition interaction
-\(F(1, 5955.67) = 13.92\), \(p < .001\) showing a lower effects of investment on the Manipulation condition compared to the Control condition in post-defection trials.
-However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.
+where participants continued to return higher proportions when receiving
+higher investments. Finally, we still find an Investment by Condition
+interaction
+\(F(1, 5955.67) = 13.92\), \(p < .001\)
+showing a lower effects of investment on the Manipulation condition
+compared to the Control condition in post-defection trials. However, the
+absence of a Condition by Phase effect indicates there was no difference
+between conditions on participants' reaction to a one-off low investment
+by the co-player.
 
@@
 
-Was the HMM's strategy similar between pre and post phases in the control condition?
-Was participants' behavior post exposure differentiated enough to induce a different reaction from the HMM?
-To answer these questions, we test for differences in the HMM agent's investment by Phase, Condition and RS using a linear mixed-effects model as described in the methods section.
-As seen in Figure \ref{fig:boxPlots}, we find no main or interaction effects, indicating the HMM's behavior was on aggregate similar across levels of Phase, Condition and RS.
-This consistency in the investor's behavior is a desirable feature of the HMM agent when the participants behavior is largely simialr between phases.
-More importantly, it indicates that the lower returns of participants in the post phase of the manipulation condition were not differentiated enough to make the HMM react by transitioning to lower latent trust states.
-It is also noteworthy that the HMM agent was relatively successful in imitating human behavior in this paradigm: When asked during debrief whether they thought the investors they faced were human or not, \(41\)\% of participants thought they were either facing a human or were not sure of the nature of the co-player.
-When asked to justify their choice, many answers reflected participants projecting human traits such as ``spitefulness'' or ``greed'' onto the artificial co-player's behavior.
+Was the HMM's strategy similar between pre and post phases in the
+control condition? Was participants' behavior post exposure
+differentiated enough to induce a different reaction from the HMM? To
+answer these questions, we test for differences in the HMM agent's
+investment by Phase, Condition and RS using a linear mixed-effects model
+as described in the methods section. As seen in Figure
+\ref{fig:boxPlots}, we find no main or interaction effects, indicating
+the HMM's behavior was on aggregate similar across levels of Phase,
+Condition and RS. This consistency in the investor's behavior is a
+desirable feature of the HMM agent when the participants behavior is
+largely simialr between phases. More importantly, it indicates that the
+lower returns of participants in the post phase of the manipulation
+condition were not differentiated enough to make the HMM react by
+transitioning to lower latent trust states. It is also noteworthy that
+the HMM agent was relatively successful in imitating human behavior in
+this paradigm: When asked during debrief whether they thought the
+investors they faced were human or not, \(41\)\% of participants thought
+they were either facing a human or were not sure of the nature of the
+co-player. When asked to justify their choice, many answers reflected
+participants projecting human traits such as ``spitefulness'' or ``greed''
+onto the artificial co-player's behavior.
 
@@
 So far we focused on analysing behavior for the pre and post phases.
-Here, we look at returns and investments in the exposure phase.
-The linear mixed effects model of participants' returns in the exposure phase does not show a main effect of Condition on returns.
-There was a main effect of Investment, \(F(1, 4117.20) = 233.19\), \(p < .001\), with participants positively reciprocating higher investments, an interaction effect between Condition and Investment \(F(1, 4117.20) = 45.93\), \(p < .001\), showing a stronger positive reciprocity in the Control condition, and finally a three way interaction between the RS group, Condition and Investment \(F(1, 4117.20) = 4.21\), \(p = .040\), showing that this stronger positive reciprocity to investment in the Control condition is higher for participants with high RS.
-The linear mixed effects model of the HMM investments shows a main effect of Condition \(F(1, 202) = 197.64\), \(p < .001\), suggesting higher overall investments for the forgiving HMM compared to the human-like HMM, but no difference in investments when facing low and high RS groups.
+Here, we look at returns and investments in the exposure phase. The
+linear mixed effects model of participants' returns in the exposure
+phase does not show a main effect of Condition on returns. There was a
+main effect of Investment,
+\(F(1, 4117.20) = 233.19\), \(p < .001\), with
+participants positively reciprocating higher investments, an interaction
+effect between Condition and Investment
+\(F(1, 4117.20) = 45.93\), \(p < .001\),
+showing a stronger positive reciprocity in the Control condition, and
+finally a three way interaction between the RS group, Condition and
+Investment
+\(F(1, 4117.20) = 4.21\), \(p = .040\),
+showing that this stronger positive reciprocity to investment in the
+Control condition is higher for participants with high RS. The linear
+mixed effects model of the HMM investments shows a main effect of
+Condition \(F(1, 202) = 197.64\), \(p < .001\),
+suggesting higher overall investments for the forgiving HMM compared to
+the human-like HMM, but no difference in investments when facing low and
+high RS groups.
 
-In summary, despite the forgiving HMM sending overall higher investments in the exposure phase, participants returned similar proportions of the multiplied investments as those facing the human-like HMM.
-The positive reciprocity of returns to investments was higher in the Control condition with this relationship stronger for the high RS group.
+In summary, despite the forgiving HMM sending overall higher investments
+in the exposure phase, participants returned similar proportions of the
+multiplied investments as those facing the human-like HMM. The positive
+reciprocity of returns to investments was higher in the Control
+condition with this relationship stronger for the high RS group.
 
@@
 
-Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity Questionnaire score (RSQ), Spearman's \(r_{\mathrm{s}} = .52\), \(p < 0.001\), there was no correlation between these questionnaire scores and participant's return or overall task performance.
+Whilst we found a significant correlation between participant's Levels
+of Personality Functioning Score (LPFS) and the Rejection Sensitivity
+Questionnaire score (RSQ), Spearman's \(r_{\mathrm{s}} = .52\), \(p < 0.001\),
+there was no correlation between these questionnaire scores and
+participant's return or overall task performance.
 
@@
 
-Figure \ref{fig:plotRatings} shows participants' ratings of each player they faced by condition and RS group.
-We will focus on two contrasts to analyse the ratings by Condition and RS group.
-The first is between the rating in the first phase (``pre'') when participants phase the human-like HMM, and the average rating during the exposure phase (average of ``expo1'', ``expo2'' and ``expo3'') where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition).
-The second contrast is between the ``pre'' and ``post'' phases of the experiment where in both conditions participants face the same human-like HMM.
+Figure \ref{fig:plotRatings} shows participants' ratings of each player
+they faced by condition and RS group. We will focus on two contrasts to
+analyse the ratings by Condition and RS group. The first is between the
+rating in the first phase (``pre'') when participants phase the human-like
+HMM, and the average rating during the exposure phase (average of
+``expo1'', ``expo2'' and ``expo3'') where they either face the forgiving HMM
+(Manipulation condition) or the human-like HMM again (Control
+condition). The second contrast is between the ``pre'' and ``post'' phases
+of the experiment where in both conditions participants face the same
+human-like HMM.
 
@@
 
-For those with high RS, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative \(\Delta M = 2.57\), 95\% CI \([0.84, 4.30]\), \(t(808) = 2.91\), \(p = .004\).
-There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
-Those in the Control condition rated the investors faced in the exposure group (same HMM) as less cooperative \(\Delta M = -2.65\), 95\% CI \([-4.37, -0.94]\), \(t(808) = -3.04\), \(p = .002\), less forgiving \(\Delta M = -2.19\), 95\% CI \([-4.00, -0.39]\), \(t(808) = -2.38\), \(p = .017\), and were less keen on facing them again \(\Delta M = -3.62\), 95\% CI \([-5.73, -1.50]\), \(t(808) = -3.36\), \(p < .001\).
+For those with high RS, participants in the Manipulation condition rated
+the investors they faced in the exposure phase (the forgiving HMM) as
+more Cooperative
+\(\Delta M = 2.57\), 95\% CI \([0.84, 4.30]\), \(t(808) = 2.91\), \(p = .004\).
+There was no difference in ratings on forgiveness and whether they would
+like to face the co-players again. Those in the Control condition rated
+the investors faced in the exposure group (same HMM) as less cooperative
+\(\Delta M = -2.65\), 95\% CI \([-4.37, -0.94]\), \(t(808) = -3.04\), \(p = .002\),
+less forgiving
+\(\Delta M = -2.19\), 95\% CI \([-4.00, -0.39]\), \(t(808) = -2.38\), \(p = .017\),
+and were less keen on facing them again
+\(\Delta M = -3.62\), 95\% CI \([-5.73, -1.50]\), \(t(808) = -3.36\), \(p < .001\).
 
-For those with low RS, there was no difference in any of the ratings between the ``pre'' and ``exposure'' phases for the Manipulation condition.
-For the Control condition, participants indicated less willingness to face the exposure co-player compared to the ``pre'' player, \(\Delta M = -2.43\), 95\% CI \([-4.52, -0.34]\), \(t(808) = -2.29\), \(p = .023\) but also did not differ on cooperation and forgiveness ratings.
-In summary, those with low RS had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving.
-In contrast, we see that exposing participants with high RS to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.
+For those with low RS, there was no difference in any of the ratings
+between the ``pre'' and ``exposure'' phases for the Manipulation condition.
+For the Control condition, participants indicated less willingness to
+face the exposure co-player compared to the ``pre'' player,
+\(\Delta M = -2.43\), 95\% CI \([-4.52, -0.34]\), \(t(808) = -2.29\), \(p = .023\)
+but also did not differ on cooperation and forgiveness ratings. In
+summary, those with low RS had a mostly undifferentiated perception of
+players between the pre and exposure phases, even when the co-player was
+in fact designed to be more forgiving. In contrast, we see that exposing
+participants with high RS to a more cooperative and more forgiving agent
+has compensated for the decrease in ratings that would have occurred if
+faced with a co-player with the exact same strategy.
 
@@
 
-Participants high on RS in the Manipulation condition rated the co-players in the ``post'' phase similarly on cooperation, lower on
-forgiveness \(\Delta M = -0.88\), 95\% CI \([-1.63, -0.14]\), \(t(808) = -2.33\), \(p = .020\),
-and lower on willingness to face them again \(\Delta M = -1.29\), 95\% CI \([-2.16, -0.42]\), \(t(808) = -2.92\), \(p = .004\).
-Those in the Control condition rated the investors post the exposure phase lower on all three attributes
-(Cooperation: \(\Delta M = -0.96\), 95\% CI \([-1.66, -0.26]\), \(t(808) = -2.70\), \(p = .007\),
-Forgiveness: \(\Delta M = -1.00\), 95\% CI \([-1.74, -0.26]\), \(t(808) = -2.66\), \(p = .008\),
-Play again: \(\Delta M = -1.21\), 95\% CI \([-2.07, -0.35]\), \(t(808) = -2.76\), \(p = .006\)).
+Participants high on RS in the Manipulation condition rated the
+co-players in the ``post'' phase similarly on cooperation, lower on
+forgiveness
+\(\Delta M = -0.88\), 95\% CI \([-1.63, -0.14]\), \(t(808) = -2.33\), \(p = .020\),
+and lower on willingness to face them again
+\(\Delta M = -1.29\), 95\% CI \([-2.16, -0.42]\), \(t(808) = -2.92\), \(p = .004\).
+Those in the Control condition rated the investors post the exposure
+phase lower on all three attributes (Cooperation:
+\(\Delta M = -0.96\), 95\% CI \([-1.66, -0.26]\), \(t(808) = -2.70\), \(p = .007\),
+Forgiveness:
+\(\Delta M = -1.00\), 95\% CI \([-1.74, -0.26]\), \(t(808) = -2.66\), \(p = .008\),
+Play again:
+\(\Delta M = -1.21\), 95\% CI \([-2.07, -0.35]\), \(t(808) = -2.76\), \(p = .006\)).
 
-For participants low on RS, those in the Manipulation condition rated the co-players in the ``post'' phase lower on all three attributes
+For participants low on RS, those in the Manipulation condition rated
+the co-players in the ``post'' phase lower on all three attributes
 (Cooperation:\(\Delta M = -0.74\), 95\% CI \([-1.45, -0.03]\), \(t(808) = -2.03\), \(p = .042\),
-Forgiveness: \(\Delta M = -1.14\), 95\% CI \([-1.89, -0.39]\), \(t(808) = -2.98\), \(p = .003\),
-Play again: \(\Delta M = -1.30\), 95\% CI \([-2.18, -0.42]\), \(t(808) = -2.90\), \(p = .004\)).
-Those in the Control condition did not differ in their ratings of the ``pre'' and ``post'' phase players.
+Forgiveness:
+\(\Delta M = -1.14\), 95\% CI \([-1.89, -0.39]\), \(t(808) = -2.98\), \(p = .003\),
+Play again:
+\(\Delta M = -1.30\), 95\% CI \([-2.18, -0.42]\), \(t(808) = -2.90\), \(p = .004\)).
+Those in the Control condition did not differ in their ratings of the
+``pre'' and ``post'' phase players.
 
-In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the Control condition.
-In contrast, the high RS group shows a negative bias towards the co-players after the ``pre'' phase in the Control condition even though the player continues to use the same strategy.
-After exposure to the forgiving HMM, both groups rate the ``post'' co-player worse than the ``pre'' even though they are the same.
+In summary, we again see that those with low RS accurately perceive the
+co-player as similar on all attributes throughout the phases in the
+Control condition. In contrast, the high RS group shows a negative bias
+towards the co-players after the ``pre'' phase in the Control condition
+even though the player continues to use the same strategy. After
+exposure to the forgiving HMM, both groups rate the ``post'' co-player
+worse than the ``pre'' even though they are the same.
 
@@
 
-We used a randomized controlled online experiment where participants played a RTG with artificial agents designed to simulate human-like trust-building scenarios.
-Participants were then exposed to either forgiving or human-like HMM agents before playing another RTG.
-Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation.
-Instead, participants reduced their returns overall whilst the returns of those in the Control group did not change between the pre and post phase of the experiment.
-Neither did participants show more forgiving behavior in the Manipulation condition after the one-off defection by the agent.
-Why did participants reduce their returns even though they were repeatedly exposed to more cooperative and more forgiving artificial agents?
-A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the Manipulation condition.
+We used a randomized controlled online experiment where participants
+played a RTG with artificial agents designed to simulate human-like
+trust-building scenarios. Participants were then exposed to either
+forgiving or human-like HMM agents before playing another RTG. We found
+that RS did not moderate participants' returns as trustees in the
+repeated trust game. While previous research has shown that RS affects
+\emph{trust} formation, appraisal and repair, its impact on \emph{reciprocity} in
+repeated economic exchanges has been less explored. Our results suggest
+a potential dissociation between RS's known effects on broader social
+behavior and its limited influence on reciprocity in structured,
+repeated interactions, challenging assumptions about the pervasive
+influence of RS on social behavior and highlighting the complexity of
+factors influencing reciprocity in economic exchanges.
 
-Those exposed to the forgiving agent rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure.
-One possible explanation for this drop in rating is that participants exhibited a negative contrast effect.
-This occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people.
-If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favourable or even negative by comparison (Kobre and Lipsitt 1972).
-As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving agents (Zentall 2005).
-If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.
+Contrary to our hypothesis, exposure to forgiving agents did not
+increase participant's reciprocity or cooperation, nor did it prompt the
+artificial agent to increase their trust in participants through higher
+investments. Instead, participants reduced their returns overall whilst
+the returns of those in the Control group did not change between the pre
+and post phase of the experiment. Why did participants reduce their
+returns even though they were repeatedly exposed to more cooperative and
+more forgiving artificial agents? A look at how the participants rated
+their co-players might shed some light on what might be driving this
+reduction in returns for those in the Manipulation condition. Those
+exposed to the forgiving agent rated their opponent in the post-exposure
+phase lower on all attributes even though they faced the same dynamic
+human-like HMM as pre-exposure. One possible explanation for this drop
+in rating is that participants exhibited a negative contrast effect.
+This occurs when the evaluation of a person, object, or situation is
+influenced by comparisons with recently encountered contrasting objects
+or people. If we've repeatedly interacted with someone exceptionally
+nice, our perception of a normal level of niceness might be skewed,
+making typical behavior seem less favourable or even negative by
+comparison (Kobre \& Lipsitt, 1972). As the most recently faced opponents
+were more forgiving and cooperative, this negative contrast effect may
+have trumped any learning transfer from being repeatedly exposed to
+forgiving agents (Zentall, 2005). If this contrast effect
+is indeed replicable, then an avenue for future research would be to use
+it to our benefit by making the participants play agents with low
+cooperation perception.
 
-It is worth highlighting that RS was not correlated to overall performance in the task.
-Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
-However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their HMM co-players: They accurately perceived increased cooperativeness in the forgiving agent, showing an ability to adjust their judgments based on interaction patterns.
-This may indicate that perceived support from cooperative partners can mitigate their rejection concerns.
-However, their declining ratings of similar co-players over time, despite consistent agent behavior, suggests that occasional pre-programmed defections may have amplified their perceptions of untrustworthiness and reluctance for future interactions.
-These nuanced perceptions align with the heightened social awareness often associated with high rejection sensitivity (Downey and Feldman 1996).
-Individuals with low RS were less affected by the agent's behavioral variations, showing a less volatile baseline of social perception.
-They still exhibit behavior consistent with a contrast effect in the Manipulation condition.
-This stability in perception might reflect a lower propensity to scrutinize social cues or a more robust internal model of social interactions that is less easily perturbed by short-term variations in partner behavior.
+This effect appears to operate independently of RS, as both groups
+exhibited it. However, RS did influence how participants perceived and
+reacted to their co-players' behavior. High RS individuals demonstrated
+a heightened attunement to changes in their co-players' behavior,
+accurately perceiving increased cooperativeness in the forgiving agent.
+Paradoxically, this same sensitivity may explain their decreased
+reactivity to investment levels and declining ratings of similar
+co-players over time in the Control condition. Occasional pre-programmed
+defections may have amplified their perceptions of untrustworthiness,
+leading to a more cautious approach and less responsiveness to
+investment variations. This aligns with the heightened social awareness
+and vigilance often associated with high RS (Downey \& Feldman, 1996).
+This finding suggests that individuals with high RS might benefit from
+interventions that focus on calibrating their perceptions of social cues
+in economic exchanges. In contrast, low RS individuals were less
+affected by the agent's behavioral variations, showing a more stable
+baseline of social perception. This stability in perception might
+reflect a lower propensity to scrutinize social cues or a more robust
+internal model of social interactions that is less easily perturbed by
+short-term variations in partner behavior. Their increased
+responsiveness to investments post-exposure in the Manipulation
+condition suggests they may have internalized the cooperative norms
+experienced, carrying over this increased reciprocity to subsequent
+interactions.
 
-Despite these differences in perception, the actual cooperative behavior -- as measured by returns in the trust game -- did not significantly differ between RS groups.
-This suggests a potential dissociation between the cognitive-perceptual processes involved in evaluating social partners and the behavioral output in economic games.
-There are multiple potential explanations for this: First, the ratings may reflect explicit, conscious evaluations of social partners, while behavior in the trust game might be guided more by implicit, automatic processes that are less differentiated between RS groups.
-This interpretation aligns with dual-process theories of social cognition (Lieberman 2007).
-Second, the structured nature of the trust game might provide a context where the behavioral expression of rejection sensitivity is attenuated.
-In contrast, the more open-ended nature of providing ratings might allow for greater expression of RS-related perceptual biases.
+Despite these differences in perception and investment reactivity, the
+actual cooperative behavior -- as measured by returns in the trust game --
+did not significantly differ between RS groups. This suggests a
+potential dissociation between the cognitive-perceptual processes
+involved in evaluating social partners and the behavioral output in
+economic games. There are multiple potential explanations for this:
+First, the ratings may reflect explicit, conscious evaluations of social
+partners, while behavior in the trust game might be guided more by
+implicit, automatic processes that are less differentiated between RS
+groups. This interpretation aligns with dual-process theories of social
+cognition (Lieberman, 2007). Second, the structured nature of the
+trust game might provide a context where the behavioral expression of RS
+is attenuated. In contrast, the more open-ended nature of providing
+ratings might allow for greater expression of RS-related perceptual
+biases.
 
-These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged.
-Conversely, therapies targeting these perceptual processes might not necessarily translate into immediate behavioral changes.
-Future research should investigate whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods.
-Additionally, exploring specific conditions or thresholds under which perceptual differences begin to manifest in behavioral divergences could provide valuable insights into the dynamics of rejection sensitivity.
-Simply exposing individuals to highly cooperative or forgiving partners may not suffice to induce lasting changes in behavior.
-Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time.
-Combining periods of forgiveness with clear communication about mutual expectations for cooperation might be more effective (Balliet, Mulder, and Van Lange 2011).
-Future studies should also consider the temporal dynamics of these effects, examining whether longer exposure periods or different patterns of forgiving behavior could lead to more positive outcomes.
-Moreover, investigating the role of explicit communication and meta-cognitive reflection on the cooperative process could offer insights into leveraging positive interactions to promote enduring cooperative tendencies.
+These findings have important implications for interventions aimed at
+promoting trust and cooperation. Simply exposing individuals to
+artificially highly cooperative or forgiving partners may not suffice to
+induce lasting changes in behavior, and may even lead to negative
+contrast effects. Instead, interventions should focus on developing
+adaptive strategies for navigating the complexities of real-world social
+interactions, which often involve a mix of cooperative and
+non-cooperative behaviors. Such approaches could gradually shape
+expectations and behaviors over time, taking into account individual
+differences in RS. For individuals with high RS, interventions might
+specifically target the calibration of social perceptions in economic
+exchanges, helping to bridge the gap between their heightened
+sensitivity to social cues and their actual cooperative behavior. Future
+research should investigate whether the observed dissociation between
+perception and behavior persists in more naturalistic social
+interactions or over longer periods. Additionally, exploring the
+temporal dynamics of these effects and the role of explicit
+communication in the cooperative process could offer valuable insights.
+Interventions combining realistic cooperative scenarios with clear
+communication about mutual expectations might prove more effective in
+fostering enduring cooperative tendencies that are robust across various
+social contexts (Balliet et al., 2011).
+
+\subsection{Limitations}\label{limitations}
+
+While this study offers valuable insights into trust and cooperation
+dynamics in social interactions mediated by artificial agents, several
+limitations warrant consideration. The artificial nature of the
+interactions, while enabling high experimental control, may limit
+ecological validity, particularly regarding the paradigm's ability to
+trigger rejection sensitivity in participants. Additionally, the
+relatively brief exposure to forgiving HMM agents during the
+manipulation phase may have been insufficient to induce lasting changes
+in participants' reciprocity behavior. The online format of the
+experiment eliminates important social cues present in face-to-face
+interactions, which might influence trust and cooperation differently.
+Although the RSQ is a well-validated measure of rejection sensitivity
+(RS), future studies could benefit from a multi-method approach,
+potentially incorporating behavioral measures or additional self-report
+scales for a more comprehensive RS assessment. Furthermore, while our
+use of HMM agents to simulate human-like behavior in economic games is
+innovative, some participants may have suspected they were interacting
+with AI, potentially confounding the results. Despite these limitations,
+our findings provide a foundation for future research. Subsequent
+studies could address these constraints by incorporating face-to-face
+interactions, utilizing multiple measures of RS, examining the long-term
+effects of exposure to forgiving interaction partners, and developing
+more convincing cover stories or explicitly comparing responses to human
+and computerised partners.
+
+\subsection{Constraints on Generality}\label{constraints-on-generality}
+
+The results may be specific to adults with high or low RS recruited from
+online platforms, and may not generalize to clinical populations,
+children, or older adults. Our use of a computerized Repeated Trust Game
+with HMM agents, while allowing for high experimental control, may limit
+generalizability to face-to-face interactions or games with different
+economic structures. The brief exposure phase and pre-programmed
+defections are specific to our design and may not reflect real-world
+trust-building scenarios. The online context may not capture all aspects
+of high-stakes or information-rich interactions. We believe the core
+finding of decreased cooperation after exposure to forgiving agents
+should generalize across different populations and contexts, though the
+effect's strength and its interaction with RS may vary. While the
+specific economic game, agent representation, and perception assessment
+questions could be varied, the use of artificial agents with consistent
+behavior, an exposure phase with more forgiving behavior, and assessment
+of both behavior and perceptions should remain constant to preserve the
+results. Future studies could systematically vary these factors to
+establish the boundaries of generalizability for our findings.
+Additionally, cultural differences in norms of cooperation and trust may
+influence the generalizability of these findings, necessitating
+cross-cultural replications to establish the universality of the
+observed effects.
 
@@
 
-This randomised controlled experiment enabled us to uncover unexpected effects of exposure to forgiving behavior on subsequent cooperation, particularly in relation to rejection sensitivity.
-These findings challenge existing assumptions about fostering cooperative behavior and suggest the need for more nuanced interventions.
-Importantly, the use of HMM-based artificial agents in this study represents a significant methodological advancement.
-By providing a balance between experimental control and realistic, adaptive behavior, these agents allowed for a nuanced exploration of trust dynamics that would be challenging to achieve with human confederates or simplistic computer algorithms.
-This approach opens up new possibilities for studying complex social interactions in controlled settings, potentially bridging the gap between laboratory experiments and real-world social dynamics.
+This randomised controlled experiment enabled us to uncover unexpected
+effects of exposure to forgiving behavior on subsequent cooperation,
+particularly in relation to RS. These findings challenge existing
+assumptions about fostering cooperative behavior and suggest the need
+for more nuanced interventions. Importantly, the use of HMM-based
+artificial agents in this study represents a significant methodological
+advancement. By providing a balance between experimental control and
+realistic, adaptive behavior, these agents allowed for a nuanced
+exploration of trust dynamics that would be challenging to achieve with
+human confederates or simplistic computer algorithms. This approach
+opens up new possibilities for studying complex social interactions in
+controlled settings, potentially bridging the gap between laboratory
+experiments and real-world social dynamics.
 
@@
 
-I.
-Guennouni, G.
-Koppe and C.
-Korn.
-designed and developed the study concept.
-Experiment design, testing and data collection were performed by I.
-Guennouni.
-I.
-Guennouni analysed and interpreted the data under the supervision of G.
-Koppe and C.
-Korn.
-All authors jointly wrote and approved the final version of the manuscript for submission.
+I. Guennouni, G. Koppe and C. Korn. designed and developed the study
+concept. Experiment design, testing and data collection were performed
+by I. Guennouni. I. Guennouni analysed and interpreted the data under
+the supervision of G. Koppe and C. Korn. All authors jointly wrote and
+approved the final version of the manuscript for submission.
 
@@
 
-This publication was supported through state funds approved by the State Parliament of Baden-Württemberg for the Innovation Campus Health + Life Science alliance Heidelberg Mannheim.
-I.
-Guennouni was supported by a research fellowship from the AI Health Innovation Cluster
+This study was supported by the Federal Ministry of Science, Education,
+and Culture (MWK) of the state of Baden-Württemberg within the AI Health
+Innovation Cluster Initiative, the German Research Foundation (DFG)
+within the Excellence Strategy EXC 2181/1 -- 390900948 (STRUCTURES), and
+the Hector II foundation. I. Guennouni was supported by a research
+fellowship from the AI Health Innovation Cluster.
 
@@
 
-The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.
+The author(s) declared that there were no conflicts of interest with
+respect to the authorship or the publication of this article.
+
+\section*{Acknowledgements}\label{acknowledgements}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+We are grateful to Tobias Nolte, Andreas Hula and Read Montague's team
+at Virginia Tech for sharing with us the data on the RTG allowing us to
+fit the HMM to investor data. We're also grateful for Samuel Dupret for
+help designing the online platform used for the experiment and Maarten
+Speekenbrink for his guidance on the use of HMM models.
 
@@
 
-All correspondence and requests for materials should be addressed to I. Guennouni.
+All correspondence and requests for materials should be addressed to I.
+Guennouni.
 
@@
 
-Preregistration: The hypotheses and methods were not preregistered.The analysis plan was not preregistered.
-Materials: All study materials are publicly available (\url{https://github.com/ismailg/exposure-public}).
-Data: All primary data are publicly available (\url{https://github.com/ismailg/exposure-public}).
-Analysis scripts: All analysis scripts are publicly available (\url{https://github.com/ismailg/exposure-public}).
+Preregistration: The hypotheses and methods were not preregistered.The
+analysis plan was not preregistered. Materials: All study materials are
+publicly available (\url{https://github.com/ismailg/exposure-public}). Data:
+All primary data are publicly available
+(\url{https://github.com/ismailg/exposure-public}). Analysis scripts: All
+analysis scripts are publicly available
+(\url{https://github.com/ismailg/exposure-public}).
 
@@
 \bibitem[\citeproctext]{ref-abramov_influence_2022-1}
-Abramov, Gamze, Jason Kautz, Sebastien Miellet, and Frank P.
-Deane.
-2022.
-{``The {Influence} of {Attachment Style}, {Self-protective Beliefs}, and {Feelings} of {Rejection} on the {Decline} and {Growth} of {Trust} as a {Function} of {Borderline Personality Disorder Trait Count}.''} \emph{Journal of Psychopathology and Behavioral Assessment} 44 (3): 773--86.
-\url{https://doi.org/10.1007/s10862-022-09965-9}.
+Abramov, G., Kautz, J., Miellet, S., \& Deane, F.
+P.
+(2022).
+The {Influence} of {Attachment Style}, {Self-protective Beliefs}, and {Feelings} of {Rejection} on the {Decline} and {Growth} of {Trust} as a {Function} of {Borderline Personality Disorder Trait Count}.
+\emph{Journal of Psychopathology and Behavioral Assessment}, \emph{44}(3), 773--786.
+\url{https://doi.org/10.1007/s10862-022-09965-9}
 
 \bibitem[\citeproctext]{ref-almaatouq_empirica_2021}
-Almaatouq, Abdullah, Joshua Becker, James P.
-Houghton, Nicolas Paton, Duncan J.
-Watts, and Mark E.
-Whiting.
-2021.
-{``Empirica: A Virtual Lab for High-Throughput Macro-Level Experiments.''} \emph{Behavior Research Methods} 53 (5): 2158--71.
-\url{https://doi.org/10.3758/s13428-020-01535-9}.
+Almaatouq, A., Becker, J., Houghton, J.
+P., Paton, N., Watts, D.
+J., \& Whiting, M.
+E.
+(2021).
+Empirica: A virtual lab for high-throughput macro-level experiments.
+\emph{Behavior Research Methods}, \emph{53}(5), 2158--2171.
+\url{https://doi.org/10.3758/s13428-020-01535-9}
 
@@
 American Psychiatric Association.
-2013.
-\emph{Diagnostic and {Statistical Manual} of {Mental Disorders}}.
-Fifth Edition.
+(2013).
+\emph{Diagnostic and {Statistical Manual} of {Mental Disorders}} (Fifth Edition).
 American Psychiatric Association.
-\url{https://doi.org/10.1176/appi.books.9780890425596}.
+\url{https://doi.org/10.1176/appi.books.9780890425596}
 
 \bibitem[\citeproctext]{ref-balliet_reward_2011}
-Balliet, Daniel, Laetitia B.
-Mulder, and Paul A.
+Balliet, D., Mulder, L.
+B., \& Van Lange, P.
+A.
 M.
-Van Lange.
-2011.
-{``Reward, Punishment, and Cooperation: {A} Meta-Analysis.''} \emph{Psychological Bulletin} 137 (4): 594--615.
-\url{https://doi.org/10.1037/a0023489}.
+(2011).
+Reward, punishment, and cooperation: {A} meta-analysis.
+\emph{Psychological Bulletin}, \emph{137}(4), 594--615.
+\url{https://doi.org/10.1037/a0023489}
 
 \bibitem[\citeproctext]{ref-bandura_social_1977}
-Bandura, Albert. 1977. \emph{Social {Learning Theory}}. Prentice Hall.
+Bandura, A. (1977). \emph{Social {Learning Theory}}. Prentice Hall.
+
+\bibitem[\citeproctext]{ref-berenson_rejection_2009}
+Berenson, K.
+R., Gyurak, A., Ayduk, Ö., Downey, G., Garner, M.
+J., Mogg, K., Bradley, B.
+P., \& Pine, D.
+S.
+(2009).
+Rejection sensitivity and disruption of attention by social threat cues.
+\emph{Journal of Research in Personality}, \emph{43}(6), 1064--1072.
+\url{https://doi.org/10.1016/j.jrp.2009.07.007}
 
 \bibitem[\citeproctext]{ref-bowlby_attachment_1978}
-Bowlby, John. 1978. {``Attachment Theory and Its Therapeutic Implications.''} \emph{Adolescent Psychiatry} 6: 5--33.
+Bowlby, J. (1978). Attachment theory and its therapeutic implications. \emph{Adolescent Psychiatry}, \emph{6}, 5--33.
+
+\bibitem[\citeproctext]{ref-bowlby_secure_1988}
+Bowlby, J. (1988). \emph{A secure base: Parent-child attachment and healthy human development}. Basic Books.
 
 \bibitem[\citeproctext]{ref-charness_investment_2008}
-Charness, Gary, Ramón Cobo-Reyes, and Natalia Jiménez.
-2008.
-{``An Investment Game with Third-Party Intervention.''} \emph{Journal of Economic Behavior \& Organization} 68 (1): 18--28.
-\url{https://doi.org/10.1016/j.jebo.2008.02.006}.
+Charness, G., Cobo-Reyes, R., \& Jiménez, N.
+(2008).
+An investment game with third-party intervention.
+\emph{Journal of Economic Behavior \& Organization}, \emph{68}(1), 18--28.
+\url{https://doi.org/10.1016/j.jebo.2008.02.006}
 
 \bibitem[\citeproctext]{ref-downey_implications_1996}
-Downey, Geraldine, and Scott I.
-Feldman.
-1996.
-{``Implications of Rejection Sensitivity for Intimate Relationships.''} \emph{Journal of Personality and Social Psychology} 70 (6): 1327--43.
-\url{https://doi.org/10.1037/0022-3514.70.6.1327}.
+Downey, G., \& Feldman, S.
+I.
+(1996).
+Implications of rejection sensitivity for intimate relationships.
+\emph{Journal of Personality and Social Psychology}, \emph{70}(6), 1327--1343.
+\url{https://doi.org/10.1037/0022-3514.70.6.1327}
 
 \bibitem[\citeproctext]{ref-downey_early_1997}
-Downey, Geraldine, Hala Khouri, and Scott I.
-Feldman.
-1997.
-{``Early Interpersonal Trauma and Later Adjustment: {The} Mediational Role of Rejection Sensitivity.''} In \emph{Developmental Perspectives on Trauma: {Theory}, Research, and Intervention}, 85--114.
-Rochester Symposium on Developmental Psychology, {Vol}.
-8.
-Rochester, NY, US: University of Rochester Press.
+Downey, G., Khouri, H., \& Feldman, S.
+I.
+(1997).
+Early interpersonal trauma and later adjustment: {The} mediational role of rejection sensitivity.
+In \emph{Developmental perspectives on trauma: {Theory}, research, and intervention} (pp.
+85--114).
+University of Rochester Press.
+
+\bibitem[\citeproctext]{ref-faul_statistical_2009}
+Faul, F., Erdfelder, E., Buchner, A., \& Lang, A.-G.
+(2009).
+Statistical power analyses using {G}*{Power} 3.1: {Tests} for correlation and regression analyses.
+\emph{Behavior Research Methods}, \emph{41}(4), 1149--1160.
+\url{https://doi.org/10.3758/BRM.41.4.1149}
 
 \bibitem[\citeproctext]{ref-fiedler_social_2011}
-Fiedler, Marina, Ernan Haruvy, and Sherry Xin Li.
-2011.
-{``Social Distance in a Virtual World Experiment.''} \emph{Games and Economic Behavior} 72 (2): 400--426.
-\url{https://doi.org/10.1016/j.geb.2010.09.004}.
+Fiedler, M., Haruvy, E., \& Li, S. X. (2011). Social distance in a virtual world experiment. \emph{Games and Economic Behavior}, \emph{72}(2), 400--426. \url{https://doi.org/10.1016/j.geb.2010.09.004}
 
 \bibitem[\citeproctext]{ref-fonagy_mentalizing_2017}
-Fonagy, Peter, and Chloe Campbell.
-2017.
-{``\href{https://www.ncbi.nlm.nih.gov/pubmed/29135441}{Mentalizing, Attachment and Epistemic Trust: How Psychotherapy Can Promote Resilience}.''} \emph{Psychiatria Hungarica: A Magyar Pszichiatriai Tarsasag Tudomanyos Folyoirata} 32 (3): 283--87.
+Fonagy, P., \& Campbell, C.
+(2017).
+\href{https://www.ncbi.nlm.nih.gov/pubmed/29135441}{Mentalizing, attachment and epistemic trust: How psychotherapy can promote resilience}.
+\emph{Psychiatria Hungarica: A Magyar Pszichiatriai Tarsasag Tudomanyos Folyoirata}, \emph{32}(3), 283--287.
 
 \bibitem[\citeproctext]{ref-fowler_cooperative_2010}
-Fowler, James H., and Nicholas A.
-Christakis.
-2010.
-{``Cooperative Behavior Cascades in Human Social Networks.''} \emph{Proceedings of the National Academy of Sciences} 107 (12): 5334--38.
-\url{https://doi.org/10.1073/pnas.0913149107}.
+Fowler, J.
+H., \& Christakis, N.
+A.
+(2010).
+Cooperative behavior cascades in human social networks.
+\emph{Proceedings of the National Academy of Sciences}, \emph{107}(12), 5334--5338.
+\url{https://doi.org/10.1073/pnas.0913149107}
 
 \bibitem[\citeproctext]{ref-gao_associations_2017}
-Gao, Shuling, Mark Assink, Andrea Cipriani, and Kangguang Lin.
-2017.
-{``Associations Between Rejection Sensitivity and Mental Health Outcomes: {A} Meta-Analytic Review.''} \emph{Clinical Psychology Review} 57 (November): 59--74.
-\url{https://doi.org/10.1016/j.cpr.2017.08.007}.
+Gao, S., Assink, M., Cipriani, A., \& Lin, K.
+(2017).
+Associations between rejection sensitivity and mental health outcomes: {A} meta-analytic review.
+\emph{Clinical Psychology Review}, \emph{57}, 59--74.
+\url{https://doi.org/10.1016/j.cpr.2017.08.007}
 
 \bibitem[\citeproctext]{ref-hepp_prosociality_2022}
-Hepp, Johanna, and Inga Niedtfeld.
-2022.
-{``Prosociality in Personality Disorders: {Status} Quo and Research Agenda.''} \emph{Current Opinion in Psychology} 44 (April): 208--14.
-\url{https://doi.org/10.1016/j.copsyc.2021.09.013}.
+Hepp, J., \& Niedtfeld, I.
+(2022).
+Prosociality in personality disorders: {Status} quo and research agenda.
+\emph{Current Opinion in Psychology}, \emph{44}, 208--214.
+\url{https://doi.org/10.1016/j.copsyc.2021.09.013}
 
 \bibitem[\citeproctext]{ref-herpertz_social-cognitive_2014}
-Herpertz, Sabine C., and Katja Bertsch.
-2014.
-{``The Social-Cognitive Basis of Personality Disorders.''} \emph{Current Opinion in Psychiatry} 27 (1): 73--77.
-\url{https://doi.org/10.1097/YCO.0000000000000026}.
+Herpertz, S.
+C., \& Bertsch, K.
+(2014).
+The social-cognitive basis of personality disorders.
+\emph{Current Opinion in Psychiatry}, \emph{27}(1), 73--77.
+\url{https://doi.org/10.1097/YCO.0000000000000026}
 
 \bibitem[\citeproctext]{ref-joyce_trust_1995}
-Joyce, Berg, John Dickhaut, and Kevin McCabe. 1995. {``Trust, {Reciprocity}, and {Social History}.''} \emph{Games and Economic Behavior} 10 (1): 122--42.
+Joyce, B., Dickhaut, J., \& McCabe, K. (1995). Trust, {Reciprocity}, and {Social History}. \emph{Games and Economic Behavior}, \emph{10}(1), 122--142.
 
 \bibitem[\citeproctext]{ref-king-casas_getting_2005}
-King-Casas, Brooks, Damon Tomlin, Cedric Anen, Colin F.
-Camerer, Steven R.
-Quartz, and P.
-Read Montague.
-2005.
-{``Getting to {Know You}: {Reputation} and {Trust} in a {Two-Person Economic Exchange}.''} \emph{Science} 308 (5718): 78--83.
-\url{https://doi.org/10.1126/science.1108062}.
+King-Casas, B., Tomlin, D., Anen, C., Camerer, C.
+F., Quartz, S.
+R., \& Montague, P.
+R.
+(2005).
+Getting to {Know You}: {Reputation} and {Trust} in a {Two-Person Economic Exchange}.
+\emph{Science}, \emph{308}(5718), 78--83.
+\url{https://doi.org/10.1126/science.1108062}
 
 \bibitem[\citeproctext]{ref-kobre_negative_1972}
-Kobre, Kenneth R, and Lewis P Lipsitt.
-1972.
-{``A Negative Contrast Effect in Newborns.''} \emph{Journal of Experimental Child Psychology} 14 (1): 81--91.
-\url{https://doi.org/10.1016/0022-0965(72)90033-1}.
+Kobre, K.
+R., \& Lipsitt, L.
+P.
+(1972).
+A negative contrast effect in newborns.
+\emph{Journal of Experimental Child Psychology}, \emph{14}(1), 81--91.
+\url{https://doi.org/10.1016/0022-0965(72)90033-1}
 
 \bibitem[\citeproctext]{ref-lieberman_social_2007}
-Lieberman, Matthew D.
-2007.
-{``Social {Cognitive Neuroscience}: {A Review} of {Core Processes}.''} \emph{Annual Review of Psychology} 58 (1): 259--89.
-\url{https://doi.org/10.1146/annurev.psych.58.110405.085654}.
+Lieberman, M.
+D.
+(2007).
+Social {Cognitive Neuroscience}: {A Review} of {Core Processes}.
+\emph{Annual Review of Psychology}, \emph{58}(1), 259--289.
+\url{https://doi.org/10.1146/annurev.psych.58.110405.085654}
 
 \bibitem[\citeproctext]{ref-matuschek_balancing_2017}
-Matuschek, Hannes, Reinhold Kliegl, Shravan Vasishth, Harald Baayen, and Douglas Bates.
-2017.
-{``Balancing {Type I} Error and Power in Linear Mixed Models.''} \emph{Journal of Memory and Language} 94 (June): 305--15.
-\url{https://doi.org/10.1016/j.jml.2017.01.001}.
+Matuschek, H., Kliegl, R., Vasishth, S., Baayen, H., \& Bates, D.
+(2017).
+Balancing {Type I} error and power in linear mixed models.
+\emph{Journal of Memory and Language}, \emph{94}, 305--315.
+\url{https://doi.org/10.1016/j.jml.2017.01.001}
 
 \bibitem[\citeproctext]{ref-miano_rejection_2013}
-Miano, Annemarie, Eric A.
-Fertuck, Arnoud Arntz, and Barbara Stanley.
-2013.
-{``Rejection {Sensitivity Is} a {Mediator Between Borderline Personality Disorder Features} and {Facial Trust Appraisal}.''} \emph{Journal of Personality Disorders} 27 (4): 442--56.
-\url{https://doi.org/10.1521/pedi_2013_27_096}.
+Miano, A., Fertuck, E.
+A., Arntz, A., \& Stanley, B.
+(2013).
+Rejection {Sensitivity Is} a {Mediator Between Borderline Personality Disorder Features} and {Facial Trust Appraisal}.
+\emph{Journal of Personality Disorders}, \emph{27}(4), 442--456.
+\url{https://doi.org/10.1521/pedi_2013_27_096}
 
 \bibitem[\citeproctext]{ref-mikulincer_attachment_1998}
-Mikulincer, Mario.
-1998.
-{``Attachment Working Models and the Sense of Trust: {An} Exploration of Interaction Goals and Affect Regulation.''} \emph{Journal of Personality and Social Psychology} 74 (5): 1209--24.
-\url{https://doi.org/10.1037/0022-3514.74.5.1209}.
+Mikulincer, M.
+(1998).
+Attachment working models and the sense of trust: {An} exploration of interaction goals and affect regulation.
+\emph{Journal of Personality and Social Psychology}, \emph{74}(5), 1209--1224.
+\url{https://doi.org/10.1037/0022-3514.74.5.1209}
 
@@
 Mulder, R.
-T., P.
-R.
-Joyce, P.
-F.
-Sullivan, C.
-M.
-Bulik, and F.
+T., Joyce, P.
+R., Sullivan, P.
+F., Bulik, C.
+M., \& Carter, F.
 A.
-Carter.
-1999.
-{``The Relationship Among Three Models of Personality Psychopathology: {DSM-III-R} Personality Disorder, {TCI} Scores and {DSQ} Defences.''} \emph{Psychological Medicine} 29 (4): 943--51.
-\url{https://doi.org/10.1017/S0033291799008533}.
+(1999).
+The relationship among three models of personality psychopathology: {DSM-III-R} personality disorder, {TCI} scores and {DSQ} defences.
+\emph{Psychological Medicine}, \emph{29}(4), 943--951.
+\url{https://doi.org/10.1017/S0033291799008533}
 
 \bibitem[\citeproctext]{ref-richetin_emotional_2018}
-Richetin, Juliette, Anita Poggi, Paola Ricciardelli, Eric A.
-Fertuck, and Emanuele Preti.
-2018.
-{``The Emotional Components of Rejection Sensitivity as a Mediator Between {Borderline Personality Disorder} and Biased Appraisal of Trust in Faces.''} \emph{Clinical Neuropsychiatry: Journal of Treatment Evaluation} 15 (4): 200--205.
+Richetin, J., Poggi, A., Ricciardelli, P., Fertuck, E.
+A., \& Preti, E.
+(2018).
+The emotional components of rejection sensitivity as a mediator between {Borderline Personality Disorder} and biased appraisal of trust in faces.
+\emph{Clinical Neuropsychiatry: Journal of Treatment Evaluation}, \emph{15}(4), 200--205.
 
 \bibitem[\citeproctext]{ref-singmann_afex_2022}
-Singmann, Henrik, Ben Bolker, Jake Westfall, Frederik Aust, Mattan S. Ben-Shachar, Søren Højsgaard, John Fox, et al. 2022. {``Afex: {Analysis} of {Factorial Experiments}.''}
+Singmann, H., Bolker, B., Westfall, J., Aust, F., Ben-Shachar, M.
+S., Højsgaard, S., Fox, J., Lawrence, M.
+A., Mertens, U., Love, J., Lenth, R., \& Christensen, R.
+H.
+B.
+(2022).
+\emph{Afex: {Analysis} of {Factorial Experiments}}.
 
 \bibitem[\citeproctext]{ref-weekers_level_2019}
-Weekers, Laura C., Joost Hutsebaut, and Jan H.
-Kamphuis.
-2019.
-{``The {Level} of {Personality Functioning Scale}-{Brief Form} 2.0: {Update} of a Brief Instrument for Assessing Level of Personality Functioning.''} \emph{Personality and Mental Health} 13 (1): 3--14.
-\url{https://doi.org/10.1002/pmh.1434}.
+Weekers, L.
+C., Hutsebaut, J., \& Kamphuis, J.
+H.
+(2019).
+The {Level} of {Personality Functioning Scale}-{Brief Form} 2.0: {Update} of a brief instrument for assessing level of personality functioning.
+\emph{Personality and Mental Health}, \emph{13}(1), 3--14.
+\url{https://doi.org/10.1002/pmh.1434}
 
 \bibitem[\citeproctext]{ref-zentall_within-trial_2005}
-Zentall, Thomas R.
-2005.
-{``A {Within-trial Contrast Effect} and Its {Implications} for {Several Social Psychological Phenomena}.''} \emph{International Journal of Comparative Psychology} 18 (4).
-\url{https://doi.org/10.46867/ijcp.2005.18.04.08}.
+Zentall, T.
+R.
+(2005).
+A {Within-trial Contrast Effect} and its {Implications} for {Several Social Psychological Phenomena}.
+\emph{International Journal of Comparative Psychology}, \emph{18}(4).
+\url{https://doi.org/10.46867/ijcp.2005.18.04.08}
 

### 2024-10-10 article.Rmd
format changes JPCS

@@
 title: "When Forgiveness Backfires: Rejection Sensitivity and Cooperative Behavior Following Exposure to Adaptive Forgiving Agents"
-
 author:
-  "Ismail Guennouni^1^*, Georgia Koppe^2^, Christoph Korn^3^"
-
+  "Ismail Guennouni^1,2,3,4^*, Georgia Koppe^1,2,3^$^\\dagger$, Christoph Korn^4^$^\\dagger$"
 abstract: |
-  Can exposure to forgiving behavior improve cooperative outcomes, and does Rejection Sensitivity (RS) moderate this effect?
-This randomized controlled experiment explores these questions using a novel approach with Hidden Markov Model (HMM)-based artificial agents in repeated trust games (RTGs).
-Participants (N = 206), pre-screened for high or low RS, interacted with either consistently behaving or more forgiving co-players, which were in reality HMM agents.
-Unexpectedly, exposure to forgiving agents led to reduced cooperation in subsequent interactions.
-Control group participants maintained consistent behavior.
-RS levels influenced perceptions of the agent's cooperativeness but did not significantly affect behavioral outcomes, revealing a perception-behavior dissociation.
-These findings challenge assumptions about fostering cooperation through simple exposure to positive interactions and highlight the importance of accounting for negative contrast effects when designing such interventions.
-They also demonstrate the potential of HMM-based artificial agents for studying interactive social dynamics, offering a methodological advancement for future research.
-Our results suggest the need for nuanced interventions that gradually shape expectations and behaviors to foster cooperative outcomes, particularly for individuals with high rejection sensitivity.
-
 bibliography: "bib/exposure.bib"
-biblio-style: spphys
+csl: apa.csl
 output:
@@
   - \usepackage{lscape}
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
+
 \small
-^1^ *Central Institute of Mental Health, Medical Faculty Mannheim/Heidelberg University, Germany*
-
-^2^ *Faculty of Mathematics and Computer Science, Heidelberg University, Germany*
-
-^3^ *Department of General Psychiatry, Section Social Neuroscience, Heidelberg University, Germany*
-
-^\*^ *Corresponding author. Address: Central institute of Mental Health, J5, Mannheim, Germany. Email: ismail.guennouni@zi-mannheim.de* <br>
+
+^1^ *Department of Psychiatry and Psychotherapy, Central Institute of
+Mental Health, Medical Faculty Mannheim, Heidelberg University,
+Mannheim, Germany*
+
+^2^ *Interdisciplinary Center for Scientific Computing, Faculty of
+Mathematics and Computer Science, Heidelberg University, Heidelberg,
+Germany*
+
+^3^ *Hector Institute for AI in Psychiatry, Central Institute of Mental
+Health, Medical Faculty Mannheim, Heidelberg University, Mannheim
+Germany*
+
+^4^ *Department of General Psychiatry, Section Social Neuroscience,
+Heidelberg University, Germany*
+
+$^\dagger$ *Joint last author*
+
+^\*^ *Corresponding author. Address: Central institute of Mental Health,
+J5, Mannheim, Germany. Email:
+[ismail.guennouni\@zi-mannheim.de](mailto:ismail.guennouni@zi-mannheim.de){.email}*
+
+\pagebreak
+
+**Abstract:**
+
+Can exposure to forgiving behavior improve cooperative outcomes, and
+does Rejection Sensitivity (RS) moderate this effect? This randomized
+controlled experiment explores these questions using a novel approach
+with Hidden Markov Model (HMM)-based artificial agents in repeated trust
+games (RTGs). Participants (N = 206), pre-screened for high or low RS,
+interacted with either consistently behaving or more forgiving
+co-players, which were in reality HMM agents. Unexpectedly, exposure to
+forgiving agents led to reduced cooperation in subsequent interactions.
+Control group participants maintained consistent behavior. RS levels
+influenced perceptions of the agent's cooperativeness but did not
+significantly affect behavioral outcomes, revealing a
+perception-behavior dissociation. These findings challenge assumptions
+about fostering cooperation through simple exposure to positive
+interactions and highlight the importance of accounting for negative
+contrast effects when designing such interventions. They also
+demonstrate the potential of HMM-based artificial agents for studying
+interactive social dynamics, offering a methodological advancement for
+future research. Our results suggest the need for nuanced interventions
+that gradually shape expectations and behaviors to foster cooperative
+outcomes, particularly addressing the discrepancy between perceptions
+and behavior in individuals with high rejection sensitivity.
 
@@
 
-**Keywords** : Social Trust; Forgiveness Intervention; Rejection Sensitivity; Interpersonal Dynamics; Hidden Markov Models
+**Keywords** : Interpersonal functioning; Rejection Sensitivity;
+Forgiveness Intervention; Trust-based Cooperation; Hidden Markov Models
 
@@
 
-**Highlights:**
-
-- Exposure to forgiving artificial agents unexpectedly led to reduced cooperation in subsequent interactions, challenging assumptions about fostering trust through positive experiences
-
-- Rejection sensitivity influenced perceptions of artificial agents' cooperativeness but not behavioral outcomes, revealing a dissociation between perception and behavior
-
-- Novel Hidden Markov Model-based artificial agents provide adaptive, human-like behavior in trust games, allowing for precise control of "cooperative" and "forgiving" strategies while maintaining experimental control of co-player strategies
+**General Scientific Summary:**
+
+This study reveals that exposing individuals to forgiving behavior in
+economic games can unexpectedly decrease their subsequent cooperation.
+Interestingly, while people with high rejection sensitivity more
+accurately perceived their partners' cooperativeness, this didn't
+translate into more cooperative behavior. These findings challenge
+assumptions about fostering cooperation and highlight the need for more
+nuanced interventions to promote reciprocity, especially for individuals
+prone to expecting rejection.
 
@@
 
-Trust is fundamental to human social interactions, facilitating seamless relations at both interpersonal and intergroup levels.
-The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders [@fonagy_mentalizing_2017].
-Individuals with personality disorders (PD) often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors – a marker for the severity of PD symptoms [@herpertz_social-cognitive_2014; @mulder_relationship_1999].
-
-One explanation for such social challenges lies in early caregiver experiences.
-Attachment theory [@bowlby_attachment_1978] suggests that the quality of these relationships shapes our capacity for secure attachments and trust.
-Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998].
-These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection [@downey_implications_1996; @downey_early_1997].
-RS has been linked to the development of various mental health conditions, including depression, anxiety, personality disorders, and self-harm [@gao_associations_2017].
-A recent meta-analysis revealed prosocial behavior and interpersonal trust as two key processes of interpersonal functioning that are markedly impaired in PDs and which are likely to contribute to interpersonal dysfunction in this population [@hepp_prosociality_2022].
-The interaction of rejection sensitivity and trust-based constructs has been explored, particularly in Borderline Personality Disorder (BPD).
-@miano_rejection_2013 and @richetin_emotional_2018 found that rejection sensitivity mediated the relationship between BPD features and lower trust appraisal.
-@abramov_influence_2022-1 found that higher baseline feelings of rejection in individuals with BPD predict slower trust formation and less pronounced declines in trust following trust violations during the trust game.
-However, the interaction between trustworthiness and rejection sensitivity hasn't been studied as extensively in other personality disorders or in non-clinical populations, leaving a gap in our understanding of how these constructs might interplay.
-Could an ingrained sensitivity to rejection contribute to reduced pro-social behavior such as a failure to show trustworthiness?
-
-If indeed RS contributes to interpersonal dysfunction through failure to show trustworthiness, could exposing those who exhibit it to cooperative and forgiving interaction partners lead to improved interpersonal outcomes?
-Research in the fields of behavioral economics and psychology has explored how positive social interactions influence cooperation and pro-social behavior.
-The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995].
-In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005].
-@fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior.
-Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior.
-These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust and potentially fostering trustworthy behavior.
-
-In this study, we use a randomized controlled online experiment to test whether exposing participants with varying RS levels to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation.
-To simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players' data.
-A key aspect of these agents is that their actions depend on a latent "trust state" which reacts dynamically to the trustees' returns, simulating real-life trust-building scenarios.
-An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively, or the propensity to retaliate after breakdowns of cooperation.
-
-We pre-screened participants for high or low rejection sensitivity (RS) using a validated questionnaire, then assigned them exclusively to the trustee role in a series of trust games.
-After playing a 15-round RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition.
-In the Manipulation condition, participants were exposed over three RTGs to HMM investors designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.
-In the Control condition, participants played three RTGs against the same human-like HMM.
-After this exposure phase, all participants played another 15-round RTG with a human-like HMM investor.
-We expect those exhibiting high RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
-We also hypothesize that those in the Manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase, with those in the higher RS group showing a bigger effect.
+Trust is fundamental to human social interactions, facilitating seamless
+relations at both interpersonal and intergroup levels. The study of
+psychopathology has linked deficits in trust-based constructs to the
+development of mental health disorders [@fonagy_mentalizing_2017].
+Individuals with personality disorders (PD) often struggle to form and
+maintain social connections, a difficulty reflected in uncooperative
+behaviors – a marker for the severity of PD symptoms
+[@herpertz_social-cognitive_2014; @mulder_relationship_1999].
+
+One explanation for such social challenges lies in early caregiver
+experiences. Attachment theory [@bowlby_attachment_1978] suggests that
+the quality of these relationships shapes our capacity for secure
+attachments and trust. Individuals with higher levels of insecure
+attachment may recall negative trust-related experiences more easily,
+report fewer positive trust experiences, and use less constructive
+coping strategies when trust is broken [@mikulincer_attachment_1998].
+These insecure attachment patterns are often associated with heightened
+rejection sensitivity (RS), a tendency to anxiously expect, readily
+perceive, and intensely react to rejection [@downey_implications_1996;
+@downey_early_1997]. RS has been linked to the development of various
+mental health conditions, including depression, anxiety, personality
+disorders, and self-harm [@gao_associations_2017]. Individuals high in
+RS show attentional biases towards social threat cues, which may
+contribute to difficulties in social interactions
+[@berenson_rejection_2009]. A recent meta-analysis revealed prosocial
+behavior and interpersonal trust as two key processes of interpersonal
+functioning that are markedly impaired in PDs and which are likely to
+contribute to interpersonal dysfunction in this population
+[@hepp_prosociality_2022]. The interaction of RS and trust-based
+constructs has been explored, particularly in Borderline Personality
+Disorder (BPD). @miano_rejection_2013 and @richetin_emotional_2018 found
+that RS mediated the relationship between BPD features and lower trust
+appraisal. @abramov_influence_2022-1 found that higher baseline feelings
+of rejection in individuals with BPD predict slower trust formation and
+less pronounced declines in trust following trust violations during the
+trust game. However, the interaction between *reciprocity* and RS hasn't
+been studied as extensively, leaving a gap in our understanding of how
+these constructs might interplay. Could an ingrained sensitivity to
+rejection contribute to reduced pro-social behavior such as a failure to
+show reciprocity as a result of perceived rejection by the interaction
+partner?
+
+Given that RS may be a manifestation of maladaptive attachment styles,
+it's crucial to explore whether exposure to consistently forgiving
+interaction partners could help reshape these interpersonal expectations
+and behaviors. If RS indeed influences reciprocity in repeated
+interactions, could such exposure serve as a corrective experience,
+potentially improving interpersonal outcomes? This approach aligns with
+attachment theory, which suggests that new, positive relational
+experiences can modify internal working models of relationships
+[@bowlby_secure_1988]. Research in the fields of behavioral economics
+and psychology has also explored how positive social interactions
+influence cooperation and pro-social behavior. The use of the repeated
+trust game (RTG), a well-established experimental approach, has allowed
+for the analysis of the development of trust through ongoing
+interactions [@joyce_trust_1995]. In this paradigm, cycles of mutual
+trust, where each party's trust is reciprocated with trustworthiness,
+have the effect of enhancing cooperative behaviors even among
+individuals who are initially inclined to be distrustful
+[@king-casas_getting_2005]. @fowler_cooperative_2010 studied behavior in
+social networks interacting in a public goods game and found that
+cooperative behavior tends to cluster, suggesting that exposure to
+cooperative peers can lead to more cooperative behavior. Similarly,
+research on social learning theory [@bandura_social_1977] has long
+demonstrated that individuals learn and model the behavior of those
+around them, indicating that if someone is consistently exposed to
+cooperative and positive individuals, they're likely to emulate this
+behavior. These insights highlight that engaging with compassionate and
+forgiving others can be an effective method for mitigating deeply
+ingrained mistrust and potentially fostering trustworthy behavior.
+
+In this study, we use a randomized controlled online experiment to test
+whether exposing participants with varying RS levels to forgiving and
+more cooperative co-players results in more trustworthy behavior and a
+repair of potential breakdowns in RTG cooperation. To simulate realistic
+social interaction while maintaining a high degree of experimental
+control, we take a novel paradigmatic approach: We use generative models
+of how humans play the RTG to design an agent that plays the role of the
+investor, based on Hidden Markov Models (HMMs) fitted to real players'
+data. A key aspect of these agents is that their actions depend on a
+latent "trust state" which reacts dynamically to the trustees' returns,
+simulating real-life trust-building scenarios. An advantage of having
+such a generative model of behavior is the possibility of controlling
+different aspects of the agent's strategy such as its general policy,
+the propensity to cooperate actively, or the propensity to trust again
+after breakdowns of cooperation. To further mimic real-world
+interactions and examine participants' responses to one-off breakdowns
+of cooperation, we incorporate occasional pre-programmed low investments
+by the agent.
+
+We pre-screened participants for high or low RS using a validated
+questionnaire, then assigned them exclusively to the trustee role in a
+series of trust games. After playing a 15-round RTG with a human-like
+HMM investor, they were randomly assigned to either a Control or
+Manipulation condition. In the Manipulation condition, participants were
+exposed over three RTGs to HMM investors designed with a limited
+propensity for retaliation, potentially mitigating ingrained mistrust.
+In the Control condition, participants played three RTGs against the
+same human-like HMM. After this exposure phase, all participants played
+another 15-round RTG with a human-like HMM investor, similar to the one
+in the pre-exposure phase. We expect those exhibiting high RS to be more
+responsive to the pre-programmed change in the agent's cooperativeness
+and retaliation propensity. We also hypothesize that those in the
+Manipulation condition would behave more cooperatively and have a lower
+propensity to retaliate to the co-player's defection after the exposure
+phase, with those in the higher RS group showing a bigger effect.
 
@@
 
-To have participants with large differences in RS, a total of 1195 participants were pre-screened on the Prolific Academic platform (prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10) totalling 206 participants (56% female).
-These were then invited through prolific to take part in the main experiment study.
-The required sample size was determined using an a priori power analysis to have an 80% probability to detect a small effect size with a 5% type I error rate in a repeated measures ANOVA with between factors.
-The mean age of participants was 34.6 years, with an 11.9 years standard deviation.
-Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5.
+To have participants with large differences in RS, a total of 1195
+participants were pre-screened on the Prolific Academic platform
+(prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to
+finally select two similarly sized groups: One with high RS (RSQ score
+\> 15, N=103) and the other with low RS (RSQ score \< 10, N=103)
+totalling 206 participants (56% female). These were then invited through
+prolific to take part in the main experiment. The required sample size
+was determined using an *a priori* power analysis to have an 80%
+probability to detect a small effect size (Cohen’s f = 0.10) for a
+within-between interaction with a 5% type I error rate in a repeated
+measures ANOVA. The sample size calculation assumed 2 groups, 2
+measurement per group and was performed using the G\*Power software
+[@faul_statistical_2009]. The mean age of participants was 34.6 years,
+with an 11.9 years standard deviation. The majority of participants
+identified ethnically as White ($80$%). The online cohort registered 30
+unique countries of birth with the most frequent being the U.K ($33$%)
+followed by Poland ($10$%) and Portugal ($10$%). Participants were paid
+a fixed fee of £6 plus a bonus payment dependent on their performance
+that averaged £0.5. Data was collected over multiple sessions between
+December 2023 and February 2024.
 
@@
 
-The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : High or Low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A).
-Participants within each pre-screened group were randomly assigned to one of the two levels of the Condition factor.
-The games were designed and implemented online using Empirica v1 [@almaatouq_empirica_2021].
-The planned experiment received approval from the University of Heidelberg's Medical Faculty ethics commission (ID:S-708/2023) and the experiment was performed in accordance with the ethics board guidelines and regulations.
-All participants provided informed consent prior to their participation.
+The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS :
+High or Low) by 2 (Phase: Trust-Game Pre Manipulation, Trust-Game Post
+Manipulation) design, with repeated measures on the Phase factor (Figure
+\@ref(fig:HMMPanels).A). Participants within each pre-screened RS group
+were randomly assigned to one of the two levels of the Condition factor,
+resulting in 101 participants in the Manipulation condition and 105 in
+the Control condition. The games were designed and implemented online
+using Empirica v1 [@almaatouq_empirica_2021]. The planned experiment
+received approval from the University of Heidelberg's Medical Faculty
+ethics commission (ID:S-708/2023) and the experiment was performed in
+accordance with the ethics board guidelines and regulations. All
+participants provided informed consent prior to their participation.
 
@@
 
-### Repeated Trust Game
-
-Participants played a 15-round RTG [@joyce_trust_1995] in the trustee role against a computer-programmed investor.
-On each round the investor is endowed with 20 units and decides how much of that endowment to invest.
-This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor.
-If the trustee returns more than one third of the amount, the investor makes a gain.
-Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right.
-The participants were able to choose the icon that represents them at the start of the experiment.
-The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner.
-Participants were not told they were facing computerised co-players.
-We chose to simulate the behavior of a human interaction partner through allowing for a delay whilst pairing with new opponents as the start of each game as well as programming the agents to respond during each round after a varying time lapse (randomly chosen between 5 and 10 seconds).
-
-The computerised investor consisted of a hidden Markov model (HMM) trained on an independent existing behavioral RTG data set of human investors.
-This data-driven approach thus sought to learn an investor strategy that mimics human-like interactions.
-The data set used for training consists of 388 ten round games with the same player (full details can be found in the Supplementary Information).
-On this data set, the HMM was inferred with three latent states that could be interpreted as reflecting a “low-trust”, a “medium-trust”, and a “high-trust” state.
-A separate output distribution, that maps each HMM state onto possible investments from 0 to 20 separately, is learned (Figure \@ref(fig:HMMPanels).B).
-In analogy to the latent states, these distributions can be interpreted as reflecting “low-trust”, “medium-trust”, or “high-trust” dispositions.
-Finally, the HMM is specified by transition probabilities that describe the transition between states.
-The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C)).
-The initial state for the HMM investor in each instance of the game was set to the “mid-trust” state.
-Details on how the HMM state conditional probabilities and transition functions are specified can be found in the supplement.
-
-In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair it, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) and round 13 (post-manipulation).
-On all other rounds, the investor’s actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds).
+### Repeated Trust Game and HMM Investor
+
+Participants played a 15-round RTG [@joyce_trust_1995] in the trustee
+role against a computer-programmed investor. On each round the investor
+is endowed with 20 units and decides how much of that endowment to
+invest. This investment is tripled and the trustee then decides how to
+split this tripled amount between them and the investor. If the trustee
+returns more than one third of the amount, the investor makes a gain.
+Each player was represented with an icon with the participant always on
+the left of the screen and the co-player on the right. The participants
+were able to choose the icon that represents them at the start of the
+experiment. The icon representing the co-player changed at the start of
+each new game, to simulate a new interaction partner. Participants were
+not told they were facing computerised co-players. We chose to simulate
+the behavior of a human interaction partner through allowing for a delay
+whilst pairing with new opponents as the start of each game as well as
+programming the agents to respond during each round after a varying time
+lapse (randomly chosen between 5 and 10 seconds).
+
+The computerised investor consisted of a hidden Markov model (HMM)
+trained on an independent existing behavioral RTG data set of human
+investors. This data-driven approach thus sought to learn an investor
+strategy that mimics human-like interactions. The data set used for
+training consists of 388 ten round games with the same player (full
+details can be found in the Supplementary Information). On this data
+set, the HMM was inferred with three latent states that could be
+interpreted as reflecting a “low-trust”, a “medium-trust”, and a
+“high-trust” state. A separate output distribution, that maps each HMM
+state onto possible investments from 0 to 20 separately, is learned
+(Figure \@ref(fig:HMMPanels).B). In analogy to the latent states, these
+distributions can be interpreted as reflecting “low-trust”,
+“medium-trust”, or “high-trust” dispositions. Finally, the HMM is
+specified by transition probabilities that describe the transition
+between states. The probability of these transitions was modelled as a
+function of their net return (i.e return - investment) in the previous
+round (see Figure \@ref(fig:HMMPanels).C)). The initial state for the
+HMM investor in each instance of the game was set to the “mid-trust”
+state. Details on how the HMM state conditional probabilities and
+transition functions are specified can be found in the supplement.
+
+In order to instigate a potential breakdown of trust, thereby allowing
+us to probe efforts to repair it, the computerised agent was programmed
+to provide a low investment on round 12 (pre-manipulation) and round 13
+(post-manipulation). On all other rounds, the investor’s actions were
+determined by randomly drawing an investment from the state-conditional
+distribution, with the state over rounds determined by randomly drawing
+the next state from the state-transition distribution as determined from
+the net return on the previous round (disregarding the net return
+immediately after the pre-programmed low investment rounds).
 
@@
 
-In all phases of the RTG other than the ‘Exposure phase’ (Figure \ref{fig:HMMPanels}.A), participants interacted with this human-like HMM.
-In the ‘Manipulation’ Condition of the exposure phase, however, the parameters of this HMM were adjusted to design a ‘forgiving’ and ultimately more cooperative agent.
-To achieve this, we changed the state transition probabilities of the HMM such that it becomes impossible for it to remain in a low trust state, effectively setting the transition probability for remaining in a “low-trust” state to 0.
-The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
-The policies conditional on the latent states and the transition function in the other latent states remain unchanged.
-
+In all phases of the RTG other than the ‘Exposure phase’ (Figure
+\@ref(fig:HMMPanels).A), participants interacted with this human-like
+HMM. In the ‘Manipulation’ Condition of the exposure phase, however, the
+parameters of this HMM were adjusted to design a ‘forgiving’ and
+ultimately more cooperative agent. To achieve this, we changed the state
+transition probabilities of the HMM such that it becomes impossible for
+it to remain in a low trust state, effectively setting the transition
+probability for remaining in a “low-trust” state to 0. The resulting
+transition function is shown in Figure \@ref(fig:HMMPanels).D. The
+policies conditional on the latent states and the transition function in
+the other latent states remain unchanged.
+
+## Procedure
+
+At the start of the experiment, participants provided informed consent
+and were instructed the study would consist of three phases in which
+they would face a different other player. Participants were told their
+goal was to maximise the number of points in all phases. They were not
+told the number of rounds of each phase. Participants were randomly
+assigned to either a Control or Manipulation condition. The timeline of
+the experiment is shown in Figure \@ref(fig:HMMPanels).A. Phase one
+("pre") consisted of a 15 round RTG in which participants took the role
+of trustee, facing the same investor over all 15 rounds. On each round,
+after being informed about the amount sent by the investor participants
+decided how much of the tripled investment to return to the investor,
+before continuing to the next round. Phase 2 ("exposure") consisted of
+three 7-round RTGs. Participants in the Manipulation condition faced the
+forgiving HMM investor and rated the agent on the same attributes as in
+the pre-manipulation phase. Those in the Control condition faced the
+same human-like HMM agent as in the "pre" phase and rated each co-player
+on the same attributes. To keep the experience similar to the "pre"
+phase, the agent in the Control condition was also designed to send a
+very low investment in round 5 of each of the three games. In the
+post-manipulation phase ("post"), participants in both conditions faced
+the same human-like HMM as in "pre" phase.
+
+At the beginning of each game in all three phases, participants were
+told they would face a new player and had to wait to be paired with an
+available co-player. This simulated the waiting time in real social
+interaction tasks. After completing each RTG in each phase, participants
+rated how cooperative and forgiving they perceived the co-player to be,
+and whether they would like to play with them again (all on a scale from
+1 to 10 with 10 being the most positive rating). After completing the
+three game phases, participants then completed the Levels of Personality
+Functioning Scale Brief-Form (LPFS-BF) questionnaire
+[@weekers_level_2019]. This is a self-report measure designed to assess
+core elements of personality functioning as defined in the Alternative
+Model for Personality Disorders in the DSM-5
+[@american_psychiatric_association_diagnostic_2013], and provides a
+dimensional assessment of personality functioning, which complements the
+categorical approach of RS. Finally, participants were asked whether
+they thought the other players were human or computer agents, to probe
+how well the agent can mimic human behavior, then debriefed and thanked
+for their participation.
+
+````{=tex}
 \begin{landscape}
@@
 \end{landscape}
-
-## Procedure
-
-At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player.
-Participants were told their goal was to maximise the number of points in all phases.
-They were not told the number of rounds of each phase.
-Participants were randomly assigned to either a Control or Manipulation condition.
-The timeline of the experiment is shown in Figure \@ref(fig:HMMPanels).A.
-Phase one ("pre") consisted of a 15 round RTG in which participants took the role of trustee, facing the same investor over all 15 rounds.
-On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round.
-Phase 2 ("exposure") consisted of three 7-round RTGs.
-Participants in the Manipulation condition faced the forgiving HMM investor and rated the agent on the same attributes as in the pre-manipulation phase.
-Those in the Control condition faced the same human-like HMM agent as in the "pre" phase and rated each co-player on the same attributes.
-To keep the experience similar to the "pre" phase, the agent in the Control condition was also designed to send a very low investment in round 5 of each of the three games.
-In the post-manipulation phase ("post"), participants in both conditions faced the same human-like HMM as in "pre" phase.
-
-At the beginning of each game in all three phases, participants were told they would face a new player and had to wait to be paired with an available co-player.
-This simulated the waiting time in real social interaction tasks.
-After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-playerr to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
-After completing the three game phases, participants then completed the Levels of Personality Functioning Scale Brief-Form (LPFS-BF) questionnaire [@weekers_level_2019].
-This is a self-report measure designed to assess core elements of personality functioning as defined in the Alternative Model for Personality Disorders in the DSM-5 [@american_psychiatric_association_diagnostic_2013], and provides a dimensional assessment of personality functioning, which complements the categorical approach of RS.
-Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.
+````
 
@@
 
-To explore whether participants behaved differently in the RTG after the manipulation compared to the Control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.
-post-manipulation), Condition (manipulation vs.
-Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase.
-The full specification of the statistical model can be found in the supplement.
-To test whether participants behaved differently between conditions in the Exposure phase, we fit a linear mixed effects model to participants returns only in the exposure phase, with Condition (manipulation vs.
-Control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts.
-Finally, to test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model.
-
-The model was estimated using the `afex` package [@singmann_afex_2022] in R.
-More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek_balancing_2017].
-A similar process was used to establish the random effects structures of linear mixed-effects models used to analyse the HMM agent investments as well as the participants' ratings of the co-players.
-There is no agreed upon way to calculate effect sizes for mixed effects models.
-Instead, we will report on testing differences in marginal means.
-For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex".
-We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
+To test whether participants behaved differently in the RTG after the
+manipulation compared to the Control group, we model the percentage
+return (percentage of tripled investment returned to investor) using a
+linear mixed-effects model with Phase (RTG game pre vs.
+post-manipulation), Condition (manipulation vs. Control), Investment,
+and RS (High vs Low RS group) as well as their interactions as fixed
+effects, and player-wise random intercepts and slopes for Phase. The
+full specification of the statistical model can be found in the
+supplement. To test whether participants behaved differently between
+conditions in the Exposure phase, we fit a linear mixed effects model to
+participants returns only in the exposure phase, with Condition
+(manipulation vs. Control), Investment, and RS (High vs Low RS group) as
+well as their interactions as fixed effects, and player-wise random
+intercepts. Finally, to test whether the HMM agent's behavior differed
+between Phases, Conditions and RS groups, we estimate a linear
+mixed-effects model of investments sent by the computerised HMM agent
+with Condition, Phase and RS and their interaction as fixed effects, and
+a similar random effects structure to the returns model.
+
+The model was estimated using the `afex` package [@singmann_afex_2022]
+in R. More complex models with additional random effects could not be
+estimated reliably, and as such the estimated model can be considered to
+include the optimal random effects structure
+[@matuschek_balancing_2017]. A similar process was used to establish the
+random effects structures of linear mixed-effects models used to analyse
+the HMM agent investments as well as the participants' ratings of the
+co-players. There is no agreed upon way to calculate effect sizes for
+mixed effects models. Instead, we will report on testing differences in
+marginal means. For the $F$-tests, we used the Kenward-Roger
+approximation to the degrees of freedom, as implemented in the R package
+"afex". We Z-transform the Investment variable (subtract the overall
+investment mean and divide by overall standard deviation) as centering
+is beneficial to interpreting the main effects more easily in the
+presence of interactions. To probe significant interactions, we
+conducted planned contrasts using the `emmeans` package in R. Given that
+we were testing multiple pre-planned comparisons, we applied the "Sidak"
+correction to control for familywise error rate while maintaining
+reasonable statistical power. This approach allowed us to investigate
+specific hypotheses about the differential effects of our manipulation
+across phases and RS groups, while protecting against inflated Type I
+error rates.
 
@@
 ```{r}
+# Number of participants in manipulation condition
+temp_manip <- avg_ret_df %>% filter(condition.f =="Manipulation")
+
+length(unique(temp_manip$playerId))
+
+# Number of participants with high RS
+
+temp_RS <- avg_ret_df %>% filter(high_RS =="high RS")
+length(unique(temp_RS$playerId))
+
+```
+
+```{r}
 # average investment for first 10 rounds pre manipulation
@@
 <!-- EFFECT SIZES -->
+
 ```{r}
@@
 
-On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60% of the endowment for investments and 35-50% of the total yield for returns, as reported in previous studies [@charness_investment_2008; @fiedler_social_2011].
+On average, investments and returns, as shown in Figure
+\@ref(fig:gamesPlot), fell within the documented range of 40-60% of the
+endowment for investments and 35-50% of the total yield for returns, as
+reported in previous studies [@charness_investment_2008;
+@fiedler_social_2011].
 
@@
 
-Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs.
-Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the Pre RTG compared to the Post.
-Importantly, we also find an interaction between Condition and Phase (RTG pre- vs.
-post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
-As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the Manipulation condition, pre - post,
-`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the Control condition.
-We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.
+Mixed-effects analysis on the percentage returns shows a significant
+main effect of Phase (Pre vs. Post RTG game),
+`r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with
+higher percentage returns in the Pre RTG compared to the Post.
+Importantly, we also find an interaction between Condition and Phase
+(RTG pre- vs. post-manipulation),
+`r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
+As shown in Figure \@ref(fig:boxPlots), post-hoc tests confirm that,
+contrary to our expectations, there was a decrease in the percentage
+returned only in the Manipulation condition, pre - post,
+`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`,
+but no change in the Control condition. We find no interaction between
+Phase, Condition and RS, suggesting that there was no difference between
+RS groups for this interaction.
 
@@
 `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`,
-such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, reflected that returns were more affected by investments in the Control condition.
-We also find a three way interaction between Phase, Investment and RS, `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_inv_scaled_high_RS)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation).
-Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
-For completeness, these effects are discussed in more detail in the supplement.
+such that higher investments were associated with higher percentage
+returns indicating positive reciprocity. An Investment by Condition
+interaction,
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`,
+reflected that returns were more affected by investments in the Control
+condition. We also find a three way interaction between Phase,
+Investment and RS,
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_inv_scaled_high_RS)`,
+showing that the differentiated effect of the investment on the
+proportion returned by RS group is itself moderated by the Phase (pre-
+vs post manipulation). Finally, we find a four-way interaction between
+Condition, Phase, Investment and RS
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
 
@@
 
+```
+
+```{r}
+library(emmeans)
+
+# First, get the estimated marginal means
+emm <- emmeans(mod_returns_pct_RS, ~ inv_scaled | phase.f * condition.f * high_RS)
+
+# Then, get the slopes (effects of investment) for each combination
+slopes <- emtrends(mod_returns_pct_RS, ~ phase.f * condition.f * high_RS, var = "inv_scaled")
+slopes
+
+# Define the contrasts
+corrected_contrasts <- list(
+  "Diff in inv effect (Post - Pre, Manip, Low RS)" = c(0, 0, -1, 1, 0, 0, 0, 0),
+  "Diff in inv effect (Post - Pre, Manip, High RS)" = c(0, 0, 0, 0, 0, 0, -1, 1),
+  "Diff in (Post - Pre) between High and Low RS" = c(0, 0, 1, -1, 0, 0, -1, 1)
+)
+
+# Test the contrasts with multiple comparison correction
+results_4w <- contrast(slopes, method = corrected_contrasts, adjust = "sidak")
+
+# View the results
+summary(results_4w)
+
+```
+
+To examine this four-way interaction, we conducted a contrast analysis
+of how the effect of investment on returns changed from pre- to
+post-manipulation for different RS groups in both conditions. Starting
+with the Manipulation condition, for participants with low RS, the
+effect of investment on returns increased significantly from pre- to
+post-manipulation,
+`r papaja::apa_print(results_4w)$full_result$DiffininveffectPost_PreManipLowRS`.
+This suggests that after the manipulation, low RS participants became
+more responsive to their co-player's investments, returning
+proportionally more as investments increased. In contrast, for
+participants with high RS, the effect of investment on returns decreased
+significantly from pre- to post-manipulation,
+`r papaja::apa_print(results_4w)$full_result$DiffininveffectPost_PreManipHighRS`.
+This indicates that high RS participants became less responsive to their
+co-player's investments after the manipulation, with smaller increases
+in returns as investments increased. The difference in these pre-post
+changes between high and low RS groups was significant,
+`r papaja::apa_print(results_4w)$full_result$DiffinPost_PrebetweenHighandLowRS`.
+This result suggests that the manipulation had significantly different
+effects on how low and high RS participants responded to their
+co-player's investments.
+
+In the Control condition, we observed no significant changes in how
+participants responded to their co-player's investments between the pre
+and post phases, regardless of their RS level.
+
+```{r}
+# Define the contrasts for both conditions
+contrasts_ctrl <- list(
+  # Control condition contrasts
+  "Diff in inv effect (Post - Pre, Control, Low RS)" = c(-1, 1, 0, 0, 0, 0, 0, 0),
+  "Diff in inv effect (Post - Pre, Control, High RS)" = c(0, 0, 0, 0, -1, 1, 0, 0),
+  "Diff in (Post - Pre) between High and Low RS in Control" = c(1, -1, 0, 0, -1, 1, 0, 0)
+)
+
+# Test the contrasts with multiple comparison correction
+results_ctrl <- contrast(slopes, method = contrasts_ctrl, adjust = "sidak")
+
+# View the results
+print(summary(results_ctrl))
 ```
@@
 The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns.
-The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.",fig.align="center", fig.width=6, fig.height = 8}
+The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.
+RS did not affect returns and investments by Phase and Condition.",fig.align="center", fig.width=6, fig.height = 8}
 
@@
 <!-- Focusing on post defection only:  -->
+
 ### Post Defection Trials
@@
 
-Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor?
-To explore this question, we restrict the analysis to the trials following the pre-programmed defection by the HMM agent in both the "pre" (trials 12 to 15) and the "post" phases (trials 13 to 15).
-We fit the same mixed effects model as for all the trials with the exception of the RS variable.
-This is because RS did not show main or interaction effects in the main model, and also due to the necessity of running a simpler model to accommodate the low number of trials.
-We find a significant main effect of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` with returns lower in the second game post defection trials compared to the first.
-We also find a main effect of Investment
+Did participants learn to be more forgiving and cooperative after
+witnessing the pre-programmed defection by the HMM investor? To explore
+this question, we restrict the analysis to the trials following the
+pre-programmed defection by the HMM agent in both the "pre" (trials 12
+to 15) and the "post" phases (trials 13 to 15). We fit the same mixed
+effects model as for all the trials with the exception of the RS
+variable. This is because RS did not show main effects in the previous
+model, and also due to the necessity of running a simpler model to
+accommodate the low number of trials. We find a significant main effect
+of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f`
+with returns lower in the second game post defection trials compared to
+the first. We also find a main effect of Investment
 `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_retPostDef_RS)$full_result$inv_scaled)`
-where participants continued to return higher proportions when receiving higher investments. 
-Finally, we still find an Investment by Condition interaction 
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the Manipulation condition compared to the Control condition in post-defection trials.
-However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.
+where participants continued to return higher proportions when receiving
+higher investments. Finally, we still find an Investment by Condition
+interaction
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`
+showing a lower effects of investment on the Manipulation condition
+compared to the Control condition in post-defection trials. However, the
+absence of a Condition by Phase effect indicates there was no difference
+between conditions on participants' reaction to a one-off low investment
+by the co-player.
 
@@
 ### HMM Investor in Pre and Post Phases
+
 <!-- Linear mixed effects model for investments -->
-Was the HMM's strategy similar between pre and post phases in the control condition?
-Was participants' behavior post exposure differentiated enough to induce a different reaction from the HMM?
-To answer these questions, we test for differences in the HMM agent's investment by Phase, Condition and RS using a linear mixed-effects model as described in the methods section.
-As seen in Figure \@ref(fig:boxPlots), we find no main or interaction effects, indicating the HMM's behavior was on aggregate similar across levels of Phase, Condition and RS.
-This consistency in the investor's behavior is a desirable feature of the HMM agent when the participants behavior is largely simialr between phases.
-More importantly, it indicates that the lower returns of participants in the post phase of the manipulation condition were not differentiated enough to make the HMM react by transitioning to lower latent trust states.
-It is also noteworthy that the HMM agent was relatively successful in imitating human behavior in this paradigm: When asked during debrief whether they thought the investors they faced were human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player.
-When asked to justify their choice, many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.
+
+Was the HMM's strategy similar between pre and post phases in the
+control condition? Was participants' behavior post exposure
+differentiated enough to induce a different reaction from the HMM? To
+answer these questions, we test for differences in the HMM agent's
+investment by Phase, Condition and RS using a linear mixed-effects model
+as described in the methods section. As seen in Figure
+\@ref(fig:boxPlots), we find no main or interaction effects, indicating
+the HMM's behavior was on aggregate similar across levels of Phase,
+Condition and RS. This consistency in the investor's behavior is a
+desirable feature of the HMM agent when the participants behavior is
+largely simialr between phases. More importantly, it indicates that the
+lower returns of participants in the post phase of the manipulation
+condition were not differentiated enough to make the HMM react by
+transitioning to lower latent trust states. It is also noteworthy that
+the HMM agent was relatively successful in imitating human behavior in
+this paradigm: When asked during debrief whether they thought the
+investors they faced were human or not, $41$% of participants thought
+they were either facing a human or were not sure of the nature of the
+co-player. When asked to justify their choice, many answers reflected
+participants projecting human traits such as "spitefulness" or "greed"
+onto the artificial co-player's behavior.
 
@@
 ```
+
 So far we focused on analysing behavior for the pre and post phases.
-Here, we look at returns and investments in the exposure phase.
-The linear mixed effects model of participants' returns in the exposure phase does not show a main effect of Condition on returns.
-There was a main effect of Investment, `r papaja::apa_print(mod_returns_expo)$full_result$inv_scaled`, with participants positively reciprocating higher investments, an interaction effect between Condition and Investment `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2",papaja::apa_print(mod_returns_expo)$full_result$condition_f_inv_scaled)`, showing a stronger positive reciprocity in the Control condition, and finally a three way interaction between the RS group, Condition and Investment `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2",papaja::apa_print(mod_returns_expo)$full_result$condition_f_inv_scaled_high_RS)`, showing that this stronger positive reciprocity to investment in the Control condition is higher for participants with high RS.
-The linear mixed effects model of the HMM investments shows a main effect of Condition `r papaja::apa_print(mod_invs_expo)$full_result$condition_f`, suggesting higher overall investments for the forgiving HMM  compared to the human-like HMM, but no difference in investments when facing low and high RS groups.
-
-In summary, despite the forgiving HMM sending overall higher investments in the exposure phase, participants returned similar proportions of the multiplied investments as those facing the human-like HMM.
-The positive reciprocity of returns to investments was higher in the Control condition with this relationship stronger for the high RS group.
+Here, we look at returns and investments in the exposure phase. The
+linear mixed effects model of participants' returns in the exposure
+phase does not show a main effect of Condition on returns. There was a
+main effect of Investment,
+`r papaja::apa_print(mod_returns_expo)$full_result$inv_scaled`, with
+participants positively reciprocating higher investments, an interaction
+effect between Condition and Investment
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2",papaja::apa_print(mod_returns_expo)$full_result$condition_f_inv_scaled)`,
+showing a stronger positive reciprocity in the Control condition, and
+finally a three way interaction between the RS group, Condition and
+Investment
+`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2",papaja::apa_print(mod_returns_expo)$full_result$condition_f_inv_scaled_high_RS)`,
+showing that this stronger positive reciprocity to investment in the
+Control condition is higher for participants with high RS. The linear
+mixed effects model of the HMM investments shows a main effect of
+Condition `r papaja::apa_print(mod_invs_expo)$full_result$condition_f`,
+suggesting higher overall investments for the forgiving HMM compared to
+the human-like HMM, but no difference in investments when facing low and
+high RS groups.
+
+In summary, despite the forgiving HMM sending overall higher investments
+in the exposure phase, participants returned similar proportions of the
+multiplied investments as those facing the human-like HMM. The positive
+reciprocity of returns to investments was higher in the Control
+condition with this relationship stronger for the high RS group.
 
@@
 
-Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity  Questionnaire score (RSQ), Spearman's `r spearman_rho`, $p < 0.001$, there was no correlation between these questionnaire scores and participant's return or overall task performance.
+Whilst we found a significant correlation between participant's Levels
+of Personality Functioning Score (LPFS) and the Rejection Sensitivity
+Questionnaire score (RSQ), Spearman's `r spearman_rho`, $p < 0.001$,
+there was no correlation between these questionnaire scores and
+participant's return or overall task performance.
 
@@
 
-Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition and RS group.
-We will focus on two contrasts to analyse the ratings by Condition and RS group.
-The first is between the rating in the first phase ("pre") when participants phase the human-like HMM, and the average rating during the exposure phase (average of "expo1", "expo2" and "expo3") where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition).
-The second contrast is between the "pre" and "post" phases of the experiment where in both conditions participants face the same human-like HMM.
+Figure \@ref(fig:plotRatings) shows participants' ratings of each player
+they faced by condition and RS group. We will focus on two contrasts to
+analyse the ratings by Condition and RS group. The first is between the
+rating in the first phase ("pre") when participants phase the human-like
+HMM, and the average rating during the exposure phase (average of
+"expo1", "expo2" and "expo3") where they either face the forgiving HMM
+(Manipulation condition) or the human-like HMM again (Control
+condition). The second contrast is between the "pre" and "post" phases
+of the experiment where in both conditions participants face the same
+human-like HMM.
 
@@
 
-For those with high RS, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`.
-There was no difference in ratings on forgiveness and whether they would like to face the co-players again.
-Those in the Control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
-
-For those with low RS, there was no difference in any of the ratings between the "pre" and "exposure" phases for the Manipulation condition.
-For the Control condition, participants indicated less willingness to face the exposure co-player compared to the "pre" player, `r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo` but also did not differ on cooperation and forgiveness ratings.
-In summary, those with low RS had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving.
-In contrast, we see that exposing participants with high RS to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.
+For those with high RS, participants in the Manipulation condition rated
+the investors they faced in the exposure phase (the forgiving HMM) as
+more Cooperative
+`r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`.
+There was no difference in ratings on forgiveness and whether they would
+like to face the co-players again. Those in the Control condition rated
+the investors faced in the exposure group (same HMM) as less cooperative
+`r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`,
+less forgiving
+`r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`,
+and were less keen on facing them again
+`r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
+
+For those with low RS, there was no difference in any of the ratings
+between the "pre" and "exposure" phases for the Manipulation condition.
+For the Control condition, participants indicated less willingness to
+face the exposure co-player compared to the "pre" player,
+`r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo`
+but also did not differ on cooperation and forgiveness ratings. In
+summary, those with low RS had a mostly undifferentiated perception of
+players between the pre and exposure phases, even when the co-player was
+in fact designed to be more forgiving. In contrast, we see that exposing
+participants with high RS to a more cooperative and more forgiving agent
+has compensated for the decrease in ratings that would have occurred if
+faced with a co-player with the exact same strategy.
 
@@
 ### Comparing Pre and Post Ratings
+
 <!-- Comparing post-exposure to exposure ratings, participants in the Manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
+
 <!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
@@
 
-Participants high on RS in the Manipulation condition rated the co-players in the "post" phase similarly on cooperation, lower on 
-forgiveness `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
-and lower on willingness to face them again `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`. 
-Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
-(Cooperation: `r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
-Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
-Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).
-
-For participants low on RS, those in the Manipulation condition rated the co-players in the "post" phase lower on all three attributes 
+Participants high on RS in the Manipulation condition rated the
+co-players in the "post" phase similarly on cooperation, lower on
+forgiveness
+`r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
+and lower on willingness to face them again
+`r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`.
+Those in the Control condition rated the investors post the exposure
+phase lower on all three attributes (Cooperation:
+`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`,
+Forgiveness:
+`r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`,
+Play again:
+`r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).
+
+For participants low on RS, those in the Manipulation condition rated
+the co-players in the "post" phase lower on all three attributes
 (Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`,
-Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
-Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`). 
-Those in the Control condition did not differ in their ratings of the "pre" and "post" phase players.
-
-In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the Control condition.
-In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the Control condition even though the player continues to use the same strategy.
-After exposure to the forgiving HMM, both groups rate the "post" co-player worse than the "pre" even though they are the same.
+Forgiveness:
+`r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`,
+Play again:
+`r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`).
+Those in the Control condition did not differ in their ratings of the
+"pre" and "post" phase players.
+
+In summary, we again see that those with low RS accurately perceive the
+co-player as similar on all attributes throughout the phases in the
+Control condition. In contrast, the high RS group shows a negative bias
+towards the co-players after the "pre" phase in the Control condition
+even though the player continues to use the same strategy. After
+exposure to the forgiving HMM, both groups rate the "post" co-player
+worse than the "pre" even though they are the same.
 
@@
 
-We used a randomized controlled online experiment where participants played a RTG with artificial agents designed to simulate human-like trust-building scenarios.
-Participants were then exposed to either forgiving or human-like HMM agents before playing another RTG.
-Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation.
-Instead, participants reduced their returns overall whilst the returns of those in the Control group did not change between the pre and post phase of the experiment.
-Neither did participants show more forgiving behavior in the Manipulation condition after the one-off defection by the agent.
-Why did participants reduce their returns even though they were repeatedly exposed to more cooperative and more forgiving artificial agents?
-A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the Manipulation condition.
-
-Those exposed to the forgiving agent rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure.
-One possible explanation for this drop in rating is that participants exhibited a negative contrast effect.
-This occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people.
-If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favourable or even negative by comparison [@kobre_negative_1972].
-As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving agents [@zentall_within-trial_2005].
-If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.
-
-It is worth highlighting that RS was not correlated to overall performance in the task.
-Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
-However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their HMM co-players: They accurately perceived increased cooperativeness in the forgiving agent, showing an ability to adjust their judgments based on interaction patterns.
-This may indicate that perceived support from cooperative partners can mitigate their rejection concerns.
-However, their declining ratings of similar co-players over time, despite consistent agent behavior, suggests that occasional pre-programmed defections may have amplified their perceptions of untrustworthiness and reluctance for future interactions.
-These nuanced perceptions align with the heightened social awareness often associated with high rejection sensitivity [@downey_implications_1996].
-Individuals with low RS were less affected by the agent's behavioral variations, showing a less volatile baseline of social perception.
-They still exhibit behavior consistent with a contrast effect in the Manipulation condition.
-This stability in perception might reflect a lower propensity to scrutinize social cues or a more robust internal model of social interactions that is less easily perturbed by short-term variations in partner behavior.
-
-Despite these differences in perception, the actual cooperative behavior – as measured by returns in the trust game – did not significantly differ between RS groups.
-This suggests a potential dissociation between the cognitive-perceptual processes involved in evaluating social partners and the behavioral output in economic games.
-There are multiple potential explanations for this: First, the ratings may reflect explicit, conscious evaluations of social partners, while behavior in the trust game might be guided more by implicit, automatic processes that are less differentiated between RS groups.
-This interpretation aligns with dual-process theories of social cognition [@lieberman_social_2007].
-Second, the structured nature of the trust game might provide a context where the behavioral expression of rejection sensitivity is attenuated.
-In contrast, the more open-ended nature of providing ratings might allow for greater expression of RS-related perceptual biases.
-
-These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged.
-Conversely, therapies targeting these perceptual processes might not necessarily translate into immediate behavioral changes.
-Future research should investigate whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods.
-Additionally, exploring specific conditions or thresholds under which perceptual differences begin to manifest in behavioral divergences could provide valuable insights into the dynamics of rejection sensitivity.
-Simply exposing individuals to highly cooperative or forgiving partners may not suffice to induce lasting changes in behavior.
-Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time.
-Combining periods of forgiveness with clear communication about mutual expectations for cooperation might be more effective [@balliet_reward_2011].
-Future studies should also consider the temporal dynamics of these effects, examining whether longer exposure periods or different patterns of forgiving behavior could lead to more positive outcomes.
-Moreover, investigating the role of explicit communication and meta-cognitive reflection on the cooperative process could offer insights into leveraging positive interactions to promote enduring cooperative tendencies.
-
-<!-- This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of generative models of human behavior to design artificial agents that can interact in economic games.
-The use of these agents in the RTG led to similar investment and returns to those recorded in human dyadic interactions.
-Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.
-This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control.
--->
+We used a randomized controlled online experiment where participants
+played a RTG with artificial agents designed to simulate human-like
+trust-building scenarios. Participants were then exposed to either
+forgiving or human-like HMM agents before playing another RTG. We found
+that RS did not moderate participants' returns as trustees in the
+repeated trust game. While previous research has shown that RS affects
+*trust* formation, appraisal and repair, its impact on *reciprocity* in
+repeated economic exchanges has been less explored. Our results suggest
+a potential dissociation between RS's known effects on broader social
+behavior and its limited influence on reciprocity in structured,
+repeated interactions, challenging assumptions about the pervasive
+influence of RS on social behavior and highlighting the complexity of
+factors influencing reciprocity in economic exchanges.
+
+Contrary to our hypothesis, exposure to forgiving agents did not
+increase participant's reciprocity or cooperation, nor did it prompt the
+artificial agent to increase their trust in participants through higher
+investments. Instead, participants reduced their returns overall whilst
+the returns of those in the Control group did not change between the pre
+and post phase of the experiment. Why did participants reduce their
+returns even though they were repeatedly exposed to more cooperative and
+more forgiving artificial agents? A look at how the participants rated
+their co-players might shed some light on what might be driving this
+reduction in returns for those in the Manipulation condition. Those
+exposed to the forgiving agent rated their opponent in the post-exposure
+phase lower on all attributes even though they faced the same dynamic
+human-like HMM as pre-exposure. One possible explanation for this drop
+in rating is that participants exhibited a negative contrast effect.
+This occurs when the evaluation of a person, object, or situation is
+influenced by comparisons with recently encountered contrasting objects
+or people. If we've repeatedly interacted with someone exceptionally
+nice, our perception of a normal level of niceness might be skewed,
+making typical behavior seem less favourable or even negative by
+comparison [@kobre_negative_1972]. As the most recently faced opponents
+were more forgiving and cooperative, this negative contrast effect may
+have trumped any learning transfer from being repeatedly exposed to
+forgiving agents [@zentall_within-trial_2005]. If this contrast effect
+is indeed replicable, then an avenue for future research would be to use
+it to our benefit by making the participants play agents with low
+cooperation perception.
+
+This effect appears to operate independently of RS, as both groups
+exhibited it. However, RS did influence how participants perceived and
+reacted to their co-players' behavior. High RS individuals demonstrated
+a heightened attunement to changes in their co-players' behavior,
+accurately perceiving increased cooperativeness in the forgiving agent.
+Paradoxically, this same sensitivity may explain their decreased
+reactivity to investment levels and declining ratings of similar
+co-players over time in the Control condition. Occasional pre-programmed
+defections may have amplified their perceptions of untrustworthiness,
+leading to a more cautious approach and less responsiveness to
+investment variations. This aligns with the heightened social awareness
+and vigilance often associated with high RS [@downey_implications_1996].
+This finding suggests that individuals with high RS might benefit from
+interventions that focus on calibrating their perceptions of social cues
+in economic exchanges. In contrast, low RS individuals were less
+affected by the agent's behavioral variations, showing a more stable
+baseline of social perception. This stability in perception might
+reflect a lower propensity to scrutinize social cues or a more robust
+internal model of social interactions that is less easily perturbed by
+short-term variations in partner behavior. Their increased
+responsiveness to investments post-exposure in the Manipulation
+condition suggests they may have internalized the cooperative norms
+experienced, carrying over this increased reciprocity to subsequent
+interactions.
+
+Despite these differences in perception and investment reactivity, the
+actual cooperative behavior – as measured by returns in the trust game –
+did not significantly differ between RS groups. This suggests a
+potential dissociation between the cognitive-perceptual processes
+involved in evaluating social partners and the behavioral output in
+economic games. There are multiple potential explanations for this:
+First, the ratings may reflect explicit, conscious evaluations of social
+partners, while behavior in the trust game might be guided more by
+implicit, automatic processes that are less differentiated between RS
+groups. This interpretation aligns with dual-process theories of social
+cognition [@lieberman_social_2007]. Second, the structured nature of the
+trust game might provide a context where the behavioral expression of RS
+is attenuated. In contrast, the more open-ended nature of providing
+ratings might allow for greater expression of RS-related perceptual
+biases.
+
+These findings have important implications for interventions aimed at
+promoting trust and cooperation. Simply exposing individuals to
+artificially highly cooperative or forgiving partners may not suffice to
+induce lasting changes in behavior, and may even lead to negative
+contrast effects. Instead, interventions should focus on developing
+adaptive strategies for navigating the complexities of real-world social
+interactions, which often involve a mix of cooperative and
+non-cooperative behaviors. Such approaches could gradually shape
+expectations and behaviors over time, taking into account individual
+differences in RS. For individuals with high RS, interventions might
+specifically target the calibration of social perceptions in economic
+exchanges, helping to bridge the gap between their heightened
+sensitivity to social cues and their actual cooperative behavior. Future
+research should investigate whether the observed dissociation between
+perception and behavior persists in more naturalistic social
+interactions or over longer periods. Additionally, exploring the
+temporal dynamics of these effects and the role of explicit
+communication in the cooperative process could offer valuable insights.
+Interventions combining realistic cooperative scenarios with clear
+communication about mutual expectations might prove more effective in
+fostering enduring cooperative tendencies that are robust across various
+social contexts [@balliet_reward_2011].
+
+## Limitations
+
+While this study offers valuable insights into trust and cooperation
+dynamics in social interactions mediated by artificial agents, several
+limitations warrant consideration. The artificial nature of the
+interactions, while enabling high experimental control, may limit
+ecological validity, particularly regarding the paradigm's ability to
+trigger rejection sensitivity in participants. Additionally, the
+relatively brief exposure to forgiving HMM agents during the
+manipulation phase may have been insufficient to induce lasting changes
+in participants' reciprocity behavior. The online format of the
+experiment eliminates important social cues present in face-to-face
+interactions, which might influence trust and cooperation differently.
+Although the RSQ is a well-validated measure of rejection sensitivity
+(RS), future studies could benefit from a multi-method approach,
+potentially incorporating behavioral measures or additional self-report
+scales for a more comprehensive RS assessment. Furthermore, while our
+use of HMM agents to simulate human-like behavior in economic games is
+innovative, some participants may have suspected they were interacting
+with AI, potentially confounding the results. Despite these limitations,
+our findings provide a foundation for future research. Subsequent
+studies could address these constraints by incorporating face-to-face
+interactions, utilizing multiple measures of RS, examining the long-term
+effects of exposure to forgiving interaction partners, and developing
+more convincing cover stories or explicitly comparing responses to human
+and computerised partners.
+
+## Constraints on Generality
+
+The results may be specific to adults with high or low RS recruited from
+online platforms, and may not generalize to clinical populations,
+children, or older adults. Our use of a computerized Repeated Trust Game
+with HMM agents, while allowing for high experimental control, may limit
+generalizability to face-to-face interactions or games with different
+economic structures. The brief exposure phase and pre-programmed
+defections are specific to our design and may not reflect real-world
+trust-building scenarios. The online context may not capture all aspects
+of high-stakes or information-rich interactions. We believe the core
+finding of decreased cooperation after exposure to forgiving agents
+should generalize across different populations and contexts, though the
+effect's strength and its interaction with RS may vary. While the
+specific economic game, agent representation, and perception assessment
+questions could be varied, the use of artificial agents with consistent
+behavior, an exposure phase with more forgiving behavior, and assessment
+of both behavior and perceptions should remain constant to preserve the
+results. Future studies could systematically vary these factors to
+establish the boundaries of generalizability for our findings.
+Additionally, cultural differences in norms of cooperation and trust may
+influence the generalizability of these findings, necessitating
+cross-cultural replications to establish the universality of the
+observed effects.
 
@@
 # Author contributions statement {.unnumbered}
-I.
-Guennouni, G.
-Koppe and C.
-Korn.
-designed and developed the study concept.
-Experiment design, testing and data collection were performed by I.
+
+I. Guennouni, G. Koppe and C. Korn. designed and developed the study
+concept. Experiment design, testing and data collection were performed
+by I. Guennouni. I. Guennouni analysed and interpreted the data under
+the supervision of G. Koppe and C. Korn. All authors jointly wrote and
+approved the final version of the manuscript for submission.
+
+# Funding {.unnumbered}
+
+This study was supported by the Federal Ministry of Science, Education,
+and Culture (MWK) of the state of Baden-Württemberg within the AI Health
+Innovation Cluster Initiative, the German Research Foundation (DFG)
+within the Excellence Strategy EXC 2181/1 – 390900948 (STRUCTURES), and
+the Hector II foundation. I. Guennouni was supported by a research
+fellowship from the AI Health Innovation Cluster.
+
+# Competing interests statement {.unnumbered}
+
+The author(s) declared that there were no conflicts of interest with
+respect to the authorship or the publication of this article.
+
+# Acknowledgements {.unnumbered}
+
+We are grateful to Tobias Nolte, Andreas Hula and Read Montague's team
+at Virginia Tech for sharing with us the data on the RTG allowing us to
+fit the HMM to investor data. We're also grateful for Samuel Dupret for
+help designing the online platform used for the experiment and Maarten
+Speekenbrink for his guidance on the use of HMM models.
+
+# Additional information {.unnumbered}
+
+## Correspondence {.unnumbered}
+
+All correspondence and requests for materials should be addressed to I.
 Guennouni.
-I.
-Guennouni analysed and interpreted the data under the supervision of G.
-Koppe and C.
-Korn.
-All authors jointly wrote and approved the final version of the manuscript for submission.
-
-# Funding {.unnumbered}
-This publication was supported through state funds approved by the State Parliament of Baden-Württemberg for the Innovation Campus Health + Life Science alliance Heidelberg Mannheim.
-I.
-Guennouni was supported by a research fellowship from the AI Health Innovation Cluster
-
-# Competing interests statement {.unnumbered}
-The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.
-
-# Additional information  {.unnumbered}
-
-## Correspondence {.unnumbered}
-
-All correspondence and requests for materials should be addressed to I. Guennouni.
 
@@
 
-Preregistration: The hypotheses and methods were not preregistered.The analysis plan was not preregistered.
-Materials: All study materials are publicly available (https://github.com/ismailg/exposure-public).
-Data: All primary data are publicly available (https://github.com/ismailg/exposure-public).
-Analysis scripts: All analysis scripts are publicly available (https://github.com/ismailg/exposure-public).
+Preregistration: The hypotheses and methods were not preregistered.The
+analysis plan was not preregistered. Materials: All study materials are
+publicly available (<https://github.com/ismailg/exposure-public>). Data:
+All primary data are publicly available
+(<https://github.com/ismailg/exposure-public>). Analysis scripts: All
+analysis scripts are publicly available
+(<https://github.com/ismailg/exposure-public>).
 

### 2024-10-10 article_doc.Rmd
format changes JPCS

--- 2024-07-09 article.Rmd
@@
 author:
-  "Ismail Guennouni^1^*, Georgia Koppe^2^, Christoph Korn^3^"
+  "Ismail Guennouni^1,2,3,4^*, Georgia Koppe^1,2,3^$^\\dagger$, Christoph Korn^4^$^\\dagger$"
 
@@
 bibliography: "bib/exposure.bib"
-biblio-style: spphys
+csl: apa.csl
+
 output:
-  bookdown::pdf_document2:
-    keep_tex: true
-    toc: false
-header-includes:
-  - \usepackage{lscape}
+  bookdown::word_document2:
+    fig_caption: true
+    number_sections: false 
 ---
 \small
-^1^ *Central Institute of Mental Health, Medical Faculty Mannheim/Heidelberg University, Germany*
-
-^2^ *Faculty of Mathematics and Computer Science, Heidelberg University, Germany*
-
-^3^ *Department of General Psychiatry, Section Social Neuroscience, Heidelberg University, Germany*
-
-^\*^ *Corresponding author. Address: Central institute of Mental Health, J5, Mannheim, Germany. Email: ismail.guennouni@zi-mannheim.de* <br>
+^1^ *Department of Psychiatry and Psychotherapy, Central Institute of Mental Health, Medical Faculty Mannheim, Heidelberg University, Mannheim, Germany*
+
+^2^ *Interdisciplinary Center for Scientific Computing, Faculty of Mathematics and Computer Science, Heidelberg University, Heidelberg, Germany*
+
+^3^ *Hector Institute for AI in Psychiatry, Central Institute of Mental Health, Medical Faculty Mannheim, Heidelberg University, Mannheim Germany*
+
+^4^ *Department of General Psychiatry, Section Social Neuroscience, Heidelberg University, Germany*
+
+$^\dagger$  *Joint last author*
+
+^\*^ *Corresponding author. Address: Central institute of Mental Health, J5, Mannheim, Germany. Email: ismail.guennouni@zi-mannheim.de*
 
@@
 An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively, or the propensity to retaliate after breakdowns of cooperation.
+To further mimic real-world interactions and examine participants' responses to one-off breakdowns of cooperation, we incorporate occasional pre-programmed low investments by the agent.
 
@@
 In the Control condition, participants played three RTGs against the same human-like HMM.
-After this exposure phase, all participants played another 15-round RTG with a human-like HMM investor.
+After this exposure phase, all participants played another 15-round RTG with a human-like HMM investor, similar to the one in the pre-exposure phase.
 We expect those exhibiting high RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.
@@
 To have participants with large differences in RS, a total of 1195 participants were pre-screened on the Prolific Academic platform (prolific.co) using the Rejection Sensitivity Questionnaire (RSQ) to finally select two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10) totalling 206 participants (56% female).
-These were then invited through prolific to take part in the main experiment study.
-The required sample size was determined using an a priori power analysis to have an 80% probability to detect a small effect size with a 5% type I error rate in a repeated measures ANOVA with between factors.
+These were then invited through prolific to take part in the main experiment.
+The required sample size was determined using an *a priori* power analysis to have an 80% probability to detect a small effect size with a 5% type I error rate in a repeated measures ANOVA with between factors.
 The mean age of participants was 34.6 years, with an 11.9 years standard deviation.
@@
 
-The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : High or Low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A).
+The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : High or Low) by 2 (Phase: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the Phase factor (Figure \@ref(fig:HMMPanels).A).
 Participants within each pre-screened group were randomly assigned to one of the two levels of the Condition factor.
@@
 
-### Repeated Trust Game
+### Repeated Trust Game and HMM Investor
 
@@
 
-In all phases of the RTG other than the ‘Exposure phase’ (Figure \ref{fig:HMMPanels}.A), participants interacted with this human-like HMM.
+In all phases of the RTG other than the ‘Exposure phase’ (Figure \@ref(fig:HMMPanels).A), participants interacted with this human-like HMM.
 In the ‘Manipulation’ Condition of the exposure phase, however, the parameters of this HMM were adjusted to design a ‘forgiving’ and ultimately more cooperative agent.
 To achieve this, we changed the state transition probabilities of the HMM such that it becomes impossible for it to remain in a low trust state, effectively setting the transition probability for remaining in a “low-trust” state to 0.
-The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D.
+The resulting transition function is shown in Figure \@ref(fig:HMMPanels).D.
 The policies conditional on the latent states and the transition function in the other latent states remain unchanged.
-
-\begin{landscape}
-```{r HMMPanels, fig.cap= "\\small{A: A timeline of the experiment for both conditions.
-The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor.
-The investor is endowed with 20 units at the start of each round.
-They need to decide how much of that endowment they want to invest with the trustee.
-The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor.
-The difference between conditions is the type of agents participants are exposed to in the exposure phase.
-Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG against human trustees.
-From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B.
-The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C.
-Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
-Panel D shows the transition probabilities of the forgiving HMM agent in the low trust state.
-Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state, as a proxy for coaxing behavior.
-Transitions in the medium and high trust states were identical for both agents.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
-
-library(magick)
-
-# Read in the images
-image1 <- image_read("figures/timeline.png")
-image2 <- image_read("figures/HMMinvPolicy2.png")
-image3 <- image_read("figures/HMMinvTrans2.png")
-image4 <- image_read("figures/HMM_switch.png")
-
-# Determine the target width for all images, for example, the width of the widest image
-target_width <- max(image_info(image1)$width, image_info(image2)$width, 
-                    image_info(image3)$width, image_info(image4)$width)
-
-target_height <- max(image_info(image1)$height, image_info(image2)$height, 
-                    image_info(image3)$height, image_info(image4)$height)
-
-# Scale images to have the same width
-image1 <- image_scale(image1, paste0(target_width, "x", target_height))
-image2 <- image_scale(image2, paste0(target_width, "x", target_height))
-image3 <- image_scale(image3, paste0(target_width, "x", target_height))
-image4 <- image_scale(image4, paste0(target_width, "x", target_height))
-
-# Annotate each image with a label
-image1 <- image_annotate(image1, text = "A", location = "+0+10", size = 90, color = "black", gravity = "northwest")
-image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
-image3 <- image_annotate(image3, text = "C", location = "-0+10", size = 90, color = "black", gravity = "northwest")
-image4 <- image_annotate(image4, text = "D", location = "+0+10", size = 90, color = "black", gravity = "northwest")
-
-# Assuming a 300 DPI resolution for good print quality
-dpi <- 300
-
-# Spacing of 0.1 inches converted to pixels
-spacing <- 0.05 * dpi
-
-# Calculating the width and height for each image
-width_per_image <- ((12 - 0.1) / 2) * dpi
-height_per_image <- ((7.4 - 0.1) / 2) * dpi
-# The geometry string for each image, including spacing (assuming no borders)
-geometry_string <- paste0(width_per_image, 'x', height_per_image, '+', spacing, '+', spacing)
-
-# Now create the montage with the specified geometry
-combined_image <- image_montage(
-  c(image1, image2, image3, image4), 
-  tile = "2x2",
-  geometry = geometry_string
-)
-
-# Display the combined image
-plot(combined_image)
-```
-\end{landscape}
 
@@
 This simulated the waiting time in real social interaction tasks.
-After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-playerr to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
+After completing each RTG in each phase, participants rated how cooperative and forgiving they perceived the co-player to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).
 After completing the three game phases, participants then completed the Levels of Personality Functioning Scale Brief-Form (LPFS-BF) questionnaire [@weekers_level_2019].
 This is a self-report measure designed to assess core elements of personality functioning as defined in the Alternative Model for Personality Disorders in the DSM-5 [@american_psychiatric_association_diagnostic_2013], and provides a dimensional assessment of personality functioning, which complements the categorical approach of RS.
-Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.
+Finally, participants were asked whether they thought the other players were human or computer agents, to probe how well the agent can mimic human behavior, then debriefed and thanked for their participation.
+
+```{r HMMPanels, fig.cap= "Panel A: a timeline of the experiment for both conditions.
+The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor.
+The investor is endowed with 20 units at the start of each round.
+They need to decide how much of that endowment they want to invest with the trustee.
+The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor.
+The difference between conditions is the type of agents participants are exposed to in the exposure phase.
+Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG.
+From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B.
+The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C.
+Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
+Panel D shows the transition probabilities of the forgiving HMM agent in the low trust state.
+Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state, as a proxy for coaxing behavior.
+Transitions in the medium and high trust states were identical for both agents.", include=T, fig.width=11, fig.height=8.5, out.width="100%", dpi=300, dev='png'}
+
+library(magick)
+
+# Read in the images
+image1 <- image_read("figures/timeline.png")
+image2 <- image_read("figures/HMMinvPolicy2.png")
+image3 <- image_read("figures/HMMinvTrans2.png")
+image4 <- image_read("figures/HMM_switch.png")
+
+# Function to scale image while maintaining aspect ratio
+scale_image <- function(img, target_width, target_height) {
+  img_info <- image_info(img)
+  aspect_ratio <- img_info$width / img_info$height
+
+  if (aspect_ratio > (target_width / target_height)) {
+    new_width <- target_width
+    new_height <- target_width / aspect_ratio
+  } else {
+    new_height <- target_height
+    new_width <- target_height * aspect_ratio
+  }
+
+  image_scale(img, paste0(new_width, "x", new_height))
+}
+
+# Calculate target dimensions (97% of half the page to leave minimal margin)
+page_width <- 11 * 300  # 11 inches at 300 DPI
+page_height <- 8.5 * 300  # 8.5 inches at 300 DPI
+target_width <- (page_width * 0.97) / 2
+target_height <- (page_height * 0.97) / 2
+
+# Scale images
+image1 <- scale_image(image1, target_width, target_height)
+image2 <- scale_image(image2, target_width, target_height)
+image3 <- scale_image(image3, target_width, target_height)
+image4 <- scale_image(image4, target_width, target_height)
+
+# Annotate each image with a label
+label_size <- 72
+image1 <- image_annotate(image1, text = "A", location = "+10+10", size = label_size, color = "black", font = "Arial-Bold")
+image2 <- image_annotate(image2, text = "B", location = "+10+10", size = label_size, color = "black", font = "Arial-Bold")
+image3 <- image_annotate(image3, text = "C", location = "+10+10", size = label_size, color = "black", font = "Arial-Bold")
+image4 <- image_annotate(image4, text = "D", location = "+10+10", size = label_size, color = "black", font = "Arial-Bold")
+
+# Add a small border to each image to create spacing
+border_size <- 5
+image1 <- image_border(image1, "white", paste0(border_size, "x", border_size))
+image2 <- image_border(image2, "white", paste0(border_size, "x", border_size))
+image3 <- image_border(image3, "white", paste0(border_size, "x", border_size))
+image4 <- image_border(image4, "white", paste0(border_size, "x", border_size))
+
+# Create the montage
+combined_image <- image_append(
+  c(image_append(c(image1, image2), stack = TRUE),
+    image_append(c(image3, image4), stack = TRUE))
+)
+
+# Display the combined image
+plot(combined_image)
+```
 
@@
 
-To explore whether participants behaved differently in the RTG after the manipulation compared to the Control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.
+To test whether participants behaved differently in the RTG after the manipulation compared to the Control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs.
 post-manipulation), Condition (manipulation vs.
@@
 We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.
+To probe significant interactions, we conducted planned contrasts using the `emmeans` package in R.
+Given that we were testing multiple pre-planned comparisons, we applied the "Sidak" correction to control for familywise error rate while maintaining reasonable statistical power.
+This approach allowed us to investigate specific hypotheses about the differential effects of our manipulation across phases and RS groups, while protecting against inflated Type I error rates.
 
@@
 
-On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60% of the endowment for investments and 35-50% of the total yield for returns, as reported in previous studies [@charness_investment_2008; @fiedler_social_2011].
+On average, investments and returns, as shown in Figure \@ref(fig:gamesPlot), fell within the documented range of 40-60% of the endowment for investments and 35-50% of the total yield for returns, as reported in previous studies [@charness_investment_2008; @fiedler_social_2011].
 
@@
 post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
-As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the Manipulation condition, pre - post,
+As shown in Figure \@ref(fig:boxPlots), post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the Manipulation condition, pre - post,
 `r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the Control condition.
@@
 Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
-For completeness, these effects are discussed in more detail in the supplement.
 
@@
 
+```
+
+```{r}
+library(emmeans)
+
+# First, get the estimated marginal means
+emm <- emmeans(mod_returns_pct_RS, ~ inv_scaled | phase.f * condition.f * high_RS)
+
+# Then, get the slopes (effects of investment) for each combination
+slopes <- emtrends(mod_returns_pct_RS, ~ phase.f * condition.f * high_RS, var = "inv_scaled")
+slopes
+
+# Define the contrasts
+corrected_contrasts <- list(
+  "Diff in inv effect (Post - Pre, Manip, Low RS)" = c(0, 0, -1, 1, 0, 0, 0, 0),
+  "Diff in inv effect (Post - Pre, Manip, High RS)" = c(0, 0, 0, 0, 0, 0, -1, 1),
+  "Diff in (Post - Pre) between High and Low RS" = c(0, 0, 1, -1, 0, 0, -1, 1)
+)
+
+# Test the contrasts with multiple comparison correction
+results_4w <- contrast(slopes, method = corrected_contrasts, adjust = "sidak")
+
+# View the results
+summary(results_4w)
+
+```
+To examine this four-way interaction, we conducted a contrast analysis of how the effect of investment on returns changed from pre- to post-manipulation for different rejection sensitivity (RS) groups in both conditions.
+Starting with the Manipulation condition, for participants with low RS, the effect of investment on returns increased significantly from pre- to post-manipulation,
+`r papaja::apa_print(results_4w)$full_result$DiffininveffectPost_PreManipLowRS`.
+This suggests that after the manipulation, low RS participants became more responsive to their co-player's investments, returning proportionally more as investments increased.
+In contrast, for participants with high RS, the effect of investment on returns decreased significantly from pre- to post-manipulation,
+`r papaja::apa_print(results_4w)$full_result$DiffininveffectPost_PreManipHighRS`.
+This indicates that high RS participants became less responsive to their co-player's investments after the manipulation, with smaller increases in returns as investments increased.
+The difference in these pre-post changes between high and low RS groups was significant,
+`r papaja::apa_print(results_4w)$full_result$DiffinPost_PrebetweenHighandLowRS`.
+This result suggests that the manipulation had significantly different effects on how low and high RS participants responded to their co-player's investments.
+
+In the Control condition, we observed no significant changes in how participants responded to their co-player's investments between the pre and post phases, regardless of their RS level.
+
+```{r}
+# Define the contrasts for both conditions
+contrasts_ctrl <- list(
+  # Control condition contrasts
+  "Diff in inv effect (Post - Pre, Control, Low RS)" = c(-1, 1, 0, 0, 0, 0, 0, 0),
+  "Diff in inv effect (Post - Pre, Control, High RS)" = c(0, 0, 0, 0, -1, 1, 0, 0),
+  "Diff in (Post - Pre) between High and Low RS in Control" = c(1, -1, 0, 0, -1, 1, 0, 0)
+)
+
+# Test the contrasts with multiple comparison correction
+results_ctrl <- contrast(slopes, method = contrasts_ctrl, adjust = "sidak")
+
+# View the results
+print(summary(results_ctrl))
 ```
@@
 We fit the same mixed effects model as for all the trials with the exception of the RS variable.
-This is because RS did not show main or interaction effects in the main model, and also due to the necessity of running a simpler model to accommodate the low number of trials.
+This is because RS did not show main effects in the previous model, and also due to the necessity of running a simpler model to accommodate the low number of trials.
 We find a significant main effect of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` with returns lower in the second game post defection trials compared to the first.
@@
 `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the Manipulation condition compared to the Control condition in post-defection trials.
-However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.
+However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off low investment by the co-player.
 
@@
 
-It is worth highlighting that RS was not correlated to overall performance in the task.
-Neither did it moderate the change in returns between Conditions or have an effect on participant returns.
+This effect appears to operate independently of RS, as both groups exhibited it.
+However, RS did influence how participants perceived and reacted to their co-players' behavior.
+High RS individuals demonstrated a heightened attunement to changes in their co-players' behavior, accurately perceiving increased cooperativeness in the forgiving agent.
+Paradoxically, this same sensitivity may explain their decreased reactivity to investment levels and declining ratings of similar co-players over time in the Control condition.
+Occasional pre-programmed defections may have amplified their perceptions of untrustworthiness, leading to a more cautious approach and less responsiveness to investment variations.
+This aligns with the heightened social awareness and vigilance often associated with high rejection sensitivity [@downey_implications_1996].
+In contrast, low RS individuals were less affected by the agent's behavioral variations, showing a more stable baseline of social perception.
+This stability in perception might reflect a lower propensity to scrutinize social cues or a more robust internal model of social interactions that is less easily perturbed by short-term variations in partner behavior.
+Their increased responsiveness to investments post-exposure in the Manipulation condition suggests they may have internalized the cooperative norms experienced, carrying over this increased reciprocity to subsequent interactions.
+
+<!-- It is worth highlighting that RS was not correlated to overall performance in the task.
+Neither did it moderate the change in returns between Conditions or have an effect on average participant returns.
 However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their HMM co-players: They accurately perceived increased cooperativeness in the forgiving agent, showing an ability to adjust their judgments based on interaction patterns.
@@
 This stability in perception might reflect a lower propensity to scrutinize social cues or a more robust internal model of social interactions that is less easily perturbed by short-term variations in partner behavior.
-
-Despite these differences in perception, the actual cooperative behavior – as measured by returns in the trust game – did not significantly differ between RS groups.
+-->
+
+Despite these differences in perception and investment reactivity, the actual cooperative behavior – as measured by returns in the trust game – did not significantly differ between RS groups.
 This suggests a potential dissociation between the cognitive-perceptual processes involved in evaluating social partners and the behavioral output in economic games.
@@
 
-These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged.
+<!-- These findings suggest that interventions focused solely on altering behavior might be insufficient, as the underlying perceptual and cognitive processes may remain unchanged.
 Conversely, therapies targeting these perceptual processes might not necessarily translate into immediate behavioral changes.
@@
 Moreover, investigating the role of explicit communication and meta-cognitive reflection on the cooperative process could offer insights into leveraging positive interactions to promote enduring cooperative tendencies.
-
-<!-- This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of generative models of human behavior to design artificial agents that can interact in economic games.
-The use of these agents in the RTG led to similar investment and returns to those recorded in human dyadic interactions.
-Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism.
-This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control.
 -->
+
+These findings have important implications for interventions aimed at promoting trust and cooperation.
+Simply exposing individuals to highly cooperative or forgiving partners may not suffice to induce lasting changes in behavior, and may even lead to negative contrast effects.
+Instead, interventions may need to incorporate more nuanced approaches that gradually shape expectations and behaviors over time, taking into account individual differences in rejection sensitivity.
+Future research should investigate whether this dissociation between perception and behavior persists in more naturalistic social interactions or over longer periods.
+Additionally, exploring the temporal dynamics of these effects and the role of explicit communication in the cooperative process could offer valuable insights.
+Combining periods of forgiveness with clear communication about mutual expectations for cooperation might prove more effective in fostering enduring cooperative tendencies [@balliet_reward_2011].
+
+While this study provides valuable insights into the dynamics of trust and cooperation in social interactions mediated by AI agents, several limitations should be considered.
+First, the artificial nature of the interactions, while allowing for high experimental control, may limit the ecological validity of our findings.
+Second, the relatively short-term exposure to forgiving AI agents during the manipulation phase may not have been sufficient to induce lasting changes in participants' trust and cooperative behavior.
+Future studies could benefit from longer-term interventions to examine more enduring effects.
+Furthermore, the online nature of the experiment eliminates important social cues present in face-to-face interactions, which might influence trust and cooperation differently.
+Finally, while our use of HMM agents to simulate human-like behavior in economic games is innovative, it's important to acknowledge that some participants may have suspected they were interacting with AI, potentially introducing confound to the results.
+Despite these limitations, our findings provide a foundation for future research in this area.
+Subsequent studies could address these limitations by incorporating face-to-face interactions, using multiple measures of rejection sensitivity, and examining the long-term effects of exposure to forgiving interaction partners.
 
@@
 # Funding {.unnumbered}
-This publication was supported through state funds approved by the State Parliament of Baden-Württemberg for the Innovation Campus Health + Life Science alliance Heidelberg Mannheim.
+This study was supported by the Federal Ministry of Science, Education, and Culture (MWK) of the state of Baden-Württemberg within the AI Health Innovation Cluster Initiative, the German Research Foundation (DFG) within the Excellence Strategy EXC 2181/1 – 390900948 (STRUCTURES), and the Hector II foundation.
 I.
-Guennouni was supported by a research fellowship from the AI Health Innovation Cluster
+Guennouni was supported by a research fellowship from the AI Health Innovation Cluster.
 
@@
 The author(s) declared that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+# Acknowledgements {.unnumbered}
+
+We are grateful to Tobias Nolte, Andreas Hula and Read Montague's team at Virginia Teach for sharing with us the data on the RTG allowing us to fit the HMM.
+We're also grateful for Samuel Dupret for help designing the online platform used for the experiment and Maarten Speekenbrink for his guidance on the use of HMM models.
 

### 2025-12-05 article.Rmd
Revise manuscript: add event study analysis, strengthen limitations, update terminology
- Add event study and pre-low-investment analyses to test contrast vs betrayal aversion
- Acknowledge confound between forgiveness and low investment exposure in Limitations
- Change "defection" to "low investment" terminology throughout
- Reframe RS findings: intact detection but impaired recovery in high RS
- Condense Player Ratings section; revise abstract and discussion

@@
   bookdown::pdf_document2:
+   # bookdown::word_document2:
     keep_tex: true
@@
 
-Can exposure to forgiving behavior improve cooperative outcomes, and
-does Rejection Sensitivity (RS) moderate this effect? This randomized
-controlled experiment explores these questions using a novel approach
-with Hidden Markov Model (HMM)-based artificial agents in repeated trust
-games (RTGs). Participants (N = 206), pre-screened for high or low RS,
-interacted with either consistently behaving or more forgiving
-co-players, which were in reality HMM agents. Unexpectedly, exposure to
-forgiving agents led to reduced cooperation in subsequent interactions.
-Control group participants maintained consistent behavior. RS levels
-influenced perceptions of the agent's cooperativeness but did not
-significantly affect behavioral outcomes, revealing a
-perception-behavior dissociation. These findings challenge assumptions
-about fostering cooperation through simple exposure to positive
-interactions and highlight the importance of accounting for negative
-contrast effects when designing such interventions. They also
-demonstrate the potential of HMM-based artificial agents for studying
-interactive social dynamics, offering a methodological advancement for
-future research. Our results suggest the need for nuanced interventions
-that gradually shape expectations and behaviors to foster cooperative
-outcomes, particularly addressing the discrepancy between perceptions
-and behavior in individuals with high rejection sensitivity.
+Can exposure to forgiving partners improve interpersonal cooperation?
+Attachment theory suggests positive relational experiences can correct negative internal working models, but individuals high in Rejection Sensitivity (RS)—characterized by anxious expectations of rejection—may struggle to benefit from such experiences.
+We tested this using a randomized experiment (N = 206) in which participants played repeated trust games with HMM-based artificial agents that simulate human-like trust dynamics.
+After a baseline game, participants were exposed to either forgiving agents (Manipulation) or standard agents (Control), then played a final game with a standard agent.
+Contrary to predictions, forgiveness exposure *reduced* subsequent cooperation—participants appeared to perceive the standard post-exposure agent as less cooperative by comparison (a negative contrast effect).
+Critically, RS moderated specific behavioral patterns but not overall cooperation levels: high RS participants failed to recover cooperation after trust violations and became *less* responsive to partner behavior following exposure, whereas low RS participants showed normal recovery and became *more* responsive.
+These findings suggest that positive relational experiences do not universally promote cooperation, and that high RS individuals may require interventions targeting their capacity to update expectations rather than simply providing positive experiences.
 
@@
 
-This study reveals that exposing individuals to forgiving behavior in
-economic games can unexpectedly decrease their subsequent cooperation.
-Interestingly, while people with high rejection sensitivity more
-accurately perceived their partners' cooperativeness, this didn't
-translate into more cooperative behavior. These findings challenge
-assumptions about fostering cooperation and highlight the need for more
-nuanced interventions to promote reciprocity, especially for individuals
-prone to expecting rejection.
+This study found that exposing people to forgiving partners in economic games unexpectedly decreased their subsequent cooperation—likely because normal partners seemed less cooperative by comparison.
+Individuals high in rejection sensitivity showed a distinct pattern: while they detected trust violations as readily as others, they failed to restore cooperation afterward and became less responsive to their partner's behavior.
+In contrast, those low in rejection sensitivity appeared to learn from the positive exposure and became more reciprocal.
+These findings suggest that simply providing positive social experiences may not benefit everyone equally, and that individuals prone to rejection sensitivity may need targeted support to translate positive experiences into lasting behavioral change.
 
@@
 library(magick)
+
+# Helper function to format emmeans contrast results in APA style
+format_emmeans_apa <- function(result_row) {
+  est <- result_row$estimate
+  se <- result_row$SE
+  df <- result_row$df
+  t_val <- result_row$t.ratio
+  p_val <- result_row$p.value
+
+  # Format p-value
+  if (p_val < .001) {
+    p_str <- "p < .001"
+  } else {
+    p_str <- paste0("p = ", papaja::printp(p_val))
+  }
+
+  # Format the string
+  sprintf("$\\Delta M$ = %.2f, $SE$ = %.2f, $t$(%.1f) = %.2f, %s",
+          est, se, df, t_val, p_str)
+}
 
@@
 been studied as extensively, leaving a gap in our understanding of how
-these constructs might interplay. Could an ingrained sensitivity to
-rejection contribute to reduced pro-social behavior such as a failure to
-show reciprocity as a result of perceived rejection by the interaction
-partner?
+these constructs might interplay.
 
 Given that RS may be a manifestation of maladaptive attachment styles,
-it's crucial to explore whether exposure to consistently forgiving
-interaction partners could help reshape these interpersonal expectations
-and behaviors. If RS indeed influences reciprocity in repeated
-interactions, could such exposure serve as a corrective experience,
-potentially improving interpersonal outcomes? This approach aligns with
-attachment theory, which suggests that new, positive relational
-experiences can modify internal working models of relationships
-[@bowlby_secure_1988]. Research in the fields of behavioral economics
-and psychology has also explored how positive social interactions
-influence cooperation and pro-social behavior. The use of the repeated
-trust game (RTG), a well-established experimental approach, has allowed
-for the analysis of the development of trust through ongoing
-interactions [@joyce_trust_1995]. In this paradigm, cycles of mutual
-trust, where each party's trust is reciprocated with trustworthiness,
-have the effect of enhancing cooperative behaviors even among
-individuals who are initially inclined to be distrustful
-[@king-casas_getting_2005]. @fowler_cooperative_2010 studied behavior in
-social networks interacting in a public goods game and found that
-cooperative behavior tends to cluster, suggesting that exposure to
-cooperative peers can lead to more cooperative behavior. Similarly,
-research on social learning theory [@bandura_social_1977] has long
-demonstrated that individuals learn and model the behavior of those
-around them, indicating that if someone is consistently exposed to
-cooperative and positive individuals, they're likely to emulate this
-behavior. These insights highlight that engaging with compassionate and
-forgiving others can be an effective method for mitigating deeply
-ingrained mistrust and potentially fostering trustworthy behavior.
+it is important to explore whether exposure to consistently forgiving
+interaction partners could reshape interpersonal expectations and
+behaviors. The *corrective experience hypothesis*, rooted in attachment
+theory, suggests that new positive relational experiences can modify
+internal working models of relationships [@bowlby_secure_1988]. Research
+on social learning [@bandura_social_1977] similarly demonstrates that
+individuals model the behavior of those around them, and exposure to
+cooperative peers promotes cooperative behavior
+[@fowler_cooperative_2010]. In the repeated trust game (RTG) paradigm,
+cycles of reciprocated trust enhance cooperative behaviors even among
+initially distrustful individuals [@king-casas_getting_2005]. This
+perspective predicts that exposure to forgiving partners should increase
+subsequent cooperation, as participants internalize more positive
+expectations about social interactions. Given that high RS individuals
+show heightened sensitivity to social cues, they might be particularly
+responsive to such corrective experiences.
 
@@
 exposed over three RTGs to HMM investors designed with a limited
-propensity for retaliation, potentially mitigating ingrained mistrust.
-In the Control condition, participants played three RTGs against the
-same human-like HMM. After this exposure phase, all participants played
-another 15-round RTG with a human-like HMM investor, similar to the one
-in the pre-exposure phase. We expect those exhibiting high RS to be more
-responsive to the pre-programmed change in the agent's cooperativeness
-and retaliation propensity. We also hypothesize that those in the
-Manipulation condition would behave more cooperatively and have a lower
-propensity to retaliate to the co-player's defection after the exposure
-phase, with those in the higher RS group showing a bigger effect.
+propensity for retaliation. In the Control condition, participants
+played three RTGs against the same human-like HMM. After this exposure
+phase, all participants played another 15-round RTG with a human-like
+HMM investor, similar to the one in the pre-exposure phase.
+
+Based on the corrective experience hypothesis, we predicted that
+forgiveness exposure would increase subsequent cooperation, with high RS
+individuals potentially benefiting from positive relational experiences
+that challenge their negative expectations. We examined both overall
+effects of the manipulation and differential responses based on RS, with
+particular attention to how participants respond to trust violations
+before and after the exposure phase.
 
@@
 The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS :
-High or Low) by 2 (Phase: Trust-Game Pre Manipulation, Trust-Game Post
-Manipulation) design, with repeated measures on the Phase factor (Figure
+High or Low) by 2 (Phase: Trust-Game Pre-Exposure, Trust-Game Post-Exposure) design, with repeated measures on the Phase factor (Figure
 \@ref(fig:HMMPanels).A). Participants within each pre-screened RS group
@@
 us to probe efforts to repair it, the computerised agent was programmed
-to provide a low investment on round 12 (pre-manipulation) and round 13
-(post-manipulation). On all other rounds, the investor’s actions were
+to provide a low investment on round 12 (pre-exposure) and round 13
+(post-exposure). On all other rounds, the investor’s actions were
 determined by randomly drawing an investment from the state-conditional
@@
 forgiving HMM investor and rated the agent on the same attributes as in
-the pre-manipulation phase. Those in the Control condition faced the
+the pre-exposure phase. Those in the Control condition faced the
 same human-like HMM agent as in the "pre" phase and rated each co-player
@@
 very low investment in round 5 of each of the three games. In the
-post-manipulation phase ("post"), participants in both conditions faced
+post-exposure phase ("post"), participants in both conditions faced
 the same human-like HMM as in "pre" phase.
@@
 \begin{landscape}
-```{r HMMPanels, fig.cap= "\\small{A: A timeline of the experiment for both conditions.
-The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor.
-The investor is endowed with 20 units at the start of each round.
-They need to decide how much of that endowment they want to invest with the trustee.
-The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor.
-The difference between conditions is the type of agents participants are exposed to in the exposure phase.
-Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG.
-From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B.
-The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C.
-Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
-Panel D shows the transition probabilities of the forgiving HMM agent in the low trust state.
-Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state, as a proxy for coaxing behavior.
-Transitions in the medium and high trust states were identical for both agents.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}
+```{r HMMPanels, fig.cap= "A: Experiment timeline.
+Participants (trustees) played RTGs with HMM investor agents.
+The investor sends investments (multiplied by 3) and participants decide returns.
+Conditions differ in exposure phase agents.
+B-D: The artificial investor is a three-state HMM fitted to human data.
+B: Investment distributions by latent state.
+C: Transition probabilities to states at t+1 as a function of net return at t; each panel shows a different starting state.
+D: Forgiving HMM transitions from low-trust state---unlike the human-like HMM, it always exits low-trust and favors high-trust transitions.", include=T, out.width='0.95\\linewidth', fig.align='center'}
 
@@
 
-# Calculating the width and height for each image
-width_per_image <- ((12 - 0.1) / 2) * dpi
-height_per_image <- ((7.4 - 0.1) / 2) * dpi
+# Calculating the width and height for each image (with shorter caption, can use more space)
+width_per_image <- ((10.5 - 0.1) / 2) * dpi
+height_per_image <- ((5.8 - 0.1) / 2) * dpi
 # The geometry string for each image, including spacing (assuming no borders)
@@
 
-To test whether participants behaved differently in the RTG after the
-manipulation compared to the Control group, we model the percentage
-return (percentage of tripled investment returned to investor) using a
-linear mixed-effects model with Phase (RTG game pre vs.
-post-manipulation), Condition (manipulation vs. Control), Investment,
-and RS (High vs Low RS group) as well as their interactions as fixed
-effects, and player-wise random intercepts and slopes for Phase. The
-full specification of the statistical model can be found in the
-supplement. To test whether participants behaved differently between
-conditions in the Exposure phase, we fit a linear mixed effects model to
-participants returns only in the exposure phase, with Condition
-(manipulation vs. Control), Investment, and RS (High vs Low RS group) as
-well as their interactions as fixed effects, and player-wise random
-intercepts. Finally, to test whether the HMM agent's behavior differed
-between Phases, Conditions and RS groups, we estimate a linear
-mixed-effects model of investments sent by the computerised HMM agent
-with Condition, Phase and RS and their interaction as fixed effects, and
-a similar random effects structure to the returns model.
-
-The model was estimated using the `afex` package [@singmann_afex_2022]
-in R. More complex models with additional random effects could not be
-estimated reliably, and as such the estimated model can be considered to
-include the optimal random effects structure
-[@matuschek_balancing_2017]. A similar process was used to establish the
-random effects structures of linear mixed-effects models used to analyse
-the HMM agent investments as well as the participants' ratings of the
-co-players. There is no agreed upon way to calculate effect sizes for
-mixed effects models. Instead, we will report on testing differences in
-marginal means. For the $F$-tests, we used the Kenward-Roger
-approximation to the degrees of freedom, as implemented in the R package
-"afex". We Z-transform the Investment variable (subtract the overall
-investment mean and divide by overall standard deviation) as centering
-is beneficial to interpreting the main effects more easily in the
-presence of interactions. To probe significant interactions, we
-conducted planned contrasts using the `emmeans` package in R. Given that
-we were testing multiple pre-planned comparisons, we applied the "Sidak"
-correction to control for familywise error rate while maintaining
-reasonable statistical power. This approach allowed us to investigate
-specific hypotheses about the differential effects of our manipulation
-across phases and RS groups, while protecting against inflated Type I
-error rates.
+We analyzed participants' behavior in the RTG using linear mixed-effects models.
+First, to examine the effect of the manipulation, we modeled the percentage return (percentage of tripled investment returned to investor) as a function of Phase (RTG game pre vs.
+post-exposure), Condition (Manipulation vs.
+Control), Investment, and RS (High vs Low RS group), including all interactions as fixed effects.
+This model included player-wise random intercepts and slopes for Phase.
+Second, we analyzed behavior during the Exposure phase specifically, modeling returns with Condition, Investment, and RS and their interactions as fixed effects, along with player-wise random intercepts.
+Third, to verify the consistency of the HMM agent, we modeled the investments sent by the computerized agent using Condition, Phase, and RS and their interactions as fixed effects.
+To isolate effects occurring prior to any pre-programmed low investment, we also analyzed returns in rounds preceding the low investment trials only (rounds 1-11 in the pre-exposure phase and rounds 1-12 in the post-exposure phase) using the same model specification.
+Finally, to rigorously assess participants' reactions to and recovery from the specific instance of pre-programmed low investment, we conducted an event study analysis centered on the low investment round ($t=0$).
+We analyzed percentage returns in a three-round window ($t-1$ to $t+1$) using a linear mixed-effects model with Phase, Condition, Time Point, and RS Group as fixed effects.
+We specifically examined two key behavioral responses: the Drop (change in return from $t-1$ to the low investment round) and the Recovery (change in return from the low investment round to $t+1$).
+The full specification of all statistical models can be found in the supplement.
+
+All models were estimated using the `afex` package [@singmann_afex_2022] in R.
+We determined the random effects structure by starting with the maximal model and simplifying until convergence was achieved, ensuring the optimal structure [@matuschek_balancing_2017].
+A similar process was applied to the models analyzing HMM agent investments and participant ratings.
+We report differences in marginal means rather than effect sizes, as there is no consensus on effect size calculation for mixed models.
+$F$-tests used the Kenward-Roger approximation for degrees of freedom.
+The Investment variable was Z-transformed to facilitate the interpretation of main effects in the presence of interactions.
+Significant interactions were probed using planned contrasts with the `emmeans` package.
+We applied the "Sidak" correction for multiple comparisons to control the familywise error rate while maintaining statistical power.
 
@@
 ```{r}
-# average investment for first 10 rounds pre manipulation
+# average investment for first 10 rounds pre-exposure
 avg_pct <- avg_ret_df %>% filter(roundNum < 11 , gameNum.f == "pre") %>% summarise(avg_inv = mean(inv_pct), avg_ret = mean(ret_pct_na, na.rm=TRUE))
@@
 The dotted lines identify the rounds where the pre-programmed one-off low investment occurs.
-We note lower average returns post vs pre in the Manipulation condition, whilst returns in the Control condition are similar between the two phases.",fig.align="center", fig.width=6, fig.height = 4}
+We note lower average returns post vs pre in the Manipulation condition, whilst returns in the Control condition are similar between the two phases.",fig.align="center", fig.width=5, fig.height = 3.3}
 
@@
   xintercept = c(12, 13),
-  Defection_round = factor(c("Pre-manipulation", "Post-manipulation"), levels = c("Pre-manipulation", "Post-manipulation"))
+  Low_inv_round = factor(c("Pre-phase", "Post-phase"), levels = c("Pre-phase", "Post-phase"))
 )
@@
                alpha = 0.3, linetype = 0) +
-  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
-  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
+  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Low_inv_round), color = "black", linewidth = 0.6) +
+  scale_linetype_manual(values = c("Pre-phase" = "dotted", "Post-phase" = "dashed"),name = "Low investment round") +
   labs(x = "Round",
@@
         legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10)) +
+        legend.title = element_text(size = 9),
+        legend.text = element_text(size = 8)) +
   scale_color_manual(values = c("darkblue", "darkgreen"),
-                     labels = c("Pre-manipulation", "Post-manipulation")) +
+                     labels = c("Pre-phase", "Post-phase")) +
   scale_fill_manual(values = c("darkblue", "darkgreen"),
@@
   guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
-         linetype = guide_legend(title = "Defection round")) +
+         linetype = guide_legend(title = "Low investment round")) +
   facet_wrap(~condition.f)
@@
 
-Mixed-effects analysis on the percentage returns shows a significant
-main effect of Phase (Pre vs. Post RTG game),
-`r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with
-higher percentage returns in the Pre RTG compared to the Post.
-Importantly, we also find an interaction between Condition and Phase
-(RTG pre- vs. post-manipulation),
-`r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`.
-As shown in Figure \@ref(fig:boxPlots), post-hoc tests confirm that,
-contrary to our expectations, there was a decrease in the percentage
-returned only in the Manipulation condition, pre - post,
-`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`,
-but no change in the Control condition. We find no interaction between
-Phase, Condition and RS, suggesting that there was no difference between
-RS groups for this interaction.
-
-There was also a significant main effect of Investment,
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`,
-such that higher investments were associated with higher percentage
-returns indicating positive reciprocity. An Investment by Condition
-interaction,
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`,
-reflected that returns were more affected by investments in the Control
-condition. We also find a three way interaction between Phase,
-Investment and RS,
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_inv_scaled_high_RS)`,
-showing that the differentiated effect of the investment on the
-proportion returned by RS group is itself moderated by the Phase (pre-
-vs post manipulation). Finally, we find a four-way interaction between
-Condition, Phase, Investment and RS
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.
+Participants returned higher percentages in the Pre phase compared to
+the Post phase
+(`r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`). This
+effect was moderated by Condition
+(`r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`):
+contrary to our expectations, participants in the Manipulation condition
+decreased their returns from pre to post
+(`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`),
+while those in the Control condition showed no change (Figure
+\@ref(fig:boxPlots)). RS did not moderate this Condition × Phase
+interaction.
+
+Higher investments elicited higher percentage returns, indicating
+positive reciprocity
+(`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`).
+This relationship was stronger in the Control condition than in the
+Manipulation condition
+(`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`).
+The effect of investment on returns varied by RS group and Phase
+(`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_inv_scaled_high_RS)`),
+and a four-way interaction indicated that these patterns further
+differed across Conditions
+(`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`).
+
+```{r invMod, cache=TRUE}
+
+mod_invs_RS <- mixed( investment ~ phase.f*condition.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
+
+summary(mod_invs_RS)
+```
+
+```{r}
+saveRDS(summary(mod_invs_RS), "data/mod_invs_RS.RDS")
+```
+
+```{r violPlotsDef, cache=TRUE}
+library(afex)
+library(ggplot2)
+library(ggsignif)
+
+# Define colors for Phase (different from event study's Condition colors)
+phase_cols <- c("pre" = "#E69F00", "post" = "#0072B2")
+
+# Generate the plot data from afex_plot
+ret_all_data <- afex_plot(mod_returns_pct_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
+                     mapping = c("linetype", "shape", "fill"),
+                     factor_levels = list(condition.f = c("Control", "Manipulation")),
+                     legend_title = "Phase",
+                     emmeans_arg = list(p.adjust.method = "tukey"))
+
+# Extract the data used in afex_plot
+plot_data_ret <- ret_all_data$data
+
+# Generate the bar chart for returns
+ret_all <- ggplot(plot_data_ret, aes(x = condition.f, y = y, fill = phase.f)) +
+  geom_col(position = position_dodge(width = 0.7), width = 0.6, color = "black", size = 0.3) +
+  geom_errorbar(aes(ymin = y - 1.96 * SE, ymax = y + 1.96 * SE),
+                position = position_dodge(width = 0.7), width = 0.2, size = 0.5) +
+  scale_fill_manual(values = phase_cols, labels = c("Pre-phase", "Post-phase")) +
+  theme_bw(base_size = 11) +
+  labs(y = "% Returns", x = "Condition", fill = "Trust Game Phase") +
+  theme(
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 11)
+  ) +
+  coord_cartesian(ylim = c(0, 0.6)) +
+  # Significance bracket for Manipulation pre vs post (moved higher to avoid overlap)
+  geom_segment(aes(x = 1.825, xend = 2.175, y = 0.54, yend = 0.54), inherit.aes = FALSE, size = 0.5) +
+  geom_segment(aes(x = 1.825, xend = 1.825, y = 0.54, yend = 0.52), inherit.aes = FALSE, size = 0.5) +
+  geom_segment(aes(x = 2.175, xend = 2.175, y = 0.54, yend = 0.52), inherit.aes = FALSE, size = 0.5) +
+  geom_text(aes(x = 2, y = 0.56, label = "**"), inherit.aes = FALSE, size = 5)
+
+######################################################################################################
+
+# Generate the plot data from afex_plot for investments
+inv_all_data <- afex_plot(mod_invs_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
+                     mapping = c("linetype", "shape", "fill"),
+                     factor_levels = list(condition.f = c("Control", "Manipulation")),
+                     legend_title = "Phase")
+
+# Extract the data used in afex_plot
+plot_data_inv <- inv_all_data$data
+
+# Generate the bar chart for investments
+inv_all <- ggplot(plot_data_inv, aes(x = condition.f, y = y, fill = phase.f)) +
+  geom_col(position = position_dodge(width = 0.7), width = 0.6, color = "black", size = 0.3) +
+  geom_errorbar(aes(ymin = y - 1.96 * SE, ymax = y + 1.96 * SE),
+                position = position_dodge(width = 0.7), width = 0.2, size = 0.5) +
+  scale_fill_manual(values = phase_cols, labels = c("Pre-phase", "Post-phase")) +
+  theme_bw(base_size = 11) +
+  labs(y = "HMM Investment", x = "Condition", fill = "Trust Game Phase") +
+  theme(
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 11)
+  ) +
+  coord_cartesian(ylim = c(0, 12))
+
+```
+
+```{r boxPlots, include=TRUE, fig.pos="H", fig.cap="Marginal means of percentage returns (top) and HMM investments (bottom) by Phase and Condition.
+Bars show estimated marginal means; error bars represent 95\\% confidence intervals.
+Participants in the Manipulation condition returned lower proportions post-exposure compared to pre-exposure (** p < .01), while Control participants showed no change.
+HMM investment did not differ across Phases or Conditions.", fig.align="center", fig.width=5, fig.height=4.5}
+
+library(patchwork)
+
+ret_all / inv_all + plot_layout(guides = 'collect') & theme(legend.position = "bottom", legend.title = element_text(size = 9), legend.text = element_text(size = 8))
+
+```
 
@@
 corrected_contrasts <- list(
-  "Diff in inv effect (Post - Pre, Manip, Low RS)" = c(0, 0, -1, 1, 0, 0, 0, 0),
-  "Diff in inv effect (Post - Pre, Manip, High RS)" = c(0, 0, 0, 0, 0, 0, -1, 1),
-  "Diff in (Post - Pre) between High and Low RS" = c(0, 0, 1, -1, 0, 0, -1, 1)
+  "ManipLowRS" = c(0, 0, -1, 1, 0, 0, 0, 0),
+  "ManipHighRS" = c(0, 0, 0, 0, 0, 0, -1, 1),
+  "HighVsLowRS" = c(0, 0, 1, -1, 0, 0, -1, 1)
 )
@@
 of how the effect of investment on returns changed from pre- to
-post-manipulation for different RS groups in both conditions. Starting
+post-exposure for different RS groups in both conditions (Figure \@ref(fig:fourwayInteractionPlot)). Starting
 with the Manipulation condition, for participants with low RS, the
 effect of investment on returns increased significantly from pre- to
-post-manipulation,
-`r papaja::apa_print(results_4w)$full_result$DiffininveffectPost_PreManipLowRS`.
+post-phase,
+`r papaja::apa_print(results_4w)$full_result$ManipLowRS`.
 This suggests that after the manipulation, low RS participants became
@@
 participants with high RS, the effect of investment on returns decreased
-significantly from pre- to post-manipulation,
-`r papaja::apa_print(results_4w)$full_result$DiffininveffectPost_PreManipHighRS`.
+significantly from pre- to post-exposure,
+`r papaja::apa_print(results_4w)$full_result$ManipHighRS`.
 This indicates that high RS participants became less responsive to their
@@
 changes between high and low RS groups was significant,
-`r papaja::apa_print(results_4w)$full_result$DiffinPost_PrebetweenHighandLowRS`.
+`r papaja::apa_print(results_4w)$full_result$HighVsLowRS`.
 This result suggests that the manipulation had significantly different
@@
 
-```{r fourwayInteractionPlot, include=F, fig.cap ="Visualising the four way interaction between Phase, RS group, Condition and Investment.
-The plot shows the effect of the investment on participant returns for each condition, phase and RS group.
-The interesting effect is in the Manipulation Condition: " }
+```{r fourwayInteractionPlot, include=TRUE, fig.pos="H", fig.width=5.5, fig.height=3.15, fig.cap ="Marginal effect of investment on percentage returns by Phase, Condition, and RS group.
+Bars show estimated slopes (change in returns per SD increase in investment) from the mixed model; error bars represent 95\\% confidence intervals.
+In the Manipulation condition, Low RS participants became more responsive to investments post-exposure (* p < .05), while High RS participants became less responsive (* p < .05).
+No significant changes were observed in the Control condition."}
 
@@
 
-# 
-# ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
-#   # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
-#   geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
-#   labs(title = "Effect of inv_scaled on Outcome by Condition",
-#        x = "inv_scaled", y = "Outcome") +
-#   theme_minimal() +
-#   facet_wrap(~high_RS)
-
-# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
-ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
-  # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
-  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
-  labs(x = "Investment Z-values", y = "Return proportion", color = "Experiment Phase") +
-  theme_minimal() +
-  facet_wrap(~condition.f*high_RS)+
+# Convert emtrends output to data frame for plotting
+slopes_df <- as.data.frame(summary(slopes))
+
+# Rename columns for clarity
+slopes_df <- slopes_df %>%
+  mutate(
+    Phase = factor(phase.f, levels = c("pre", "post"), labels = c("Pre", "Post")),
+    Condition = factor(condition.f, levels = c("Control", "Manipulation")),
+    RS_Group = factor(high_RS, levels = c("low RS", "high RS"), labels = c("Low RS", "High RS"))
+  )
+
+# Define colors matching Figure 4 (Phase colors)
+phase_cols <- c("Pre" = "#E69F00", "Post" = "#0072B2")
+
+# Create the bar chart
+ggplot(slopes_df, aes(x = Condition, y = inv_scaled.trend, fill = Phase)) +
+  geom_col(position = position_dodge(width = 0.7), width = 0.6, color = "black", size = 0.3) +
+  geom_errorbar(
+    aes(ymin = inv_scaled.trend - 1.96 * SE, ymax = inv_scaled.trend + 1.96 * SE),
+    position = position_dodge(width = 0.7),
+    width = 0.2,
+    size = 0.5
+  ) +
+  facet_wrap(~ RS_Group, ncol = 2) +
+  scale_fill_manual(values = phase_cols, labels = c("Pre-phase", "Post-phase")) +
+  labs(
+    x = "Condition",
+    y = "Effect of Investment on Returns\n(Slope ± 95% CI)",
+    fill = "Phase"
+  ) +
+  theme_bw(base_size = 11) +
+  theme(
+    legend.position = "bottom",
+    strip.text = element_text(size = 11, face = "bold"),
+    panel.grid.minor = element_blank(),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 11)
+  ) +
+  # Add significance bracket for Low RS panel (Manipulation: Pre vs Post)
+  geom_segment(
+    data = data.frame(RS_Group = factor("Low RS", levels = c("Low RS", "High RS"))),
+    aes(x = 1.825, xend = 2.175, y = 0.068, yend = 0.068),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_segment(
+    data = data.frame(RS_Group = factor("Low RS", levels = c("Low RS", "High RS"))),
+    aes(x = 1.825, xend = 1.825, y = 0.068, yend = 0.065),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_segment(
+    data = data.frame(RS_Group = factor("Low RS", levels = c("Low RS", "High RS"))),
+    aes(x = 2.175, xend = 2.175, y = 0.068, yend = 0.065),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_text(
+    data = data.frame(RS_Group = factor("Low RS", levels = c("Low RS", "High RS"))),
+    aes(x = 2, y = 0.072, label = "*"),
+    inherit.aes = FALSE, size = 6
+  ) +
+  # Add significance bracket for High RS panel (Manipulation: Pre vs Post)
+  geom_segment(
+    data = data.frame(RS_Group = factor("High RS", levels = c("Low RS", "High RS"))),
+    aes(x = 1.825, xend = 2.175, y = 0.058, yend = 0.058),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_segment(
+    data = data.frame(RS_Group = factor("High RS", levels = c("Low RS", "High RS"))),
+    aes(x = 1.825, xend = 1.825, y = 0.058, yend = 0.055),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_segment(
+    data = data.frame(RS_Group = factor("High RS", levels = c("Low RS", "High RS"))),
+    aes(x = 2.175, xend = 2.175, y = 0.058, yend = 0.055),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_text(
+    data = data.frame(RS_Group = factor("High RS", levels = c("Low RS", "High RS"))),
+    aes(x = 2, y = 0.062, label = "*"),
+    inherit.aes = FALSE, size = 6
+  )
+
+```
+
+### Returns Prior to Pre-Programmed Low Investment Trials
+
+```{r preDefectionData, include=FALSE}
+# Prepare data for rounds prior to pre-programmed low investment
+# Pre-phase: rounds 1-11 (low investment at round 12)
+# Post-phase: rounds 1-12 (low investment at round 13)
+predef_df <- avg_ret_df %>%
+  filter(
+    (gameNum.f == "pre" & roundNum < 12) |
+    (gameNum.f == "post" & roundNum < 13)
+  )
+```
+
+```{r preDefectionModel, include=FALSE, cache=TRUE}
+# Same model as main analysis but restricted to rounds prior to low investment
+mod_predef <- mixed(ret_pct_na ~ phase.f * condition.f * inv_scaled * high_RS + (1 + phase.f | playerId),
+                    data = predef_df, REML = TRUE, method = "KR")
+```
+
+```{r preDefectionContrasts, include=FALSE, cache=TRUE}
+# Pre vs Post within each Condition
+prepost_bycond_predef <- pairs(emmeans(mod_predef, "phase.f", by = "condition.f", pbkrtest.limit = 5000))
+```
+
+To distinguish between contrast effects and betrayal aversion as explanations for reduced cooperation in the Manipulation condition, we examined returns in the rounds preceding the pre-programmed low investment.
+If contrast effects were operating, participants in the Manipulation condition should already show reduced returns before experiencing any low investment in the post-exposure phase.
+Conversely, if betrayal aversion were the primary mechanism, group differences should only emerge after the low investment.
+
+The Phase $\times$ Condition interaction was significant in rounds prior to the low investment (`r papaja::apa_print(mod_predef)$full_result$phase_f_condition_f`).
+Participants in the Manipulation condition significantly decreased their returns from pre- to post-exposure phase even before encountering the low investment (`r papaja::apa_print(prepost_bycond_predef)$full_result$Manipulation_Pre_post`), whereas those in the Control condition showed no change (`r papaja::apa_print(prepost_bycond_predef)$full_result$Control_Pre_post`).
+The four-way interaction also remained significant (`r papaja::apa_print(mod_predef)$full_result$phase_f_condition_f_inv_scaled_high_RS`), suggesting that the differential responsiveness to investments observed in the full analysis was likewise present before the low investment occurred.
+
+<!-- Focusing on post defection only:  -->
+
+### Reaction to Pre-programmed Low Investment: Event Study Analysis
+
+To understand how participants reacted to and recovered from the pre-programmed low investment, an event study analysis was conducted centered on the low investment round (Figure \@ref(fig:eventStudyPlot)).
+Two behavioral responses were examined: the Drop (change in returns at the low investment round relative to $t-1$, where negative values indicate reduced returns reflecting punishment of low trust) and Recovery (change in returns at $t+1$ relative to the low investment round, where positive values indicate restored cooperation).
+
+```{r eventStudyData, include=FALSE}
+# Define defection rounds for each phase
+def_rounds <- data.frame(
+  gameNum.f = c("pre", "post"),
+  def_round = c(12, 13)
+)
+
+# Prepare data: Calculate relative round (t=0 is defection)
+event_data <- avg_ret_df %>%
+  filter(gameNum.f %in% c("pre", "post")) %>%
+  mutate(gameNum.f = as.character(gameNum.f)) %>% # Ensure matching types for join
+  left_join(def_rounds, by = "gameNum.f") %>%
+  mutate(
+    rel_round = roundNum - def_round,
+    # Create a factor for plotting order with nice labels
+    time_point = factor(rel_round, levels = -2:2, labels = c("t_minus_2", "t_minus_1", "Defection", "t_plus_1", "t_plus_2"))
+  ) %>%
+  filter(rel_round >= -2 & rel_round <= 2)
+```
+
+```{r eventStudyModel, include=FALSE, cache=TRUE}
+# Fit mixed model with RS moderation
+# We use a full interaction model to test if the trajectory (time_point) differs by Condition, Phase, and RS
+mod_event <- mixed(ret_pct_na ~ phase.f * condition.f * time_point * high_RS + (1 | playerId), 
+                   data = event_data, 
+                   REML = TRUE, 
+                   method = "KR")
+
+# Setup contrasts for "Drop" (t-1 to Defection) and "Recovery" (Defection to t+1)
+emm_event <- emmeans(mod_event, ~ time_point | phase.f * condition.f * high_RS)
+
+custom_contrasts <- list(
+  "Drop"     = c(0, -1, 1, 0, 0),    # Defection - t_minus_1 (negative when returns decrease)
+  "Recovery" = c(0, 0, -1, 1, 0)     # t_plus_1 - Defection (positive when returns recover)
+)
+
+# Calculate specific contrasts
+event_contrasts <- contrast(emm_event, custom_contrasts)
+
+# --- NEW: Calculate contrasts irrespective of RS (Condition x Phase) ---
+# We average over RS levels to get the main effect of Condition in each Phase
+emm_event_noRS <- emmeans(mod_event, ~ time_point | phase.f * condition.f)
+event_contrasts_noRS <- contrast(emm_event_noRS, custom_contrasts)
+cond_diffs_post_noRS <- contrast(event_contrasts_noRS, interaction = c("identity", "pairwise", "identity"), by = "phase.f")
+
+# Test differences between conditions in the Post phase (RS specific)
+cond_diffs_post <- contrast(event_contrasts, interaction = c("identity", "pairwise", "identity"), by = "phase.f")
+
+# Extract specific p-values and estimates for reporting
+# We need to grab these values to use in inline R
+# Helper function or just manual extraction
+get_p_val <- function(contrasts_obj, row_idx) {
+  s <- summary(contrasts_obj)
+  p <- s$p.value[row_idx]
+  if(p < .001) return("< .001")
+  return(paste("=", round(p, 3)))
+}
+```
+
+```{r eventStudyPlot, echo=FALSE, include=TRUE, fig.pos="H", fig.width=5.5, fig.height=3.15, fig.cap="Drop and Recovery responses to pre-programmed low investment in the post-exposure phase (pre-exposure phase not shown as no between-condition differences were observed).
+Drop = change in returns from low investment round minus t-1 (negative values indicate reduced returns); Recovery = change from t+1 minus low investment round (positive values indicate restored returns).
+Bars show estimated marginal means from the mixed model; error bars represent 95\\% confidence intervals.
+The bracket shows the significant between-condition difference in Recovery for High RS participants.
+** p < .01."}
+
+# Extract contrast estimates from the model for Post phase only
+contrast_summary <- summary(event_contrasts) %>%
+  as.data.frame() %>%
+  filter(phase.f == "post") %>%
+  mutate(
+    # Clean up names for plotting
+    Contrast = factor(contrast, levels = c("Drop", "Recovery")),
+    Condition = factor(condition.f, levels = c("Control", "Manipulation")),
+    RS_Group = factor(high_RS, levels = c("low RS", "high RS"))
+  )
+
+# Extract between-condition significance tests for Post phase
+cond_diffs_summary <- summary(cond_diffs_post) %>%
+  as.data.frame() %>%
+  filter(phase.f == "post") %>%
+  mutate(
+    sig_label = case_when(
+      p.value < 0.001 ~ "***",
+      p.value < 0.01 ~ "**",
+      p.value < 0.05 ~ "*",
+      TRUE ~ "n.s."
+    )
+  )
+
+# Define colors
+cols <- c("Control" = "#F8766D", "Manipulation" = "#00BFC4")
+
+# Create the bar chart
+p <- ggplot(contrast_summary, aes(x = Contrast, y = estimate, fill = Condition)) +
+  geom_col(position = position_dodge(width = 0.7), width = 0.6, color = "black", size = 0.3) +
+  geom_errorbar(
+    aes(ymin = estimate - 1.96 * SE, ymax = estimate + 1.96 * SE),
+    position = position_dodge(width = 0.7),
+    width = 0.2,
+    size = 0.5
+  ) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.5) +
+  facet_wrap(~ RS_Group, ncol = 2) +
+  scale_fill_manual(values = cols) +
+  labs(
+    x = "Behavioral Response",
+    y = "Change in Percentage Return",
+    fill = "Condition"
+  ) +
+  theme_bw(base_size = 11) +
+  theme(
+    legend.position = "bottom",
+    strip.text = element_text(size = 11, face = "bold"),
+    panel.grid.minor = element_blank(),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 11)
+  ) +
+  # Add significance bracket for Recovery in High RS panel
+  # Get the y-position for the bracket (above the higher bar)
+  geom_segment(
+    data = data.frame(RS_Group = factor("high RS", levels = c("low RS", "high RS"))),
+    aes(x = 1.825, xend = 2.175, y = 0.18, yend = 0.18),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_segment(
+    data = data.frame(RS_Group = factor("high RS", levels = c("low RS", "high RS"))),
+    aes(x = 1.825, xend = 1.825, y = 0.18, yend = 0.17),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_segment(
+    data = data.frame(RS_Group = factor("high RS", levels = c("low RS", "high RS"))),
+    aes(x = 2.175, xend = 2.175, y = 0.18, yend = 0.17),
+    inherit.aes = FALSE, size = 0.5
+  ) +
+  geom_text(
+    data = data.frame(RS_Group = factor("high RS", levels = c("low RS", "high RS"))),
+    aes(x = 2, y = 0.20, label = "**"),
+    inherit.aes = FALSE, size = 5
+  )
+
+print(p)
+```
+
+In the pre-exposure phase, there were no significant differences between conditions in either the Drop (`r papaja::apa_print(cond_diffs_post_noRS)$full_result$Drop_Control_Manipulation_Pre`) or Recovery (`r papaja::apa_print(cond_diffs_post_noRS)$full_result$Recovery_Control_Manipulation_Pre`), confirming that both groups started with equivalent behavioral patterns.
+In the post-exposure phase, a divergence emerged.
+The Control group showed a significantly larger Drop than the Manipulation group (`r papaja::apa_print(cond_diffs_post_noRS)$full_result$Drop_Control_Manipulation_Post`), indicating that participants exposed to the forgiving agent showed a blunted immediate reaction to the low investment and did not reduce their returns as sharply.
+The subsequent Recovery did not differ significantly between conditions overall (`r papaja::apa_print(cond_diffs_post_noRS)$full_result$Recovery_Control_Manipulation_Post`).
+
+When examining moderation by RS, the pattern appeared to be driven primarily by high RS participants.
+Low RS participants showed no significant difference in Recovery between conditions (`r papaja::apa_print(cond_diffs_post)$full_result$Recovery_Control_Manipulation_Low_RS_Post`), with both groups displaying modest, similar recovery patterns after the low investment.
+The larger Drop observed in the Control condition for this group was partly attributable to their elevated cooperation level at $t-1$.
+
+High RS participants showed a different pattern.
+In the Control condition, they demonstrated trust repair by significantly increasing their returns after the low investment (`r papaja::apa_print(event_contrasts)$full_result$Post_Control_High_RS_Recovery`).
+However, those in the Manipulation condition failed to recover, showing no significant increase in returns at $t+1$ (`r papaja::apa_print(event_contrasts)$full_result$Post_Manipulation_High_RS_Recovery`).
+The difference in Recovery between conditions was significant (`r papaja::apa_print(cond_diffs_post)$full_result$Recovery_Control_Manipulation_High_RS_Post`).
+
+In summary, the forgiveness intervention appeared to dampen reciprocal responsiveness, hindering the re-establishment of cooperation following a temporary withdrawal of trust.
+This effect was more pronounced among high RS individuals.
+While high RS participants in the Control condition demonstrated active reciprocity by reducing returns sharply when trust was withdrawn and increasing them when trust was restored, those exposed to the forgiving agent exhibited a disengaged pattern characterized by a muted reaction to the low investment and a failure to reinstate high returns afterward.
+
+<!-- Old Event Study line plot - replaced by bar chart above
+```{r eventStudy, eval=FALSE, include=FALSE, fig.cap="Event study of participant returns centered on the defection round (t=0).
+Comparison of Control and Manipulation conditions across Pre and Post phases."}
+# Define defection rounds for each phase
+def_rounds <- data.frame(
+  gameNum.f = c("pre", "post"),
+  def_round = c(12, 13)
+)
+
+# Prepare data: Calculate relative round (t=0 is defection)
+event_data <- avg_ret_df %>%
+  filter(gameNum.f %in% c("pre", "post")) %>%
+  mutate(gameNum.f = as.character(gameNum.f)) %>% # Ensure matching types for join
+  left_join(def_rounds, by = "gameNum.f") %>%
+  mutate(
+    rel_round = roundNum - def_round,
+    # Create a factor for plotting order with nice labels
+    time_point = factor(rel_round, levels = -2:2, labels = c("t-2", "t-1", "Betrayal", "t+1", "t+2"))
+  ) %>%
+  filter(rel_round >= -2 & rel_round <= 2) # Filter for the window of interest
+
+# Plot the event study
+ggplot(event_data, aes(x = time_point, y = ret_pct_na, color = condition.f, group = condition.f)) +
+  stat_summary(fun = mean, geom = "point", size = 2) +
+  stat_summary(fun = mean, geom = "line", size = 1) +
+  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
+  facet_wrap(~phase.f) +
+  labs(
+    title = "Reaction to Defection (Event Study)",
+    x = "Trial relative to Defection",
+    y = "Percentage Return",
+    color = "Condition"
+  ) +
+  theme_bw() +
   theme(legend.position = "bottom")
-
-```
-
-```{r invMod, cache=TRUE}
-
-mod_invs_RS <- mixed( investment ~ phase.f*condition.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
-
-summary(mod_invs_RS)
-```
-
-```{r}
-saveRDS(summary(mod_invs_RS), "data/mod_invs_RS.RDS")
-```
-
-```{r violPlotsDef, cache=TRUE}
-library(afex)
-library(ggplot2)
-library(ggsignif)
-
-# Generate the plot with points and error bars
-ret_all <- afex_plot(mod_returns_pct_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
-                     mapping = c("linetype", "shape", "fill"),
-                     factor_levels = list(condition.f = c("Control", "Manipulation")),
-                     legend_title = "Phase",
-                     emmeans_arg = list(p.adjust.method = "tukey"))
-
-# Extract the data used in afex_plot
-plot_data <- ret_all$data
-
-# Generate the plot
-ret_all <- ggplot(plot_data, aes(x = condition.f, y = y, color = phase.f, group = phase.f)) +
-  geom_point(position = position_dodge(width = 0.8)) +
-  geom_errorbar(aes(ymin = y - SE, ymax = y + SE), width = 0.2, position = position_dodge(width = 0.8)) +
-  theme_bw() +
-  labs(y = "% Returns", x = "Condition", color = "Trust Game Phase") +
-  theme(plot.title = element_text(size = 16), legend.position = "bottom") +
-  geom_signif(
-              annotations = "***", 
-              xmin = 1.8, xmax = 2.2,
-              y_position = 0.47, 
-              tip_length = 0.03, 
-              textsize = 3, 
-              color = "black")
-
-######################################################################################################
-
-# Generate the plot with points and error bars
-inv_all <- afex_plot(mod_invs_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
-                     mapping = c("linetype", "shape", "fill"),
-                     factor_levels = list(condition.f = c("Control", "Manipulation")),
-                     legend_title = "Phase")
-
-# Extract the data used in afex_plot
-plot_data <- inv_all$data
-
-# Generate the plot
-inv_all <- ggplot(plot_data, aes(x = condition.f, y = y, color = phase.f, group = phase.f)) +
-  geom_point(position = position_dodge(width = 0.8)) +
-  geom_errorbar(aes(ymin = y - SE, ymax = y + SE), width = 0.2, position = position_dodge(width = 0.8)) +
-  theme_bw() +
-  labs(y = "HMM Investment", x = "Condition", color = "Trust Game Phase") +
-  theme(plot.title = element_text(size = 16), legend.position = "bottom")+
-  ylim(9, 11)
-
-```
-
-```{r boxPlots, include=T,fig.cap="Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition.
-The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns.
-The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.
-RS did not affect returns and investments by Phase and Condition.",fig.align="center", fig.width=6, fig.height = 8}
-
-library(patchwork)
-
-ret_all/inv_all +  plot_layout(guides = 'collect') & theme(legend.position = "bottom")
-
-```
-
-<!-- Focusing on post defection only:  -->
-
+```
+-->
+
+<!-- 
 ### Post Defection Trials
@@
 
-Did participants learn to be more forgiving and cooperative after
-witnessing the pre-programmed defection by the HMM investor? To explore
-this question, we restrict the analysis to the trials following the
-pre-programmed defection by the HMM agent in both the "pre" (trials 12
-to 15) and the "post" phases (trials 13 to 15). We fit the same mixed
-effects model as for all the trials with the exception of the RS
-variable. This is because RS did not show main effects in the previous
-model, and also due to the necessity of running a simpler model to
-accommodate the low number of trials. We find a significant main effect
-of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f`
-with returns lower in the second game post defection trials compared to
-the first. We also find a main effect of Investment
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_retPostDef_RS)$full_result$inv_scaled)`
-where participants continued to return higher proportions when receiving
-higher investments. Finally, we still find an Investment by Condition
-interaction
-`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`
-showing a lower effects of investment on the Manipulation condition
-compared to the Control condition in post-defection trials. However, the
-absence of a Condition by Phase effect indicates there was no difference
-between conditions on participants' reaction to a one-off low investment
-by the co-player.
+<!-- Did participants learn to be more forgiving and cooperative after -->
+<!-- witnessing the pre-programmed defection by the HMM investor? To explore -->
+<!-- this question, we restrict the analysis to the trials following the -->
+<!-- pre-programmed defection by the HMM agent in both the "pre" (trials 12 -->
+<!-- to 15) and the "post" phases (trials 13 to 15). We fit the same mixed -->
+<!-- effects model as for all the trials with the exception of the RS -->
+<!-- variable. This is because RS did not show main effects in the previous -->
+<!-- model, and also due to the necessity of running a simpler model to -->
+<!-- accommodate the low number of trials. We find a significant main effect -->
+<!-- of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` -->
+<!-- with returns lower in the second game post defection trials compared to -->
+<!-- the first. We also find a main effect of Investment -->
+<!-- `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_retPostDef_RS)$full_result$inv_scaled)` -->
+<!-- where participants continued to return higher proportions when receiving -->
+<!-- higher investments. Finally, we still find an Investment by Condition -->
+<!-- interaction -->
+<!-- `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` -->
+<!-- showing a lower effects of investment on the Manipulation condition -->
+<!-- compared to the Control condition in post-defection trials. However, the -->
+<!-- absence of a Condition by Phase effect indicates there was no difference -->
+<!-- between conditions on participants' reaction to a one-off low investment -->
+<!-- by the co-player. -->
+<!-- -->
 
@@
   # geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
-  # scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
+  # scale_linetype_manual(values = c("Pre-phase" = "dotted", "Post-phase" = "dashed"),name = "Defection round") +
   labs(x = "Round",
@@
 
-comp1_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp1_coop_interaction <- contrast(emm_coop_interaction, list("PrePost" = contrast_matrix[1,]))
 comp2_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
-# comp3_coop_interaction <- contrast(emm_coop_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Extract summary for easier inline reporting
+coop_summary <- summary(comp1_coop_interaction)
+coop_results <- list(
+  ControlHighRS = coop_summary[coop_summary$condition.f == "Control" & coop_summary$high_RS == "high RS", ],
+  ControlLowRS = coop_summary[coop_summary$condition.f == "Control" & coop_summary$high_RS == "low RS", ],
+  ManipHighRS = coop_summary[coop_summary$condition.f == "Manipulation" & coop_summary$high_RS == "high RS", ],
+  ManipLowRS = coop_summary[coop_summary$condition.f == "Manipulation" & coop_summary$high_RS == "low RS", ]
+)
 
@@
 comp1_coop_interaction
-print("Pre vs Expo")
-test(comp2_coop_interaction, adjust="bonferroni")
-test(comp2_coop_interaction)
-
-# test_comp3_coop_interaction
 ```
@@
 
-comp1_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp1_forg_interaction <- contrast(emm_forg_interaction, list("PrePost" = contrast_matrix[1,]))
 comp2_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
-# comp3_forg_interaction <- contrast(emm_forg_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Extract summary for easier inline reporting
+forg_summary <- summary(comp1_forg_interaction)
+forg_results <- list(
+  ControlHighRS = forg_summary[forg_summary$condition.f == "Control" & forg_summary$high_RS == "high RS", ],
+  ControlLowRS = forg_summary[forg_summary$condition.f == "Control" & forg_summary$high_RS == "low RS", ],
+  ManipHighRS = forg_summary[forg_summary$condition.f == "Manipulation" & forg_summary$high_RS == "high RS", ],
+  ManipLowRS = forg_summary[forg_summary$condition.f == "Manipulation" & forg_summary$high_RS == "low RS", ]
+)
 
@@
 comp1_forg_interaction
-print("Pre vs Expo")
-comp2_forg_interaction
-
-# test_comp3_forg_interaction
 ```
@@
 
-comp1_again_interaction <- contrast(emm_again_interaction, list("Pre vs Post" = contrast_matrix[1,]))
+comp1_again_interaction <- contrast(emm_again_interaction, list("PrePost" = contrast_matrix[1,]))
 comp2_again_interaction <- contrast(emm_again_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
-# comp3_again_interaction <- contrast(emm_again_interaction, list("Post vs Expo" = contrast_matrix[3,]))
+
+# Extract summary for easier inline reporting
+again_summary <- summary(comp1_again_interaction)
+again_results <- list(
+  ControlHighRS = again_summary[again_summary$condition.f == "Control" & again_summary$high_RS == "high RS", ],
+  ControlLowRS = again_summary[again_summary$condition.f == "Control" & again_summary$high_RS == "low RS", ],
+  ManipHighRS = again_summary[again_summary$condition.f == "Manipulation" & again_summary$high_RS == "high RS", ],
+  ManipLowRS = again_summary[again_summary$condition.f == "Manipulation" & again_summary$high_RS == "low RS", ]
+)
 
@@
 comp1_again_interaction
-print("Pre vs Expo")
-comp2_again_interaction
-
-# test_comp3_again_interaction
 ```
@@
 
-Figure \@ref(fig:plotRatings) shows participants' ratings of each player
-they faced by condition and RS group. We will focus on two contrasts to
-analyse the ratings by Condition and RS group. The first is between the
-rating in the first phase ("pre") when participants phase the human-like
-HMM, and the average rating during the exposure phase (average of
-"expo1", "expo2" and "expo3") where they either face the forgiving HMM
-(Manipulation condition) or the human-like HMM again (Control
-condition). The second contrast is between the "pre" and "post" phases
-of the experiment where in both conditions participants face the same
-human-like HMM.
-
-### Comparing Pre and Exposure Ratings
-
-For those with high RS, participants in the Manipulation condition rated
-the investors they faced in the exposure phase (the forgiving HMM) as
-more Cooperative
-`r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`.
-There was no difference in ratings on forgiveness and whether they would
-like to face the co-players again. Those in the Control condition rated
-the investors faced in the exposure group (same HMM) as less cooperative
-`r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`,
-less forgiving
-`r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`,
-and were less keen on facing them again
-`r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.
-
-For those with low RS, there was no difference in any of the ratings
-between the "pre" and "exposure" phases for the Manipulation condition.
-For the Control condition, participants indicated less willingness to
-face the exposure co-player compared to the "pre" player,
-`r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo`
-but also did not differ on cooperation and forgiveness ratings. In
-summary, those with low RS had a mostly undifferentiated perception of
-players between the pre and exposure phases, even when the co-player was
-in fact designed to be more forgiving. In contrast, we see that exposing
-participants with high RS to a more cooperative and more forgiving agent
-has compensated for the decrease in ratings that would have occurred if
-faced with a co-player with the exact same strategy.
-
-```{r plotRatings,include=T, echo=FALSE, fig.pos='H', fig.cap="Averages and standard errors of the participants ratings of the opponent (y-axis) by each game and condition for each phase (x-axis).
-Pre and post are the 15 round RTGs before and after the exposure phase respectively.
-The games titled expo 1 to 3 are the three 7 round games during the exposure phase.
-The blue line represents participants' perception of co-player cooperation, the red line indicates perceived co-player forgiveness, and the green line shows the participants' willingness to play again with the same co-player.
-Cooperation, forgiveness, and willingness to play again ratings remained relatively stable for the low RS group except post-manipulation, where they were lower.
-In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the Manipulation condition, but decreased over time in the Control group.
-Ratings were also markedly lower in the post-manipulation game for this group.",fig.align="center", fig.width=6, fig.height = 6}
+Figure \@ref(fig:plotRatings) shows participants' ratings of co-players across phases.
+We examined two contrasts: pre-exposure versus exposure phase ratings, and pre-exposure versus post-exposure ratings.
+
+High RS participants showed more differentiated perceptions of the agents.
+In the Manipulation condition, they rated the forgiving agents as more cooperative during exposure (`r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_High_RS_Pre_vs_Expo`).
+In the Control condition, however, high RS participants rated the same human-like HMM progressively more negatively—lower on cooperation (`r papaja::apa_print(comp2_coop_interaction)$full_result$Control_High_RS_Pre_vs_Expo`), forgiveness (`r papaja::apa_print(comp2_forg_interaction)$full_result$Control_High_RS_Pre_vs_Expo`), and willingness to play again (`r papaja::apa_print(comp2_again_interaction)$full_result$Control_High_RS_Pre_vs_Expo`)—despite the agent's strategy remaining unchanged.
+Low RS participants showed largely undifferentiated perceptions between pre and exposure phases regardless of condition.
+
+Comparing pre to post-exposure ratings revealed a contrast effect: after experiencing the forgiving agent, both RS groups in the Manipulation condition rated the post-exposure agent (identical to pre-exposure) more negatively on forgiveness (High RS: `r format_emmeans_apa(forg_results$ManipHighRS)`; Low RS: `r format_emmeans_apa(forg_results$ManipLowRS)`) and willingness to play again (High RS: `r format_emmeans_apa(again_results$ManipHighRS)`; Low RS: `r format_emmeans_apa(again_results$ManipLowRS)`).
+Low RS participants in the Control condition showed stable ratings, accurately perceiving the consistent agent strategy, while high RS participants in the Control condition continued their negative drift (Cooperation: `r format_emmeans_apa(coop_results$ControlHighRS)`).
+These rating patterns converge with the behavioral findings, suggesting high RS individuals are particularly sensitive to relative comparisons between interaction partners.
+
+```{r plotRatings,include=T, echo=FALSE, fig.pos='H', fig.cap="Participants' ratings of co-players by phase, condition, and RS group.
+Blue: perceived cooperation; red: perceived forgiveness; green: willingness to play again.
+Low RS participants showed stable ratings in the Control condition.
+High RS participants showed declining ratings in the Control condition despite unchanged agent strategy, and more differentiated perceptions in the Manipulation condition.", fig.align="center", fig.width=5, fig.height = 5}
 
@@
 ```{r}
-# temporarily paste code from Figure 2 to compare styles: 
-# Plot
-ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
-  stat_summary(fun = mean, geom = "line") +
-  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
-               alpha = 0.3, linetype = 0) +
-  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
-  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
-  labs(x = "Round",
-       y = "Percentage Return",
-       color = "Trust Game Phase") +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        legend.box = "vertical", # This will stack the legends vertically
-        legend.key = element_blank(),
-        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
-        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10)) +
-  scale_color_manual(values = c("darkblue", "darkgreen"),
-                     labels = c("Pre-manipulation", "Post-manipulation")) +
-  scale_fill_manual(values = c("darkblue", "darkgreen"),
-                    guide = "none") + # Hide the fill legend
-  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
-  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
-         linetype = guide_legend(title = "Defection round")) +
-  facet_wrap(~condition.f)
-
-```
-
-```{r}
-library(ggplot2)
-
-ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f))) +
-  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem, color = condition.f), 
-                width = .1, position = position_dodge(width = 0.25)) +
-  geom_line(aes(color = condition.f), linetype = "dashed", position = position_dodge(width = 0.25)) +
-  geom_point(aes(color = condition.f), position = position_dodge(width = 0.25), size = 3) +
-  facet_grid(high_RS ~ rating_type, scales = "free_x") +
-  scale_color_manual(values = c("Control" = "blue", "Manipulation" = "red")) +
-  labs(x = "Phase", y = "Mean Rating", color = "Condition") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-```
-
-### Comparing Pre and Post Ratings
-
-<!-- Comparing post-exposure to exposure ratings, participants in the Manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
-
-<!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`.
-Ratings for those in the Control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.
--->
-
-Participants high on RS in the Manipulation condition rated the
-co-players in the "post" phase similarly on cooperation, lower on
-forgiveness
-`r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
-and lower on willingness to face them again
-`r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`.
-Those in the Control condition rated the investors post the exposure
-phase lower on all three attributes (Cooperation:
-`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`,
-Forgiveness:
-`r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`,
-Play again:
-`r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).
-
-For participants low on RS, those in the Manipulation condition rated
-the co-players in the "post" phase lower on all three attributes
-(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`,
-Forgiveness:
-`r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`,
-Play again:
-`r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`).
-Those in the Control condition did not differ in their ratings of the
-"pre" and "post" phase players.
-
-In summary, we again see that those with low RS accurately perceive the
-co-player as similar on all attributes throughout the phases in the
-Control condition. In contrast, the high RS group shows a negative bias
-towards the co-players after the "pre" phase in the Control condition
-even though the player continues to use the same strategy. After
-exposure to the forgiving HMM, both groups rate the "post" co-player
-worse than the "pre" even though they are the same.
-
-```{r}
 d_finished %>% 
@@
 trust-building scenarios. Participants were then exposed to either
-forgiving or human-like HMM agents before playing another RTG. We found
+forgiving HMM agents (which, by design, were also more cooperative due to their inability to remain in a low-trust state) or standard human-like HMM agents before playing another RTG. We found
 that RS did not moderate participants' returns as trustees in the
@@
 and post phase of the experiment. Why did participants reduce their
-returns even though they were repeatedly exposed to more cooperative and
-more forgiving artificial agents? A look at how the participants rated
+returns even though they were repeatedly exposed to agents designed to be more forgiving? A look at how the participants rated
 their co-players might shed some light on what might be driving this
@@
 comparison [@kobre_negative_1972]. As the most recently faced opponents
-were more forgiving and cooperative, this negative contrast effect may
+were more forgiving (and consequently more cooperative), this negative contrast effect may
 have trumped any learning transfer from being repeatedly exposed to
@@
 
-This effect appears to operate independently of RS, as both groups
-exhibited it. However, RS did influence how participants perceived and
-reacted to their co-players' behavior. High RS individuals demonstrated
-a heightened attunement to changes in their co-players' behavior,
-accurately perceiving increased cooperativeness in the forgiving agent.
-Paradoxically, this same sensitivity may explain their decreased
-reactivity to investment levels and declining ratings of similar
-co-players over time in the Control condition. Occasional pre-programmed
-defections may have amplified their perceptions of untrustworthiness,
-leading to a more cautious approach and less responsiveness to
-investment variations. This aligns with the heightened social awareness
-and vigilance often associated with high RS [@downey_implications_1996].
-This finding suggests that individuals with high RS might benefit from
-interventions that focus on calibrating their perceptions of social cues
-in economic exchanges. In contrast, low RS individuals were less
-affected by the agent's behavioral variations, showing a more stable
-baseline of social perception. This stability in perception might
-reflect a lower propensity to scrutinize social cues or a more robust
-internal model of social interactions that is less easily perturbed by
-short-term variations in partner behavior. Their increased
-responsiveness to investments post-exposure in the Manipulation
-condition suggests they may have internalized the cooperative norms
-experienced, carrying over this increased reciprocity to subsequent
-interactions.
-
-Despite these differences in perception and investment reactivity, the
-actual cooperative behavior – as measured by returns in the trust game –
-did not significantly differ between RS groups. This suggests a
-potential dissociation between the cognitive-perceptual processes
-involved in evaluating social partners and the behavioral output in
-economic games. There are multiple potential explanations for this:
-First, the ratings may reflect explicit, conscious evaluations of social
-partners, while behavior in the trust game might be guided more by
-implicit, automatic processes that are less differentiated between RS
-groups. This interpretation aligns with dual-process theories of social
-cognition [@lieberman_social_2007]. Second, the structured nature of the
-trust game might provide a context where the behavioral expression of RS
-is attenuated. In contrast, the more open-ended nature of providing
-ratings might allow for greater expression of RS-related perceptual
-biases.
-
-These findings have important implications for interventions aimed at
-promoting trust and cooperation. Simply exposing individuals to
-artificially highly cooperative or forgiving partners may not suffice to
-induce lasting changes in behavior, and may even lead to negative
-contrast effects. Instead, interventions should focus on developing
-adaptive strategies for navigating the complexities of real-world social
-interactions, which often involve a mix of cooperative and
-non-cooperative behaviors. Such approaches could gradually shape
-expectations and behaviors over time, taking into account individual
-differences in RS. For individuals with high RS, interventions might
-specifically target the calibration of social perceptions in economic
-exchanges, helping to bridge the gap between their heightened
-sensitivity to social cues and their actual cooperative behavior. Future
-research should investigate whether the observed dissociation between
-perception and behavior persists in more naturalistic social
-interactions or over longer periods. Additionally, exploring the
-temporal dynamics of these effects and the role of explicit
-communication in the cooperative process could offer valuable insights.
-Interventions combining realistic cooperative scenarios with clear
-communication about mutual expectations might prove more effective in
-fostering enduring cooperative tendencies that are robust across various
-social contexts [@balliet_reward_2011].
+An alternative explanation relates to differences in exposure-phase experiences: the Control agent performed pre-programmed low investments (round 5 of each game), whereas the Manipulation agent did not.
+Consequently, Manipulation participants experienced 21 consecutive rounds without any low investments before encountering the post-phase low investment—a substantially longer period of consistently positive interactions than Control participants.
+This raises the possibility that reduced cooperation in the Manipulation condition reflects heightened betrayal aversion rather than negative contrast effects.
+
+However, the analysis of returns prior to the pre-programmed low investment provides evidence against betrayal aversion as the primary mechanism.
+If betrayal aversion were driving the effect, group differences should only emerge after participants encountered the low investment in the post-exposure phase.
+Instead, Manipulation participants had already significantly reduced their returns in rounds 1-12 of the post-exposure phase—before any low investment occurred.
+This pattern is consistent with contrast effects operating from the beginning of the post-exposure phase, as participants immediately perceived the human-like agent as less cooperative compared to the forgiving agent they had just experienced.
+While betrayal aversion may contribute to specific aspects of the observed patterns, such as the impaired recovery following the low investment in high RS participants, it cannot account for the overall reduction in cooperation that was already evident before any low investment.
+
+While the negative contrast effect operated across RS groups, the pattern of responses to trust violations differed in ways that align with clinical models of rejection sensitivity.
+In the event study analysis, high and low RS participants showed comparable immediate reactions to the low investment (the Drop), indicating intact detection of trust violations regardless of RS level.
+However, the groups diverged in their subsequent recovery patterns: high RS participants in the Manipulation condition failed to restore cooperation following the low investment, whereas low RS participants and Control participants showed recovery.
+This dissociation between intact rejection detection and impaired relationship repair is consistent with research on social learning difficulties in individuals with elevated RS and related clinical presentations.
+Studies of borderline personality disorder, where RS is characteristically elevated, have documented specific deficits in updating social expectations following positive interpersonal experiences [@staebler_facial_2011; @schuster_ambiguous_2021].
+Similarly, research on depression has identified "cognitive immunization" processes whereby negative schemas resist modification despite contradictory evidence [@kube_distorted_2020].
+The high RS participants' failure to recover cooperation, despite prior exposure to consistently forgiving behavior, may reflect analogous difficulties in leveraging positive social experiences to update expectations and restore trust.
+
+The four-way interaction findings further support this interpretation.
+High RS participants showed decreased responsiveness to their co-player's investments following the forgiveness manipulation, a pattern suggestive of withdrawal from contingent social exchange.
+This reduced sensitivity to partner behavior parallels the self-silencing and social withdrawal documented in high RS populations, where anticipatory self-protection can paradoxically undermine relationship maintenance [@ayduk_rejection_2000; @romero-canyas_rejection_2010].
+In contrast, low RS participants showed increased responsiveness to investments post-phase, suggesting they internalized the cooperative norms experienced during exposure and carried this forward to subsequent interactions.
+This differential capacity to benefit from positive social experiences maps onto broader findings that RS impedes the acquisition and transfer of adaptive interpersonal strategies [@pietrzak_appearance-rejection_2005].
+
+The combination of blunted responsiveness to investments and absent recovery in high RS participants suggests a pattern of passive disengagement rather than active retaliation.
+While overall return levels did not differ between RS groups, these specific behavioral signatures indicate that RS does modulate particular aspects of cooperative behavior.
+The ratings data complement these findings: high RS participants showed more negative explicit evaluations of their co-players, rating them lower on forgiveness and willingness to interact again.
+This convergence between explicit ratings and behavioral patterns suggests that the effects of RS on social exchange are expressed across multiple response systems.
+The structured nature of the trust game may constrain RS effects to specific behavioral signatures (such as contingent responding and recovery) rather than overall cooperation levels, while the more open-ended nature of rating tasks allows for broader expression of RS-related evaluative biases [@lieberman_social_2007].
+
+These findings have implications for interventions aimed at promoting trust and cooperation.
+The present results suggest that exposure to positive social models alone may be insufficient for high RS individuals, and may even produce iatrogenic effects through negative contrast.
+The specific deficits observed—impaired recovery from trust violations and reduced sensitivity to partner behavior—point to potential intervention targets.
+Approaches that focus on enhancing the capacity to update expectations following interpersonal ruptures may be more effective than simply providing positive experiences.
+This could include explicit training in recognizing repair attempts, practicing graduated trust restoration, or developing metacognitive awareness of the tendency toward disengagement following perceived rejection.
+Future research should examine whether these behavioral patterns generalize to naturalistic social contexts and whether targeted interventions can modify the updating and recovery deficits observed in high RS participants [@balliet_reward_2011].
 
@@
 While this study offers valuable insights into trust and cooperation
-dynamics in social interactions mediated by artificial agents, several
-limitations warrant consideration. The artificial nature of the
-interactions, while enabling high experimental control, may limit
-ecological validity, particularly regarding the paradigm's ability to
-trigger rejection sensitivity in participants. Additionally, the
-relatively brief exposure to forgiving HMM agents during the
-manipulation phase may have been insufficient to induce lasting changes
-in participants' reciprocity behavior. The online format of the
-experiment eliminates important social cues present in face-to-face
-interactions, which might influence trust and cooperation differently.
-Although the RSQ is a well-validated measure of rejection sensitivity
-(RS), future studies could benefit from a multi-method approach,
-potentially incorporating behavioral measures or additional self-report
-scales for a more comprehensive RS assessment. Furthermore, while our
-use of HMM agents to simulate human-like behavior in economic games is
-innovative, some participants may have suspected they were interacting
-with AI, potentially confounding the results. Despite these limitations,
-our findings provide a foundation for future research. Subsequent
-studies could address these constraints by incorporating face-to-face
-interactions, utilizing multiple measures of RS, examining the long-term
-effects of exposure to forgiving interaction partners, and developing
-more convincing cover stories or explicitly comparing responses to human
-and computerised partners.
+dynamics, several limitations warrant consideration. First, and most importantly, the Control and Manipulation conditions
+differed not only in the forgiveness of exposure-phase agents but also
+in whether participants experienced pre-programmed low investments
+during exposure (Control agents sent low investments at round 5 of each exposure game; Manipulation agents did not).
+This design feature means we cannot fully disentangle the effects of agent forgiveness per se from the effects of experiencing an uninterrupted sequence of positive interactions.
+Although the analysis of pre-low investment returns provides evidence favoring contrast effects over betrayal aversion as the primary mechanism, and both conditions ultimately faced the same human-like agent in the post-exposure phase, caution is warranted in attributing the observed effects specifically to "forgiveness exposure." Future replications
+should equate these experiences across conditions to isolate the effects
+of agent forgiveness from the absence of negative experiences. Second, our extreme groups design for RS (selecting
+participants with RSQ scores > 15 or < 10) maximized power to detect
+moderation effects but may inflate effect sizes and limits
+generalizability to individuals with moderate RS. Future research should
+examine RS as a continuous variable. Third, the brief exposure phase
+(three 7-round games) may have been insufficient to induce lasting
+changes. Fourth, the online format eliminates social cues present in
+face-to-face interactions. Notably, while 41% of participants believed
+they faced humans and a similar proportion were unsure, the observed
+effects emerged even among those who suspected AI, suggesting robustness
+of the findings. Despite these limitations, subsequent studies could
+address these constraints by incorporating face-to-face interactions,
+longer exposure periods, and continuous RS measurement.
 
@@
 economic structures. The brief exposure phase and pre-programmed
-defections are specific to our design and may not reflect real-world
+low investments are specific to our design and may not reflect real-world
 trust-building scenarios. The online context may not capture all aspects
@@
 
-Preregistration: The hypotheses and methods were not preregistered.The
-analysis plan was not preregistered. Materials: All study materials are
+Preregistration: The hypotheses and methods were not preregistered.
+The primary hypothesis, based on attachment theory, predicted that forgiveness exposure would increase cooperation.
+The alternative predictions regarding contrast effects and RS-specific updating biases were incorporated into the theoretical framework following initial data analysis, though both perspectives were grounded in existing literature.
+The analysis plan was not preregistered.
+Materials: All study materials are
 publicly available (<https://github.com/ismailg/exposure-public>). Data:

### 2025-12-05 article.tex
Revise manuscript: add event study analysis, strengthen limitations, update terminology
- Add event study and pre-low-investment analyses to test contrast vs betrayal aversion
- Acknowledge confound between forgiveness and low investment exposure in Limitations
- Change "defection" to "low investment" terminology throughout
- Reframe RS findings: intact detection but impaired recovery in high RS
- Condense Player Ratings section; revise abstract and discussion

@@
 \PassOptionsToPackage{hyphens}{url}
-%
 \documentclass[
 ]{article}
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
 \usepackage{amsmath,amssymb}
+\setcounter{secnumdepth}{5}
 \usepackage{iftex}
@@
 \makeatother
-\usepackage{xcolor}
-\usepackage[margin=1in]{geometry}
 \usepackage{longtable,booktabs,array}
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
 \makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
 % definitions for citeproc citations
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \usepackage{lscape}
@@
 \usepackage{xcolor}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
@@
 
-Can exposure to forgiving behavior improve cooperative outcomes, and
-does Rejection Sensitivity (RS) moderate this effect? This randomized
-controlled experiment explores these questions using a novel approach
-with Hidden Markov Model (HMM)-based artificial agents in repeated trust
-games (RTGs). Participants (N = 206), pre-screened for high or low RS,
-interacted with either consistently behaving or more forgiving
-co-players, which were in reality HMM agents. Unexpectedly, exposure to
-forgiving agents led to reduced cooperation in subsequent interactions.
-Control group participants maintained consistent behavior. RS levels
-influenced perceptions of the agent's cooperativeness but did not
-significantly affect behavioral outcomes, revealing a
-perception-behavior dissociation. These findings challenge assumptions
-about fostering cooperation through simple exposure to positive
-interactions and highlight the importance of accounting for negative
-contrast effects when designing such interventions. They also
-demonstrate the potential of HMM-based artificial agents for studying
-interactive social dynamics, offering a methodological advancement for
-future research. Our results suggest the need for nuanced interventions
-that gradually shape expectations and behaviors to foster cooperative
-outcomes, particularly addressing the discrepancy between perceptions
-and behavior in individuals with high rejection sensitivity.
+Can exposure to forgiving partners improve interpersonal cooperation?
+Attachment theory suggests positive relational experiences can correct negative internal working models, but individuals high in Rejection Sensitivity (RS)---characterized by anxious expectations of rejection---may struggle to benefit from such experiences.
+We tested this using a randomized experiment (N = 206) in which participants played repeated trust games with HMM-based artificial agents that simulate human-like trust dynamics.
+After a baseline game, participants were exposed to either forgiving agents (Manipulation) or standard agents (Control), then played a final game with a standard agent.
+Contrary to predictions, forgiveness exposure \emph{reduced} subsequent cooperation---participants appeared to perceive the standard post-exposure agent as less cooperative by comparison (a negative contrast effect).
+Critically, RS moderated specific behavioral patterns but not overall cooperation levels: high RS participants failed to recover cooperation after trust violations and became \emph{less} responsive to partner behavior following exposure, whereas low RS participants showed normal recovery and became \emph{more} responsive.
+These findings suggest that positive relational experiences do not universally promote cooperation, and that high RS individuals may require interventions targeting their capacity to update expectations rather than simply providing positive experiences.
 
@@
 
-This study reveals that exposing individuals to forgiving behavior in
-economic games can unexpectedly decrease their subsequent cooperation.
-Interestingly, while people with high rejection sensitivity more
-accurately perceived their partners' cooperativeness, this didn't
-translate into more cooperative behavior. These findings challenge
-assumptions about fostering cooperation and highlight the need for more
-nuanced interventions to promote reciprocity, especially for individuals
-prone to expecting rejection.
+This study found that exposing people to forgiving partners in economic games unexpectedly decreased their subsequent cooperation---likely because normal partners seemed less cooperative by comparison.
+Individuals high in rejection sensitivity showed a distinct pattern: while they detected trust violations as readily as others, they failed to restore cooperation afterward and became less responsive to their partner's behavior.
+In contrast, those low in rejection sensitivity appeared to learn from the positive exposure and became more reciprocal.
+These findings suggest that simply providing positive social experiences may not benefit everyone equally, and that individuals prone to rejection sensitivity may need targeted support to translate positive experiences into lasting behavioral change.
 
@@
 been studied as extensively, leaving a gap in our understanding of how
-these constructs might interplay. Could an ingrained sensitivity to
-rejection contribute to reduced pro-social behavior such as a failure to
-show reciprocity as a result of perceived rejection by the interaction
-partner?
+these constructs might interplay.
 
 Given that RS may be a manifestation of maladaptive attachment styles,
-it's crucial to explore whether exposure to consistently forgiving
-interaction partners could help reshape these interpersonal expectations
-and behaviors. If RS indeed influences reciprocity in repeated
-interactions, could such exposure serve as a corrective experience,
-potentially improving interpersonal outcomes? This approach aligns with
-attachment theory, which suggests that new, positive relational
-experiences can modify internal working models of relationships
-(Bowlby, 1988). Research in the fields of behavioral economics
-and psychology has also explored how positive social interactions
-influence cooperation and pro-social behavior. The use of the repeated
-trust game (RTG), a well-established experimental approach, has allowed
-for the analysis of the development of trust through ongoing
-interactions (Joyce et al., 1995). In this paradigm, cycles of mutual
-trust, where each party's trust is reciprocated with trustworthiness,
-have the effect of enhancing cooperative behaviors even among
-individuals who are initially inclined to be distrustful
-(King-Casas et al., 2005). Fowler \& Christakis (2010) studied behavior in
-social networks interacting in a public goods game and found that
-cooperative behavior tends to cluster, suggesting that exposure to
-cooperative peers can lead to more cooperative behavior. Similarly,
-research on social learning theory (Bandura, 1977) has long
-demonstrated that individuals learn and model the behavior of those
-around them, indicating that if someone is consistently exposed to
-cooperative and positive individuals, they're likely to emulate this
-behavior. These insights highlight that engaging with compassionate and
-forgiving others can be an effective method for mitigating deeply
-ingrained mistrust and potentially fostering trustworthy behavior.
+it is important to explore whether exposure to consistently forgiving
+interaction partners could reshape interpersonal expectations and
+behaviors. The \emph{corrective experience hypothesis}, rooted in attachment
+theory, suggests that new positive relational experiences can modify
+internal working models of relationships (Bowlby, 1988). Research
+on social learning (Bandura, 1977) similarly demonstrates that
+individuals model the behavior of those around them, and exposure to
+cooperative peers promotes cooperative behavior
+(Fowler \& Christakis, 2010). In the repeated trust game (RTG) paradigm,
+cycles of reciprocated trust enhance cooperative behaviors even among
+initially distrustful individuals (King-Casas et al., 2005). This
+perspective predicts that exposure to forgiving partners should increase
+subsequent cooperation, as participants internalize more positive
+expectations about social interactions. Given that high RS individuals
+show heightened sensitivity to social cues, they might be particularly
+responsive to such corrective experiences.
 
@@
 The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS :
-High or Low) by 2 (Phase: Trust-Game Pre Manipulation, Trust-Game Post
-Manipulation) design, with repeated measures on the Phase factor (Figure
+High or Low) by 2 (Phase: Trust-Game Pre-Exposure, Trust-Game Post-Exposure) design, with repeated measures on the Phase factor (Figure
 \ref{fig:HMMPanels}.A). Participants within each pre-screened RS group
@@
 us to probe efforts to repair it, the computerised agent was programmed
-to provide a low investment on round 12 (pre-manipulation) and round 13
-(post-manipulation). On all other rounds, the investor's actions were
+to provide a low investment on round 12 (pre-exposure) and round 13
+(post-exposure). On all other rounds, the investor's actions were
 determined by randomly drawing an investment from the state-conditional
@@
 forgiving HMM investor and rated the agent on the same attributes as in
-the pre-manipulation phase. Those in the Control condition faced the
+the pre-exposure phase. Those in the Control condition faced the
 same human-like HMM agent as in the ``pre'' phase and rated each co-player
@@
 very low investment in round 5 of each of the three games. In the
-post-manipulation phase (``post''), participants in both conditions faced
+post-exposure phase (``post''), participants in both conditions faced
 the same human-like HMM as in ``pre'' phase.
@@
 
-{\centering \includegraphics[width=1\linewidth,height=1\textheight]{article_files/figure-latex/HMMPanels-1}
+{\centering \includegraphics[width=0.95\linewidth]{article_files/figure-latex/HMMPanels-1}
 
@@
 
-\caption{\small{A: A timeline of the experiment for both conditions.
-The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor.
-The investor is endowed with 20 units at the start of each round.
-They need to decide how much of that endowment they want to invest with the trustee.
-The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor.
-The difference between conditions is the type of agents participants are exposed to in the exposure phase.
-Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG.
-From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B.
-The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C.
-Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1.
-Panel D shows the transition probabilities of the forgiving HMM agent in the low trust state.
-Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state, as a proxy for coaxing behavior.
-Transitions in the medium and high trust states were identical for both agents.}}\label{fig:HMMPanels}
+\caption{A: Experiment timeline.
+Participants (trustees) played RTGs with HMM investor agents.
+The investor sends investments (multiplied by 3) and participants decide returns.
+Conditions differ in exposure phase agents.
+B-D: The artificial investor is a three-state HMM fitted to human data.
+B: Investment distributions by latent state.
+C: Transition probabilities to states at t+1 as a function of net return at t; each panel shows a different starting state.
+D: Forgiving HMM transitions from low-trust state---unlike the human-like HMM, it always exits low-trust and favors high-trust transitions.}\label{fig:HMMPanels}
 \end{figure}
@@
 
-To test whether participants behaved differently in the RTG after the
-manipulation compared to the Control group, we model the percentage
-return (percentage of tripled investment returned to investor) using a
-linear mixed-effects model with Phase (RTG game pre vs.
-post-manipulation), Condition (manipulation vs.~Control), Investment,
-and RS (High vs Low RS group) as well as their interactions as fixed
-effects, and player-wise random intercepts and slopes for Phase. The
-full specification of the statistical model can be found in the
-supplement. To test whether participants behaved differently between
-conditions in the Exposure phase, we fit a linear mixed effects model to
-participants returns only in the exposure phase, with Condition
-(manipulation vs.~Control), Investment, and RS (High vs Low RS group) as
-well as their interactions as fixed effects, and player-wise random
-intercepts. Finally, to test whether the HMM agent's behavior differed
-between Phases, Conditions and RS groups, we estimate a linear
-mixed-effects model of investments sent by the computerised HMM agent
-with Condition, Phase and RS and their interaction as fixed effects, and
-a similar random effects structure to the returns model.
-
-The model was estimated using the \texttt{afex} package (Singmann et al., 2022)
-in R. More complex models with additional random effects could not be
-estimated reliably, and as such the estimated model can be considered to
-include the optimal random effects structure
-(Matuschek et al., 2017). A similar process was used to establish the
-random effects structures of linear mixed-effects models used to analyse
-the HMM agent investments as well as the participants' ratings of the
-co-players. There is no agreed upon way to calculate effect sizes for
-mixed effects models. Instead, we will report on testing differences in
-marginal means. For the \(F\)-tests, we used the Kenward-Roger
-approximation to the degrees of freedom, as implemented in the R package
-``afex''. We Z-transform the Investment variable (subtract the overall
-investment mean and divide by overall standard deviation) as centering
-is beneficial to interpreting the main effects more easily in the
-presence of interactions. To probe significant interactions, we
-conducted planned contrasts using the \texttt{emmeans} package in R. Given that
-we were testing multiple pre-planned comparisons, we applied the ``Sidak''
-correction to control for familywise error rate while maintaining
-reasonable statistical power. This approach allowed us to investigate
-specific hypotheses about the differential effects of our manipulation
-across phases and RS groups, while protecting against inflated Type I
-error rates.
+We analyzed participants' behavior in the RTG using linear mixed-effects models.
+First, to examine the effect of the manipulation, we modeled the percentage return (percentage of tripled investment returned to investor) as a function of Phase (RTG game pre vs.~post-exposure), Condition (Manipulation vs.~Control), Investment, and RS (High vs Low RS group), including all interactions as fixed effects.
+This model included player-wise random intercepts and slopes for Phase.
+Second, we analyzed behavior during the Exposure phase specifically, modeling returns with Condition, Investment, and RS and their interactions as fixed effects, along with player-wise random intercepts.
+Third, to verify the consistency of the HMM agent, we modeled the investments sent by the computerized agent using Condition, Phase, and RS and their interactions as fixed effects.
+To isolate effects occurring prior to any pre-programmed low investment, we also analyzed returns in rounds preceding the low investment trials only (rounds 1-11 in the pre-exposure phase and rounds 1-12 in the post-exposure phase) using the same model specification.
+Finally, to rigorously assess participants' reactions to and recovery from the specific instance of pre-programmed low investment, we conducted an event study analysis centered on the low investment round (\(t=0\)).
+We analyzed percentage returns in a three-round window (\(t-1\) to \(t+1\)) using a linear mixed-effects model with Phase, Condition, Time Point, and RS Group as fixed effects.
+We specifically examined two key behavioral responses: the Drop (change in return from \(t-1\) to the low investment round) and the Recovery (change in return from the low investment round to \(t+1\)).
+The full specification of all statistical models can be found in the supplement.
+
+All models were estimated using the \texttt{afex} package (Singmann et al., 2022) in R.
+We determined the random effects structure by starting with the maximal model and simplifying until convergence was achieved, ensuring the optimal structure (Matuschek et al., 2017).
+A similar process was applied to the models analyzing HMM agent investments and participant ratings.
+We report differences in marginal means rather than effect sizes, as there is no consensus on effect size calculation for mixed models.
+\(F\)-tests used the Kenward-Roger approximation for degrees of freedom.
+The Investment variable was Z-transformed to facilitate the interpretation of main effects in the presence of interactions.
+Significant interactions were probed using planned contrasts with the \texttt{emmeans} package.
+We applied the ``Sidak'' correction for multiple comparisons to control the familywise error rate while maintaining statistical power.
 
@@
 
-Mixed-effects analysis on the percentage returns shows a significant
-main effect of Phase (Pre vs.~Post RTG game),
-\(F(1, 201.63) = 5.81\), \(p = .017\), with
-higher percentage returns in the Pre RTG compared to the Post.
-Importantly, we also find an interaction between Condition and Phase
-(RTG pre- vs.~post-manipulation),
-\(F(1, 201.63) = 4.38\), \(p = .038\).
-As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that,
-contrary to our expectations, there was a decrease in the percentage
-returned only in the Manipulation condition, pre - post,
-\(\Delta M = 0.03\), 95\% CI \([0.01, 0.05]\), \(t(201.50) = 3.15\), \(p = .002\),
-but no change in the Control condition. We find no interaction between
-Phase, Condition and RS, suggesting that there was no difference between
-RS groups for this interaction.
-
-There was also a significant main effect of Investment,
-\(F(1, 5955.67) = 325.35\), \(p < .001\),
-such that higher investments were associated with higher percentage
-returns indicating positive reciprocity. An Investment by Condition
-interaction,
-\(F(1, 5955.67) = 13.92\), \(p < .001\),
-reflected that returns were more affected by investments in the Control
-condition. We also find a three way interaction between Phase,
-Investment and RS,
-\(F(1, 5864.62) = 7.84\), \(p = .005\),
-showing that the differentiated effect of the investment on the
-proportion returned by RS group is itself moderated by the Phase (pre-
-vs post manipulation). Finally, we find a four-way interaction between
-Condition, Phase, Investment and RS
-\(F(1, 5864.62) = 9.24\), \(p = .002\).
+Participants returned higher percentages in the Pre phase compared to
+the Post phase
+(\(F(1, 201.63) = 5.81\), \(p = .017\)). This
+effect was moderated by Condition
+(\(F(1, 201.63) = 4.38\), \(p = .038\)):
+contrary to our expectations, participants in the Manipulation condition
+decreased their returns from pre to post
+(\(\Delta M = 0.03\), 95\% CI \([0.01, 0.05]\), \(t(201.50) = 3.15\), \(p = .002\)),
+while those in the Control condition showed no change (Figure
+\ref{fig:boxPlots}). RS did not moderate this Condition × Phase
+interaction.
+
+Higher investments elicited higher percentage returns, indicating
+positive reciprocity
+(\(F(1, 5955.67) = 325.35\), \(p < .001\)).
+This relationship was stronger in the Control condition than in the
+Manipulation condition
+(\(F(1, 5955.67) = 13.92\), \(p < .001\)).
+The effect of investment on returns varied by RS group and Phase
+(\(F(1, 5864.62) = 7.84\), \(p = .005\)),
+and a four-way interaction indicated that these patterns further
+differed across Conditions
+(\(F(1, 5864.62) = 9.24\), \(p = .002\)).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=\textwidth]{article_files/figure-latex/boxPlots-1}
+
+}
+
+\caption{Marginal means of percentage returns (top) and HMM investments (bottom) by Phase and Condition.
+Bars show estimated marginal means; error bars represent 95\% confidence intervals.
+Participants in the Manipulation condition returned lower proportions post-exposure compared to pre-exposure (** p < .01), while Control participants showed no change.
+HMM investment did not differ across Phases or Conditions.}\label{fig:boxPlots}
+\end{figure}
 
@@
 of how the effect of investment on returns changed from pre- to
-post-manipulation for different RS groups in both conditions. Starting
+post-exposure for different RS groups in both conditions (Figure \ref{fig:fourwayInteractionPlot}). Starting
 with the Manipulation condition, for participants with low RS, the
 effect of investment on returns increased significantly from pre- to
-post-manipulation,
+post-phase,
 \(\Delta M = 0.03\), \(95\%\ \mathrm{CI}_\mathrm{\scriptstyle Sidak(3)}\) \([0.01, 0.05]\), \(t(5881.28) = 3.15\), \(p_\mathrm{\scriptstyle Sidak(3)} = .005\).
@@
 participants with high RS, the effect of investment on returns decreased
-significantly from pre- to post-manipulation,
+significantly from pre- to post-exposure,
 \(\Delta M = -0.02\), \(95\%\ \mathrm{CI}_\mathrm{\scriptstyle Sidak(3)}\) \([-0.04, 0.00]\), \(t(5891.81) = -2.67\), \(p_\mathrm{\scriptstyle Sidak(3)} = .023\).
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{article_files/figure-latex/boxPlots-1}
-
-}
-
-\caption{Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition.
-The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns.
-The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.
-RS did not affect returns and investments by Phase and Condition.}\label{fig:boxPlots}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{article_files/figure-latex/fourwayInteractionPlot-1} \caption{Marginal effect of investment on percentage returns by Phase, Condition, and RS group.
+Bars show estimated slopes (change in returns per SD increase in investment) from the mixed model; error bars represent 95\% confidence intervals.
+In the Manipulation condition, Low RS participants became more responsive to investments post-exposure (* p < .05), while High RS participants became less responsive (* p < .05).
+No significant changes were observed in the Control condition.}\label{fig:fourwayInteractionPlot}
 \end{figure}
 
-\subsubsection{Post Defection Trials}\label{post-defection-trials}
-
-Did participants learn to be more forgiving and cooperative after
-witnessing the pre-programmed defection by the HMM investor? To explore
-this question, we restrict the analysis to the trials following the
-pre-programmed defection by the HMM agent in both the ``pre'' (trials 12
-to 15) and the ``post'' phases (trials 13 to 15). We fit the same mixed
-effects model as for all the trials with the exception of the RS
-variable. This is because RS did not show main effects in the previous
-model, and also due to the necessity of running a simpler model to
-accommodate the low number of trials. We find a significant main effect
-of Phase \(F(1, 227.47) = 4.76\), \(p = .030\)
-with returns lower in the second game post defection trials compared to
-the first. We also find a main effect of Investment
-\(F(1, 1296.14) = 156.41\), \(p < .001\)
-where participants continued to return higher proportions when receiving
-higher investments. Finally, we still find an Investment by Condition
-interaction
-\(F(1, 5955.67) = 13.92\), \(p < .001\)
-showing a lower effects of investment on the Manipulation condition
-compared to the Control condition in post-defection trials. However, the
-absence of a Condition by Phase effect indicates there was no difference
-between conditions on participants' reaction to a one-off low investment
-by the co-player.
+\subsubsection{Returns Prior to Pre-Programmed Low Investment Trials}\label{returns-prior-to-pre-programmed-low-investment-trials}
+
+To distinguish between contrast effects and betrayal aversion as explanations for reduced cooperation in the Manipulation condition, we examined returns in the rounds preceding the pre-programmed low investment.
+If contrast effects were operating, participants in the Manipulation condition should already show reduced returns before experiencing any low investment in the post-exposure phase.
+Conversely, if betrayal aversion were the primary mechanism, group differences should only emerge after the low investment.
+
+The Phase \(\times\) Condition interaction was significant in rounds prior to the low investment (\(F(1, 204.05) = 4.86\), \(p = .029\)).
+Participants in the Manipulation condition significantly decreased their returns from pre- to post-exposure phase even before encountering the low investment (\(\Delta M = 0.03\), 95\% CI \([0.01, 0.06]\), \(t(202.02) = 2.83\), \(p = .005\)), whereas those in the Control condition showed no change (\(\Delta M = 0.00\), 95\% CI \([-0.02, 0.02]\), \(t(201.17) = -0.15\), \(p = .878\)).
+The four-way interaction also remained significant (\(F(1, 4454.34) = 7.12\), \(p = .008\)), suggesting that the differential responsiveness to investments observed in the full analysis was likewise present before the low investment occurred.
+
+\subsubsection{Reaction to Pre-programmed Low Investment: Event Study Analysis}\label{reaction-to-pre-programmed-low-investment-event-study-analysis}
+
+To understand how participants reacted to and recovered from the pre-programmed low investment, an event study analysis was conducted centered on the low investment round (Figure \ref{fig:eventStudyPlot}).
+Two behavioral responses were examined: the Drop (change in returns at the low investment round relative to \(t-1\), where negative values indicate reduced returns reflecting punishment of low trust) and Recovery (change in returns at \(t+1\) relative to the low investment round, where positive values indicate restored cooperation).
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{article_files/figure-latex/eventStudyPlot-1} \caption{Drop and Recovery responses to pre-programmed low investment in the post-exposure phase (pre-exposure phase not shown as no between-condition differences were observed).
+Drop = change in returns from low investment round minus t-1 (negative values indicate reduced returns); Recovery = change from t+1 minus low investment round (positive values indicate restored returns).
+Bars show estimated marginal means from the mixed model; error bars represent 95\% confidence intervals.
+The bracket shows the significant between-condition difference in Recovery for High RS participants.
+** p < .01.}\label{fig:eventStudyPlot}
+\end{figure}
+
+In the pre-exposure phase, there were no significant differences between conditions in either the Drop (\(M = -0.02\), 95\% CI \([-0.09, 0.05]\), \(t(1773.82) = -0.61\), \(p = .543\)) or Recovery (\(M = 0.00\), 95\% CI \([-0.06, 0.07]\), \(t(1773.56) = 0.11\), \(p = .914\)), confirming that both groups started with equivalent behavioral patterns.
+In the post-exposure phase, a divergence emerged.
+The Control group showed a significantly larger Drop than the Manipulation group (\(M = -0.08\), 95\% CI \([-0.15, -0.02]\), \(t(1773.93) = -2.46\), \(p = .014\)), indicating that participants exposed to the forgiving agent showed a blunted immediate reaction to the low investment and did not reduce their returns as sharply.
+The subsequent Recovery did not differ significantly between conditions overall (\(M = 0.07\), 95\% CI \([0.00, 0.13]\), \(t(1774.02) = 1.92\), \(p = .055\)).
+
+When examining moderation by RS, the pattern appeared to be driven primarily by high RS participants.
+Low RS participants showed no significant difference in Recovery between conditions (\(M = 0.00\), 95\% CI \([-0.09, 0.10]\), \(t(1774.24) = 0.06\), \(p = .954\)), with both groups displaying modest, similar recovery patterns after the low investment.
+The larger Drop observed in the Control condition for this group was partly attributable to their elevated cooperation level at \(t-1\).
+
+High RS participants showed a different pattern.
+In the Control condition, they demonstrated trust repair by significantly increasing their returns after the low investment (\(\Delta M = 0.14\), 95\% CI \([0.08, 0.21]\), \(t(1774.05) = 4.19\), \(p < .001\)).
+However, those in the Manipulation condition failed to recover, showing no significant increase in returns at \(t+1\) (\(\Delta M = 0.01\), 95\% CI \([-0.05, 0.08]\), \(t(1773.54) = 0.42\), \(p = .676\)).
+The difference in Recovery between conditions was significant (\(M = 0.13\), 95\% CI \([0.03, 0.22]\), \(t(1773.79) = 2.66\), \(p = .008\)).
+
+In summary, the forgiveness intervention appeared to dampen reciprocal responsiveness, hindering the re-establishment of cooperation following a temporary withdrawal of trust.
+This effect was more pronounced among high RS individuals.
+While high RS participants in the Control condition demonstrated active reciprocity by reducing returns sharply when trust was withdrawn and increasing them when trust was restored, those exposed to the forgiving agent exhibited a disengaged pattern characterized by a muted reaction to the low investment and a failure to reinstate high returns afterward.
 
@@
 
-Figure \ref{fig:plotRatings} shows participants' ratings of each player
-they faced by condition and RS group. We will focus on two contrasts to
-analyse the ratings by Condition and RS group. The first is between the
-rating in the first phase (``pre'') when participants phase the human-like
-HMM, and the average rating during the exposure phase (average of
-``expo1'', ``expo2'' and ``expo3'') where they either face the forgiving HMM
-(Manipulation condition) or the human-like HMM again (Control
-condition). The second contrast is between the ``pre'' and ``post'' phases
-of the experiment where in both conditions participants face the same
-human-like HMM.
-
-\subsubsection{Comparing Pre and Exposure Ratings}\label{comparing-pre-and-exposure-ratings}
-
-For those with high RS, participants in the Manipulation condition rated
-the investors they faced in the exposure phase (the forgiving HMM) as
-more Cooperative
-\(\Delta M = 2.57\), 95\% CI \([0.84, 4.30]\), \(t(808) = 2.91\), \(p = .004\).
-There was no difference in ratings on forgiveness and whether they would
-like to face the co-players again. Those in the Control condition rated
-the investors faced in the exposure group (same HMM) as less cooperative
-\(\Delta M = -2.65\), 95\% CI \([-4.37, -0.94]\), \(t(808) = -3.04\), \(p = .002\),
-less forgiving
-\(\Delta M = -2.19\), 95\% CI \([-4.00, -0.39]\), \(t(808) = -2.38\), \(p = .017\),
-and were less keen on facing them again
-\(\Delta M = -3.62\), 95\% CI \([-5.73, -1.50]\), \(t(808) = -3.36\), \(p < .001\).
-
-For those with low RS, there was no difference in any of the ratings
-between the ``pre'' and ``exposure'' phases for the Manipulation condition.
-For the Control condition, participants indicated less willingness to
-face the exposure co-player compared to the ``pre'' player,
-\(\Delta M = -2.43\), 95\% CI \([-4.52, -0.34]\), \(t(808) = -2.29\), \(p = .023\)
-but also did not differ on cooperation and forgiveness ratings. In
-summary, those with low RS had a mostly undifferentiated perception of
-players between the pre and exposure phases, even when the co-player was
-in fact designed to be more forgiving. In contrast, we see that exposing
-participants with high RS to a more cooperative and more forgiving agent
-has compensated for the decrease in ratings that would have occurred if
-faced with a co-player with the exact same strategy.
+Figure \ref{fig:plotRatings} shows participants' ratings of co-players across phases.
+We examined two contrasts: pre-exposure versus exposure phase ratings, and pre-exposure versus post-exposure ratings.
+
+High RS participants showed more differentiated perceptions of the agents.
+In the Manipulation condition, they rated the forgiving agents as more cooperative during exposure (\(\Delta M = 2.57\), 95\% CI \([0.84, 4.30]\), \(t(808) = 2.91\), \(p = .004\)).
+In the Control condition, however, high RS participants rated the same human-like HMM progressively more negatively---lower on cooperation (\(\Delta M = -2.65\), 95\% CI \([-4.37, -0.94]\), \(t(808) = -3.04\), \(p = .002\)), forgiveness (\(\Delta M = -2.19\), 95\% CI \([-4.00, -0.39]\), \(t(808) = -2.38\), \(p = .017\)), and willingness to play again (\(\Delta M = -3.62\), 95\% CI \([-5.73, -1.50]\), \(t(808) = -3.36\), \(p < .001\))---despite the agent's strategy remaining unchanged.
+Low RS participants showed largely undifferentiated perceptions between pre and exposure phases regardless of condition.
+
+Comparing pre to post-exposure ratings revealed a contrast effect: after experiencing the forgiving agent, both RS groups in the Manipulation condition rated the post-exposure agent (identical to pre-exposure) more negatively on forgiveness (High RS: \(\Delta M\) = -0.88, \(SE\) = 0.38, \(t\)(808.0) = -2.33, p = .020; Low RS: \(\Delta M\) = -1.14, \(SE\) = 0.38, \(t\)(808.0) = -2.98, p = .003) and willingness to play again (High RS: \(\Delta M\) = -1.29, \(SE\) = 0.44, \(t\)(808.0) = -2.92, p = .004; Low RS: \(\Delta M\) = -1.30, \(SE\) = 0.45, \(t\)(808.0) = -2.90, p = .004).
+Low RS participants in the Control condition showed stable ratings, accurately perceiving the consistent agent strategy, while high RS participants in the Control condition continued their negative drift (Cooperation: \(\Delta M\) = -0.96, \(SE\) = 0.36, \(t\)(808.0) = -2.70, p = .007).
+These rating patterns converge with the behavioral findings, suggesting high RS individuals are particularly sensitive to relative comparisons between interaction partners.
 
@@
 
-\caption{Averages and standard errors of the participants ratings of the opponent (y-axis) by each game and condition for each phase (x-axis).
-Pre and post are the 15 round RTGs before and after the exposure phase respectively.
-The games titled expo 1 to 3 are the three 7 round games during the exposure phase.
-The blue line represents participants' perception of co-player cooperation, the red line indicates perceived co-player forgiveness, and the green line shows the participants' willingness to play again with the same co-player.
-Cooperation, forgiveness, and willingness to play again ratings remained relatively stable for the low RS group except post-manipulation, where they were lower.
-In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the Manipulation condition, but decreased over time in the Control group.
-Ratings were also markedly lower in the post-manipulation game for this group.}\label{fig:plotRatings}
+\caption{Participants' ratings of co-players by phase, condition, and RS group.
+Blue: perceived cooperation; red: perceived forgiveness; green: willingness to play again.
+Low RS participants showed stable ratings in the Control condition.
+High RS participants showed declining ratings in the Control condition despite unchanged agent strategy, and more differentiated perceptions in the Manipulation condition.}\label{fig:plotRatings}
 \end{figure}
-
-\subsubsection{Comparing Pre and Post Ratings}\label{comparing-pre-and-post-ratings}
-
-Participants high on RS in the Manipulation condition rated the
-co-players in the ``post'' phase similarly on cooperation, lower on
-forgiveness
-\(\Delta M = -0.88\), 95\% CI \([-1.63, -0.14]\), \(t(808) = -2.33\), \(p = .020\),
-and lower on willingness to face them again
-\(\Delta M = -1.29\), 95\% CI \([-2.16, -0.42]\), \(t(808) = -2.92\), \(p = .004\).
-Those in the Control condition rated the investors post the exposure
-phase lower on all three attributes (Cooperation:
-\(\Delta M = -0.96\), 95\% CI \([-1.66, -0.26]\), \(t(808) = -2.70\), \(p = .007\),
-Forgiveness:
-\(\Delta M = -1.00\), 95\% CI \([-1.74, -0.26]\), \(t(808) = -2.66\), \(p = .008\),
-Play again:
-\(\Delta M = -1.21\), 95\% CI \([-2.07, -0.35]\), \(t(808) = -2.76\), \(p = .006\)).
-
-For participants low on RS, those in the Manipulation condition rated
-the co-players in the ``post'' phase lower on all three attributes
-(Cooperation:\(\Delta M = -0.74\), 95\% CI \([-1.45, -0.03]\), \(t(808) = -2.03\), \(p = .042\),
-Forgiveness:
-\(\Delta M = -1.14\), 95\% CI \([-1.89, -0.39]\), \(t(808) = -2.98\), \(p = .003\),
-Play again:
-\(\Delta M = -1.30\), 95\% CI \([-2.18, -0.42]\), \(t(808) = -2.90\), \(p = .004\)).
-Those in the Control condition did not differ in their ratings of the
-``pre'' and ``post'' phase players.
-
-In summary, we again see that those with low RS accurately perceive the
-co-player as similar on all attributes throughout the phases in the
-Control condition. In contrast, the high RS group shows a negative bias
-towards the co-players after the ``pre'' phase in the Control condition
-even though the player continues to use the same strategy. After
-exposure to the forgiving HMM, both groups rate the ``post'' co-player
-worse than the ``pre'' even though they are the same.
 
@@
 comparison (Kobre \& Lipsitt, 1972). As the most recently faced opponents
-were more forgiving and cooperative, this negative contrast effect may
+were more forgiving (and consequently more cooperative), this negative contrast effect may
 have trumped any learning transfer from being repeatedly exposed to
@@
 
-This effect appears to operate independently of RS, as both groups
-exhibited it. However, RS did influence how participants perceived and
-reacted to their co-players' behavior. High RS individuals demonstrated
-a heightened attunement to changes in their co-players' behavior,
-accurately perceiving increased cooperativeness in the forgiving agent.
-Paradoxically, this same sensitivity may explain their decreased
-reactivity to investment levels and declining ratings of similar
-co-players over time in the Control condition. Occasional pre-programmed
-defections may have amplified their perceptions of untrustworthiness,
-leading to a more cautious approach and less responsiveness to
-investment variations. This aligns with the heightened social awareness
-and vigilance often associated with high RS (Downey \& Feldman, 1996).
-This finding suggests that individuals with high RS might benefit from
-interventions that focus on calibrating their perceptions of social cues
-in economic exchanges. In contrast, low RS individuals were less
-affected by the agent's behavioral variations, showing a more stable
-baseline of social perception. This stability in perception might
-reflect a lower propensity to scrutinize social cues or a more robust
-internal model of social interactions that is less easily perturbed by
-short-term variations in partner behavior. Their increased
-responsiveness to investments post-exposure in the Manipulation
-condition suggests they may have internalized the cooperative norms
-experienced, carrying over this increased reciprocity to subsequent
-interactions.
-
-Despite these differences in perception and investment reactivity, the
-actual cooperative behavior -- as measured by returns in the trust game --
-did not significantly differ between RS groups. This suggests a
-potential dissociation between the cognitive-perceptual processes
-involved in evaluating social partners and the behavioral output in
-economic games. There are multiple potential explanations for this:
-First, the ratings may reflect explicit, conscious evaluations of social
-partners, while behavior in the trust game might be guided more by
-implicit, automatic processes that are less differentiated between RS
-groups. This interpretation aligns with dual-process theories of social
-cognition (Lieberman, 2007). Second, the structured nature of the
-trust game might provide a context where the behavioral expression of RS
-is attenuated. In contrast, the more open-ended nature of providing
-ratings might allow for greater expression of RS-related perceptual
-biases.
-
-These findings have important implications for interventions aimed at
-promoting trust and cooperation. Simply exposing individuals to
-artificially highly cooperative or forgiving partners may not suffice to
-induce lasting changes in behavior, and may even lead to negative
-contrast effects. Instead, interventions should focus on developing
-adaptive strategies for navigating the complexities of real-world social
-interactions, which often involve a mix of cooperative and
-non-cooperative behaviors. Such approaches could gradually shape
-expectations and behaviors over time, taking into account individual
-differences in RS. For individuals with high RS, interventions might
-specifically target the calibration of social perceptions in economic
-exchanges, helping to bridge the gap between their heightened
-sensitivity to social cues and their actual cooperative behavior. Future
-research should investigate whether the observed dissociation between
-perception and behavior persists in more naturalistic social
-interactions or over longer periods. Additionally, exploring the
-temporal dynamics of these effects and the role of explicit
-communication in the cooperative process could offer valuable insights.
-Interventions combining realistic cooperative scenarios with clear
-communication about mutual expectations might prove more effective in
-fostering enduring cooperative tendencies that are robust across various
-social contexts (Balliet et al., 2011).
+An alternative explanation relates to differences in exposure-phase experiences: the Control agent performed pre-programmed low investments (round 5 of each game), whereas the Manipulation agent did not.
+Consequently, Manipulation participants experienced 21 consecutive rounds without any low investments before encountering the post-phase low investment---a substantially longer period of consistently positive interactions than Control participants.
+This raises the possibility that reduced cooperation in the Manipulation condition reflects heightened betrayal aversion rather than negative contrast effects.
+
+However, the analysis of returns prior to the pre-programmed low investment provides evidence against betrayal aversion as the primary mechanism.
+If betrayal aversion were driving the effect, group differences should only emerge after participants encountered the low investment in the post-exposure phase.
+Instead, Manipulation participants had already significantly reduced their returns in rounds 1-12 of the post-exposure phase---before any low investment occurred.
+This pattern is consistent with contrast effects operating from the beginning of the post-exposure phase, as participants immediately perceived the human-like agent as less cooperative compared to the forgiving agent they had just experienced.
+While betrayal aversion may contribute to specific aspects of the observed patterns, such as the impaired recovery following the low investment in high RS participants, it cannot account for the overall reduction in cooperation that was already evident before any low investment.
+
+While the negative contrast effect operated across RS groups, the pattern of responses to trust violations differed in ways that align with clinical models of rejection sensitivity.
+In the event study analysis, high and low RS participants showed comparable immediate reactions to the low investment (the Drop), indicating intact detection of trust violations regardless of RS level.
+However, the groups diverged in their subsequent recovery patterns: high RS participants in the Manipulation condition failed to restore cooperation following the low investment, whereas low RS participants and Control participants showed recovery.
+This dissociation between intact rejection detection and impaired relationship repair is consistent with research on social learning difficulties in individuals with elevated RS and related clinical presentations.
+Studies of borderline personality disorder, where RS is characteristically elevated, have documented specific deficits in updating social expectations following positive interpersonal experiences (Schuster et al., 2021; Staebler et al., 2011).
+Similarly, research on depression has identified ``cognitive immunization'' processes whereby negative schemas resist modification despite contradictory evidence (Kube et al., 2020).
+The high RS participants' failure to recover cooperation, despite prior exposure to consistently forgiving behavior, may reflect analogous difficulties in leveraging positive social experiences to update expectations and restore trust.
+
+The four-way interaction findings further support this interpretation.
+High RS participants showed decreased responsiveness to their co-player's investments following the forgiveness manipulation, a pattern suggestive of withdrawal from contingent social exchange.
+This reduced sensitivity to partner behavior parallels the self-silencing and social withdrawal documented in high RS populations, where anticipatory self-protection can paradoxically undermine relationship maintenance (Ayduk et al., 2000; Romero-Canyas et al., 2010).
+In contrast, low RS participants showed increased responsiveness to investments post-phase, suggesting they internalized the cooperative norms experienced during exposure and carried this forward to subsequent interactions.
+This differential capacity to benefit from positive social experiences maps onto broader findings that RS impedes the acquisition and transfer of adaptive interpersonal strategies (Pietrzak et al., 2005).
+
+The combination of blunted responsiveness to investments and absent recovery in high RS participants suggests a pattern of passive disengagement rather than active retaliation.
+While overall return levels did not differ between RS groups, these specific behavioral signatures indicate that RS does modulate particular aspects of cooperative behavior.
+The ratings data complement these findings: high RS participants showed more negative explicit evaluations of their co-players, rating them lower on forgiveness and willingness to interact again.
+This convergence between explicit ratings and behavioral patterns suggests that the effects of RS on social exchange are expressed across multiple response systems.
+The structured nature of the trust game may constrain RS effects to specific behavioral signatures (such as contingent responding and recovery) rather than overall cooperation levels, while the more open-ended nature of rating tasks allows for broader expression of RS-related evaluative biases (Lieberman, 2007).
+
+These findings have implications for interventions aimed at promoting trust and cooperation.
+The present results suggest that exposure to positive social models alone may be insufficient for high RS individuals, and may even produce iatrogenic effects through negative contrast.
+The specific deficits observed---impaired recovery from trust violations and reduced sensitivity to partner behavior---point to potential intervention targets.
+Approaches that focus on enhancing the capacity to update expectations following interpersonal ruptures may be more effective than simply providing positive experiences.
+This could include explicit training in recognizing repair attempts, practicing graduated trust restoration, or developing metacognitive awareness of the tendency toward disengagement following perceived rejection.
+Future research should examine whether these behavioral patterns generalize to naturalistic social contexts and whether targeted interventions can modify the updating and recovery deficits observed in high RS participants (Balliet et al., 2011).
 
@@
 While this study offers valuable insights into trust and cooperation
-dynamics in social interactions mediated by artificial agents, several
-limitations warrant consideration. The artificial nature of the
-interactions, while enabling high experimental control, may limit
-ecological validity, particularly regarding the paradigm's ability to
-trigger rejection sensitivity in participants. Additionally, the
-relatively brief exposure to forgiving HMM agents during the
-manipulation phase may have been insufficient to induce lasting changes
-in participants' reciprocity behavior. The online format of the
-experiment eliminates important social cues present in face-to-face
-interactions, which might influence trust and cooperation differently.
-Although the RSQ is a well-validated measure of rejection sensitivity
-(RS), future studies could benefit from a multi-method approach,
-potentially incorporating behavioral measures or additional self-report
-scales for a more comprehensive RS assessment. Furthermore, while our
-use of HMM agents to simulate human-like behavior in economic games is
-innovative, some participants may have suspected they were interacting
-with AI, potentially confounding the results. Despite these limitations,
-our findings provide a foundation for future research. Subsequent
-studies could address these constraints by incorporating face-to-face
-interactions, utilizing multiple measures of RS, examining the long-term
-effects of exposure to forgiving interaction partners, and developing
-more convincing cover stories or explicitly comparing responses to human
-and computerised partners.
+dynamics, several limitations warrant consideration. First, and most importantly, the Control and Manipulation conditions
+differed not only in the forgiveness of exposure-phase agents but also
+in whether participants experienced pre-programmed low investments
+during exposure (Control agents sent low investments at round 5 of each exposure game; Manipulation agents did not).
+This design feature means we cannot fully disentangle the effects of agent forgiveness per se from the effects of experiencing an uninterrupted sequence of positive interactions.
+Although the analysis of pre-low investment returns provides evidence favoring contrast effects over betrayal aversion as the primary mechanism, and both conditions ultimately faced the same human-like agent in the post-exposure phase, caution is warranted in attributing the observed effects specifically to ``forgiveness exposure.'' Future replications
+should equate these experiences across conditions to isolate the effects
+of agent forgiveness from the absence of negative experiences. Second, our extreme groups design for RS (selecting
+participants with RSQ scores \textgreater{} 15 or \textless{} 10) maximized power to detect
+moderation effects but may inflate effect sizes and limits
+generalizability to individuals with moderate RS. Future research should
+examine RS as a continuous variable. Third, the brief exposure phase
+(three 7-round games) may have been insufficient to induce lasting
+changes. Fourth, the online format eliminates social cues present in
+face-to-face interactions. Notably, while 41\% of participants believed
+they faced humans and a similar proportion were unsure, the observed
+effects emerged even among those who suspected AI, suggesting robustness
+of the findings. Despite these limitations, subsequent studies could
+address these constraints by incorporating face-to-face interactions,
+longer exposure periods, and continuous RS measurement.
 
@@
 
-Preregistration: The hypotheses and methods were not preregistered.The
-analysis plan was not preregistered. Materials: All study materials are
+Preregistration: The hypotheses and methods were not preregistered.
+The primary hypothesis, based on attachment theory, predicted that forgiveness exposure would increase cooperation.
+The alternative predictions regarding contrast effects and RS-specific updating biases were incorporated into the theoretical framework following initial data analysis, though both perspectives were grounded in existing literature.
+The analysis plan was not preregistered.
+Materials: All study materials are
 publicly available (\url{https://github.com/ismailg/exposure-public}). Data:
@@
 
-\phantomsection\label{refs}
+\protect\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
@@
 \url{https://doi.org/10.1176/appi.books.9780890425596}
+
+\bibitem[\citeproctext]{ref-ayduk_rejection_2000}
+Ayduk, O., Downey, G., \& Kim, M.
+(2000).
+Rejection sensitivity and depressive symptoms in women.
+\emph{Personality and Social Psychology Bulletin}, \emph{26}(8), 909--919.
+\url{https://doi.org/10.1177/0146167200269001}
 
@@
 \url{https://doi.org/10.1016/0022-0965(72)90033-1}
+
+\bibitem[\citeproctext]{ref-kube_distorted_2020}
+Kube, T., Schwarting, R., Rozenkrantz, L., Glombiewski, J.
+A., \& Rief, W.
+(2020).
+Distorted cognitive processes in major depression: A predictive processing perspective.
+\emph{Biological Psychiatry}, \emph{87}(5), 388--398.
+\url{https://doi.org/10.1016/j.biopsych.2019.07.017}
 
@@
 
+\bibitem[\citeproctext]{ref-pietrzak_appearance-rejection_2005}
+Pietrzak, R.
+H., Downey, G., \& Ayduk, O.
+(2005).
+Appearance-rejection sensitivity predicts body dysmorphic disorder symptoms and cosmetic surgery.
+\emph{Annals of Clinical Psychiatry}, \emph{17}(4), 213--219.
+\url{https://doi.org/10.1080/10401230500295471}
+
 \bibitem[\citeproctext]{ref-richetin_emotional_2018}
@@
 \emph{Clinical Neuropsychiatry: Journal of Treatment Evaluation}, \emph{15}(4), 200--205.
+
+\bibitem[\citeproctext]{ref-romero-canyas_rejection_2010}
+Romero-Canyas, R., Downey, G., Berenson, K.
+R., Ayduk, O., \& Kang, N.
+J.
+(2010).
+Rejection sensitivity and the rejection-hostility link in romantic relationships.
+\emph{Journal of Personality}, \emph{78}(1), 119--148.
+\url{https://doi.org/10.1111/j.1467-6494.2009.00611.x}
+
+\bibitem[\citeproctext]{ref-schuster_ambiguous_2021}
+Schuster, F., Hoerz-Sagstetter, S., Seidl, E., Mauer, C., Zeiss, M., Reinhard, M.
+A., Padberg, F., \& Jobst, A.
+(2021).
+Ambiguous social rejection from a close other affects neural and behavioral responses in borderline personality disorder.
+\emph{Personality Disorders: Theory, Research, and Treatment}, \emph{12}(6), 583--594.
+\url{https://doi.org/10.1037/per0000454}
 
@@
 
+\bibitem[\citeproctext]{ref-staebler_facial_2011}
+Staebler, K., Renneberg, B., Stopsack, M., Fiedler, P., Weiler, M., \& Roepke, S.
+(2011).
+Facial emotional expression in reaction to social exclusion in borderline personality disorder.
+\emph{Psychological Medicine}, \emph{41}(9), 1929--1938.
+\url{https://doi.org/10.1017/S0033291711000080}
+
 \bibitem[\citeproctext]{ref-weekers_level_2019}
